### 2016-01-03 paper.tex
add files

\documentclass{acm_sen_article}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
%\renewcommand{\baselinestretch}{0.93}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Software Model Checking in Embedded Systems}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{Lucas C. Cordeiro \\
Electronic and Information Research Center\\
Federal University of Amazonas, Brazil\\
Email: lucascordeiro@ufam.edu.br
\and
Eddie B. de Lima Filho \\
Center for Science, Technology, and Innovation \\ for the Industrial Pole of Manaus, Brazil \\
Email: eddie@ctpim.org.br}
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678--2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Our reliance on the correct functioning of embedded systems is growing rapidly. Such systems are used in a wide range of applications such as micro-grids, airbag control systems, mobile phones, and high-end television sets. These systems are becoming more and more complex and require multi-core processors with scalable shared memory to meet the increasing computational power demands. The reliability of the embedded (distributed) software is thus a key issue in the system development. State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification and high-level processor model during simulation. However, verification of embedded systems raises additional challenges: meet time and energy constraints, handle concurrent software, platform-dependent software, and legacy designs (usually written in low-level languages). This paper describes challenges, problems, and recent advances to ensure the correctness and the timeliness of embedded systems. Reliability issues in the development of modern micro-grids are considered as a prominent model checking application.
\end{abstract}

%\IEEEpeerreviewmaketitle

%--------------------------------------
\section{Introduction}
%--------------------------------------

Embedded computer systems are used in a variety of sophisticated applications, which range from safety-critical systems such as nuclear reactors and automotive controllers, to entertainment software such as games and graphics animation~\cite{Kopetz11}. Embedded systems are ubiquitous in modern day information systems and are becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance. As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, success of such system depends on the relevance and quality of service it can provide.

A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy. Additionally, the increasing computational power and decreasing size and cost of processors is enabling system designers to move more functionalities to software. This leads to difficulties in verifying design correctness since stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered.

%--------------------------------------
\section{Verification Challenges}
%--------------------------------------

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation. However, verification of embedded systems raises additional challenges, which include: meet time and energy constraints, handle concurrent software, platform-dependent software, and legacy designs (usually written in low-level languages). Thus, the main challenge for verifying embedded systems relies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification. These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.

%--------------------------------------
\section{Research Problem (RP)}
%--------------------------------------

This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving the correctness and the timeliness of embedded systems, by taking into account stringent constraints imposed by the hardware.

%--------------------------------------
\section{Bounded Model Checking of Embedded Systems}
%--------------------------------------

Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and to find subtle bugs in real programs~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}. The idea behind the BMC techniques is to check for the violation of a given property at a given depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition (VC) $ \psi $ such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
less than or equal to \textit{k}.

In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the theories supported by SMT solvers ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency). Moreover, BMC techniques are only able to falsify properties up to the given depth \textit{k}; they are not able to prove the correctness of the system, unless an upper bound of \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth. In particular, BMC techniques limit the visited regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$. This limits the state space that needs to be explored during verification, leaving enough that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found; BMC tools are, however, susceptible to exhaustion of time or memory limits for programs with loops whose bounds are too large or cannot be determined statically.

%--------------------------------------
\subsection{Induction-based Verification of C/C++ Programs}
%--------------------------------------

One technique typically used to prove properties, for any given depth, is mathematical induction. In particular, the algorithm called \textit{k}-induction was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and to respect time constraints specified during the design phase of a system~\cite{EenS03}. Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied due to the monolithic transition relation present 
in hardware designs~\cite{EenS03,GrosseLD09,Sheera00}. Recently, Gadelha {\it et al.} and Beyer {\it et al.} contribute a new algorithm to prove correctness of (a large set of) C programs by mathematical induction in a completely automatic way ({\it i.e.}, the user does not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence in the literature that model checking embedded software using \textit{k}-induction (and invariants) can be exploited efficiently in embedded systems verification~\cite{Rocha15}.

%--------------------------------------
\section{Achievements}
\label{achievements}
%--------------------------------------

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based context-BMC for full C/C++ programs (ESBMC)~\cite{Cordeiro12,ECBS13}. \textbf{(RP2)} This approach was further extended to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}. These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark’s creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC). \textbf{(RP3)} The SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems ({\it e.g.}, digital filters and controllers)~\cite{esbmc_controller,dsv_spin2015,esbmc_filter}. \textbf{(RP3)} Finally, novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in real-life embedded applications~\cite{Gadelha15,Rocha15}. Last but not least, ESBMC represents one of the most prominent BMC tools in the international software verification competition~\cite{MorseCNF13,MorseRCN014}.

%--------------------------------------
\section{Conclusions}
\label{conclusions}
%--------------------------------------

This paper presented the main challenges to verify design correctness in embedded systems. In particular, stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification. Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants) can be applied to formally verify the correctness and the timeliness of embedded systems. Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers. As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to modern micro-grids, considering small-scale versions of a centralized electricity system so that reliability and carbon emission reduction are amenable to verification.

% conference papers do not normally have an appendix

% use section* for acknowledgment
%\section*{Acknowledgment}

%The authors would like to thank...

\begin{thebibliography}{1}

\bibitem{Kopetz11}
H.~Kopetz.
\newblock {\em Real-Time Systems - Design Principles for Distributed Embedded Applications}.
\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.

\bibitem{handbook09}
Biere A., Heule M., van Maaren H., Walsh T., eds.:
\newblock Handbook of Satisfiability.
\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.

\bibitem{BarrettSST09}
Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
\newblock 26.
\newblock In: Satisfiability Modulo Theories. Volume 185 of Frontiers in
  Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.

\bibitem{Clarke04}
Clarke E., Kroening D., Lerda F.:
\newblock A tool for checking {ANSI-C} programs.
\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg,  pp. 168--176, 2004.

\bibitem{MerzFS12}
Merz F., Falke S., Sinz C.:
\newblock {LLBMC}: Bounded model checking of {C} and {C++} programs using a
  compiler ir.
\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. Springer-Verlag, pp. 146--161, 2012.

\bibitem{CordeiroF11}
Cordeiro L., Fischer B.:
\newblock Verifying multi-threaded software using {SMT}-based context-bounded
  model checking.
\newblock In: International Conference on Software Engineering. ACM,  pp. 331--340, 2011.

\bibitem{Ivancic05}
Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
\newblock Model checking {C} programs using {F-Soft}.
\newblock In: IEEE International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.

\bibitem{Cordeiro12}
Cordeiro L., Fischer B., Marques{-}Silva J.:
\newblock Smt-based bounded model checking for embedded {ANSI-C} software.
\newblock {IEEE} Trans. Software Eng. \textbf{38}(4), pp. 957--974, 2012.

\bibitem{Donaldson10}
Donaldson A., Kroening D., R{\"{u}}mmer P.:
\newblock {SCRATCH}: A tool for automatic analysis of dma races.
\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.

\bibitem{Kinductor}
Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
\newblock Software verification using k-induction.
\newblock In: International Conference on Static Analysis. Springer-Verlag, pp. 351--368, 2011.

\bibitem{EenS03}
E{\'{e}}n, N., S{\"{o}}rensson, N.:
\newblock Temporal induction by incremental {SAT} solving.
\newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.

\bibitem{GrosseLD09}
Gro{\ss}e D., Le H., Drechsler R.:
\newblock Induction-based formal verification of {SystemC TLM} designs.
\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.

\bibitem{Sheera00}
Sheeran M., Singh S., St{\aa}lmarck G.:
\newblock Checking safety properties using induction and a {SAT}-solver.
\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.

\bibitem{Gadelha15}
Gadelha M., Ismail H., Cordeiro L.:
\newblock Handling loops in bounded model checking of C programs via k-induction
\newblock In: International Journal on Software Tools for Technology Transfer (to appear), pp. 1--18, 2015.
\newblock http://dx.doi.org/10.1007/s10009-015-0407-9

\bibitem{Beyer15}
Beyer D., Dangl M., Wendler P.:
\newblock Boosting k-Induction with Continuously-Refined Invariants.
\newblock International Conference on Computer-Aided Verification, pp. 622--640, 2015.

\bibitem{Rocha15}
Rocha H., Ismail H., Cordeiro L., Barreto R.:
\newblock Model Checking Embedded C Software using k-Induction and Invariants. 
\newblock V Brazilian Symposium on Computing Systems Engineering (to appear), pp. 1--6, 2015.

\bibitem{ECBS13}
Ramalho M., Lopes M., Sousa F., Marques H., Cordeiro L., Fischer B.:
\newblock {SMT-Based Bounded Model Checking of C++ Programs}.
\newblock In: International Conference on Engineering of Computer-Based Systems, pp. 147--156, 2013.

\bibitem{Pereira15}
Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
\newblock Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking. 
\newblock To appear in ACM Symposium on Applied Computing, Software Verification and Testing, 2016.

\bibitem{esbmc_controller}
Bessa I., Abreu R., Cordeiro L., Filho J.:
\newblock {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers)}. 
\newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.

\bibitem{dsv_spin2015}
Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
\newblock {DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.

\bibitem{esbmc_filter}
Abreu R., Cordeiro L., Filho E.:
\newblock {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}. 
\newblock In: XXXI Brazilian Symposium on Telecommunications, 2013.
\newblock http://dx.doi.org/10.14209/sbrt.2013.57

\bibitem{MorseCNF13}
Morse J., Cordeiro L., Nicole D., Fischer B.:
\newblock Handling unbounded loops with {ESBMC} 1.20.
\newblock In: Tools and Algorithms for the Construction and Analysis of
  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.

\bibitem{MorseRCN014}
Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
\newblock {ESBMC} 1.22.
\newblock In: Tools and Algorithms for the Construction and Analysis of
  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.

\end{thebibliography}

% that's all folks
\end{document}

### 2016-01-03 paper.tex
initial comments

@@
 
+\usepackage{color}
 % correct bad hyphenation here
@@
 \begin{abstract}
-Our reliance on the correct functioning of embedded systems is growing rapidly.
-Such systems are used in a wide range of applications such as micro-grids, airbag control systems, mobile phones, and high-end television sets.
-These systems are becoming more and more complex and require multi-core processors with scalable shared memory to meet the increasing computational power demands.
-The reliability of the embedded (distributed) software is thus a key issue in the system development.
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification and high-level processor model during simulation.
-However, verification of embedded systems raises additional challenges: meet time and energy constraints, handle concurrent software, platform-dependent software, and legacy designs (usually written in low-level languages).
-This paper describes challenges, problems, and recent advances to ensure the correctness and the timeliness of embedded systems.
-Reliability issues in the development of modern micro-grids are considered as a prominent model checking application.
+Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, airbag control, health care, surveillance, mobile devices, and high-end television sets.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
+As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
+Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor model during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
+The present paper discuss challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
+Reliability issues, in the development of modern micro-grids, are then considered, as a prominent model checking application.
 \end{abstract}
@@
 
-Embedded computer systems are used in a variety of sophisticated applications, which range from safety-critical systems such as nuclear reactors and automotive controllers, to entertainment software such as games and graphics animation~\cite{Kopetz11}.
-Embedded systems are ubiquitous in modern day information systems and are becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, success of such system depends on the relevance and quality of service it can provide.
+Generally, embedded computer systems perform dedicated functions with high degree of reliability.
+They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance.
+As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, their success depends both on the provided service relevance and quality.
+
+Indeed, many current embedded systems become interesting solutions only if they can reliably
+performs their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems.
+For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
+In addition, wrong diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.
 
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost of processors is enabling system designers to move more functionalities to software.
-This leads to difficulties in verifying design correctness since stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered.
+Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered.
 
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation.
-However, verification of embedded systems raises additional challenges, which include: meet time and energy constraints, handle concurrent software, platform-dependent software, and legacy designs (usually written in low-level languages).
-Thus, the main challenge for verifying embedded systems relies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages.
+\textcolor{red}{dei uma modificada aqui...
+deh uma olhada...
+deixei o antigo comentado abaixo:}\textcolor{green}{In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
+For instance, such aspects are of extreme relevance in micro-grids, in order to ensure reliability, which is a key issue for (smart) cities and consumers}.
+%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
 These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.
@@
 
-This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving the correctness and the timeliness of embedded systems, by taking into account stringent constraints imposed by the hardware.
+This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \textcolor{red}{achar outra palavra aqui?:}theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware\textcolor{green}{and incorporating knowledge about system purpose and associated features}.
 

### 2016-01-03 paper.tex
a new section was added about incorporating knowledge to verification

@@
 Eddie B. de Lima Filho \\
-Center for Science, Technology, and Innovation \\ for the Industrial Pole of Manaus, Brazil \\
+Center of Science, Technology, and Innovation \\ for the Industrial Pole of Manaus, Brazil \\
 Email: eddie@ctpim.org.br}
@@
 
-This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \textcolor{red}{achar outra palavra aqui?:}theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware\textcolor{green}{and incorporating knowledge about system purpose and associated features}.
+This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09} \textcolor{red}{achar outra palavra aqui?:}theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware \textcolor{green}{and incorporating knowledge about system purpose and associated features}.
 
@@
 
-Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and to find subtle bugs in real programs~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
-The idea behind the BMC techniques is to check for the violation of a given property at a given depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition (VC) $ \psi $ such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
+Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real programs~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+The idea behind the BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition (VC) $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
 less than or equal to \textit{k}.
@@
 In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the theories supported by SMT solvers ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency).
-Moreover, BMC techniques are only able to falsify properties up to the given depth \textit{k}; they are not able to prove the correctness of the system, unless an upper bound of \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
-In particular, BMC techniques limit the visited regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$.
-This limits the state space that needs to be explored during verification, leaving enough that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found; BMC tools are, however, susceptible to exhaustion of time or memory limits for programs with loops whose bounds are too large or cannot be determined statically.
+Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}, {\it i.e.}.
+Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
+In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found.
+Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
 
@@
 One technique typically used to prove properties, for any given depth, is mathematical induction.
-In particular, the algorithm called \textit{k}-induction was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and to respect time constraints specified during the design phase of a system~\cite{EenS03}.
-Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied due to the monolithic transition relation present
-in hardware designs~\cite{EenS03,GrosseLD09,Sheera00}.
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} contribute a new algorithm to prove correctness of (a large set of) C programs by mathematical induction in a completely automatic way ({\it i.e.}, the user does not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence in the literature that model checking embedded software using \textit{k}-induction (and invariants) can be exploited efficiently in embedded systems verification~\cite{Rocha15}.
+In particular, the algorithm called \textit{k}-induction was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}.
+Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15}.
+
+%--------------------------------------
+\subsection{Incorporating System Knowledge to Verification}
+%--------------------------------------
+
+\textcolor{green}{It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correct written, but also if it will respond properly in real environments or under external problems.
+For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.}
+
+\textcolor{green}{Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Recently, Husamma {\it et al.} proposed a verification tool for digital systems...}
 
@@
 
-\bibitem{handbook09}
-Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock Handbook of Satisfiability.
-\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
-
 \bibitem{BarrettSST09}
@@
   Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
+
+\bibitem{handbook09}
+Biere A., Heule M., van Maaren H., Walsh T., eds.:
+\newblock Handbook of Satisfiability.
+\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
 

### 2016-01-03 paper.tex
address Eddie's comments

@@
 % Do not put math or special symbols in the title.
-\title{Software Model Checking in Embedded Systems}
+\title{Software Model Checking in Embedded Systems \\ (Position Paper)}
 
@@
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered.
+Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered~\cite{Kroening15}.
 
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages.
-\textcolor{red}{dei uma modificada aqui...
-deh uma olhada...
-deixei o antigo comentado abaixo:}\textcolor{green}{In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
-For instance, such aspects are of extreme relevance in micro-grids, in order to ensure reliability, which is a key issue for (smart) cities and consumers}.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages.
+In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
+For instance, such aspects are of extreme relevance in micro-grids and syber-physical system, in order to ensure reliability, which is a key issue for (smart) cities, industries and consumers.
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 
-This position paper tackles three major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09} \textcolor{red}{achar outra palavra aqui?:}theories to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, and (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware \textcolor{green}{and incorporating knowledge about system purpose and associated features}.
+This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, in particular, extend background theories typically supported by SMT solvers to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to threads interleavings, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, in order to detect system-level failures.
 
@@
 
+\bibitem{Kroening15}
+Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
+\newblock Effective verification of low-level software with nested interrupts. 
+\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
+
+\bibitem{Behrend15}
+Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
+\newblock Scalable and Optimized Hybrid Verification of Embedded Software. 
+\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
+
+\bibitem{Lettnin09}
+Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Schönknecht V., Reitemeyer S.:
+\newblock Semiformal verification of temporal properties in automotive hardware dependent software. 
+\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
+
+\bibitem{handbook09}
+Biere A., Heule M., van Maaren H., Walsh T., eds.:
+\newblock Handbook of Satisfiability.
+\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
+
 \bibitem{BarrettSST09}
@@
   Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
-
-\bibitem{handbook09}
-Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock Handbook of Satisfiability.
-\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
 

### 2016-01-03 paper.tex
implement Eddie's comments

@@
 
-Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real programs~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real (embedded systems) software~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
 The idea behind the BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition (VC) $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
@@
 In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the theories supported by SMT solvers ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency).
-Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}, {\it i.e.}.
+Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}.
 Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
@@
 %--------------------------------------
-\subsection{Induction-based Verification of C/C++ Programs}
+\subsection{Induction-based Verification of C Programs}
 %--------------------------------------
@@
 
-\textcolor{green}{It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correct written, but also if it will respond properly in real environments or under external problems.
-For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.}
+It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correctly written, but also if it will respond properly in real environments or under external problems.
+For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.
 
-\textcolor{green}{Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Husamma {\it et al.} proposed a verification tool for digital systems...}
+Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Recently, Husamma {\it et al.} proposed a verification tool for digital systems to aid enginners to check for overflow, limit cycle, error, timing, stability, and minimum phase in digital systems, considering finite word length (FWL) effects~\cite{dsv_spin2015}.
 
@@
 \textbf{(RP2)} This approach was further extended to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}.
-These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark’s creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-\textbf{(RP3)} The SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems ({\it e.g.}, digital filters and controllers)~\cite{esbmc_controller,dsv_spin2015,esbmc_filter}.
-\textbf{(RP3)} Finally, novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in real-life embedded applications~\cite{Gadelha15,Rocha15}.
+These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in real-life embedded applications~\cite{Gadelha15,Rocha15}.
+\textbf{(RP4)} The SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems ({\it e.g.}, digital filters and controllers)~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}.
 Last but not least, ESBMC represents one of the most prominent BMC tools in the international software verification competition~\cite{MorseCNF13,MorseRCN014}.
@@
 Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
-\newblock Effective verification of low-level software with nested interrupts. 
+\newblock {\em Effective verification of low-level software with nested interrupts}. 
 \newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
@@
 Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
-\newblock Scalable and Optimized Hybrid Verification of Embedded Software. 
+\newblock {\em Scalable and Optimized Hybrid Verification of Embedded Software}. 
 \newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
@@
 Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Schönknecht V., Reitemeyer S.:
-\newblock Semiformal verification of temporal properties in automotive hardware dependent software. 
+\newblock {\em Semiformal verification of temporal properties in automotive hardware dependent software}. 
 \newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
@@
 Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock Handbook of Satisfiability.
+\newblock {\em Handbook of Satisfiability}.
 \newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
@@
 Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
-\newblock 26.
-\newblock In: Satisfiability Modulo Theories. Volume 185 of Frontiers in
-  Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
+\newblock {\em Satisfiability Modulo Theories}. 
+\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
 
@@
 Clarke E., Kroening D., Lerda F.:
-\newblock A tool for checking {ANSI-C} programs.
+\newblock {\em A tool for checking {ANSI-C} programs.}
 \newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg,  pp. 168--176, 2004.
@@
 Merz F., Falke S., Sinz C.:
-\newblock {LLBMC}: Bounded model checking of {C} and {C++} programs using a
-  compiler ir.
+\newblock {\em {LLBMC}: Bounded model checking of {C} and {C++} programs using a compiler {IR}.}
 \newblock In: International Conference on Verified Software: Theories, Tools, Experiments. Springer-Verlag, pp. 146--161, 2012.
@@
 Cordeiro L., Fischer B.:
-\newblock Verifying multi-threaded software using {SMT}-based context-bounded
-  model checking.
+\newblock {\em Verifying multi-threaded software using {SMT}-based context-bounded model checking.}
 \newblock In: International Conference on Software Engineering. ACM,  pp. 331--340, 2011.
@@
 Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
-\newblock Model checking {C} programs using {F-Soft}.
+\newblock {\em Model checking {C} programs using {F-Soft}.}
 \newblock In: IEEE International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
@@
 Cordeiro L., Fischer B., Marques{-}Silva J.:
-\newblock Smt-based bounded model checking for embedded {ANSI-C} software.
+\newblock {\em {SMT}-based bounded model checking for embedded {ANSI-C} software.}
 \newblock {IEEE} Trans. Software Eng. \textbf{38}(4), pp. 957--974, 2012.
@@
 Donaldson A., Kroening D., R{\"{u}}mmer P.:
-\newblock {SCRATCH}: A tool for automatic analysis of dma races.
+\newblock {\em {SCRATCH}: A tool for automatic analysis of {DMA} races.}
 \newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
@@
 Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-\newblock Software verification using k-induction.
+\newblock {\em Software Verification using \textit{k}-Induction.}
 \newblock In: International Conference on Static Analysis. Springer-Verlag, pp. 351--368, 2011.
@@
 E{\'{e}}n, N., S{\"{o}}rensson, N.:
-\newblock Temporal induction by incremental {SAT} solving.
+\newblock {\em Temporal induction by incremental {SAT} solving.}
 \newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.
@@
 Gro{\ss}e D., Le H., Drechsler R.:
-\newblock Induction-based formal verification of {SystemC TLM} designs.
+\newblock {\em Induction-based formal verification of {SystemC TLM} designs.}
 \newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
@@
 Sheeran M., Singh S., St{\aa}lmarck G.:
-\newblock Checking safety properties using induction and a {SAT}-solver.
+\newblock {\em Checking safety properties using induction and a {SAT}-solver.}
 \newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
@@
 Gadelha M., Ismail H., Cordeiro L.:
-\newblock Handling loops in bounded model checking of C programs via k-induction
+\newblock {\em Handling loops in bounded model checking of C programs via \textit{k}-induction.}
 \newblock In: International Journal on Software Tools for Technology Transfer (to appear), pp. 1--18, 2015.
@@
 Beyer D., Dangl M., Wendler P.:
-\newblock Boosting k-Induction with Continuously-Refined Invariants.
+\newblock {\em Boosting \textit{k}-Induction with Continuously-Refined Invariants.}
 \newblock International Conference on Computer-Aided Verification, pp. 622--640, 2015.
@@
 
+\bibitem{dsv_spin2015}
+Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
+\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
+\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
+
 \bibitem{ECBS13}
 Ramalho M., Lopes M., Sousa F., Marques H., Cordeiro L., Fischer B.:
-\newblock {SMT-Based Bounded Model Checking of C++ Programs}.
+\newblock {\em {SMT-Based Bounded Model Checking of C++ Programs} }.
 \newblock In: International Conference on Engineering of Computer-Based Systems, pp. 147--156, 2013.
@@
 Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
-\newblock Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking. 
+\newblock {\em Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.} 
 \newblock To appear in ACM Symposium on Applied Computing, Software Verification and Testing, 2016.
@@
 Bessa I., Abreu R., Cordeiro L., Filho J.:
-\newblock {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers)}. 
+\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers)}.}
 \newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.
-
-\bibitem{dsv_spin2015}
-Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-\newblock {DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
-\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
 
@@
 Abreu R., Cordeiro L., Filho E.:
-\newblock {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}. 
+\newblock {\em {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}.}
 \newblock In: XXXI Brazilian Symposium on Telecommunications, 2013.
@@
 Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock Handling unbounded loops with {ESBMC} 1.20.
+\newblock {\em Handling unbounded loops with {ESBMC} 1.20.}
 \newblock In: Tools and Algorithms for the Construction and Analysis of
@@
 Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
-\newblock {ESBMC} 1.22.
+\newblock {\em {ESBMC} 1.22.}
 \newblock In: Tools and Algorithms for the Construction and Analysis of

### 2016-01-03 paper.tex
implement Eddie's comments

@@
 Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15}.
+However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
 
@@
 \newblock {\em {LLBMC}: Bounded model checking of {C} and {C++} programs using a compiler {IR}.}
-\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. Springer-Verlag, pp. 146--161, 2012.
+\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
 
@@
 \newblock {\em Software Verification using \textit{k}-Induction.}
-\newblock In: International Conference on Static Analysis. Springer-Verlag, pp. 351--368, 2011.
+\newblock In: International Conference on Static Analysis. LNCS 6887, pp. 351--368, 2011.
 
@@
 Rocha H., Ismail H., Cordeiro L., Barreto R.:
-\newblock Model Checking Embedded C Software using k-Induction and Invariants. 
+\newblock {\em Model Checking Embedded C Software using k-Induction and Invariants.}
 \newblock V Brazilian Symposium on Computing Systems Engineering (to appear), pp. 1--6, 2015.
+
+\bibitem{Brain15}
+Brain M., Joshi S., Kroening D., Schrammel P.:
+\newblock {\em Safety Verification and Refutation by k-Invariants and k-Induction.}
+\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
 

### 2016-01-04 paper.tex
minor changes

@@
 % Do not put math or special symbols in the title.
-\title{Software Model Checking in Embedded Systems \\ (Position Paper)}
+\title{Bounded Model Checking for Embedded Systems \\ (Position Paper)}
 
@@
 In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
-For instance, such aspects are of extreme relevance in micro-grids and syber-physical system, in order to ensure reliability, which is a key issue for (smart) cities, industries and consumers.
+For instance, such aspects are of extreme relevance in micro-grids and cyber-physical system, in order to ensure reliability, which is a key issue for (smart) cities, industries and consumers.
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based context-BMC for full C/C++ programs (ESBMC)~\cite{Cordeiro12,ECBS13}.
-\textbf{(RP2)} This approach was further extended to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based context-BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++ programs~\cite{ECBS13}.
+\textbf{(RP2)} This approach was further developed and extended to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}.
 These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
@@
 % use section* for acknowledgment
-%\section*{Acknowledgment}
+\section*{Acknowledgment}
 
 %The authors would like to thank...
+This research was supported by CNPq $475647$/$2013$-$0$ grant.
 
@@
 \newblock {\em Boosting \textit{k}-Induction with Continuously-Refined Invariants.}
-\newblock International Conference on Computer-Aided Verification, pp. 622--640, 2015.
+\newblock International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
 

### 2016-01-04 paper.tex
minor changes

@@
 The present paper discuss challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
-Reliability issues, in the development of modern micro-grids, are then considered, as a prominent model checking application.
+Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent model checking application.
 \end{abstract}
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems, which demand short development cycles and high-level of reliability~\cite{leeCPS}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, their success depends both on the provided service relevance and quality.
@@
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered~\cite{Kroening15}.
+Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
 
@@
 \newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
+
+\bibitem{leeCPS}
+E.~Lee 
+\newblock {\em Cyber-physical systems: Design challenges.} 
+\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
 

### 2016-01-04 paper.tex
minor changes

@@
 In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
-For instance, such aspects are of extreme relevance in micro-grids and cyber-physical system, in order to ensure reliability, which is a key issue for (smart) cities, industries and consumers.
+For instance, such aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers.
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 
-In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the theories supported by SMT solvers ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency).
+In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the (background) theories supported by SMT solvers, {\it e.g.}, real-time, memory allocation, interrupts, and concurrency.
 Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}.
@@
 
-One technique typically used to prove properties, for any given depth, is mathematical induction.
-In particular, the algorithm called \textit{k}-induction was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}.
+A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
+The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
+In the base-case, we aim to check that $\phi$ holds in all states reachable from an initial state within $k$ steps (we assume that $k \geq 0$) and in the induction-step, we aim to check that whenever $\phi$ holds in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next state $s_{k+1}$ of the system.
+In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}.
 Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically address programing language issues and general correctness, without taking into account target applications or system behavior.
 Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correctly written, but also if it will respond properly in real environments or under external problems.

### 2016-01-04 paper.tex
update conclusions

@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based context-BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++ programs~\cite{ECBS13}.
-\textbf{(RP2)} This approach was further developed and extended to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98 programs~\cite{ECBS13}.
+\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques to prune the state-space exploration~\cite{morse15}.
 These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in real-life embedded applications~\cite{Gadelha15,Rocha15}.
-\textbf{(RP4)} The SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems ({\it e.g.}, digital filters and controllers)~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}.
-Last but not least, ESBMC represents one of the most prominent BMC tools in the international software verification competition~\cite{MorseCNF13,MorseRCN014}.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}.
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}.
+Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
 
@@
 This paper presented the main challenges to verify design correctness in embedded systems.
-In particular, stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification.
-Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants) can be applied to formally verify the correctness and the timeliness of embedded systems.
+In particular, it emphasizes that stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification.
+Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
+extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems.
 Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers.
-As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to modern micro-grids, considering small-scale versions of a centralized electricity system so that reliability and carbon emission reduction are amenable to verification.
+As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification.
 
@@
 
+\bibitem{morse15}
+J.~Morse.
+\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
+\newblock {University of Southampton, PhD Thesis}, 2015.
+
 \bibitem{esbmc_controller}

### 2016-01-04 paper.tex
add a new reference

@@
 \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}.
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}.
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
@@
 
+\bibitem{JMorse15}
+Morse J., Cordeiro L., Nicole D., Fischer B.:
+\newblock {\em {Model checking LTL properties over ANSI-C programs with bounded traces.}} 
+\newblock In: Software and System Modeling 14(1): pp. 65--81, 2015.
+
 \bibitem{MorseCNF13}

### 2016-01-04 paper.tex
fix references

@@
 E.~Lee 
-\newblock {\em Cyber-physical systems: Design challenges.} 
+\newblock {\em Cyber-physical Systems: Design Challenges.} 
 \newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
@@
 Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
-\newblock {\em Effective verification of low-level software with nested interrupts}. 
+\newblock {\em Effective Verification of Low-Level Software with Nested Interrupts}. 
 \newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
@@
 Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Schönknecht V., Reitemeyer S.:
-\newblock {\em Semiformal verification of temporal properties in automotive hardware dependent software}. 
+\newblock {\em Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}. 
 \newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
@@
 Clarke E., Kroening D., Lerda F.:
-\newblock {\em A tool for checking {ANSI-C} programs.}
+\newblock {\em A Tool for Checking {ANSI-C} Programs.}
 \newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg,  pp. 168--176, 2004.
@@
 Merz F., Falke S., Sinz C.:
-\newblock {\em {LLBMC}: Bounded model checking of {C} and {C++} programs using a compiler {IR}.}
+\newblock {\em {LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}.}
 \newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
@@
 Cordeiro L., Fischer B.:
-\newblock {\em Verifying multi-threaded software using {SMT}-based context-bounded model checking.}
+\newblock {\em Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking.}
 \newblock In: International Conference on Software Engineering. ACM,  pp. 331--340, 2011.
@@
 Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
-\newblock {\em Model checking {C} programs using {F-Soft}.}
+\newblock {\em Model Checking {C} Programs using {F-Soft}.}
 \newblock In: IEEE International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
@@
 Cordeiro L., Fischer B., Marques{-}Silva J.:
-\newblock {\em {SMT}-based bounded model checking for embedded {ANSI-C} software.}
+\newblock {\em {SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software.}
 \newblock {IEEE} Trans. Software Eng. \textbf{38}(4), pp. 957--974, 2012.
@@
 Donaldson A., Kroening D., R{\"{u}}mmer P.:
-\newblock {\em {SCRATCH}: A tool for automatic analysis of {DMA} races.}
+\newblock {\em {SCRATCH}: A Tool for Automatic Analysis of {DMA} Races.}
 \newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
@@
 E{\'{e}}n, N., S{\"{o}}rensson, N.:
-\newblock {\em Temporal induction by incremental {SAT} solving.}
+\newblock {\em Temporal Induction by Incremental {SAT} Solving.}
 \newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.
@@
 Gro{\ss}e D., Le H., Drechsler R.:
-\newblock {\em Induction-based formal verification of {SystemC TLM} designs.}
+\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
 \newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
@@
 Sheeran M., Singh S., St{\aa}lmarck G.:
-\newblock {\em Checking safety properties using induction and a {SAT}-solver.}
+\newblock {\em Checking Safety Properties using Induction and a {SAT}-solver.}
 \newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
@@
 Gadelha M., Ismail H., Cordeiro L.:
-\newblock {\em Handling loops in bounded model checking of C programs via \textit{k}-induction.}
+\newblock {\em Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction.}
 \newblock In: International Journal on Software Tools for Technology Transfer (to appear), pp. 1--18, 2015.
@@
 Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {Model checking LTL properties over ANSI-C programs with bounded traces.}} 
+\newblock {\em {Model Checking LTL Properties over ANSI-C Programs with Bounded Traces.}} 
 \newblock In: Software and System Modeling 14(1): pp. 65--81, 2015.

### 2016-01-04 paper.tex
comments regarding the whole paper

@@
 Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems, which demand short development cycles and high-level of reliability~\cite{leeCPS}.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system; in particular, their success depends both on the provided service relevance and quality.
+As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success depends both on the provided service relevance and quality.
 
 Indeed, many current embedded systems become interesting solutions only if they can reliably
-performs their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems.
+perform their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
@@
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost common to computer processors is enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
+Additionally, the increasing computational power and decreasing size and cost, which are common to computer processors, are enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
 
@@
 State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages.
-In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
-For instance, such aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers.
+In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+For instance, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third is essential in systems that implement scientific models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
+However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
 
 %--------------------------------------
-\subsection{Incorporating System Knowledge to Verification}
+\subsection{Incorporating System Knowledge to Verification Procedures}
 %--------------------------------------
@@
 Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Husamma {\it et al.} proposed a verification tool for digital systems to aid enginners to check for overflow, limit cycle, error, timing, stability, and minimum phase in digital systems, considering finite word length (FWL) effects~\cite{dsv_spin2015}.
+Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
+Ultimately, if some requirements are not met with a given configuration (realization structure and hardware restrictions), the error reports provided DsVerifier may suggest another setup.
+Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAVs.
 
@@
 \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques to prune the state-space exploration~\cite{morse15}.
-These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in, standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
 \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}.
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} \textcolor{red}{depois mencionar explicitamente UAVS} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
@@
 extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems.
-Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers.
+Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers \textcolor{red}{falar sobre a incorporacao e conhecimento sobre o sistema}.
 As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification.

### 2016-01-04 paper.tex
new mods in the abstract section

@@
 \begin{abstract}
-Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, airbag control, health care, surveillance, mobile devices, and high-end television sets.
+Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control, etc.), health care, surveillance, mobile devices, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor model during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
+Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor model, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
 The present paper discuss challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.

### 2016-01-04 paper.tex
extension of some topics throughout th whole paper and general corrections

@@
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor model, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
-The present paper discuss challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
+The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
 Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent model checking application.
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems, which demand short development cycles and high-level of reliability~\cite{leeCPS}.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success depends both on the provided service relevance and quality.
 
-Indeed, many current embedded systems become interesting solutions only if they can reliably
-perform their target tasks, such as unmanned aerial vehicles (UAVs) and medical monitoring systems.
+Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems, become interesting solutions only if they can reliably
+perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
-In addition, wrong diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.
+In addition, wrong disease diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.
+
+Besides, when physical interaction with the real world is needed, which happens in CPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
+Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation.
 
@@
 In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
-For instance, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third is essential in systems that implement scientific models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
+Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encodings into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, in particular, extend background theories typically supported by SMT solvers to reason accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to threads interleavings, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, in order to detect system-level failures.
+This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level failures.
 
@@
 
-Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real (embedded systems) software~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
-The idea behind the BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition (VC) $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
+Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real (embedded system) software~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+The idea behind the BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
 less than or equal to \textit{k}.
@@
 
-A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
+A feasible alternative to check for properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
 The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
-In the base-case, we aim to check that $\phi$ holds in all states reachable from an initial state within $k$ steps (we assume that $k \geq 0$) and in the induction-step, we aim to check that whenever $\phi$ holds in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next state $s_{k+1}$ of the system.
-In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}.
-Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+In the former, the goal to check that $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
+The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
+
+In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+
+Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a \textcolor{red}{conforme conversamos, eh um novo mesmo?} new algorithm for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
 However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically address programing language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correctly written, but also if it will respond properly in real environments or under external problems.
-For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.
+It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
+For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
 
@@
 Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
-Ultimately, if some requirements are not met with a given configuration (realization structure and hardware restrictions), the error reports provided DsVerifier may suggest another setup.
-Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAVs.
+Indeed, DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+%Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.
 
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98 programs~\cite{ECBS13}.
-\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques to prune the state-space exploration~\cite{morse15}.
-These verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer in, standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15}.
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} \textcolor{red}{depois mencionar explicitamente UAVS} and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
-Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems) as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98~\cite{ECBS13}.
+\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+In particular, a specific UVA application was tackled, in order to verify its attitude controllers \textcolor{red}{que artigo do iury mencionar?}.
+Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
 
@@
 
-This paper presented the main challenges to verify design correctness in embedded systems.
-In particular, it emphasizes that stringent constraints imposed by the hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification.
+This paper presented the main challenges related to the verification of design correctness in embedded systems.
+In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification.
 Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
 extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems.
-Given that the complexity of software has increased significantly in embedded products, there are still some (recent) advances to stress and cover exhaustively the state space of the system, in order to verify low-level properties that have to meet the application's deadline, access the memory region, handle concurrency, and control the hardware registers \textcolor{red}{falar sobre a incorporacao e conhecimento sobre o sistema}.
-As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded systems implementation; special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification.
+
+Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
+Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification to another step, where not only code correctness will be addressed, but also full system reliability.
+
+As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded system implementations.
+Special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification, probably through behavior models.
 
@@
 \newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
+
+\bibitem{munir}
+Munir  S., Stankovic J. A., Liang C.-J. M., Lin S.:
+\newblock {\em Cyber Physical System Challenges for Human-in-the-Loop Control}. 
+\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
 

### 2016-01-04 paper.tex
minor changes

@@
 \begin{abstract}
-Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control, etc.), health care, surveillance, mobile devices, and consumer electronics.
+Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.

### 2016-01-04 paper.tex
minor changes

@@
 
-Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems, become interesting solutions only if they can reliably
-perform their target tasks.
+Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
@@
 
-A feasible alternative to check for properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
+A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
 The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
-In the former, the goal to check that $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
+In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
 The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
 
-In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
 Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
 
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a \textcolor{red}{conforme conversamos, eh um novo mesmo?} new algorithm for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
+Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15} because of the distinctive characteristics mentioned earlier, which influence the embedded system development and verification process.
 
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
 Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
@@
 Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+
+Scratch is also another example of software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) for the Cell BE processor~\cite{Donaldson10}.
+The tool also uses BMC to detect DMA races, and bounded model checking with k-induction to prove absence of races.
+
 %Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98~\cite{ECBS13}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}.
+
 \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+
 \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
+
 \textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UVA application was tackled, in order to verify its attitude controllers \textcolor{red}{que artigo do iury mencionar?}.
+In particular, a specific UAV application was tackled, in order to verify its attitude controllers~\cite{Bessa16}.
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
@@
 \newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
+
+\bibitem{Cordeiro09}
+Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
+\newblock {\em Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints.} 
+\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
 
@@
 
+\bibitem{Bessa16}
+Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
+\newblock {\em { Verification of Delta Form Realization in Fixed-Point Digital Controllers Using Bounded Model Checking.}} 
+\newblock In: Design Automation for Embedded Systems (to appear), pp. 49--54, 2016.
+
 \bibitem{MorseCNF13}

### 2016-01-04 paper.tex
minor changes

@@
 Scratch is also another example of software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) for the Cell BE processor~\cite{Donaldson10}.
-The tool also uses BMC to detect DMA races, and bounded model checking with k-induction to prove absence of races.
+The tool also uses BMC to detect DMA races, and bounded model checking with \textit{k}-induction to prove absence of races.
 
@@
 Rocha H., Ismail H., Cordeiro L., Barreto R.:
-\newblock {\em Model Checking Embedded C Software using k-Induction and Invariants.}
+\newblock {\em Model Checking Embedded C Software using \textit{k}-Induction and Invariants.}
 \newblock V Brazilian Symposium on Computing Systems Engineering (to appear), pp. 1--6, 2015.
@@
 Brain M., Joshi S., Kroening D., Schrammel P.:
-\newblock {\em Safety Verification and Refutation by k-Invariants and k-Induction.}
+\newblock {\em Safety Verification and Refutation by k-Invariants and \textit{k}-Induction.}
 \newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.

### 2016-01-04 paper.tex
minor changes

@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level failures.
+This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs; (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving; (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware; and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level failures.
 
@@
 %--------------------------------------
-\subsection{Incorporating System Knowledge to Verification Procedures}
+\subsection{Incorporating System Knowledge to Automated Verification Procedures}
 %--------------------------------------

### 2016-01-04 paper.tex
final corrections part 1

--- 2016-01-04 paper.tex
@@
 \begin{abstract}
-Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics.
+The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control, etc.), health care, surveillance, mobile devices, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS}.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success depends both on the provided service relevance and quality.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
+As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both provided service relevance and quality.
 
-Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
+Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems, become interesting solutions only if they can reliably
+perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
@@
 Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation.
+Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
 
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost, which are common to computer processors, are enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
+Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
 
@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level failures.
+This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
 
@@
 
-A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
+A feasible alternative to check for properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
 The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
-In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
+In the former, the goal to check that $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
 The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
 
-In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
 Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
 
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15} because of the distinctive characteristics mentioned earlier, which influence the embedded system development and verification process.
+Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a \textcolor{red}{conforme conversamos, eh um novo mesmo?} new algorithm for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
 
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
 Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
@@
 Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
-
-Scratch is also another example of software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) for the Cell BE processor~\cite{Donaldson10}.
-The tool also uses BMC to detect DMA races, and bounded model checking with k-induction to prove absence of races.
-
+Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+Indeed, DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
 %Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}.
-
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98~\cite{ECBS13}.
 \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-
 \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
-
 \textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UAV application was tackled, in order to verify its attitude controllers~\cite{Bessa16}.
+In particular, a specific UVA application was tackled, in order to verify its attitude controllers \textcolor{red}{que artigo do iury mencionar?}.
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
@@
 
-\bibitem{Cordeiro09}
-Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
-\newblock {\em Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints.} 
-\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
+\bibitem{leeCPS2}
+Lee E.:
+\newblock {\em The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}. 
+\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
 
@@
 
-\bibitem{Bessa16}
-Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
-\newblock {\em { Verification of Delta Form Realization in Fixed-Point Digital Controllers Using Bounded Model Checking.}} 
-\newblock In: Design Automation for Embedded Systems (to appear), pp. 49--54, 2016.
-
 \bibitem{MorseCNF13}

### 2016-01-04 paper.tex

--- 2016-01-04 paper.tex
@@
 \begin{abstract}
-Our dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics.
+The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS}.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success depends both on the provided service relevance and quality.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
+As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both provided service relevance and quality.
 
@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs; (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving; (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware; and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level failures.
+This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
 
@@
 \newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
+
+\bibitem{leeCPS2}
+Lee E.:
+\newblock {\em The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}. 
+\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
 

### 2016-01-05 paper.tex
final corrections part 3

@@
 Eddie B. de Lima Filho \\
-Center of Science, Technology, and Innovation \\ for the Industrial Pole of Manaus, Brazil \\
+Science, Technology, and Innovation Center\\ for the Industrial Pole of Manaus, Brazil \\
 Email: eddie@ctpim.org.br}
@@
 Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both provided service relevance and quality.
+As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
 
@@
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
+Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
 
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed with low-level languages.
 In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
@@
 Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real (embedded system) software~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
-The idea behind the BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a Verification Condition $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
+The idea behind BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a verification condition $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
 less than or equal to \textit{k}.
@@
 In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
-Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
 
 Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15} because of the distinctive characteristics mentioned earlier, which influence the embedded system development and verification process.
+However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
+That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
 
@@
 
-Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Indeed, research in software verification is now taking such considerations into account, and some schemes already use knowledge about the system to be verified and the underlying hardware.
 Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
+Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
 
-Scratch is also another example of software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) for the Cell BE processor~\cite{Donaldson10}.
-The tool also uses BMC to detect DMA races, and bounded model checking with \textit{k}-induction to prove absence of races.
+Scratch is another example of software-model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
+Such a tool also uses BMC, in order to detect DMA races, and bounded model checking with \textit{k}-induction, which aims to prove the absence of races.
 
@@
 \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
 
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UAV application was tackled, in order to verify its attitude controllers~\cite{Bessa16}.
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
+
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
@@
 Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
-extends BMC-based approaches from falsification to verification, can be applied to formally verify the correctness and the timeliness of embedded systems.
+extends BMC-based approaches from falsification to verification, can be applied to formally verify correctness and timeliness of embedded systems.
 
@@
 
-As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving a correct-by-construction embedded system implementations.
+As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving correct-by-construction embedded system implementations.
 Special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification, probably through behavior models.

### 2016-01-05 paper.tex
minor changes

@@
 The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
-Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent model checking application.
+Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent (bounded) model checking application.
 \end{abstract}
@@
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
-In addition, wrong disease diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.
+In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
 
@@
 Scratch is another example of software-model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
-Such a tool also uses BMC, in order to detect DMA races, and bounded model checking with \textit{k}-induction, which aims to prove the absence of races.
+Such a tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
 
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt programs~\cite{Sousa15}.
 
@@
 \bibitem{Kopetz11}
-H.~Kopetz.
+H.~Kopetz:
 \newblock {\em Real-Time Systems - Design Principles for Distributed Embedded Applications}.
@@
 \bibitem{leeCPS}
-E.~Lee 
+Lee E.:
 \newblock {\em Cyber-physical Systems: Design Challenges.} 
@@
 
+\bibitem{Sousa15}
+Sousa F., Cordeiro L., Lima Filho E.: 
+\newblock {\em { Bounded Model Checking of C++ Programs Based on the Qt Framework.}} 
+\newblock In: Global Conference on Consumer Electronics, pp. 179--180, 2015.
+
 \bibitem{Pereira15}
@@
 Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
-\newblock {\em { Verification of Delta Form Realization in Fixed-Point Digital Controllers Using Bounded Model Checking.}} 
+\newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
 \newblock In: Design Automation for Embedded Systems (to appear), pp. 49--54, 2016.

### 2016-03-18 paper.tex
include new reference

@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
@@
 \newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
+
+\bibitem{xu15}
+Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
+\newblock {\em Hierarchical energy management system for multi-source multi-product microgrids}
+\newblock In: Renewable Energy, v. 78, pp. 621-–630, 2015.
 

### 2016-03-18 paper.tex
include new reference

@@
 
-Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
+Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
@@
 \newblock In: Sensors 15(3): pp. 4837--4869, 2015.
+
+\bibitem{groza2015formal}
+Groza A., Letia I., Goron A., Zaporojan S.: 
+\newblock {\em A formal approach for identifying assurance deficits in unmanned aerial vehicle software}
+\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
 

### 2016-03-18 paper.tex
remove comma
fix reference

@@
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
-In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
+In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health with serious consequences.
 
@@
 \bibitem{Lettnin09}
-Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Schönknecht V., Reitemeyer S.:
+Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
 \newblock {\em Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}. 

### 2016-03-18 paper.tex
review verification challenges

@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed with low-level languages.
-In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}.
+In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness; additionally, embedded system verification raises additional challenges, such as
+%
+\begin{enumerate}
+	\item time and energy constraints;
+	\item handling of concurrent software;
+	\item platform restrictions; and
+	\item legacy designs.%, which are usually developed with low-level languages.
+\end{enumerate}
+
 Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
+The fourth aspect is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
+
 %In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
@@
 %--------------------------------------
-\section{Bounded Model Checking of Embedded Systems}
+\section{Bounded Model Checking for Embedded Systems}
 %--------------------------------------

### 2016-03-18 paper.tex
review research problem

@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems: (RP1) providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs, (RP2) exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving, (RP3) proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware, and (RP4) incorporating knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+This position paper tackles four major problems in computer-aided verification for embedded systems, which are (partially) open in present research.
+
+\textbf{(RP1)} providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+
+\textbf{(RP2)} exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
+
+\textbf{(RP3)} proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
+
+\textbf{(RP4)} incorporating knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+
+Section~\ref{achievements} outlines contributions toward these research problems.
+
+%--------------------------------------
+\section{Achievements}
+\label{achievements}
+%--------------------------------------
+
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt programs~\cite{Sousa15}.
+
+\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
+
+\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
+
+Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
 
@@
 %Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.
-
-%--------------------------------------
-\section{Achievements}
-\label{achievements}
-%--------------------------------------
-
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt programs~\cite{Sousa15}.
-
-\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
-
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
-
-Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
 

### 2016-03-18 paper.tex
review background on BMC

@@
 % Do not put math or special symbols in the title.
-\title{Bounded Model Checking for Embedded Systems \\ (Position Paper)}
+\title{SMT-Based Context-Bounded Model Checking for Embedded Systems (Position Paper)}
 
@@
 %--------------------------------------
+\section{Bounded Model Checking (BMC)}
+%--------------------------------------
+
+Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed by Biere {\it et al.} to verify hardware designs~\cite{Biere99,handbook09}.
+The authors were able to successfully verify digital circuits with $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a SAT solver.
+BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06}.
+
+The basic idea behind BMC techniques is to check the violation of a given property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ representing code correctness and/or system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in software, BMC thus unfolds the system \textit{k} times, in order to convert it into a verification condition $\psi$, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+
+SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of software systems, including single- and multi-threaded programs.
+In general, those BMC techniques were able to find subtle bugs in real embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+However, the main criticism with respect to BMC techniques relies on completeness since they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
+Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
+
+%In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the (background) theories supported by SMT solvers, {\it e.g.}, real-time, memory allocation, interrupts, and concurrency.
+Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}.
+Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
+In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found.
+Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
+
+%--------------------------------------
+\subsection{Induction-based Verification of C Programs}
+%--------------------------------------
+
+A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
+The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
+In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
+The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
+
+In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+
+Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
+That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
+
+%--------------------------------------
+\subsection{Incorporating System Knowledge to Automated Verification Procedures}
+%--------------------------------------
+
+It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
+For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
+
+Indeed, research in software verification is now taking such considerations into account, and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
+Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+
+Scratch is another example of software-model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
+Such a tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
+
+%--------------------------------------
 \section{Verification Challenges}
@@
 Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
-
-%--------------------------------------
-\section{Bounded Model Checking for Embedded Systems}
-%--------------------------------------
-
-Bounded Model Checking (BMC) techniques based on Boolean Satisfiability (SAT)~\cite{handbook09} or SMT~\cite{BarrettSST09} have been successfully applied to verify single- and multi-threaded programs and find subtle bugs in real (embedded system) software~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
-The idea behind BMC techniques is to check the violation of a given property at a predetermined depth, {\it i.e.}, given a transition system \textit{M}, a property $ \phi $, and a limit of iterations \textit{k}, BMC unfolds the system \textit{k} times and converts it into a verification condition $ \psi $, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth
-less than or equal to \textit{k}.
-
-In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the (background) theories supported by SMT solvers, {\it e.g.}, real-time, memory allocation, interrupts, and concurrency.
-Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}.
-Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
-In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found.
-Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
-
-%--------------------------------------
-\subsection{Induction-based Verification of C Programs}
-%--------------------------------------
-
-A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
-The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
-In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
-The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
-
-In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
-Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
-
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
-That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
-
-%--------------------------------------
-\subsection{Incorporating System Knowledge to Automated Verification Procedures}
-%--------------------------------------
-
-It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
-For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
-
-Indeed, research in software verification is now taking such considerations into account, and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
-Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
-
-Scratch is another example of software-model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
-Such a tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
 
@@
 \newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
+
+\bibitem{Biere99}
+Biere A., Cimatti A., Clarke E., Zhu Y.:
+\newblock {\em Symbolic Model Checking without BDDs}. 
+\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
+
+\bibitem{Armando06}
+Armando A., Mantovani J., Platania L.:
+\newblock {\em Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}. 
+\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
 

### 2016-03-18 paper.tex
review BMC background

@@
 
-The basic idea behind BMC techniques is to check the violation of a given property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ representing code correctness and/or system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in software, BMC thus unfolds the system \textit{k} times, in order to convert it into a verification condition $\psi$, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+The basic idea behind BMC techniques is to check the violation of a given property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ representing program correctness and/or system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in software, BMC thus unfolds the system \textit{k} times, in order to convert it into a verification condition $\psi$, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
-SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of software systems, including single- and multi-threaded programs.
-In general, those BMC techniques were able to find subtle bugs in real embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits, single- and multi-threaded programs.
+In general, those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
 However, the main criticism with respect to BMC techniques relies on completeness since they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
-Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
+Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuits implementation or programs with loops whose bounds are too large or cannot be statically determined.
 

### 2016-03-18 paper.tex
review induction-based verification

@@
 
-A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
-The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}.
-In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed).
-The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.
+One possible way to achieve completeness in BMC techniques is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
+However, the main limitation of such approach relies on producing or strengthening inductive assertions.
+In particular, loop invariants annotated in the program must be inductive (and not just invariant) for the corresponding VCs to be valid~\cite{Bradley07}.
+As a consequence, \textit{k}-induction procedures could even produce such assertions, which are indeed invariant, but if those assertions not inductive, then program correctness cannot be proved.
 
-In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
-Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+There are invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory alignment in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+The current literature even report successful applications of induction-based verification for hardware and software systems, in some cases using invariants.
+In particular, \textit{k}-induction based verification was already applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+\textit{k}-induction is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
 
@@
 Cordeiro L., Fischer B., Marques{-}Silva J.:
-\newblock {\em {SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software.}
+\newblock {\em {SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}.
 \newblock {IEEE} Trans. Software Eng. \textbf{38}(4), pp. 957--974, 2012.
+
+\bibitem{Bradley07}
+Bradley A., Manna Z.:
+\newblock {\em The calculus of computation - decision procedures with applications to verification}. 
+\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
 
@@
 \newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
+
+\bibitem{pips:2013}
+ParisTech M (2013) PIPS: Automatic Parallelizer and Code Transformation Framework. Accessed 21 February 2016
+
+\bibitem{Henry:2012}
+ Henry J, Monniaux D, Moy M (2012) PAGAI: A Path Sensitive Static Analyser. Electronic Notes in Theoretical Computer Science. Elsevier, pp. 15--25
 

### 2016-03-18 paper.tex
review 2.1 Induction-based Verification of C Programs

@@
 One possible way to achieve completeness in BMC techniques is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
-However, the main limitation of such approach relies on producing or strengthening inductive assertions.
-In particular, loop invariants annotated in the program must be inductive (and not just invariant) for the corresponding VCs to be valid~\cite{Bradley07}.
-As a consequence, \textit{k}-induction procedures could even produce such assertions, which are indeed invariant, but if those assertions not inductive, then program correctness cannot be proved.
+The main challenge of such approach relies on computing and strengthening inductive invariants from the program.
+In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding VCs to be valid~\cite{Bradley07}.
+As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those assertions must be inductive to prove program correctness.
 
-There are invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory alignment in programming languages with pointers~\cite{pips:2013,Henry:2012}.
-The current literature even report successful applications of induction-based verification for hardware and software systems, in some cases using invariants.
-In particular, \textit{k}-induction based verification was already applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
-\textit{k}-induction is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+There are several invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+The current literature even reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening~\cite{Beyer15,Rocha15,Brain15}.
 
-Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
-However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.
+More specifically, in previous work, \textit{k}-induction based verification was already applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
+Recently, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+
+The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15,Brain15}.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
@@
 
+\bibitem{Brain15}
+Brain M., Joshi S., Kroening D., Schrammel P.:
+\newblock {\em Safety Verification and Refutation by k-Invariants and k-Induction.} 
+\newblock In: Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+
 \bibitem{Beyer15}
@@
 \newblock {\em Boosting \textit{k}-Induction with Continuously-Refined Invariants.}
-\newblock International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
+\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
 
@@
 \newblock {\em Model Checking Embedded C Software using \textit{k}-Induction and Invariants.}
-\newblock V Brazilian Symposium on Computing Systems Engineering (to appear), pp. 1--6, 2015.
+\newblock V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
 

### 2016-03-18 paper.tex
review achievements

@@
 The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
-However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15,Brain15}.
+However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Beyer15,Brain15}.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt programs~\cite{Sousa15}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt-based consumer electronics applications~\cite{Sousa15}.
+Other efficient SMT-based BMC approaches have been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing approaches are still limited for specific class of programs, especially for those that contain floating-point arithmetic and manipulate the heap memory~\cite{BeyerSVCOMP15}.
 
-\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al.} was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
+Recent advances to verify multi-threaded C programs have been made by Inverso {\it et al.}~\cite{Inverso14}, but the class of programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates its effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15,Brain15}.
+However, the main challenge still remains to compute and strengthen loop invariants to prove program correctness.
+Invariant generation algorithms must still evolve to efficiently discover inductive assertions of programs or to continuously refine invariants, as described by Beyer {\it et al.}~\cite{Beyer15}.
 
@@
 \bibitem{pips:2013}
-ParisTech M (2013) PIPS: Automatic Parallelizer and Code Transformation Framework. Accessed 21 February 2016
+ParisTech M.:
+\newblock {\em PIPS: Automatic Parallelizer and Code Transformation Framework.} 
+\newblock Accessed 21 February 2016
 
 \bibitem{Henry:2012}
- Henry J, Monniaux D, Moy M (2012) PAGAI: A Path Sensitive Static Analyser. Electronic Notes in Theoretical Computer Science. Elsevier, pp. 15--25
+Henry J., Monniaux D., Moy M.: 
+\newblock {\em PAGAI: A Path Sensitive Static Analyser.} 
+\newblock In: Electronic Notes in Theoretical Computer Science, pp. 15--25, 2012.
+
+\bibitem{BeyerSVCOMP15}
+D.~Beyer:
+\newblock {\em Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).} 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
 
@@
 \newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+
+\bibitem{Inverso14}
+Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
+\newblock {\em Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}
+\newblock In: Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
 

### 2016-03-19 paper.tex
fix reference
review background section

@@
 % Do not put math or special symbols in the title.
-\title{SMT-Based Context-Bounded Model Checking for Embedded Systems (Position Paper)}
+\title{Challenges and Recent Advances to SMT-Based Context-Bounded Model Checking for Embedded Systems}
 
@@
 Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed by Biere {\it et al.} to verify hardware designs~\cite{Biere99,handbook09}.
-The authors were able to successfully verify digital circuits with $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a SAT solver.
+The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a SAT solver.
 BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06}.
 
-The basic idea behind BMC techniques is to check the violation of a given property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ representing program correctness and/or system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in software, BMC thus unfolds the system \textit{k} times, in order to convert it into a verification condition $\psi$, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+BMC techniques aims to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in the program, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
@@
 The main challenge of such approach relies on computing and strengthening inductive invariants from the program.
-In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding VCs to be valid~\cite{Bradley07}.
-As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those assertions must be inductive to prove program correctness.
+In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding verification conditions to be valid~\cite{Bradley07}.
+As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive so that verifiers can automatically prove program correctness.
 
 There are several invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
-The current literature even reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening~\cite{Beyer15,Rocha15,Brain15}.
+The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening, mostly based on interval analysis.
+In particular, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15}.
 
-More specifically, in previous work, \textit{k}-induction based verification was already applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}.
-Recently, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}.
+Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
+The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
 
-The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
-However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Beyer15,Brain15}.
+However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
+Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and facts about memory and variable aliasing~\cite{Bradley07}.
 
@@
 \newblock {\em Hierarchical energy management system for multi-source multi-product microgrids}
-\newblock In: Renewable Energy, v. 78, pp. 621-–630, 2015.
+\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
 

### 2016-03-19 paper.tex
review achievements

@@
 % Do not put math or special symbols in the title.
-\title{Challenges and Recent Advances to SMT-Based Context-Bounded Model Checking for Embedded Systems}
+\title{SMT-Based Context-Bounded Model Checking for Embedded Systems: State of the Art and Future Trends}
 
@@
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
-Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and facts about memory and variable aliasing~\cite{Bradley07}.
+Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
 
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13} and Qt-based consumer electronics applications~\cite{Sousa15}.
-Other efficient SMT-based BMC approaches have been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing approaches are still limited for specific class of programs, especially for those that contain floating-point arithmetic and manipulate the heap memory~\cite{BeyerSVCOMP15}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing approaches are still limited for specific class of programs, especially for those that contain floating-point arithmetic and manipulate the heap~\cite{BeyerSVCOMP15}.
 
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al.} was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).
-Recent advances to verify multi-threaded C programs have been made by Inverso {\it et al.}~\cite{Inverso14}, but the class of programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prunes the state space exploration~\cite{Inverso14,civl15}.
+However, the class of concurrent (embedded systems) programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrates its effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15,Brain15}.
-However, the main challenge still remains to compute and strengthen loop invariants to prove program correctness.
-Invariant generation algorithms must still evolve to efficiently discover inductive assertions of programs or to continuously refine invariants, as described by Beyer {\it et al.}~\cite{Beyer15}.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15,Brain15}.
+However, the main challenge still remains to compute and strengthen loop invariants to prove program correctness and timeliness.
+Invariant generation algorithms have substantially evolved over the years to efficiently discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them~\cite{Beyer15}.
+However, there is still a lack of studies for exploiting the combination of different invariant generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities), in order to ensure system robustness w.r.t.
+implementation aspects.
 
-\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+\textbf{(RP4)} State of the art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
 In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
+In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; in particular, micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
 
-Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.
+Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent ones for verifying C programs, as observed in the Intl.
+Competition on Software Verification~\cite{BeyerSVCOMP15}.
 
@@
 \newblock {\em Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking.}
-\newblock In: International Conference on Software Engineering. ACM,  pp. 331--340, 2011.
+\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
 
@@
 \newblock {\em Model Checking {C} Programs using {F-Soft}.}
-\newblock In: IEEE International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
+\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
 
@@
 \newblock {\em Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction.}
-\newblock In: International Journal on Software Tools for Technology Transfer (to appear), pp. 1--18, 2015.
+\newblock In: International Journal on Software Tools for Technology Transfer (to appear), 2015.
 \newblock http://dx.doi.org/10.1007/s10009-015-0407-9
@@
 \newblock {\em Safety Verification and Refutation by k-Invariants and k-Induction.} 
-\newblock In: Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
 
@@
 \newblock {\em Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}
-\newblock In: Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
+\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
 
@@
 \newblock {\em Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.} 
-\newblock To appear in ACM Symposium on Applied Computing, Software Verification and Testing, 2016.
+\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016.
 
@@
 \newblock {University of Southampton, PhD Thesis}, 2015.
+
+\bibitem{civl15}
+Zheng M, Rogers M, Luo Z, Dwyer M, Siegel S.
+\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
+\newblock In: International Conference on Automated Software Engineering, pp. 830--835, 2015.
 
@@
 \newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
-\newblock In: Design Automation for Embedded Systems (to appear), pp. 49--54, 2016.
+\newblock In: Design Automation for Embedded Systems (to appear), 2016.
 

### 2016-03-20 paper.tex
review ACHIEVEMENTS

@@
 \textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
-Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing approaches are still limited for specific class of programs, especially for those that contain floating-point arithmetic and manipulate the heap~\cite{BeyerSVCOMP15}.
+Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing BMC approaches are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{BeyerSVCOMP15}.
+One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.
 
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al.} was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prunes the state space exploration~\cite{Inverso14,civl15}.
-However, the class of concurrent (embedded systems) programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al}.
+was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state space exploration~\cite{Inverso14,civl15}.
+However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour.
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15,Brain15}.
-However, the main challenge still remains to compute and strengthen loop invariants to prove program correctness and timeliness.
-Invariant generation algorithms have substantially evolved over the years to efficiently discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them~\cite{Beyer15}.
-However, there is still a lack of studies for exploiting the combination of different invariant generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities), in order to ensure system robustness w.r.t.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15,Brain15}.
+However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way in order to be competitive with other model checking approaches.
+In particular, invariant generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
+However, there is still a lack of studies for exploiting the combination of different invariant generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
 implementation aspects.
@@
 Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent ones for verifying C programs, as observed in the Intl.
-Competition on Software Verification~\cite{BeyerSVCOMP15}.
+Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}.
 
@@
 
+\bibitem{KahlonWG09}
+Kahlon V, Wang C, Gupta A.
+\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order
+  Reduction Technique}}.
+\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
+
+\bibitem{Beyer14}
+D.~Beyer 
+\newblock {Status report on software verification - (competition summary {SV-COMP})}. 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
+
 \end{thebibliography}

### 2016-03-20 paper.tex
fix references

@@
 The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening, mostly based on interval analysis.
-In particular, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15}.
+In particular, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
 
@@
 
+%1
 \bibitem{Kopetz11}
@@
 
+%2
 \bibitem{xu15}
@@
 
+%3
 \bibitem{leeCPS}
@@
 
+%4
 \bibitem{leeCPS2}
@@
 
+%5
 \bibitem{groza2015formal}
@@
 
+%6
 \bibitem{Cordeiro09}
@@
 
+%7
 \bibitem{munir}
@@
 
+%8
 \bibitem{Kroening15}
@@
 
+%9
 \bibitem{Biere99}
@@
 
+%10
+\bibitem{handbook09}
+Biere A., Heule M., van Maaren H., Walsh T., eds.:
+\newblock {\em Handbook of Satisfiability}.
+\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
+
+%11
+\bibitem{BarrettSST09}
+Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
+\newblock {\em Satisfiability Modulo Theories}. 
+\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
+
+%12
 \bibitem{Armando06}
@@
 
-\bibitem{Behrend15}
-Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
-\newblock {\em Scalable and Optimized Hybrid Verification of Embedded Software}. 
-\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
-
-\bibitem{Lettnin09}
-Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
-\newblock {\em Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}. 
-\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
-
-\bibitem{handbook09}
-Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock {\em Handbook of Satisfiability}.
-\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
-
-\bibitem{BarrettSST09}
-Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
-\newblock {\em Satisfiability Modulo Theories}. 
-\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
-
+%13
 \bibitem{Clarke04}
@@
 
+%14
 \bibitem{MerzFS12}
@@
 
+%15
 \bibitem{CordeiroF11}
@@
 
+%16
 \bibitem{Ivancic05}
@@
 
+%17
 \bibitem{Cordeiro12}
@@
 
+%18
+\bibitem{EenS03}
+E{\'{e}}n, N., S{\"{o}}rensson, N.:
+\newblock {\em Temporal Induction by Incremental {SAT} Solving.}
+\newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.
+
+%19
+\bibitem{Sheera00}
+Sheeran M., Singh S., St{\aa}lmarck G.:
+\newblock {\em Checking Safety Properties using Induction and a {SAT}-solver.}
+\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
+
+%20
 \bibitem{Bradley07}
@@
 
-\bibitem{Donaldson10}
-Donaldson A., Kroening D., R{\"{u}}mmer P.:
-\newblock {\em {SCRATCH}: A Tool for Automatic Analysis of {DMA} Races.}
-\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
-
-\bibitem{Kinductor}
-Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-\newblock {\em Software Verification using \textit{k}-Induction.}
-\newblock In: International Conference on Static Analysis. LNCS 6887, pp. 351--368, 2011.
-
-\bibitem{EenS03}
-E{\'{e}}n, N., S{\"{o}}rensson, N.:
-\newblock {\em Temporal Induction by Incremental {SAT} Solving.}
-\newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.
-
-\bibitem{GrosseLD09}
-Gro{\ss}e D., Le H., Drechsler R.:
-\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
-\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
-
-\bibitem{Sheera00}
-Sheeran M., Singh S., St{\aa}lmarck G.:
-\newblock {\em Checking Safety Properties using Induction and a {SAT}-solver.}
-\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
-
+%21
 \bibitem{pips:2013}
@@
 
+%22
 \bibitem{Henry:2012}
@@
 
-\bibitem{BeyerSVCOMP15}
-D.~Beyer:
-\newblock {\em Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).} 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
-
+%23
 \bibitem{Gadelha15}
@@
 
-\bibitem{Brain15}
-Brain M., Joshi S., Kroening D., Schrammel P.:
-\newblock {\em Safety Verification and Refutation by k-Invariants and k-Induction.} 
-\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
-
+%24
 \bibitem{Beyer15}
@@
 
+%25
+\bibitem{Brain15}
+Brain M., Joshi S., Kroening D., Schrammel P.:
+\newblock {\em Safety Verification and Refutation by k-Invariants and \textit{k}-Induction.}
+\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+
+%26
 \bibitem{Rocha15}
@@
 
-\bibitem{Brain15}
-Brain M., Joshi S., Kroening D., Schrammel P.:
-\newblock {\em Safety Verification and Refutation by k-Invariants and \textit{k}-Induction.}
-\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+%27
+\bibitem{Donaldson10}
+Donaldson A., Kroening D., R{\"{u}}mmer P.:
+\newblock {\em {SCRATCH}: A Tool for Automatic Analysis of {DMA} Races.}
+\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
 
+%28
+\bibitem{Kinductor}
+Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
+\newblock {\em Software Verification using \textit{k}-Induction.}
+\newblock In: International Conference on Static Analysis. LNCS 6887, pp. 351--368, 2011.
+
+%29
+\bibitem{Behrend15}
+Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
+\newblock {\em Scalable and Optimized Hybrid Verification of Embedded Software}. 
+\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
+
+%30
+\bibitem{Lettnin09}
+Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
+\newblock {\em Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}. 
+\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
+
+%31
+\bibitem{GrosseLD09}
+Gro{\ss}e D., Le H., Drechsler R.:
+\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
+\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
+
+%32
+\bibitem{BeyerSVCOMP15}
+D.~Beyer:
+\newblock {\em Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).} 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
+
+%33
 \bibitem{Inverso14}
@@
 
+%34
 \bibitem{dsv_spin2015}
@@
 
+%35
 \bibitem{ECBS13}
@@
 
+%36
 \bibitem{Sousa15}
@@
 
+%37
 \bibitem{Pereira15}
@@
 
+%38
 \bibitem{morse15}
@@
 
+%39
 \bibitem{civl15}
@@
 
+%40
 \bibitem{esbmc_controller}
@@
 
+%41
 \bibitem{esbmc_filter}
@@
 
+%42
 \bibitem{JMorse15}
@@
 
+%43
 \bibitem{Bessa16}
@@
 
+%44
 \bibitem{MorseCNF13}
@@
 
+%45
 \bibitem{MorseRCN014}
@@
 
+%46
 \bibitem{KahlonWG09}
@@
 
+%47
 \bibitem{Beyer14}

### 2016-03-20 paper.tex
fix references

@@
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
-The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.
+The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
 
@@
 Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
-Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing BMC approaches are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{BeyerSVCOMP15}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing BMC approaches are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
 One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.
@@
 %29
+\bibitem{GrosseLD09}
+Gro{\ss}e D., Le H., Drechsler R.:
+\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
+\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
+
+%30
 \bibitem{Behrend15}
@@
 
-%30
+%31
+\bibitem{dsv_spin2015}
+Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
+\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
+\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
+
+%32
 \bibitem{Lettnin09}
@@
 
-%31
-\bibitem{GrosseLD09}
-Gro{\ss}e D., Le H., Drechsler R.:
-\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
-\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
+%33
+\bibitem{Beyer14}
+D.~Beyer 
+\newblock {Status report on software verification - (competition summary {SV-COMP})}. 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
 
-%32
+%34
 \bibitem{BeyerSVCOMP15}
@@
 
-%33
+%35
 \bibitem{Inverso14}
@@
 
-%34
-\bibitem{dsv_spin2015}
-Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
-\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
-
-%35
+%36
 \bibitem{ECBS13}
@@
 
-%36
+%37
 \bibitem{Sousa15}
@@
 
-%37
+%38
 \bibitem{Pereira15}
@@
 
-%38
+%39
 \bibitem{morse15}
@@
 
-%39
+%40
 \bibitem{civl15}
@@
 
-%40
+%41
 \bibitem{esbmc_controller}
@@
 
-%41
+%42
 \bibitem{esbmc_filter}
@@
 
-%42
+%43
 \bibitem{JMorse15}
@@
 
-%43
+%44
 \bibitem{Bessa16}
@@
 
-%44
+%45
 \bibitem{MorseCNF13}
@@
 
-%45
+%46
 \bibitem{MorseRCN014}
@@
 
-%46
+%47
 \bibitem{KahlonWG09}
@@
 
-%47
-\bibitem{Beyer14}
-D.~Beyer 
-\newblock {Status report on software verification - (competition summary {SV-COMP})}. 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
-
 \end{thebibliography}

### 2016-03-20 paper.tex
review ACHIEVEMENTS

@@
 
-\textbf{(RP4)} State of the art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}.
+\textbf{(RP4)} State of the art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
 In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
-In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; in particular, micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
+In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
+Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for engineers~\cite{Alves15}.
 
 Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent ones for verifying C programs, as observed in the Intl.
-Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}.
+Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
 
@@
 \bibitem{KahlonWG09}
-Kahlon V, Wang C, Gupta A.
-\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order
-  Reduction Technique}}.
+Kahlon V., Wang C., Gupta A.:
+\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
 \newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
+
+%48
+\bibitem{Alves15}
+Alves E. , Cordeiro L., Lima Filho E.:
+\newblock {\em{ Fault Localization in Multi-Threaded C Programs using Bounded Model Checking}}. 
+\newblock In: Brazilian Symposium on Computing Systems Engineering, pp. 96-101, 2015.
+
+%49
+\bibitem{BeyerW15}
+Beyer D., Dangl M., Dietsch D., Heizmann M., Stahlbauer A.:
+\newblock {\em{ Witness validation and stepwise testification across software verifiers}}. 
+\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 721--733, 2015.
+
+%50
+\bibitem{RochaIFM12}
+Rocha H., Barreto R., Cordeiro L., Dias Neto A.:
+\newblock {\em{ Understanding Programming Bugs in ANSI-C Software Using Bounded Model Checking Counter-Examples}}. 
+\newblock In: International Conference on Integrated Formal Methods, LNCS 7321, pp. 128-142, 2012.
 

### 2016-03-21 paper.tex
add figure

@@
 
+Fig.~\ref{intelligent-product} shows examples of embedded systems, which consist of a human-machine interface ({\it e.g.}, keyboard and LCD) and an instrumentation interface ({\it e.g.}, sensor, network, and actuator).
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
@@
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health with serious consequences.
+%
+\begin{figure}[!t]
+	\centering
+	\includegraphics[scale=0.35]{figure1.eps}
+	\caption{Embedded system is part of a well-specified larger system (intelligent product).}
+	\label{intelligent-product}
+\end{figure}
 

### 2016-03-21 paper.tex
add figure 2

--- 2016-03-20 paper.tex
@@
 
+Figure~\ref{intelligent-product} shows examples of embedded systems, which consist of a human-machine interface ({\it e.g.}, keyboard and LCD) and an instrumentation interface ({\it e.g.}, sensor, network, and actuator).
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09} as shown in Figure~\ref{verification-methodologies}.
+%
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{figure2.eps}
+	\caption{Verification methodologies for embedded systems.}
+	\label{verification-methodologies}
+\end{figure}
+
 In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness; additionally, embedded system verification raises additional challenges, such as

### 2016-03-22 paper.tex
review introduction

@@
 
-Figure~\ref{intelligent-product} shows examples of embedded systems, which consist of a human-machine interface ({\it e.g.}, keyboard and LCD) and an instrumentation interface ({\it e.g.}, sensor, network, and actuator).
+Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD) and an instrumentation interface ({\it e.g.}, sensor, network, and actuator).
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
-For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
+For instance, UAVs are a trend on military missions due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health with serious consequences.

### 2016-03-22 paper.tex
review 2. BOUNDED MODEL CHECKING (BMC)

@@
 Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed by Biere {\it et al.} to verify hardware designs~\cite{Biere99,handbook09}.
-The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a SAT solver.
-BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06}.
+The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
+BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06} to deal with increasing software verification complexity.
 
-BMC techniques aims to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of the program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches in the program, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+In general, BMC techniques aims to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of a program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
-SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits, single- and multi-threaded programs.
-In general, those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+From the practical point of view, SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits, single- and multi-threaded programs.
+Those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
 However, the main criticism with respect to BMC techniques relies on completeness since they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
+
 Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuits implementation or programs with loops whose bounds are too large or cannot be statically determined.

### 2016-03-22 paper.tex
review 2.1 Induction-based Verification of C Programs

@@
 
-One possible way to achieve completeness in BMC techniques is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
-The main challenge of such approach relies on computing and strengthening inductive invariants from the program.
+One promising approach to achieve completeness in BMC techniques is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
+The main challenge, however, of such approach relies on computing and strengthening inductive invariants from the program.
 In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding verification conditions to be valid~\cite{Bradley07}.
@@
 The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening, mostly based on interval analysis.
-In particular, Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
 
+Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
@@
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
-Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
+Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
 

### 2016-03-22 paper.tex
review 2.2 Incorporating System Knowledge to Automated Verification Procedures

@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior.
+It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing language issues and general correctness, without taking into account target applications or system behavior.
 Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
@@
 
-Indeed, research in software verification is now taking such considerations into account, and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
 Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions.
+Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space.
 Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

### 2016-03-22 paper.tex
review 4. RESEARCH PROBLEM (RP)

@@
 
-\textbf{(RP1)} providing suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+\textbf{(RP1)} provide suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
 
-\textbf{(RP2)} exploiting SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
+\textbf{(RP2)} exploit SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
 
-\textbf{(RP3)} proving correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
+\textbf{(RP3)} prove correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
 
-\textbf{(RP4)} incorporating knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+\textbf{(RP4)} incorporate knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
 

### 2016-03-22 paper.tex
fix reference
review 5. ACHIEVEMENTS

@@
 Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
-Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing BMC approaches are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
 One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.
@@
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al}.
-was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+was further developed to verify deadlock, lock acquisition ordering, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state space exploration~\cite{Inverso14,civl15}.
@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15,Brain15}.
-However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way in order to be competitive with other model checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Brain15,Rocha15}.
+However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model checking approaches.
 In particular, invariant generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
@@
 %37
+\bibitem{Pereira15}
+Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
+\newblock {\em Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.} 
+\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016.
+
+%38
 \bibitem{Sousa15}
@@
 \newblock In: Global Conference on Consumer Electronics, pp. 179--180, 2015.
-
-%38
-\bibitem{Pereira15}
-Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
-\newblock {\em Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.} 
-\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016.
 

### 2016-03-25 paper.tex
change title
fix typos

--- 2016-03-22 paper.tex
@@
 % Do not put math or special symbols in the title.
-\title{SMT-Based Context-Bounded Model Checking for Embedded Systems: State of the Art and Future Trends}
+\title{SMT-Based Context-Bounded Model Checking for Embedded Systems: Challenges and Future Trends}
 
@@
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor model, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
+Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
 The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again a high-level of reliability~\cite{leeCPS,leeCPS2}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
@@
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
-For instance, UAVs are a trend on military missions due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable.
+For instance, UAVs are a trend on military missions due to the absence of pilots; however, an incorrect plan execution may also cost human lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health with serious consequences.
@@
 
-In general, BMC techniques aims to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of a program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+In general, BMC techniques aim to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of a program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
@@
 There are several invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
-The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariants generation and strengthening, mostly based on interval analysis.
+The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariant generation and strengthening, mostly based on interval analysis.
 
@@
 
-However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification.
+However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing language issues and general correctness, without taking into account target applications or system behavior.
+It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
 Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
@@
 
-Scratch is another example of software-model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
-Such a tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
+Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
+The tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
 
@@
 
-In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness; additionally, embedded system verification raises additional challenges, such as
+In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness; additionally, embedded system verification raises additional challenges, such as
 %
@@
 \textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
-Such approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
-Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing BMC approaches are still limited for specific class of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
+This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
 One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.
@@
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al}.
-was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state space exploration~\cite{Inverso14,civl15}.
+was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}.
 However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded systems applications~\cite{Gadelha15,Rocha15,Brain15}.
-However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way in order to be competitive with other model checking approaches.
-In particular, invariant generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
-However, there is still a lack of studies for exploiting the combination of different invariant generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15}.
+However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+In particular, invariant-generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
+However, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
 implementation aspects.
 
-\textbf{(RP4)} State of the art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
+\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
 In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
@@
 Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
-Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification to another step, where not only code correctness will be addressed, but also full system reliability.
+Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability.
 

### 2016-03-25 paper.tex
remove et al.

@@
 
-Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed by Biere {\it et al.} to verify hardware designs~\cite{Biere99,handbook09}.
+Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}.
 The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
-BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06} to deal with increasing software verification complexity.
+BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
 
@@
 
-Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
@@
 Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
 Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space.
@@
 
-\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
+\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
 This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
@@
 
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al}.
-was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro and Fischer was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}.

### 2016-03-25 paper.tex
minor changes

@@
 The main challenge, however, of such approach relies on computing and strengthening inductive invariants from the program.
-In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding verification conditions to be valid~\cite{Bradley07}.
+In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive so that verifiers can automatically prove program correctness.
@@
 %--------------------------------------
-\subsection{Incorporating System Knowledge to Automated Verification Procedures}
+\subsection{Incorporating System Model to Automated Verification Procedures}
 %--------------------------------------
@@
 It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that they not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
 For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.

### 2016-03-25 paper.tex
minor changes

@@
 
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro and Fischer was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}.
@@
 %The authors would like to thank...
-This research was supported by CNPq $475647$/$2013$-$0$ grant.
+The authors thank M. Dangl for reviewing a draft version of this paper.~This research was supported by CNPq $475647$/$2013$-$0$ grant.
 

### 2016-03-25 paper.tex
minor changes

@@
 
-There are several invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
 The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariant generation and strengthening, mostly based on interval analysis.
@@
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
-The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
+Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
 
@@
 %--------------------------------------
-\subsection{Incorporating System Model to Automated Verification Procedures}
+\subsection{Incorporating System Models to Automated Verification Procedures}
 %--------------------------------------
@@
 Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space.
+Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
 Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09} as shown in Figure~\ref{verification-methodologies}.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
 %
@@
 Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
-One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.
+One possible research direction is to bridge the gap between BMC tools and SMT solvers to propose theories and develop more efficient decision procedures, in order to handle specific classes of programs.
 
@@
 However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
-One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour.
+One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour or to analyze interpolants to prove non-interference of context switches.
 
@@
 In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
-Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for engineers~\cite{Alves15}.
+Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.
 

### 2016-03-26 paper.tex
Corrections to section 1

@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again a high-level of reliability~\cite{leeCPS,leeCPS2}.
+Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
 
-Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD) and an instrumentation interface ({\it e.g.}, sensor, network, and actuator).
+Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}.
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
-For instance, UAVs are a trend on military missions due to the absence of pilots; however, an incorrect plan execution may also cost human lives, which is unacceptable.
-In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health with serious consequences.
+For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human lives, which is unacceptable.
+In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
 %
+
+\textcolor{red}{mudei o texto e coloquei a ref do Kopetz... na figura 1, falta ligar as frases ``multi-core processors'' e ``limited amount of energy'' com um ``with''}
+
 \begin{figure}[!t]
@@
 	\includegraphics[scale=0.35]{figure1.eps}
-	\caption{Embedded system is part of a well-specified larger system (intelligent product).}
+	\caption{An embedded system is part of a well-specified larger system (intelligent product).}
 	\label{intelligent-product}
@@
 Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}.
-The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
+Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
 BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.

### 2016-03-26 paper.tex
corrections to section 2

@@
 Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
-BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
+BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
 
-In general, BMC techniques aim to check the violation of a given (safety) property at a given system's depth, {\it i.e.}, given a transition system \textit{M}, which is derived from the control-flow graph of a program; a property $\phi$ that represents the program correctness and/or the system's behaviour; and a bound of iterations \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth.
+Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
-From the practical point of view, SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits, single- and multi-threaded programs.
-Those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
-However, the main criticism with respect to BMC techniques relies on completeness since they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
+From the practical point of view, SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits and single- and multi-threaded programs.
+Those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the available literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
+Nonetheless, the main criticism with respect to BMC techniques relies on completeness, since they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
 
-Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuits implementation or programs with loops whose bounds are too large or cannot be statically determined.
+Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuit-implementations or programs with loops, whose bounds are too large or cannot be statically determined.
 
@@
 
-One promising approach to achieve completeness in BMC techniques is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
-The main challenge, however, of such approach relies on computing and strengthening inductive invariants from the program.
-In particular, loop invariants computed from the program must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
-As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive so that verifiers can automatically prove program correctness.
+One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
+The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
+In particular, loop invariants, which were computed from programs under verifications, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
 
-There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}.
-The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariant generation and strengthening, mostly based on interval analysis.
+There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis.
 
-Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
-Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
+Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
 Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
@@
 It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that they not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
 For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
 
-Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
+Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware.
+Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
 Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
@@
 
-Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) in the Cell BE processor~\cite{Donaldson10}.
-The tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
+Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}.
+That tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
+If support to other DMA operations were added, Scratch could be adapted to different architectures, {\it i.e.}, the same techniques would be employed, but with a different system behavior/knowledge.
 

### 2016-03-26 paper.tex
corrections to section 3

@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
+State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
 %
@@
 
-In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness; additionally, embedded system verification raises additional challenges, such as
+In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+Additionally, embedded system verification raises additional challenges, such as
 %
@@
 	\item handling of concurrent software;
-	\item platform restrictions; and
-	\item legacy designs.%, which are usually developed with low-level languages.
+	\item platform restrictions;
+	\item legacy designs; and %, which are usually developed with low-level languages.
+	\item \textcolor{blue}{support to different languages and interfaces};
 \end{enumerate}
@@
 The fourth aspect is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
+\textcolor{blue}{Finally, the last one is related to the evolution of development processes and technologies, which may delay the application of suitable verification approaches.}
 

### 2016-03-26 paper.tex
corrections to section 4

@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems, which are (partially) open in present research.
+This position paper tackles four major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.
 
-\textbf{(RP1)} provide suitable encoding into the Satisfiability Modulo Theories (SMT) \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+\textbf{(RP1)} provide suitable encoding into SMT \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
 

### 2016-03-26 paper.tex
corrections to sections 5 and 6

@@
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}.
-However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
 One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour or to analyze interpolants to prove non-interference of context switches.
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15}.
-However, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
@@
 
-\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
+\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems.
+Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
 In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
@@
 Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
+
+\textcolor{red}{to pensando se o Qt nao pode ser descrito como um esforco de prover verificadores para interfaces diferentes}
 
@@
 
-This paper presented the main challenges related to the verification of design correctness in embedded systems.
+This paper presented the main challenges related to the verification of design correctness in embedded systems, and also raises some important side considerations.
 In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification.

### 2016-03-27 paper.tex
remove blue color

@@
 	\item legacy designs; and %, which are usually developed with low-level languages.
-	\item \textcolor{blue}{support to different languages and interfaces};
+	\item support to different languages and interfaces;
 \end{enumerate}
@@
 The fourth aspect is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
-\textcolor{blue}{Finally, the last one is related to the evolution of development processes and technologies, which may delay the application of suitable verification approaches.}
+Finally, the last one is related to the evolution of development processes and technologies, which may delay the application of suitable verification approaches.
 

### 2016-03-27 paper.tex
fix figure
review Incorporating System Models to Automated Verification Procedures

@@
 %
-
-\textcolor{red}{mudei o texto e coloquei a ref do Kopetz... na figura 1, falta ligar as frases ``multi-core processors'' e ``limited amount of energy'' com um ``with''}
-
 \begin{figure}[!t]
@@
 Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects.
-Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
+Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects.
+Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system.
+Here, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
+Indeed, DSVerifier is a useful test tool, which takes into account different realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
 Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

### 2016-03-27 paper.tex
add new references

@@
 
-Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}.
-Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor}.
+Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
 Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
@@
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
-One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour or to analyze interpolants to prove non-interference of context switches.
+One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05} with the goal to remove redundant behaviour or to analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.
 
@@
 %27
+\bibitem{Kinductor}
+Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
+\newblock {\em Software Verification using \textit{k}-Induction.}
+\newblock In: International Symposium on Static Analysis. LNCS 6887, pp. 351--368, 2011.
+
+%28
 \bibitem{Donaldson10}
@@
 \newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
-
-%28
-\bibitem{Kinductor}
-Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-\newblock {\em Software Verification using \textit{k}-Induction.}
-\newblock In: International Conference on Static Analysis. LNCS 6887, pp. 351--368, 2011.
 
@@
 %41
+\bibitem{Grumberg05}
+Grumberg O., Lerda F., Strichman O., Theobald M.:
+\newblock {\em { Proof-guided underapproximation-widening for multi-process systems.}} 
+\newblock In: Symposium on Principles of Programming Languages, pp. 122--131, 2005.
+
+%42
+\bibitem{McMillan11}
+K.~McMillan:
+\newblock {\em {Widening and Interpolation.}} 
+\newblock In: International Symposium on Static Analysis, LCNS 6887, pp. 1, 2011.
+
+%43
 \bibitem{esbmc_controller}
@@
 
-%42
+%44
 \bibitem{esbmc_filter}
@@
 
-%43
+%45
 \bibitem{JMorse15}
@@
 
-%44
+%46
 \bibitem{Bessa16}
@@
 
-%45
+%47
 \bibitem{MorseCNF13}
@@
 
-%46
+%48
 \bibitem{MorseRCN014}
@@
 
-%47
+%49
 \bibitem{KahlonWG09}
@@
 
-%48
+%50
 \bibitem{Alves15}
@@
 
-%49
+%51
 \bibitem{BeyerW15}
@@
 
-%50
+%52
 \bibitem{RochaIFM12}

### 2016-03-27 paper.tex
add a new figure

@@
 
-In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth.
+In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}.
 Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{figure2.eps}
+	\caption{Bounded Model Checking.}
+	\label{bounded-model-checking}
+\end{figure}
 
@@
 	\centering
-	\includegraphics[scale=0.35]{figure2.eps}
+	\includegraphics[scale=0.35]{figure3.eps}
 	\caption{Verification methodologies for embedded systems.}
@@
 %--------------------------------------
-\section{Achievements}
+\section{Achievements and Future Trends}
 \label{achievements}
@@
 Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
-One possible research direction is to bridge the gap between BMC tools and SMT solvers to propose theories and develop more efficient decision procedures, in order to handle specific classes of programs.
+One possible research direction is to bridge the gap between BMC tools and SMT solvers to propose background theories and develop more efficient decision procedures, in order to handle specific classes of programs.
 

### 2016-03-27 paper.tex
minor changes

@@
 %--------------------------------------
-\section{Achievements and Future Trends}
+\section{Current Achievements and Future Trends}
 \label{achievements}
@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
@@
 
-\textcolor{red}{to pensando se o Qt nao pode ser descrito como um esforco de prover verificadores para interfaces diferentes}
+%\textcolor{red}{to pensando se o Qt nao pode ser descrito como um esforco de prover verificadores para interfaces diferentes}
 

### 2016-03-27 paper.tex
new research problem (RP5)

@@
 
+\textcolor{blue}{\textbf{(RP5)} provide tools and approaches capable of addressing different software languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.}
+
 Section~\ref{achievements} outlines contributions toward these research problems.
@@
 Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.
+
+\textcolor{blue}{\textbf{(RP5)} Although ESMBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and languages are often developed, which then claim for suitable verification tools.
+Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check new programs.
+Some work towards that was already presented by Sousa, Cordeiro, and Filho, which employed operational models  for checking Qt-based programs.
+In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the code structure to be checked.
+Such research problem is closely related to the first one (\textbf{(RP1)}) and has the potential to devise a new paradigm in software verification.}
 

### 2016-03-27 paper.tex
corrections to the entire text

@@
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages).
+Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which were computed from programs under verifications, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+In particular, loop invariants, which were computed from programs under verification, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
@@
 
-However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification.
+Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
@@
 
-This position paper tackles four major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.
+This position paper tackles five major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.
 
@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
-In particular, invariant-generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}.
-However, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
 implementation aspects.
@@
 \textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems.
-Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}.
+Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}.
 In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
@@
 \textcolor{blue}{\textbf{(RP5)} Although ESMBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and languages are often developed, which then claim for suitable verification tools.
-Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check new programs.
+Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check different programs.
 Some work towards that was already presented by Sousa, Cordeiro, and Filho, which employed operational models for checking Qt-based programs.
-In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the code structure to be checked.
-Such research problem is closely related to the first one (\textbf{(RP1)}) and has the potential to devise a new paradigm in software verification.}
+In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
+Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.}
 
-Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent ones for verifying C programs, as observed in the Intl.
+Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl.
 Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
@@
 
-This paper presented the main challenges related to the verification of design correctness in embedded systems, and also raises some important side considerations.
+This paper presented the main challenges related to the verification of design correctness, in embedded systems, and also raised some important side considerations.
 In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification.
@@
 Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability.
+\textcolor{blue}{Finally, it seems interesting to provide behavioral models when new application interfaces or language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.}
 

### 2016-03-27 paper.tex
minor changes

@@
 
-\textcolor{blue}{\textbf{(RP5)} provide tools and approaches capable of addressing different software languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.}
+\textbf{(RP5)} provide tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
 
@@
 
-\textcolor{blue}{\textbf{(RP5)} Although ESMBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and languages are often developed, which then claim for suitable verification tools.
-Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check different programs.
-Some work towards that was already presented by Sousa, Cordeiro, and Filho, which employed operational models for checking Qt-based programs.
-In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
-Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.}
+{\textbf{(RP5)} Although ESBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and programming languages are often developed, which require suitable software verification tools.
+Indeed, it would be interesting if a new programming language model could be loaded, which along with a BMC core could check different programs.
+Some work towards that was already presented by Sousa, Cordeiro, and Filho~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics.
+In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
+Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
 
@@
 Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability.
-\textcolor{blue}{Finally, it seems interesting to provide behavioral models when new application interfaces or language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.}
+Finally, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.
 
@@
 %51
+\bibitem{Garcia16}
+Garcia M. , Sousa F., Cordeiro, L., Lima Filho E.: 
+\newblock {\em{ ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
+\newblock In: International SPIN symposium on Model Checking of Software (to appear), 2016.
+
+%52
 \bibitem{BeyerW15}
@@
 
-%52
+%53
 \bibitem{RochaIFM12}

### 2016-12-24 paper.tex
update title

@@
 % Do not put math or special symbols in the title.
-\title{SMT-Based Context-Bounded Model Checking for Embedded Systems: Challenges and Future Trends}
+\title{Verification and Synthesis of Embedded Systems Software: Advances and Challenges}
 
@@
 	\item legacy designs; and %, which are usually developed with low-level languages.
-	\item support to different languages and interfaces;
+	\item support to different programming languages and interfaces;
 \end{enumerate}

### 2017-01-01 paper.tex
review title, abstract, and introduction

@@
 % Do not put math or special symbols in the title.
-\title{Verification and Synthesis of Embedded Systems Software: Advances and Challenges}
+\title{Verification and Synthesis of Embedded Systems: \\ Advances and Challenges}
 
@@
 \author{Lucas C. Cordeiro \\
-Electronic and Information Research Center\\
-Federal University of Amazonas, Brazil\\
-Email: lucascordeiro@ufam.edu.br
+Department of Computer Science\\
+University of Oxford, UK\\
+Email: lucas.cordeiro@cs.ox.ac.uk
 \and
@@
 \begin{abstract}
-The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics.
+The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting new software architectures and legacy designs (usually written in low-level languages).
-The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
-Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent (bounded) model checking application.
+Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
+However, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting new software architectures and legacy designs (usually written in low-level languages).
+The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
+Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
 \end{abstract}
@@
 
-Generally, embedded computer systems perform dedicated functions with high degree of reliability.
+Generally, embedded computer systems perform dedicated functions with a high degree of reliability.
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous, in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again high-level of reliability~\cite{leeCPS,leeCPS2}.
+Embedded systems are ubiquitous in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again a high-level of reliability~\cite{leeCPS,leeCPS2}.
 As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
@@
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
-For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human lives, which is unacceptable.
+For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
@@
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
+In particular, these models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
 
@@
 
+We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis in Section~\ref{Preliminaries}.
+In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
+Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems, while we conclude and describe future work in Section~\ref{conclusions}.
+
 %--------------------------------------
 \section{Bounded Model Checking (BMC)}
+\label{Preliminaries}
 %--------------------------------------
@@
 Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuit-implementations or programs with loops, whose bounds are too large or cannot be statically determined.
-
-%In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the (background) theories supported by SMT solvers, {\it e.g.}, real-time, memory allocation, interrupts, and concurrency.
-Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}.
-Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth.
-In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found.
-Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.
 
@@
 \section{Verification Challenges}
+\label{Verification-Challenges} 
 %--------------------------------------
@@
 \section{Research Problem (RP)}
+\label{Research-Problem}
 %--------------------------------------
@@
 %8
+\bibitem{Abate17}
+Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
+\newblock {\em Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}.
+\newblock In: 20th International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
+
+%9
 \bibitem{Kroening15}
@@
 
-%9
+%10
 \bibitem{Biere99}
@@
 
-%10
+%11
 \bibitem{handbook09}
@@
 
-%11
+%12
 \bibitem{BarrettSST09}
@@
 
-%12
+%13
 \bibitem{Armando06}
@@
 
-%13
+%14
 \bibitem{Clarke04}
@@
 
-%14
+%15
 \bibitem{MerzFS12}
@@
 
-%15
+%16
 \bibitem{CordeiroF11}
@@
 
-%16
+%17
 \bibitem{Ivancic05}
@@
 
-%17
+%18
 \bibitem{Cordeiro12}
@@
 
-%18
+%19
 \bibitem{EenS03}
@@
 
-%19
+%20
 \bibitem{Sheera00}
@@
 
-%20
+%21
 \bibitem{Bradley07}
@@
 
-%21
+%22
 \bibitem{pips:2013}
@@
 
-%22
+%23
 \bibitem{Henry:2012}
@@
 
-%23
+%24
 \bibitem{Gadelha15}
@@
 
-%24
+%25
 \bibitem{Beyer15}
@@
 
-%25
+%26
 \bibitem{Brain15}
@@
 
-%26
+%27
 \bibitem{Rocha15}
@@
 
-%27
+%28
 \bibitem{Kinductor}
@@
 
-%28
+%29
 \bibitem{Donaldson10}
@@
 
-%29
+%30
 \bibitem{GrosseLD09}
@@
 
-%30
+%31
 \bibitem{Behrend15}
@@
 
-%31
+%32
 \bibitem{dsv_spin2015}
@@
 
-%32
+%33
 \bibitem{Lettnin09}
@@
 
-%33
+%34
 \bibitem{Beyer14}
@@
 
-%34
+%35
 \bibitem{BeyerSVCOMP15}
@@
 
-%35
+%36
 \bibitem{Inverso14}
@@
 
-%36
+%37
 \bibitem{ECBS13}
@@
 
-%38
+%39
 \bibitem{Sousa15}
@@
 
-%39
+%40
 \bibitem{morse15}
@@
 
-%40
+%41
 \bibitem{civl15}
@@
 
-%41
+%42
 \bibitem{Grumberg05}
@@
 
-%42
+%43
 \bibitem{McMillan11}
@@
 
-%43
+%44
 \bibitem{esbmc_controller}
@@
 
-%44
+%45
 \bibitem{esbmc_filter}
@@
 
-%45
+%46
 \bibitem{JMorse15}
@@
 
-%46
+%47
 \bibitem{Bessa16}
@@
 
-%47
+%48
 \bibitem{MorseCNF13}
@@
 
-%48
+%49
 \bibitem{MorseRCN014}
@@
 
-%49
+%50
 \bibitem{KahlonWG09}
@@
 
-%50
+%51
 \bibitem{Alves15}
@@
 
-%51
+%52
 \bibitem{Garcia16}
@@
 
-%52
+%53
 \bibitem{BeyerW15}
@@
 
-%53
+%54
 \bibitem{RochaIFM12}

### 2017-01-01 paper.tex
Added a new reference related to DepthK

@@
 
-There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
 The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis.
 
-Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}.
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
+%\section*{Acknowledgment}
 
 %The authors would like to thank...
-The authors thank M. Dangl for reviewing a draft version of this paper.~This research was supported by CNPq $475647$/$2013$-$0$ grant.
+%The authors thank M. Dangl for reviewing a draft version of this paper.~This research was supported by CNPq $475647$/$2013$-$0$ grant.
 
@@
 \newblock {\em Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}.
-\newblock In: 20th International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
+\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
 
@@
 \newblock {\em Model Checking Embedded C Software using \textit{k}-Induction and Invariants.}
-\newblock V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
+\newblock In: V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
 
@@
 %29
+\bibitem{Rocha17}
+Rocha, W., Rocha, H., Ismail H., Cordeiro, L., Fischer, B.:
+\newblock {\em DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}. 
+\newblock In: International Conference on Tools and Algorithms for the Construction and Analysis of Systems, 2017 (to appear).
+
+%30
 \bibitem{Donaldson10}
@@
 
-%30
+%31
 \bibitem{GrosseLD09}
@@
 
-%31
+%32
 \bibitem{Behrend15}
@@
 
-%32
+%33
 \bibitem{dsv_spin2015}
@@
 
-%33
+%34
 \bibitem{Lettnin09}
@@
 
-%34
+%35
 \bibitem{Beyer14}
@@
 
-%35
+%36
 \bibitem{BeyerSVCOMP15}
@@
 
-%36
+%37
 \bibitem{Inverso14}
@@
 
-%37
+%38
 \bibitem{ECBS13}
@@
 
-%38
+%39
 \bibitem{Pereira15}
@@
 
-%39
+%40
 \bibitem{Sousa15}
@@
 
-%40
+%41
 \bibitem{morse15}
@@
 
-%41
+%42
 \bibitem{civl15}
@@
 
-%42
+%43
 \bibitem{Grumberg05}
@@
 
-%43
+%44
 \bibitem{McMillan11}
@@
 
-%44
+%45
 \bibitem{esbmc_controller}
@@
 
-%45
+%46
 \bibitem{esbmc_filter}
@@
 
-%46
+%47
 \bibitem{JMorse15}
@@
 
-%47
+%48
 \bibitem{Bessa16}
@@
 
-%48
+%49
 \bibitem{MorseCNF13}
@@
 
-%49
+%50
 \bibitem{MorseRCN014}
@@
 
-%50
+%51
 \bibitem{KahlonWG09}
@@
 
-%51
+%52
 \bibitem{Alves15}
@@
 
-%52
+%53
 \bibitem{Garcia16}
@@
 
-%53
+%54
 \bibitem{BeyerW15}
@@
 
-%54
+%55
 \bibitem{RochaIFM12}

### 2017-01-01 paper.tex
Added new reference related to DSVerifier

@@
 %--------------------------------------
-\subsection{Incorporating System Models to Automated Verification Procedures}
+\subsection{Incorporating System Models to Automated \\ Verification and Synthesis Procedures}
 %--------------------------------------
@@
 Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects.
-Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system.
+Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.
 Here, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
@@
 %32
+\bibitem{dsv_spin2015}
+Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
+\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
+\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
+
+%33
+\bibitem{Bessa17}
+Bessa, I., Ismail, H., Palhares, R., Cordeiro, L., Chaves Filho, J.:
+\newblock {\em Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}. 
+\newblock In: IEEE Transactions on Computers, 2017 (to appear).
+
+%34
 \bibitem{Behrend15}
@@
 
-%33
-\bibitem{dsv_spin2015}
-Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
-\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
-
-%34
+%35
 \bibitem{Lettnin09}
@@
 
-%35
+%36
 \bibitem{Beyer14}
@@
 
-%36
+%37
 \bibitem{BeyerSVCOMP15}
@@
 
-%37
+%38
 \bibitem{Inverso14}
@@
 
-%38
+%39
 \bibitem{ECBS13}
@@
 
-%39
+%40
 \bibitem{Pereira15}
@@
 
-%40
+%41
 \bibitem{Sousa15}
@@
 
-%41
+%42
 \bibitem{morse15}
@@
 
-%42
+%43
 \bibitem{civl15}
@@
 
-%43
+%44
 \bibitem{Grumberg05}
@@
 
-%44
+%45
 \bibitem{McMillan11}
@@
 
-%45
+%46
 \bibitem{esbmc_controller}
 Bessa I., Abreu R., Cordeiro L., Filho J.:
-\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers)}.}
+\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers}.}
 \newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.
 
-%46
+%47
 \bibitem{esbmc_filter}
@@
 
-%47
+%48
 \bibitem{JMorse15}
@@
 
-%48
+%49
 \bibitem{Bessa16}
@@
 
-%49
+%50
 \bibitem{MorseCNF13}
@@
 
-%50
+%51
 \bibitem{MorseRCN014}
@@
 
-%51
+%52
 \bibitem{KahlonWG09}
@@
 
-%52
+%53
 \bibitem{Alves15}
@@
 
-%53
+%54
 \bibitem{Garcia16}
@@
 
-%54
+%55
 \bibitem{BeyerW15}
@@
 
-%55
+%56
 \bibitem{RochaIFM12}

### 2017-01-01 paper.tex
review Section "Incorporating System Models to Automated Verification and Synthesis Procedures"

@@
 
-Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}.
+In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}.
+In particular, a counter-example guided inductive synthesis approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets as well as FWL uncertainties in digital controllers.
+DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models taken from the literature within minutes.
+
+Apart from the control system domain, Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}.
 That tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.

### 2017-01-01 paper.tex
Added a new reference

@@
 Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects.
+Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects.
 Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.
 Here, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
-Indeed, DSVerifier is a useful test tool, which takes into account different realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
+Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
 Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
@@
 %33
+\bibitem{Monteiro16}
+F.~Monteiro:
+\newblock {\em Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}. 
+\newblock In: 24th {ACM} {SIGSOFT} International Symposium on Foundations of Software Engineering, pp. 1151--1153, 2016.
+
+%34
 \bibitem{Bessa17}
@@
 
-%34
+%35
 \bibitem{Behrend15}
@@
 
-%35
+%36
 \bibitem{Lettnin09}
@@
 
-%36
+%37
 \bibitem{Beyer14}
@@
 
-%37
+%38
 \bibitem{BeyerSVCOMP15}
@@
 
-%38
+%39
 \bibitem{Inverso14}
@@
 
-%39
+%40
 \bibitem{ECBS13}
@@
 
-%40
+%41
 \bibitem{Pereira15}
@@
 
-%41
+%42
 \bibitem{Sousa15}
@@
 
-%42
+%43
 \bibitem{morse15}
@@
 
-%43
+%44
 \bibitem{civl15}
@@
 
-%44
+%45
 \bibitem{Grumberg05}
@@
 
-%45
+%46
 \bibitem{McMillan11}
@@
 
-%46
+%47
 \bibitem{esbmc_controller}
@@
 
-%47
+%48
 \bibitem{esbmc_filter}
@@
 
-%48
+%49
 \bibitem{JMorse15}
@@
 
-%49
+%50
 \bibitem{Bessa16}
@@
 \newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
-\newblock In: Design Automation for Embedded Systems (to appear), 2016.
+\newblock In: Design Automation for Embedded Systems 20(2): pp. 95--126, 2016.
 
-%50
+%51
 \bibitem{MorseCNF13}
@@
 
-%51
+%52
 \bibitem{MorseRCN014}
@@
 
-%52
+%53
 \bibitem{KahlonWG09}
@@
 
-%53
+%54
 \bibitem{Alves15}
@@
 
-%54
+%55
 \bibitem{Garcia16}
@@
 
-%55
+%56
 \bibitem{BeyerW15}
@@
 
-%56
+%57
 \bibitem{RochaIFM12}

### 2017-01-01 paper.tex
Added a section about Program Synthesis

@@
 Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
+
+%--------------------------------------
+\subsection{Program Synthesis via Counter-Example \\ Guided Inductive Synthesis (CEGIS)}
+%--------------------------------------
+
+The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\phi$.
+In particular, program synthesis is automatically performed by engines that use a correctness specification $\phi$ as the starting point, and then produce a sequence of candidate programs that satisfy $\phi$~\cite{Alur13}.
+In this respect, a given candidate program $p$ is iteratively refined to match specification $\phi$ more closely.
+
+Counter-Example Guided Inductive Synthesis (CEGIS) is currently the most popular approach to program synthesis in practice~\cite{David15}.
+CEGIS has close connections to algorithmic debugging using counterexamples and abstraction refinement.
+These two connections, however, involve two specific steps: synthesizing a repair in the former case, and synthesizing an abstraction function in the latter.
+
+Nowadays, program synthesis engines that implement Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} can automatically produce programs for a large variety of specifications due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
 
@@
 %32
+\bibitem{Alur13}
+Alur R., Bodík R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
+\newblock {\em Syntax-guided synthesis}. 
+\newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
+
+%33
+\bibitem{sketch}
+A.~Solar-Lezama.
+\newblock {\em Program sketching}.
+\newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
+
+%34
+\bibitem{David15}
+C.~David, D.~Kroening, and M.~Lewis.
+\newblock {\em Using program synthesis for program analysis}.
+\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
+
+%35
+\bibitem{Sharma14}
+R.~Sharma and A.~Aiken.
+\newblock {\em From invariant checking to invariant inference using randomized search}.
+\newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
+
+%36
 \bibitem{dsv_spin2015}
@@
 
-%33
+%37
 \bibitem{Monteiro16}
@@
 
-%34
+%38
 \bibitem{Bessa17}
@@
 
-%35
+%39
 \bibitem{Behrend15}
@@
 
-%36
+%40
 \bibitem{Lettnin09}
@@
 
-%37
+%41
 \bibitem{Beyer14}
@@
 
-%38
+%42
 \bibitem{BeyerSVCOMP15}
@@
 
-%39
+%43
 \bibitem{Inverso14}
@@
 
-%40
+%44
 \bibitem{ECBS13}
@@
 
-%41
+%45
 \bibitem{Pereira15}
@@
 
-%42
+%46
 \bibitem{Sousa15}
@@
 
-%43
+%47
 \bibitem{morse15}
@@
 
-%44
+%48
 \bibitem{civl15}
@@
 
-%45
+%49
 \bibitem{Grumberg05}
@@
 
-%46
+%50
 \bibitem{McMillan11}
@@
 
-%47
+%51
 \bibitem{esbmc_controller}
@@
 
-%48
+%52
 \bibitem{esbmc_filter}
@@
 
-%49
+%53
 \bibitem{JMorse15}
@@
 
-%50
+%54
 \bibitem{Bessa16}
@@
 
-%51
+%55
 \bibitem{MorseCNF13}
@@
 
-%52
+%56
 \bibitem{MorseRCN014}
@@
 
-%53
+%57
 \bibitem{KahlonWG09}
@@
 
-%54
+%58
 \bibitem{Alves15}
@@
 
-%55
+%59
 \bibitem{Garcia16}
@@
 
-%56
+%60
 \bibitem{BeyerW15}
@@
 
-%57
+%61
 \bibitem{RochaIFM12}

### 2017-01-01 paper.tex
added new figures

@@
 
-Counter-Example Guided Inductive Synthesis (CEGIS) is currently the most popular approach to program synthesis in practice~\cite{David15}.
-CEGIS has close connections to algorithmic debugging using counterexamples and abstraction refinement.
-These two connections, however, involve two specific steps: synthesizing a repair in the former case, and synthesizing an abstraction function in the latter.
+Counter-Example Guided Inductive Synthesis (CEGIS) is currently one of the the most popular approaches to program synthesis in practice~\cite{David15}.
+Figures~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement.
+These connections involve two specific steps: synthesizing a repair in the former case, and synthesizing an abstraction function in the latter.
+%
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.12]{figure4.eps}
+	\caption{Counter-Example Guided Inductive Synthesis.}
+	\label{Counter-Example-Guided-Inductive-Synthesis}
+\end{figure}
 
@@
 \bibitem{Alur13}
-Alur R., Bodík R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
+Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
 \newblock {\em Syntax-guided synthesis}. 

### 2017-01-01 paper.tex
review Program Synthesis via Counter-Example Guided Inductive Synthesis (CEGIS)

--- 2017-01-01 paper.tex
@@
 % Do not put math or special symbols in the title.
-\title{Verification and Synthesis of Embedded Systems: \\ Advances and Challenges}
+\title{Automated Verification and Synthesis of Embedded Systems: Advances and Challenges}
 
@@
 Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
+
+%--------------------------------------
+\subsection{Program Synthesis via Counter-Example \\ Guided Inductive Synthesis (CEGIS)}
+%--------------------------------------
+
+The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$.
+In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then produce incrementally a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
+As a result, a given candidate program $p$ is iteratively refined to match the specification $\sigma$ more closely.
+Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis currently used in practice~\cite{David15}.
+Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.
+
+The input specification provided to the program synthesizer is of the form $\exists \vec{P} .
+\forall \vec{x}.
+\sigma(\vec{x}, \vec{P})$ where $\vec{P}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula.
+The ground terms is interpreted over some finite domain $\mathcal{D}$.
+The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally.
+Given the aforementioned specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{P}$ satisfying the specification $\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
+If {\sc synthesize} succeeds in finding a witness~$\vec{P}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x}, \vec{P})$ for all $\vec{x}\in\mathcal{D}$).
+If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
+It is worth noting that each iteration of the loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
+Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times.
+%
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.12]{figure4.eps}
+	\caption{Counter-Example Guided Inductive Synthesis.}
+	\label{Counter-Example-Guided-Inductive-Synthesis}
+\end{figure}
+
+Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
 
@@
 %32
+\bibitem{Alur13}
+Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
+\newblock {\em Syntax-guided synthesis}. 
+\newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
+
+%33
+\bibitem{sketch}
+A.~Solar-Lezama.
+\newblock {\em Program sketching}.
+\newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
+
+%34
+\bibitem{David15}
+C.~David, D.~Kroening, and M.~Lewis.
+\newblock {\em Using program synthesis for program analysis}.
+\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
+
+%35
+\bibitem{Sharma14}
+R.~Sharma and A.~Aiken.
+\newblock {\em From invariant checking to invariant inference using randomized search}.
+\newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
+
+%36
 \bibitem{dsv_spin2015}

### 2017-01-02 paper.tex
review "Verification and Synthesis Challenges for Embedded Systems"

@@
 In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}.
-In particular, a counter-example guided inductive synthesis approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets as well as FWL uncertainties in digital controllers.
-DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models taken from the literature within minutes.
+In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets as well as FWL uncertainties in digital controllers.
+DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models taken from the control literature within minutes.
 
@@
 %--------------------------------------
-\section{Verification Challenges}
+\section{Verification and Synthesis \\ Challenges for Embedded Systems}
 \label{Verification-Challenges} 
@@
 The fourth aspect is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
+In particular, software developed for that type of embedded systems has been extensively tested and verified, and also optimized for efficiency over years of development.
+Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.
 Finally, the last one is related to the evolution of development processes and technologies, which may delay the application of suitable verification approaches.
 
-%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification.
-These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.
+Those five challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
+In particular, synthesizers for those systems need to consider the effects of the quantizers (A/D and D/A converters) when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
+intelligent product in Figure~\ref{intelligent-product}).
+Additionally,  finite-precision arithmetic and their related rounding errors need to be considered when correct-by-construction code is generated for embedded systems.
 

### 2017-01-02 paper.tex
review Research Problem (RP)

@@
 
-This position paper tackles five major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.
+This position paper tackles six major problems in computer-aided verification and synthesis for embedded systems, which are (partially) open in current published research.
 
@@
 \textbf{(RP5)} provide tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
+
+\textbf{(RP6)} develop automated synthesis approaches that are algorithmically and numerically sound, in order to handle control software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
 

### 2017-01-02 paper.tex
review Current Achievements and Future Trends

@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
@@
 Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}.
+In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16,Chaves17}.
 In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
@@
 
-Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl.
-Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
+{\textbf{(RP6)} State-of-the-art synthesis approaches for embedded (control) systems typically disregard the platform in which the embedded system software operates and restrict itself to generate code that do not take into account FWL effects.
+The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe) due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}.
+Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17}.
+However, there is still little evidence whether that approach can scale for larger systems modeled by other types of representations ({\it e.g.}, state-space).
+Another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
 
-%\textcolor{red}{to pensando se o Qt nao pode ser descrito como um esforco de prover verificadores para interfaces diferentes}
-
-%Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.
+Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, SMACK~\cite{Haran15}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl.
+Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15,Beyer16}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
+The increasing number of verification tasks being added every year to the competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques.
+As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and increase further the reliability of the verified (embedded) system.
 
@@
 
-This paper presented the main challenges related to the verification of design correctness, in embedded systems, and also raised some important side considerations.
-In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification.
+This paper presented the main challenges related to the verification of design correctness, in embedded systems, and also raised some important side considerations about synthesis.
+In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification and synthesis.
 Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
@@
 %55
+\bibitem{Chaves17}
+Chaves L., Bessa I., Cordeiro L.:
+\newblock {\em { DSValidator: An Automated Counterexample Reproducibility Tool for Digital Systems}}. 
+\newblock In: CoRR abs/1610.07066, 2016.
+
+%56
 \bibitem{MorseCNF13}
@@
 
-%56
+%57
 \bibitem{MorseRCN014}
@@
 
-%57
+%58
 \bibitem{KahlonWG09}
@@
 
-%58
+%59
 \bibitem{Alves15}
@@
 
-%59
+%60
 \bibitem{Garcia16}
@@
 
-%60
+%61
+\bibitem{Jackson16}
+Jackson D., Vaziri M.:
+\newblock {\em{ Correct or usable? the limits of traditional verification (impact paper award)}}. 
+\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 11, 2016.
+
+%62
+\bibitem{Araujo16}
+Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
+\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
+\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
+
+%63
+\bibitem{Haran15}
+Haran A., Carter M., Emmi M., Lal A., Qadeer S., Rakamaric Z.:
+\newblock {\em{ SMACK+Corral: A Modular Verifier - (Competition Contribution)}}. 
+\newblock In: 21st International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 451--454, 2015.
+
+%64
+\bibitem{Beyer16}
+Beyer, D.:
+\newblock {\em{ Reliable and Reproducible Competition Results with BenchExec and Witnesses (Report on SV-COMP 2016)}}. 
+\newblock In: 22nd International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9636, pp. 887--904, 2016.
+
+%65
 \bibitem{BeyerW15}
@@
 
-%61
+%66
 \bibitem{RochaIFM12}

### 2017-01-02 paper.tex
review conclusions

@@
 Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
-Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability.
+Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
 Finally, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.
 
-As future work, the main goal of this research is to extend BMC as a design and verification tool for achieving correct-by-construction embedded system implementations.
+As future work, the main goal of this research is to extend BMC as a design, verification, and synthesis tool for achieving correct-by-construction embedded system implementations.
 Special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification, probably through behavior models.

### 2017-01-05 paper.tex
Added description about handling of non-linear and non-convex optimization problems

@@
 \begin{enumerate}
-	\item time and energy constraints;
-	\item handling of concurrent software;
-	\item platform restrictions;
-	\item legacy designs; and %, which are usually developed with low-level languages.
-	\item support to different programming languages and interfaces;
+	\item time and energy constraints.
+	\item handling of concurrent software.
+	\item platform restrictions.
+	\item legacy designs. %, which are usually developed with low-level languages.
+	\item support to different programming languages and interfaces.
+	\item handling of non-linear and non-convex optimization problems.
 \end{enumerate}
@@
 Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.
-Finally, the last one is related to the evolution of development processes and technologies, which may delay the application of suitable verification approaches.
+The fifth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches if verifiers and synthesizers do not support different programming languages and interfaces.
+Finally, the last one is related to the widespread use of embedded systems in autonomous vehicles navigation systems~\cite{Adouane16}, which demand optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
 
@@
 %--------------------------------------
-\section{Current Achievements and Future Trends}
+\section{Current Achievements and \\ Future Trends}
 \label{achievements}
@@
 However, there is still little evidence whether that approach can scale for larger systems modeled by other types of representations ({\it e.g.}, state-space).
-Another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
+In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
+As a consequence, verifiers and synthesizers need to handle a wide range of functions, including non-linear and non-convex optimization problems based on both fixed- and floating-point arithmetic.
 
@@
 H.~Kopetz:
-\newblock {\em Real-Time Systems - Design Principles for Distributed Embedded Applications}.
+\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
 \newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
@@
 Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
-\newblock {\em Hierarchical energy management system for multi-source multi-product microgrids}
+\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
 \newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
@@
 Lee E.:
-\newblock {\em Cyber-physical Systems: Design Challenges.} 
+\newblock {\em {Cyber-physical Systems: Design Challenges}}.
 \newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
@@
 Lee E.:
-\newblock {\em The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}. 
+\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
 \newblock In: Sensors 15(3): pp. 4837--4869, 2015.
@@
 Groza A., Letia I., Goron A., Zaporojan S.: 
-\newblock {\em A formal approach for identifying assurance deficits in unmanned aerial vehicle software}
+\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
 \newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
@@
 Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
-\newblock {\em Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints.} 
+\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
 \newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
@@
 Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
-\newblock {\em Cyber Physical System Challenges for Human-in-the-Loop Control}. 
+\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
 \newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
@@
 Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
-\newblock {\em Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}.
+\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
 \newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
@@
 Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
-\newblock {\em Effective Verification of Low-Level Software with Nested Interrupts}. 
+\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
 \newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
@@
 Biere A., Cimatti A., Clarke E., Zhu Y.:
-\newblock {\em Symbolic Model Checking without BDDs}. 
+\newblock {\em {Symbolic Model Checking without BDDs}}. 
 \newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
@@
 Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock {\em Handbook of Satisfiability}.
+\newblock {\em {Handbook of Satisfiability}}.
 \newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
@@
 Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
-\newblock {\em Satisfiability Modulo Theories}. 
+\newblock {\em {Satisfiability Modulo Theories}}. 
 \newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
@@
 Armando A., Mantovani J., Platania L.:
-\newblock {\em Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}. 
+\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
 \newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
@@
 Clarke E., Kroening D., Lerda F.:
-\newblock {\em A Tool for Checking {ANSI-C} Programs.}
+\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
 \newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.
@@
 Merz F., Falke S., Sinz C.:
-\newblock {\em {LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}.}
+\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
 \newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
@@
 Cordeiro L., Fischer B.:
-\newblock {\em Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking.}
+\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
 \newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
@@
 Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
-\newblock {\em Model Checking {C} Programs using {F-Soft}.}
+\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
 \newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
@@
 Cordeiro L., Fischer B., Marques{-}Silva J.:
-\newblock {\em {SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}.
-\newblock {IEEE} Trans. Software Eng. \textbf{38}(4), pp. 957--974, 2012.
+\newblock {\em {{SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}}.
+\newblock {IEEE} Trans. Software Eng. 38(4), pp. 957--974, 2012.
 
@@
 E{\'{e}}n, N., S{\"{o}}rensson, N.:
-\newblock {\em Temporal Induction by Incremental {SAT} Solving.}
-\newblock Electronic Notes in Theoretical Computer Science \textbf{89}(4), pp. 543 -- 560, 2003.
+\newblock {\em {Temporal Induction by Incremental {SAT} Solving}}.
+\newblock Electronic Notes in Theoretical Computer Science 89(4), pp. 543 -- 560, 2003.
 
@@
 Sheeran M., Singh S., St{\aa}lmarck G.:
-\newblock {\em Checking Safety Properties using Induction and a {SAT}-solver.}
+\newblock {\em {Checking Safety Properties using Induction and a {SAT}-solver}}.
 \newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
@@
 Bradley A., Manna Z.:
-\newblock {\em The calculus of computation - decision procedures with applications to verification}. 
+\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
 \newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
@@
 ParisTech M.:
-\newblock {\em PIPS: Automatic Parallelizer and Code Transformation Framework.} 
+\newblock {\em {PIPS: Automatic Parallelizer and Code Transformation Framework}}. 
 \newblock Accessed 21 February 2016
@@
 Henry J., Monniaux D., Moy M.: 
-\newblock {\em PAGAI: A Path Sensitive Static Analyser.} 
+\newblock {\em {PAGAI: A Path Sensitive Static Analyser}}. 
 \newblock In: Electronic Notes in Theoretical Computer Science, pp. 15--25, 2012.
@@
 Gadelha M., Ismail H., Cordeiro L.:
-\newblock {\em Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction.}
+\newblock {\em {Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction}}.
 \newblock In: International Journal on Software Tools for Technology Transfer (to appear), 2015.
@@
 Beyer D., Dangl M., Wendler P.:
-\newblock {\em Boosting \textit{k}-Induction with Continuously-Refined Invariants.}
+\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
 \newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
@@
 Brain M., Joshi S., Kroening D., Schrammel P.:
-\newblock {\em Safety Verification and Refutation by k-Invariants and \textit{k}-Induction.}
+\newblock {\em {Safety Verification and Refutation by k-Invariants and \textit{k}-Induction}}.
 \newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
@@
 Rocha H., Ismail H., Cordeiro L., Barreto R.:
-\newblock {\em Model Checking Embedded C Software using \textit{k}-Induction and Invariants.}
+\newblock {\em {Model Checking Embedded C Software using \textit{k}-Induction and Invariants}}.
 \newblock In: V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
@@
 Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-\newblock {\em Software Verification using \textit{k}-Induction.}
+\newblock {\em {Software Verification using \textit{k}-Induction}}.
 \newblock In: International Symposium on Static Analysis. LNCS 6887, pp. 351--368, 2011.
@@
 Rocha, W., Rocha, H., Ismail H., Cordeiro, L., Fischer, B.:
-\newblock {\em DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}. 
+\newblock {\em {DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}}. 
 \newblock In: International Conference on Tools and Algorithms for the Construction and Analysis of Systems, 2017 (to appear).
@@
 Donaldson A., Kroening D., R{\"{u}}mmer P.:
-\newblock {\em {SCRATCH}: A Tool for Automatic Analysis of {DMA} Races.}
+\newblock {\em {{SCRATCH}: A Tool for Automatic Analysis of {DMA} Races}}.
 \newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
@@
 Gro{\ss}e D., Le H., Drechsler R.:
-\newblock {\em Induction-based Formal Verification of {SystemC TLM} Designs.}
+\newblock {\em {Induction-based Formal Verification of {SystemC TLM} Designs}}.
 \newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
@@
 Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
-\newblock {\em Syntax-guided synthesis}. 
+\newblock {\em {Syntax-guided synthesis}}. 
 \newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
@@
 A.~Solar-Lezama.
-\newblock {\em Program sketching}.
+\newblock {\em {Program sketching}}.
 \newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
@@
 C.~David, D.~Kroening, and M.~Lewis.
-\newblock {\em Using program synthesis for program analysis}.
+\newblock {\em {Using program synthesis for program analysis}}.
 \newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
@@
 R.~Sharma and A.~Aiken.
-\newblock {\em From invariant checking to invariant inference using randomized search}.
+\newblock {\em {From invariant checking to invariant inference using randomized search}}.
 \newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
@@
 Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-\newblock {\em DSVerifier: A Bounded Model Checking Tool for Digital Systems}. 
+\newblock {\em {DSVerifier: A Bounded Model Checking Tool for Digital Systems}}. 
 \newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
@@
 F.~Monteiro:
-\newblock {\em Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}. 
+\newblock {\em {Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}}.
 \newblock In: 24th {ACM} {SIGSOFT} International Symposium on Foundations of Software Engineering, pp. 1151--1153, 2016.
@@
 Bessa, I., Ismail, H., Palhares, R., Cordeiro, L., Chaves Filho, J.:
-\newblock {\em Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}. 
+\newblock {\em {Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}}. 
 \newblock In: IEEE Transactions on Computers, 2017 (to appear).
@@
 Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
-\newblock {\em Scalable and Optimized Hybrid Verification of Embedded Software}. 
+\newblock {\em {Scalable and Optimized Hybrid Verification of Embedded Software}}. 
 \newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
@@
 Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
-\newblock {\em Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}. 
+\newblock {\em {Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}}. 
 \newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
@@
 %41
+\bibitem{Adouane16}
+Vilca J., Adouane L., Mezouar Y.:
+\newblock {\em {Optimal Multi-Criteria Waypoint Selection for Autonomous Vehicle Navigation in Structured Environment}}. 
+\newblock In: Journal of Intelligent and Robotic Systems 82(2): pp. 301--324, 2016.
+
+%42
 \bibitem{Beyer14}
 D.~Beyer 
-\newblock {Status report on software verification - (competition summary {SV-COMP})}. 
+\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
 \newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
 
-%42
+%43
 \bibitem{BeyerSVCOMP15}
 D.~Beyer:
-\newblock {\em Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).} 
+\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
 \newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
 
-%43
+%44
 \bibitem{Inverso14}
 Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
-\newblock {\em Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}
+\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
 \newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
 
-%44
+%45
 \bibitem{ECBS13}
@@
 
-%45
+%46
 \bibitem{Pereira15}
 Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
-\newblock {\em Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking.} 
+\newblock {\em {Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking}}.
 \newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016.
 
-%46
+%47
 \bibitem{Sousa15}
@@
 
-%47
+%48
 \bibitem{morse15}
@@
 
-%48
+%49
 \bibitem{civl15}
@@
 
-%49
+%50
 \bibitem{Grumberg05}
@@
 
-%50
+%51
 \bibitem{McMillan11}
@@
 
-%51
+%52
 \bibitem{esbmc_controller}
@@
 
-%52
+%53
 \bibitem{esbmc_filter}
@@
 
-%53
+%54
 \bibitem{JMorse15}
@@
 
-%54
+%55
 \bibitem{Bessa16}
@@
 
-%55
+%56
 \bibitem{Chaves17}
@@
 
-%56
+%57
 \bibitem{MorseCNF13}
 Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em Handling unbounded loops with {ESBMC} 1.20.}
+\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
 \newblock In: Tools and Algorithms for the Construction and Analysis of
@@
 
-%57
+%58
 \bibitem{MorseRCN014}
@@
 
-%58
+%59
 \bibitem{KahlonWG09}
@@
 
-%59
+%60
 \bibitem{Alves15}
@@
 
-%60
+%61
 \bibitem{Garcia16}
@@
 
-%61
+%62
 \bibitem{Jackson16}
@@
 
-%62
+%63
 \bibitem{Araujo16}
@@
 
-%63
+%64
 \bibitem{Haran15}
@@
 
-%64
+%65
 \bibitem{Beyer16}
@@
 
-%65
+%66
 \bibitem{BeyerW15}
@@
 
-%66
+%67
 \bibitem{RochaIFM12}

### 2017-01-07 paper.tex
review Program Synthesis via Counter-Example Guided Inductive Synthesis (CEGIS)

@@
 
-The input specification provided to the program synthesizer is of the form $\exists \vec{P} .
+The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .
 \forall \vec{x}.
-\sigma(\vec{x}, \vec{P})$ where $\vec{P}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula.
-The ground terms is interpreted over some finite domain $\mathcal{D}$.
+\sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers.
+The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can encoded using the SMT solvers bit-vectors theory.
 The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally.
-Given the aforementioned specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{P}$ satisfying the specification $\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
-If {\sc synthesize} succeeds in finding a witness~$\vec{P}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x}, \vec{P})$ for all $\vec{x}\in\mathcal{D}$).
+Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
+If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
 If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
-It is worth noting that each iteration of the loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
-Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times.
 %
@@
 \end{figure}
+
+Note that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
+Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times.
+However, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
+In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase to learn from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
 

### 2017-01-07 paper.tex
minor changes

@@
 
-\textbf{(RP6)} develop automated synthesis approaches that are algorithmically and numerically sound, in order to handle control software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
+\textbf{(RP6)} develop automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
 

### 2017-01-26 paper.tex
changes int section 1 and the beginning of section 2

@@
 The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, and consumer electronics.
-Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power demands.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-However, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous in modern day information systems, and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, and in cyber-physical systems (CPS), which demand short development cycles and again a high-level of reliability~\cite{leeCPS,leeCPS2}.
-As a consequence, human life has also become more and more dependent on the services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality.
+Embedded systems are ubiquitous in modern day information systems and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, in multipurpose mobile devices, which now help in performing many routine daily tasks, and in cyber-physical systems (CPS), which demand short development cycles and again high levels of reliability~\cite{leeCPS,leeCPS2}.
+As a consequence, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
 
@@
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
-In particular, these models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+Specifically, these models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
 
@@
 
-We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis in Section~\ref{Preliminaries}.
-In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
+We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
+In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research and Section \ref{Newapp} tackles new applications.
 Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems, while we conclude and describe future work in Section~\ref{conclusions}.
@@
 Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}.
-Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver.
+Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers.
 BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.

### 2017-02-05 paper.tex
Intro

@@
 Generally, embedded computer systems perform dedicated functions with a high degree of reliability.
+Embedded systems are ubiquitous in modern day information systems and are also becoming increasingly important in our society, and nowadays, they are employed to process, monitor, and control every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous in modern day information systems and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, in multipurpose mobile devices, which now help in performing many routine daily tasks, and in cyber-physical systems (CPS), which demand short development cycles and again high levels of reliability~\cite{leeCPS,leeCPS2}.
-As a consequence, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
+Naturally, the interaction between embedded softwares and physical processes created a different class of systems with a mixture of continuous and discrete dynamics, {\it i.e.} hybrid systems.
+The domain of embedded systems allied to the most recent communication revolution cause an intense integration between different and spread physical processes (that are usually embedded systems) and computation.
+Those complex, hybrid and integrated systems are called cyber-physical systems (CPS).
+
+The main difference between embedded systems and CPS is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last ones.
+This simple difference causes a revolution in terms of flexibility, scalability, and complexity of hybrid systems so-called CPS and adds a novel class of problems to the already known challenges of embedded systems.
+For example, micro-grids,{\it i.e.} small-scale electricity systems that groups different sources of distributed generation (DG) and loads, are emerging CPS's, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
+Note that, each DG and the majority of loads already employed embedded softwares and present several safety-critical requirements, however the integration of them constitute a CPS where arises additional challenges related to synchronism, stability, control, communication, and reliability  of entire micro-grid.
+
+Thus CPS's demand short development cycles and high levels of reliability and robustness~\cite{leeCPS,leeCPS2} and present challenges that include but is not limited to the the the embedded system challenges.
+As a consequence of the popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
 

### 2017-02-06 paper.tex
minor changes

@@
 They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Naturally, the interaction between embedded softwares and physical processes created a different class of systems with a mixture of continuous and discrete dynamics, {\it i.e.} hybrid systems.
+The interaction between embedded software and physical processes created a different class of systems with a mixture of continuous and discrete dynamics, {\it i.e.}, \textit{hybrid systems}.
 The domain of embedded systems allied to the most recent communication revolution cause an intense integration between different and spread physical processes (that are usually embedded systems) and computation.
@@
 
-The main difference between embedded systems and CPS is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last ones.
+The main difference between embedded systems and CPS is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last one.
 This simple difference causes a revolution in terms of flexibility, scalability, and complexity of hybrid systems so-called CPS and adds a novel class of problems to the already known challenges of embedded systems.
-For example, micro-grids,{\it i.e.} small-scale electricity systems that groups different sources of distributed generation (DG) and loads, are emerging CPS's, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
-Note that, each DG and the majority of loads already employed embedded softwares and present several safety-critical requirements, however the integration of them constitute a CPS where arises additional challenges related to synchronism, stability, control, communication, and reliability  of entire micro-grid.
+For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPS's, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
+Note that, each DG and the majority of loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a CPS, where additional challenges arises related to synchronism, stability, control, communication, and reliability of the entire micro-grid.
 
-Thus CPS's demand short development cycles and high levels of reliability and robustness~\cite{leeCPS,leeCPS2} and present challenges that include but is not limited to the the the embedded system challenges.
+Thus CPS's demand short development cycles and high levels of reliability and robustness~\cite{leeCPS,leeCPS2} and present challenges that include but is not limited to the embedded system challenges.
 As a consequence of the popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

### 2017-02-27 paper.tex
additions until section II

--- 2017-01-26 paper.tex
@@
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior and compliance with the target system, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 
-Generally, embedded computer systems perform dedicated functions with a high degree of reliability.
-They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
-Embedded systems are ubiquitous in modern day information systems and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, in multipurpose mobile devices, which now help in performing many routine daily tasks, and in cyber-physical systems (CPS), which demand short development cycles and again high levels of reliability~\cite{leeCPS,leeCPS2}.
-As a consequence, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
+Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real time).
+They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, and control every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.
+Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
+
+The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, {\it i.e.}, \textit{hybrid systems}.
+Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).
+
+The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last one.
+This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the already known challenges of embedded systems.
+For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
+One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a CPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.
+
+Thus, CPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems.
+As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
 
@@
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
-In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
+In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even compromise many aspects of their lives, with serious consequences.
+One the one hand, portable embedded systems are capable of monitoring and identifying conditions that are very difficult to a physician, mainly when their contact with patients happen only in medical clinics.
+On the other hand, wrong or diagnostic data may delay necessary treatments or compromise patients' health.
 %
@@
 
-Besides, when physical interaction with the real world is needed, which happens in CPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
+Besides, when physical interaction with the real world is needed, which happens in CPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
 Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation.
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
-Specifically, these models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+Specifically, such models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+
+For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if read and write operations are performed according to what is specified in the device's datasheet and setup periods are respected.
+In summary, the main goal would be ensuring that the developed driver is compliant with the underlying hardware, which goes further than code correctness.
 
@@
 Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
+Such observations expand the addressed issues and even complement what was mentioned in the last paragraph.
 
 We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
-In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research and Section \ref{Newapp} tackles new applications.
-Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems, while we conclude and describe future work in Section~\ref{conclusions}.
+In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
+Section \ref{Newapp} then tackles new applications, which take into account the new paradigm addressed here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
+Finally, we conclude and describe future work in Section~\ref{conclusions}.
 
@@
 In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}.
-Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 
@@
 
-Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification.
+Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
@@
 \sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers.
-The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can encoded using the SMT solvers bit-vectors theory.
+The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT solvers bit-vectors theory.
 The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally.
@@
 
-Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
+Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications, due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
 
@@
 Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.
-Here, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
-Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space.
-Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable.
+As a consequence, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
+Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation.
+Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space.
+finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
 
@@
 Section~\ref{achievements} outlines contributions toward these research problems.
+
+\section{New Applications: Beyond Code Correctness}
+\label{Newapp}
 
@@
 \newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
+
+\bibitem{ldd}
+Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
+\newblock {\em {Model checking concurrent linux device drivers}}.
+\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
 

### 2017-02-27 paper.tex
Sec 2

@@
 Finally, we conclude and describe future work in Section~\ref{conclusions}.
+
+%--------------------------------------
+\section{A model for Cyber-physical systems}
+\label{sec:model}
+%--------------------------------------
+
+Models are very important for sciences and engineering that employs it to analyze, predict and design the behavior of modeled systems.
+Specially, deterministic and stochastic model are the basis of the major part of engineering advances for the last centuries.
+However, these models are not enough to represent CPS because the convergence between software, communications and hardware present a lot of events that could not be deterministically or stochastically modeled, and are better represented by nondeterministic models~\cite{leeCPS15}.
+
+The early representation of hybrid systems, {\it i.e.} systems that mixes continuous and discrete dynamics, consists in representing the discrete components and discretized continuous dynamics using discrete models, considering the non-modeled effects as uncertainties or perturbations exogenous to the models.
+This approach is not suitable for complex CPS that presents a lot of nondeterminism.
+
+CPS have been modeled through different structures that present different degrees of nondeterminism and proportion between continuous and discrete dynamics, e.g., timed automata (\textcolor{red}{CITAR}) and hybrid automata (\textcolor{red}{CITAR}).
+Rungger and Tabuada (\textcolor{red}{CITAR TAC 2016}) proposed a general and wide model for CPS.
+Following this proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
 

### 2017-02-27 paper.tex
comments and addition to section IV

--- 2017-02-27 paper.tex
@@
 In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
-Additionally, embedded system verification raises additional challenges, such as
+\textcolor{red}{nao faltou completeness?
+talvez fosse bom colocar algumas palavras disso, o que fecharia com BMC e k-induction} Additionally, embedded system verification raises additional challenges, such as
 %
@@
 	\item platform restrictions.
+	\item verification of code pieces or modules that rely on larger structures.
 	\item legacy designs. %, which are usually developed with low-level languages.
-	\item support to different programming languages and interfaces.
+	\item support to different programming languages, frameworks, and interfaces.
+	\item Correct code instrumentation \textcolor{red}{vale a pena mencionar que precisamos instrumentar corretamente o codigo?}.
 	\item handling of non-linear and non-convex optimization problems.
@@
 Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
-The fourth aspect is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
+The fourth aspect deals with code that uses an existing infrastructure and must be compliant with that, such as Linux kernel modules.
+The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
 In particular, software developed for that type of embedded systems has been extensively tested and verified, and also optimized for efficiency over years of development.
 Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.
-The fifth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches if verifiers and synthesizers do not support different programming languages and interfaces.
-Finally, the last one is related to the widespread use of embedded systems in autonomous vehicles navigation systems~\cite{Adouane16}, which demand optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
+The sixth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
+Finally, the last one is related to the widespread use of embedded systems in autonomous vehicles navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
 
-Those five challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
+Those seven challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
 In particular, synthesizers for those systems need to consider the effects of the quantizers (A/D and D/A converters) when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
@@
 
+\textcolor{red}{serah demais mencionar um framework unificado, que abranja tanto propriedades de bauxo nivel e comportamento quanto descricoes de alto nivel?}
 Section~\ref{achievements} outlines contributions toward these research problems.

### 2017-02-28 paper.tex
new changes and addition to section 1 and abstract

@@
 \begin{abstract}
-The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, and consumer electronics.
-Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power demands.
+The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, and consumer electronics.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power and flexibility demands.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior and compliance with the target system, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 
-The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last one.
-This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the already known challenges of embedded systems.
+The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things - IoT) in the latter.
+This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found embedded systems domain.
 For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
@@
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
-In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even compromise many aspects of their lives, with serious consequences.
-One the one hand, portable embedded systems are capable of monitoring and identifying conditions that are very difficult to a physician, mainly when their contact with patients happen only in medical clinics.
-On the other hand, wrong or diagnostic data may delay necessary treatments or compromise patients' health.
+In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences.
+On the one hand, portable embedded systems are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics.
+On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health.
 %
@@
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
-Specifically, such models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
 
 For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if read and write operations are performed according to what is specified in the device's datasheet and setup periods are respected.
-In summary, the main goal would be ensuring that the developed driver is compliant with the underlying hardware, which goes further than code correctness.
+In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles the interaction with surrounding entities.
 
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.
-Such observations expand the addressed issues and even complement what was mentioned in the last paragraph.
+Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency) must be considered during verification~\cite{Kroening15}.
+Such observations expand the addressed issues and even complement what was mentioned in the last paragraphs.
 
-We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
-In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section \ref{Newapp} then tackles new applications, which take into account the new paradigm addressed here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
+The goals of the present article is to briefly present part of the established research and also clarify the current trends, given that we are on the verge of a paradigm shift, mainly driven by CPSs and IoT.
+We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
+In Section~\ref{Verification-Challenges}, the main challenges regarding the verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
+Section \ref{Newapp} then addresses new applications, which take into account the new paradigm mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
 Finally, we conclude and describe future work in Section~\ref{conclusions}.
@@
 
-In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
-\textcolor{red}{nao faltou completeness?
-talvez fosse bom colocar algumas palavras disso, o que fecharia com BMC e k-induction} Additionally, embedded system verification raises additional challenges, such as
+In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+Additionally, embedded system verification raises additional challenges, such as
 %
@@
 	\item support to different programming languages, frameworks, and interfaces.
-	\item Correct code instrumentation \textcolor{red}{vale a pena mencionar que precisamos instrumentar corretamente o codigo?}.
+	\item Correct code instrumentation.
 	\item handling of non-linear and non-convex optimization problems.
@@
 
-\textcolor{red}{serah demais mencionar um framework unificado, que abranja tanto propriedades de bauxo nivel e comportamento quanto descricoes de alto nivel?}
+\textbf{(RP7)} provide a unified framework, which is able to tackle high and low level properties, as well as system behavior, which can lead to another degree of completeness, in such a way that every implementation aspect of an embedded system is addressed and evaluated.
+
 Section~\ref{achievements} outlines contributions toward these research problems.

### 2017-02-28 paper.tex
removed mentions to a new paradigm (evaluate later)

@@
 
-The goals of the present article is to briefly present part of the established research and also clarify the current trends, given that we are on the verge of a paradigm shift, mainly driven by CPSs and IoT.
+The goals of the present article is to briefly present part of the established research and also clarify the current trends, which are mainly driven by CPSs and IoT.
 We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding the verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section \ref{Newapp} then addresses new applications, which take into account the new paradigm mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
+Section \ref{Newapp} then addresses new applications, which take into account the new trends mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
 Finally, we conclude and describe future work in Section~\ref{conclusions}.

### 2017-02-28 paper.tex
changes until the beginning of section 3

@@
 
-Models are very important for sciences and engineering that employs it to analyze, predict and design the behavior of modeled systems.
-Specially, deterministic and stochastic model are the basis of the major part of engineering advances for the last centuries.
-However, these models are not enough to represent CPS because the convergence between software, communications and hardware present a lot of events that could not be deterministically or stochastically modeled, and are better represented by nondeterministic models~\cite{leeCPS15}.
+Models are largely employed in the science and engineering areas, in order to analyze, predict and modify the behavior of real-world systems.
+Specially, deterministic and stochastic models are the basis of many engineering advances, in the last centuries.
+The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions.
+Nonetheless, one can argue that such models are not enough to represent CPSs, because the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
 
-The early representation of hybrid systems, {\it i.e.} systems that mixes continuous and discrete dynamics, consists in representing the discrete components and discretized continuous dynamics using discrete models, considering the non-modeled effects as uncertainties or perturbations exogenous to the models.
-This approach is not suitable for complex CPS that presents a lot of nondeterminism.
+On the one hand, the early representation of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in representing discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which may even be negligible.
+On the other hand, that approach is not suitable for complex CPSs that present a lot of nondeterminism, given that uncertainties play a central role.
 
-CPS have been modeled through different structures that present different degrees of nondeterminism and proportion between continuous and discrete dynamics, e.g., timed automata (\textcolor{red}{CITAR}) and hybrid automata (\textcolor{red}{CITAR}).
+CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata (\textcolor{red}{CITAR}) and hybrid automata (\textcolor{red}{CITAR}).
 Rungger and Tabuada (\textcolor{red}{CITAR TAC 2016}) proposed a general and wide model for CPS.
 Following this proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
+
+\textcolor{red}{seria legal falar que nondeterminism pode ser tratado com model checking}
 
@@
 In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}.
-Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents program correctness and/or a system's behavior, and an iteration bound \textit{k}, which limits loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, which is expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
 

### 2017-03-01 paper.tex
addition until section 4

@@
 The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which were computed from programs under verification, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to validate the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
 
-There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
+There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
 The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis.
+\textcolor{red}{eh bom citar essa literatura aqui}
 
-Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}.
-Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}.
-Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction and in a completely automatic way ({\it i.e.}, users do not need to provide loop invariants), were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}.
+Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during system design phases~\cite{EenS03}.
+Apart from that, \textit{k}-induction is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.
 
-Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
+It is worth noticing that \textit{k}-induction with invariants has the potential to be directly integrated into existing BMC approaches, given that the induction algorithm itself can be seen as an extension after \textit{k} unwindings and it is possible to generate program invariants with other software modules, which are then translated and instrumented into an input program \cite{Rocha15}.
+
+Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
 That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
-Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.
+Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
 
@@
 The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$.
-In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then produce incrementally a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
-As a result, a given candidate program $p$ is iteratively refined to match the specification $\sigma$ more closely.
-Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis currently used in practice~\cite{David15}.
+In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
+As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely.
+Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis that are currently used in practice~\cite{David15}.
 Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.
@@
 \forall \vec{x}.
-\sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers.
-The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT solvers bit-vectors theory.
+\sigma(\vec{x}, \vec{F})$, where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers.
+The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part.
 The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally.
-Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
-If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
-If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
+Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
+If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
+If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
 %
@@
 Note that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
-Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times.
-However, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
-In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase to learn from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
+Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
+In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
 
@@
 
-It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems.
+Currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if source code is correctly written, but also if it will properly respond in real environments or under external problems.
 For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
@@
 Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation.
-Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space.
-finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+
+In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space.
+Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
 
 In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}.
-In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets as well as FWL uncertainties in digital controllers.
-DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models taken from the control literature within minutes.
+In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets, as well as FWL uncertainties in digital controllers.
+DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models, taken from the control literature, within minutes.
 
-Apart from the control system domain, Scratch is another example of a software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}.
+Apart from the control system domain, Scratch is another example software model-checking tool, which uses knowledge about the system to be verified and the underlying hardware, with the goal of detecting races related to direct memory access (DMA), in the Cell BE processor~\cite{Donaldson10}.
 That tool also uses BMC, in order to detect DMA races, and BMC with \textit{k}-induction, which aims to prove the absence of races.
 If support to other DMA operations were added, Scratch could be adapted to different architectures, {\it i.e.}, the same techniques would be employed, but with a different system behavior/knowledge.
+
+One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also by applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks.
+For instance, regarding self-driving cars, a property could be the detection of pedestrians or animals, or even bicycles.
+Indeed, the latter is considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
+In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised \cite{BMCml}.
 
@@
 
-State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
+Generally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
 %
@@
 
-In particular, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
 Additionally, embedded system verification raises additional challenges, such as
@@
 	\item support to different programming languages, frameworks, and interfaces.
-	\item Correct code instrumentation.
+	\item correct code instrumentation.
 	\item handling of non-linear and non-convex optimization problems.
@@
 Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
-The fourth aspect deals with code that uses an existing infrastructure and must be compliant with that, such as Linux kernel modules.
+The fourth aspect deals with code that relies on existing infrastructures and must be compliant with those, such as Linux kernel modules.
 The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
@@
 The sixth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
-Finally, the last one is related to the widespread use of embedded systems in autonomous vehicles navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
+The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correct address a property.
+Finally, the last one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
 
-Those seven challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
-In particular, synthesizers for those systems need to consider the effects of the quantizers (A/D and D/A converters) when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
+Those eight challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior, whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
+In particular, synthesizers for those systems need to consider the effects of quantizers (A/D and D/A converters), when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
 intelligent product in Figure~\ref{intelligent-product}).
-Additionally,  finite-precision arithmetic and their related rounding errors need to be considered when correct-by-construction code is generated for embedded systems.
+Additionally,  finite-precision arithmetic and their related rounding errors need to be considered, when correct-by-construction code is generated for embedded systems.
 
@@
 
+\bibitem{selfcar}
+Fairley P.:
+\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
+\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
+
+\bibitem{BMCml}
+L.~Bortolussi, D. Milios, G, Sanguinetti.:
+\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
+\newblock In:. Information and Computation, 2016.
+
 %31

### 2017-03-01 paper.tex
addition until section 5

@@
 
-This position paper tackles six major problems in computer-aided verification and synthesis for embedded systems, which are (partially) open in current published research.
+This position paper tackles seven major problems in computer-aided verification and synthesis for embedded systems, which are (partially) open in current published research.
 
-\textbf{(RP1)} provide suitable encoding into SMT \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
 
-\textbf{(RP2)} exploit SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
+\textbf{(RP2)} exploitation of SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
 
-\textbf{(RP3)} prove correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
+\textbf{(RP3)} proof of correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
 
-\textbf{(RP4)} incorporate knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+\textbf{(RP4)} incorporation of knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
 
-\textbf{(RP5)} provide tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
+\textbf{(RP5)} provision of tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
 
-\textbf{(RP6)} develop automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
+\textbf{(RP6)} development of automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
 
-\textbf{(RP7)} provide a unified framework, which is able to tackle high and low level properties, as well as system behavior, which can lead to another degree of completeness, in such a way that every implementation aspect of an embedded system is addressed and evaluated.
+\textbf{(RP7)} provision of a unified framework, which is able to tackle high and low level properties, as well as system behavior, with the potential to lead to another degree of completeness, in such a way that every implementation aspect of an embedded system is addressed and evaluated.
 
 Section~\ref{achievements} outlines contributions toward these research problems.
-
-\section{New Applications: Beyond Code Correctness}
-\label{Newapp}
 
@@
 As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and increase further the reliability of the verified (embedded) system.
+
+\section{New Applications: Beyond Code Correctness}
+\label{Newapp}
 

### 2017-03-02 paper.tex
updated authors
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \author{Lucas C. Cordeiro \\
-Department of Computer Science\\
+%Department of Computer Science\\
 University of Oxford, UK\\
-Email: lucas.cordeiro@cs.ox.ac.uk
+lucas.cordeiro@cs.ox.ac.uk
 \and
 Eddie B. de Lima Filho \\
-Science, Technology, and Innovation Center\\ for the Industrial Pole of Manaus, Brazil \\
-Email: eddie@ctpim.org.br}
+Samsung Electronics, Brazil \\
+eddie.l@samsung.com
+\and
+Iury V. de Bessa \\
+Federal University of Amazonas, Brazil \\
+iury.bessa@ufam.edu.br}
 %\IEEEauthorblockN{Homer Simpson}

### 2017-03-02 paper.tex
reviewed introduction
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things - IoT) in the latter.
-This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found embedded systems domain.
+This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
 For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
@@
 
-For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if read and write operations are performed according to what is specified in the device's datasheet and setup periods are respected.
+For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected.
 In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles the interaction with surrounding entities.
@@
 Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency) must be considered during verification~\cite{Kroening15}.
-Such observations expand the addressed issues and even complement what was mentioned in the last paragraphs.
+Such observations expand the addressed issues and even complement what was mentioned in the previous paragraphs.
 

### 2017-03-02 paper.tex
added reference
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Models are largely employed in the science and engineering areas, in order to analyze, predict and modify the behavior of real-world systems.
+Models are largely employed in the science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems.
 Specially, deterministic and stochastic models are the basis of many engineering advances, in the last centuries.
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions.
-Nonetheless, one can argue that such models are not enough to represent CPSs, because the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent CPS since the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 On the one hand, the early representation of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in representing discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which may even be negligible.
-On the other hand, that approach is not suitable for complex CPSs that present a lot of nondeterminism, given that uncertainties play a central role.
+On the other hand, that approach is not suitable for complex CPS that present intense nondeterminism, given that uncertainties play a central role.
 
@@
 There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also discover the memory ``shape'', in programming languages with pointers~\cite{pips:2013,Henry:2012}.
-The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis.
-\textcolor{red}{eh bom citar essa literatura aqui}
+The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems, using invariant generation and strengthening, mostly based on interval analysis~\cite{Beyer15}.
 

### 2017-05-29 paper.tex
replaced figures
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
added comments
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
new changes to repo including machine learning

@@
 The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, and consumer electronics.
-Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power and flexibility demands.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 
+Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may be change based on data acquired during previous actions, due to model updating.
+As a consequence, system correctness verification procedures should also adapt somehow and incorporate new checks.
+
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
@@
 
-The goals of the present article is to briefly present part of the established research and also clarify the current trends, which are mainly driven by CPSs and IoT.
+The goals of the present article is to briefly present part of the established research and also clarify the current trends, which are mainly driven by CPSs, IoT, and ML.
 We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
@@
 	\centering
-	\includegraphics[scale=0.12]{figure4.eps}
+	\includegraphics[scale=0.12]{figure3.eps}
 	\caption{Counter-Example Guided Inductive Synthesis.}
@@
 	\centering
-	\includegraphics[scale=0.35]{figure3.eps}
+	\includegraphics[scale=0.35]{figure4.eps}
 	\caption{Verification methodologies for embedded systems.}
@@
 
+\textcolor{red}{We still have to mention RP7...}
+
 Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, SMACK~\cite{Haran15}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl.
@@
 \newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
+
+%1
+\bibitem{mlembe}
+S.~Virtanen:
+\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
+\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.
 

### 2017-05-29 paper.tex
minor change

@@
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, recognize the incorporation of knowledge about new problems and conditions, and supporting new software architectures and legacy designs (usually written in low-level languages).
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.

### 2017-06-18 paper.tex
changes and corrections until section 3

--- 2017-05-29 paper.tex
@@
 \begin{abstract}
-The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, and consumer electronics.
-Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
+The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things and consumer electronics.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
@@
 Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real time).
-They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, and control every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.
+They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.
 Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
 
-The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, {\it i.e.}, \textit{hybrid systems}.
+The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems}.
 Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).
 
-The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things - IoT) in the latter.
+The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, internet of things - IoT) in the latter.
 This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
@@
 
-Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}.
+Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}.
 Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
@@
 	\includegraphics[scale=0.35]{figure1.eps}
-	\caption{An embedded system is part of a well-specified larger system (intelligent product).}
+	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
 	\label{intelligent-product}
@@
 
-For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected.
+For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected, for instance.
 In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles the interaction with surrounding entities.
@@
 Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may be change based on data acquired during previous actions, due to model updating.
-As a consequence, system correctness verification procedures should also adapt somehow and incorporate new checks.
+As a consequence, system correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data.
 
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency) must be considered during verification~\cite{Kroening15}.
+Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency)~\cite{Kroening15} or even new structures provided with the goal of ensuring more computational capacity~\cite{cudalucas} must be considered during verification.
 Such observations expand the addressed issues and even complement what was mentioned in the previous paragraphs.
 
-The goals of the present article is to briefly present part of the established research and also clarify the current trends, which are mainly driven by CPSs, IoT, and ML.
-We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}.
+The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by CPSs, IoT, and ML.
+We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding the verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
@@
 Specially, deterministic and stochastic models are the basis of many engineering advances, in the last centuries.
-The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions.
-Nonetheless, one can argue that such models are not enough to represent CPS since the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
+The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions \cite{stochmodel}.
+Nonetheless, one can argue that such models are not enough to represent, CPSs since the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
 
-On the one hand, the early representation of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in representing discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which may even be negligible.
-On the other hand, that approach is not suitable for complex CPS that present intense nondeterminism, given that uncertainties play a central role.
+On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which may even be negligible.
+On the other hand, that approach is not suitable for complex CPSs that present intense nondeterminism, where uncertainties play a central role and may in validate the entire system behavior.
 
@@
 
-\textcolor{red}{seria legal falar que nondeterminism pode ser tratado com model checking}
+\textcolor{red}{seria legal falar que nao determinismo pode ser tratado com model checking e ligar com o nosso trabalho...
+inclusive , o tabuada tem um artigo falando sobre seguranca e SMT, que eu comentei abaixo}
+
+In addition, Shoukry {\it et.
+al} \cite{Shoukry} tackled another important problem related to CPSs: security.
+Indeed, the very essence of CPSs claim for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues \cite{cpsattack}, given that wrong control actions may be generated based on corrupted measurements.
+The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
+Indeed, many situations may be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT) \cite{Araujo16} \textcolor{red}{eh bom citar mais}, which indicates many links with other research areas and may lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 
@@
 Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers.
-BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
+BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
 
@@
 Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuit-implementations or programs with loops, whose bounds are too large or cannot be statically determined.
+Indeed, such an issue has pushed researchers to overcame the depth limitation and propose extensions capable of proving global correctness.
 
@@
 The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to validate the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to validate the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07} \textcolor{red}{confesso que, para mim, essa frase ficou confusa...
+eu nao consegui baixa esse livre}.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
@@
 Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
-That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes.
-Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
+That happens due to the distinctive characteristics mentioned earlier, which influence embedded system development and also employed verification processes.
+Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
 
@@
 The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$.
-In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$ as the starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
+In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$, as starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
 As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely.
@@
 The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part.
-The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally.
+The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is incrementally updated.
 Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
@@
 
-Note that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
+One may notice that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
 Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
@@
 Currently, SMT-based BMC approaches check code properties in real programs, which basically address programing-language issues and general correctness, without taking into account target applications or system behavior.
-Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if source code is correctly written, but also if it will properly respond in real environments or under external problems.
-For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.
+Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then requires schemes that do not only check if source code is correctly written, but also if it will properly respond in real environments or under external problems or corrupted data.
+For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged or even intentionally tampered \cite{Shoukry}.
 
@@
 
-In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space.
-Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.
+In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms) and other implementation restrictions, in order to explore the design space.
+Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which may even be incorporated into a given system development procedure, by providing both system verification and model refinement.
 
 In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}.
-In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets, as well as FWL uncertainties in digital controllers.
+In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with the DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets, as well as FWL uncertainties in digital controllers.
 DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models, taken from the control literature, within minutes.
+Nonetheless, system correction is of paramount importance and even with such a synthesizing tool, a final verification procedure should be executed, with the goal of ensuring that the generated solution still complies with the provided specs and other issues, which may have not been initial tackled and have the potential to compromise system reliability, are not present.
 
@@
 
-One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also by applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks.
-For instance, regarding self-driving cars, a property could be the detection of pedestrians or animals, or even bicycles.
-Indeed, the latter is considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
-In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised \cite{BMCml}.
+One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also be applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks.
+For instance, regarding self-driving cars, an important property could be the detection of pedestrians, animals, bicycles, or any obstacle present on urban and rural roads.
+Indeed, for instance, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
+In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches\cite{BMCml}.
 
@@
 
+%4
+\bibitem{cudalucas}
+Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
+\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
+\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.
+
+%1
+\bibitem{stochmodel}
+Bargmann H.:
+\newblock {\em {The role of stochastic modelling in engineering science}}.
+\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.
+
+\bibitem{Shoukry}
+Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
+\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
+\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.
+
+\bibitem{cpsattack}
+Liu Y., Ning P., Reiter M. K.:
+\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
+\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.
+
+\bibitem{Araujo16}
+Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
+\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
+\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
+
 %10
@@
 
+%25
+\bibitem{Beyer15}
+Beyer D., Dangl M., Wendler P.:
+\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
+\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
+
 %24
@@
 \newblock http://dx.doi.org/10.1007/s10009-015-0407-9
-
-%25
-\bibitem{Beyer15}
-Beyer D., Dangl M., Wendler P.:
-\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
-\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
 
@@
 
-\bibitem{selfcar}
-Fairley P.:
-\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
-\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
-
-\bibitem{BMCml}
-L.~Bortolussi, D. Milios, G, Sanguinetti.:
-\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
-\newblock In:. Information and Computation, 2016.
-
 %31
@@
 \newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
+
+%34
+\bibitem{David15}
+C.~David, D.~Kroening, and M.~Lewis.
+\newblock {\em {Using program synthesis for program analysis}}.
+\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
 
@@
 \newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
-
-%34
-\bibitem{David15}
-C.~David, D.~Kroening, and M.~Lewis.
-\newblock {\em {Using program synthesis for program analysis}}.
-\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
 
@@
 \newblock In: IEEE Transactions on Computers, 2017 (to appear).
+
+\bibitem{selfcar}
+Fairley P.:
+\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
+\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
+
+\bibitem{BMCml}
+L.~Bortolussi, D. Milios, G, Sanguinetti.:
+\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
+\newblock In:. Information and Computation, 2016.
 
@@
 %63
-\bibitem{Araujo16}
-Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
-\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
-\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
 

### 2017-06-18 paper.tex
changes to section 4

@@
 Indeed, for instance, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
-In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches\cite{BMCml}.
+In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches \cite{BMCml}.
 
@@
 
-Generally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
+Generally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use some assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
 %
@@
 In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
-Additionally, embedded system verification raises additional challenges, such as
+Additionally, embedded system verification raises additional challenges, such as:
 %
 \begin{enumerate}
-	\item time and energy constraints.
-	\item handling of concurrent software.
-	\item platform restrictions.
-	\item verification of code pieces or modules that rely on larger structures.
-	\item legacy designs. %, which are usually developed with low-level languages.
-	\item support to different programming languages, frameworks, and interfaces.
-	\item correct code instrumentation.
-	\item handling of non-linear and non-convex optimization problems.
+	\item Time and energy constraints;
+	\item Handling of concurrent software;
+	\item Platform restrictions;
+	\item Verification of code pieces or modules that rely on larger structures;
+	\item Legacy designs; %, which are usually developed with low-level languages.
+	\item Support to different programming languages, frameworks, and interfaces;
+	\item Correct code instrumentation;
+	\item Handling of non-linear and non-convex optimization problems;
+	\item Incorporation of new and adapted checks, due to system or environment evolution.
 \end{enumerate}
@@
 The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correct address a property.
-Finally, the last one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
+The eighth one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
+The ninth aspect is a consequence of changes in environment or different applications and scenarios, which may present new problems to embedded systems that did not exist or were not tackled during their development.
+Such a challenge is also closely related to ML techniques, given that they may be integrated into commercial systems and have the potential to devise new and refined control strategies, which must then be identified, checked, and validated.
+For instance, a diagnostic system may employ techniques that make it able to learn new tumor patterns and them warn their presence on patients' bodies, but a wrong updated model may compromise earlier traditional assessments and erroneously perform new ones, which should be verified and validated.
+Finally, the latter also raises another question: an embedded system must be checked only when it is developed/created or also during its life cycle, in order to ensure that improved models and actions are sound and reliable?
+That is an issue that will probably concern researchers, when ML becomes mature and widely spread.
 
-Those eight challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior, whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
+Those nine challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior (which may eventually change), whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
 In particular, synthesizers for those systems need to consider the effects of quantizers (A/D and D/A converters), when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.

### 2017-06-18 paper.tex
minor changes to section 5

@@
 
-\textbf{(RP6)} development of automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with the physical environment by considering uncertain models and FWL effects.
+\textbf{(RP6)} development of automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with physical environments, by considering uncertain models and FWL effects.
 

### 2017-06-18 paper.tex
changes to section 6

@@
 
-\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
+\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
 This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
 Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
-One possible research direction is to bridge the gap between BMC tools and SMT solvers to propose background theories and develop more efficient decision procedures, in order to handle specific classes of programs.
+One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific classes of programs.
 
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
-One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05} with the goal to remove redundant behaviour or to analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal to remove redundant behaviour or analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge remains still open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
 implementation aspects.
+Finally, optimal configurations for invariant generators are still of paramount importance, given that applications with code based on different aspects may benefit from different setups.
 
@@
 Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}.
-In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16,Chaves17}.
+In particular, a specific UAV application was tackled, with the goal of verifying its attitude controllers~\cite{Bessa16,Chaves17}.
+More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide a deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}.
 In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
@@
 In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
-Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
+\textcolor{red}{nao deveriamos falar que no pior caso poderiamos traduzir para c?
+aqui nao deveriamos mencionar o trabalho do januario?} Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
 
-{\textbf{(RP6)} State-of-the-art synthesis approaches for embedded (control) systems typically disregard the platform in which the embedded system software operates and restrict itself to generate code that do not take into account FWL effects.
-The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe) due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}.
+{\textbf{(RP6)} State-of-the-art synthesis approaches for embedded (control) systems typically disregard the platform in which the embedded system software operates and restrict themselves to generate code that do not take into account FWL effects.
+The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe), due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}.
 Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17}.
-However, there is still little evidence whether that approach can scale for larger systems modeled by other types of representations ({\it e.g.}, state-space).
+Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representations ({\it e.g.}, state-space).
 In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
@@
 
-\textcolor{red}{We still have to mention RP7...}
+\textcolor{red}{We still have to mention RP7...} \textcolor{green}{I leave that to you...}
 
@@
 
-%42
-\bibitem{Beyer14}
-D.~Beyer 
-\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
-
-%43
-\bibitem{BeyerSVCOMP15}
-D.~Beyer:
-\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
-
-%44
-\bibitem{Inverso14}
-Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
-\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
-\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
-
 %45
@@
 
+%42
+\bibitem{Beyer14}
+D.~Beyer 
+\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
+
+%43
+\bibitem{BeyerSVCOMP15}
+D.~Beyer:
+\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
+\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
+
+%59
+\bibitem{KahlonWG09}
+Kahlon V., Wang C., Gupta A.:
+\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
+\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
+
 %48
@@
 \newblock {University of Southampton, PhD Thesis}, 2015.
+
+%44
+\bibitem{Inverso14}
+Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
+\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
+\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
 
@@
 
-%57
-\bibitem{MorseCNF13}
-Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
-\newblock In: Tools and Algorithms for the Construction and Analysis of
-  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.
-
-%58
-\bibitem{MorseRCN014}
-Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {ESBMC} 1.22.}
-\newblock In: Tools and Algorithms for the Construction and Analysis of
-  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.
-
-%59
-\bibitem{KahlonWG09}
-Kahlon V., Wang C., Gupta A.:
-\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
-\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
+%60
+\bibitem{sbrtdaniel}
+de Mello D. P. M., de Freitas M.L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
+\newblock {\em{Verification of Magnitude and Phase Responses in Fixed-Point Digital Filters}}. 
+\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017.
 
@@
 %63
+
+%57
+\bibitem{MorseCNF13}
+Morse J., Cordeiro L., Nicole D., Fischer B.:
+\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
+\newblock In: Tools and Algorithms for the Construction and Analysis of
+  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.
+
+%58
+\bibitem{MorseRCN014}
+Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
+\newblock {\em {ESBMC} 1.22.}
+\newblock In: Tools and Algorithms for the Construction and Analysis of
+  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.
 

### 2017-06-18 paper.tex
additions to section 7

@@
 \label{Newapp}
+
+As already mentioned in section \ref{sec:model}, problems that are not directly related to code correctness may be tackled with formal methods, as long as the are conveniently modeled.
+Ara\'ujo {\it et.
+al} \cite{Araujo16} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
+Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
+Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory.
+
+In the same fashion, other problems may benefit from the recursive-refinement approach provided by SMT-based verification schemes.
+For instance, de Jesus {\it et.
+al} \cite{antennasbrt} developed a simplified methodology for antenna alignment that makes use of many modules already available in satellite TV receivers, along with a suitable structure for performing position correction.
+In summary, a reference signal is evaluated, until an (approximately) error-free reception is acquired, which can be measured through a simple transport stream (TS) error-indication interface and further refined with other parameters, such as carrier-to-noise ratio and signal level.
+This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignments procedures, until a counterexample is not output.
+It is worth noticing that the timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et.
+al} \cite{Araujo16}, which can be seen as an advantage for such an implementation.
 
@@
 \bibitem{sbrtdaniel}
-de Mello D. P. M., de Freitas M.L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
+de Mello D. P. M., de Freitas M. L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
 \newblock {\em{Verification of Magnitude and Phase Responses in Fixed-Point Digital Filters}}. 
@@
 
+\bibitem{antennasbrt}
+de Jesus A. S., rodrigues R. N., Ferreira A. N. G., Valente V. H., de Melo W. C., de Lima Filho E. B, da Silva J\'unior W. S.:
+\newblock {\em{Automatic Antenna Alignment System for Satellite Receivers Operating in C and Ku Bands}}. 
+\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017. (in Portuguese)
+
 \end{thebibliography}

### 2017-06-18 paper.tex
minor changes to section 7

@@
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
-Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory.
+Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory, but that can be improved, through the use of simplification techniques and and problem constraints.
 
@@
 
+Regarding behavior correctness, many other applications can be tackled, such as the ones related to automatic classification.
+Amoedo {\it et.
+al} proposed a modulation classification methodology, which the goal of spectrum sensing for opportunistic transmissions.
+Given that
 %--------------------------------------

### 2017-06-19 paper.tex
reviewed abstract
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \begin{abstract}
-The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things and consumer electronics.
+The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
@@
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs (usually written in low-level languages).
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.

### 2017-06-19 paper.tex
reviewed introduction
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real time).
+Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time).
 They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.

### 2017-06-19 paper.tex
reviewed introduction
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 In Section~\ref{Verification-Challenges}, the main challenges regarding the verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section \ref{Newapp} then addresses new applications, which take into account the new trends mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems.
+Section \ref{Newapp} then addresses new applications, which take into account the new trends mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems.
 Finally, we conclude and describe future work in Section~\ref{conclusions}.

### 2017-06-19 paper.tex
reviewed section 2
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Models are largely employed in the science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems.
-Specially, deterministic and stochastic models are the basis of many engineering advances, in the last centuries.
-The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions \cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent, CPSs since the interaction among software, communications, and hardware presents a lot of events that could not be deterministically or stochastically modeled and may be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems.
+Specially, deterministic and stochastic models are the basis of many recent engineering advances.
+The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
+Nonetheless, one can argue that such models are not enough to represent CPSs since the interaction among different entities (software and hardware components) presents several events that could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
 
-On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which may even be negligible.
-On the other hand, that approach is not suitable for complex CPSs that present intense nondeterminism, where uncertainties play a central role and may in validate the entire system behavior.
+On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible.
+On the other hand, that approach is not suitable for complex CPSs that present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
 
@@
 al} \cite{Shoukry} tackled another important problem related to CPSs: security.
-Indeed, the very essence of CPSs claim for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues \cite{cpsattack}, given that wrong control actions may be generated based on corrupted measurements.
+Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated based on corrupted measurements.
 The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
-Indeed, many situations may be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT) \cite{Araujo16} \textcolor{red}{eh bom citar mais}, which indicates many links with other research areas and may lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 

### 2017-06-19 paper.tex
explain inductive invariant
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \usepackage{color}
+\usepackage{listings}
+
 % correct bad hyphenation here
@@
 %\renewcommand{\baselinestretch}{0.93}
+
+\lstset{language=C,basicstyle=\small}
+\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
+\lstset{tabsize=2}
+\lstset{firstnumber=1}
+\lstset{frame=single}
+\lstset{
+  language={C},
+  morekeywords={assert,stdio,math,printf,assume,then,end-if,end,input,output}
+}
 
@@
 The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to validate the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07} \textcolor{red}{confesso que, para mim, essa frase ficou confusa...
-eu nao consegui baixa esse livre}.
+In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+For instance, consider the following fragment of C code as shown in Figure~\ref{figure:inductive-invariant}.
+If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
+This invariant, however, is not inductive given that $x>0$ before an iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the third iteration.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
+In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{n}-iterations.
+%
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\begin{lstlisting}
+#include <stdio.h>
+#include <assert.h>
+#include <math.h>
+int main(void) {
+ unsigned int i=1, n = -1;
+ float x=2;
+ while (i<=n) {
+   x = ((2*x) - 1);
+   if(isinf(x)) {
+     printf("\%d\n", i);
+     break;
+   }
+   i++;
+ }
+ return 0;
+}
+\end{lstlisting}
+\end{minipage}
+\caption{Motivating example for the \textit{k}-induction algorithm.}
+\label{figure:inductive-invariant}
+\end{figure}
 

### 2017-06-19 example2.c

@@ -1,0 +1,21 @@
+#include <stdio.h>
+#include <assert.h>
+#include <math.h>
+int main(void) {
+ unsigned int i=1, n = -1;
+ float x;
+ __VERIFIER_assume(x>1);
+ while (i<=n) {
+   x = ((2*x) - 1);
+   if(isinf(x)) {
+     printf("\%d\n", i);
+     break;
+   }
+   assert(x>0);
+   printf("i: \%d\n", i);   
+   i++;
+ }
+ assert(x>0);
+ return 0;
+}
+

### 2017-06-19 paper.tex
updated example of inductive invariant
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-This invariant, however, is not inductive given that $x>0$ before an iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the third iteration.
+This invariant, however, is not inductive given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
-In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{n}-iterations.
+In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{k}-iterations.
 %
@@
 \begin{lstlisting}
-#include <stdio.h>
 #include <assert.h>
-#include <math.h>
 int main(void) {
- unsigned int i=1, n = -1;
  float x=2;
- while (i<=n) {
+ while (1) {
    x = ((2*x) - 1);
-   if(isinf(x)) {
-     printf("\%d\n", i);
-     break;
-   }
-   i++;
  }

### 2017-06-19 example2.c

@@ -2,18 +2,17 @@
 #include <assert.h>
 #include <math.h>
 int main(void) {
- unsigned int i=1, n = -1;
- float x;
- __VERIFIER_assume(x>1);
- while (i<=n) {
+ unsigned int i=1;
+ float x=0.9;
+ while (1) {
    x = ((2*x) - 1);
    if(isinf(x)) {
      printf("\%d\n", i);
      break;
    }
+   printf("iii: \%d\n", i);
    assert(x>0);
-   printf("i: \%d\n", i);   
-   i++;
+   ++i;
  }
  assert(x>0);
  return 0;

### 2017-06-20 paper.tex
updated example of inductive invariant
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
minor changes in the k-induction example

@@
 In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
-For instance, consider the following fragment of C code as shown in Figure~\ref{figure:inductive-invariant}.
+For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
 If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-This invariant, however, is not inductive given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration.
+This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
-In this specific example, an inductive invariant is $x>1$ since we can check that if $x>1$ before the initial iteration, then also $x>1$ after the \textit{k}-iterations.
+In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
 %
@@
 \end{minipage}
-\caption{Motivating example for the \textit{k}-induction algorithm.}
+\caption{Motivating example for inductive invariant.}
 \label{figure:inductive-invariant}

### 2017-06-21 paper.tex
minor change in caption of figure 3
added new references about the IEEE standard
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
-If we consider the floating-point arithmetic semantics, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
+If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
 This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration.
@@
 \end{minipage}
-\caption{Motivating example for inductive invariant.}
+\caption{Motivating example for inductive invariants.}
 \label{figure:inductive-invariant}
@@
 
-%1
+%9
 \bibitem{mlembe}
@@
 
-%9
+%10
 \bibitem{Kroening15}
@@
 
-%4
+%11
 \bibitem{cudalucas}
@@
 
-%1
+%12
 \bibitem{stochmodel}
@@
 
+%13
 \bibitem{Shoukry}
@@
 
+%14
 \bibitem{cpsattack}
@@
 
+%15
 \bibitem{Araujo16}
@@
 
-%10
+%16
 \bibitem{Biere99}
@@
 
-%11
+%17
 \bibitem{handbook09}
@@
 
-%12
+%18
 \bibitem{BarrettSST09}
@@
 
-%13
+%19
 \bibitem{Armando06}
@@
 
-%14
+%20
 \bibitem{Clarke04}
@@
 
-%15
+%21
 \bibitem{MerzFS12}
@@
 
-%16
+%22
 \bibitem{CordeiroF11}
@@
 
-%17
+%23
 \bibitem{Ivancic05}
@@
 
-%18
+%24
 \bibitem{Cordeiro12}
@@
 
-%19
+%25
 \bibitem{EenS03}
@@
 
-%20
+%26
 \bibitem{Sheera00}
@@
 
-%21
+%27
 \bibitem{Bradley07}
@@
 
-%22
+%28
+\bibitem{4610935}
+{IEEE} standard for floating-point arithmetic.
+\newblock {\em IEEE Std 754-2008}, pages 1--70, Aug 2008.
+
+%29
+\bibitem{Goldberg91whatevery}
+David Goldberg.
+\newblock What every computer scientist should know about floating point
+  arithmetic.
+\newblock {\em ACM Computing Surveys}, 23(1):5--48, 1991.
+
+%30
 \bibitem{pips:2013}
@@
 
-%23
+%31
 \bibitem{Henry:2012}
@@
 
-%25
+%32
 \bibitem{Beyer15}
@@
 
-%24
+%33
 \bibitem{Gadelha15}
@@
 
-%26
+%34
 \bibitem{Brain15}
@@
 
-%27
+%35
 \bibitem{Rocha15}
@@
 
-%28
+%36
 \bibitem{Kinductor}
@@
 
-%29
+%37
 \bibitem{Rocha17}
@@
 
-%30
+%38
 \bibitem{Donaldson10}
@@
 
-%31
+%39
 \bibitem{GrosseLD09}
@@
 
-%34
+%40
 \bibitem{David15}
@@
 
-%32
+%41
 \bibitem{Alur13}
@@
 
-%33
+%42
 \bibitem{sketch}
@@
 
-%35
+%43
 \bibitem{Sharma14}
@@
 
-%36
+%44
 \bibitem{dsv_spin2015}
@@
 
-%37
+%45
 \bibitem{Monteiro16}
@@
 
-%38
+%46
 \bibitem{Bessa17}
@@
 
+%47
 \bibitem{selfcar}
@@
 
+%48
 \bibitem{BMCml}
@@
 
-%39
+%49
 \bibitem{Behrend15}
@@
 
-%40
+%50
 \bibitem{Lettnin09}
@@
 
-%41
+%51
 \bibitem{Adouane16}
@@
 
-%45
+%52
 \bibitem{ECBS13}
@@
 
-%46
+%53
 \bibitem{Pereira15}
@@
 
-%47
+%54
 \bibitem{Sousa15}
@@
 
-%42
+%55
 \bibitem{Beyer14}
@@
 
-%43
+%56
 \bibitem{BeyerSVCOMP15}
@@
 
-%59
+%57
 \bibitem{KahlonWG09}
@@
 
-%48
+%58
 \bibitem{morse15}
@@
 
-%44
+%59
 \bibitem{Inverso14}
@@
 
-%49
+%60
 \bibitem{civl15}
@@
 
-%50
+%61
 \bibitem{Grumberg05}
@@
 
-%51
+%62
 \bibitem{McMillan11}
@@
 
-%52
+%63
 \bibitem{esbmc_controller}
@@
 
-%53
+%64
 \bibitem{esbmc_filter}
@@
 
-%54
+%65
 \bibitem{JMorse15}
@@
 
-%55
+%66
 \bibitem{Bessa16}
@@
 
-%56
+%67
 \bibitem{Chaves17}
@@
 
-%60
+%68
 \bibitem{sbrtdaniel}
@@
 
-%60
+%69
 \bibitem{Alves15}
@@
 
-%61
+%70
 \bibitem{Garcia16}
@@
 
-%62
+%71
 \bibitem{Jackson16}
@@
 
-%63
-
-%57
+%72
 \bibitem{MorseCNF13}
@@
 
-%58
+%73
 \bibitem{MorseRCN014}
@@
 
-%64
+%74
 \bibitem{Haran15}
@@
 
-%65
+%75
 \bibitem{Beyer16}
@@
 
-%66
+%76
 \bibitem{BeyerW15}
@@
 
-%67
+%77
 \bibitem{RochaIFM12}
@@
 
+%78
 \bibitem{antennasbrt}

### 2017-06-21 paper.tex
added example about BMC
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-From the practical point of view, SAT-based or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits and single- and multi-threaded programs.
+Note that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of the expressions.
+A verification condition is a logical formula (constructed from the bounded program and desired correctness properties) whose validity implies that the program's behaviour agrees with its specification~\cite{Bradley07}.
+Correctness properties in programs can be specified by the user via \textit{assert} statements or automatically generated from a specification language~\cite{Thomas01}.
+If all of a bounded program's VCs are valid, then the program is in compliance with its specification up to the given bound.
+
+As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program in single static assignment (SSA) form~\cite{Appel97} is shown in Figure~\ref{figure:verification-condition}(b).
+Note that the SSA form is an intermediate representation used by compilers to facilitate optimizations and transformations of the program code.
+The common property in SSA form is that every variable has only one definition in the program text.
+This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript) such that every assignment has a unique left hand side as shown in Figure~\ref{figure:verification-condition}(b).
+%
+\begin{figure}[ht]
+\centering
+\begin{minipage}{0.45\textwidth}
+\begin{lstlisting}
+#include <assert.h>
+int x[N], a;
+...
+int main(void) {
+  a=N;
+  for(int i=0; i<N; i++)
+    if (x[i]>1)
+      a--;
+  assert(a<=N);
+  return 0;
+}
+\end{lstlisting}
+\end{minipage}
+\begin{center}
+(a)
+\end{center}
+\centering
+\begin{minipage}{0.45\textwidth}
+\begin{lstlisting}
+a1 = N
+a2 = (x[0] > 1) ? a1 - 1 : a1
+a3 = (x[1] > 1) ? a2 - 1 : a2
+a4 = (x[2] > 1) ? a3 - 1 : a3
+...
+an+1 = (x[N-1] > 1) ? an - 1 : an 
+\end{lstlisting}
+\end{minipage}
+\begin{center}
+(b)
+\end{center}
+\caption{(a) A simple C program with a \textit{for} loop. (b) The corresponding unwound C program of (a) converted into SSA form.} 
+\label{figure:verification-condition}
+\end{figure}
+
+Apart from that, this program has an exponential number of paths since each element of array $x$ can be either greater than one or less than or equal to one.
+Despite the large number of paths through the program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$ such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails.
+Note that BMC still encodes the states of the program with a size that grows linearly with N.
+More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic as follows~\cite{Bradley07}:
+%
+\begin{equation}              
+\label{verification-condition-example}
+\psi := \left [ \begin{array}{ll} 
+                a_{1} = N  \\
+                \wedge \, \, a_{2} = ite\left(x[0] > 1, a_{1} - 1, a_{1}\right) \\ 
+                \wedge \, \, a_{3} = ite\left(x[1] > 1, a_{2} - 1, a_{2}\right) \\
+                \wedge \, \, a_{4} = ite\left(x[2] > 1, a_{3} - 1, a_{3}\right) \\
+                \wedge \ldots \\
+                \wedge \, \, a_{N+1} = ite\left(x[N-1] > 1, a_{N} - 1 , a_{N}\right) \\
+                \wedge \neg \left(a_{N+1} \leq N\right) \\
+              \end{array} \right ]  \\
+              \\           
+\end{equation}
+
+The ternary operator $f \: \: ?
+\: \: t_1 : t_2$ shown in Figure~\ref{figure:verification-condition}(b) is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and depending on its value selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$).
+In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and we check whether the entire formula is satisfiable using an off-the-self SMT solver.
+Formula (\ref{verification-condition-example}) can simply be represented as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables.
+However, checking the validity of a first-order logic formula in a given background theory is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
+
+From the practical point of view, SAT- or SMT-based BMC procedures have been successfully applied to verify a large number of hardware and software systems, including digital circuits and single- and multi-threaded programs.
 Those BMC techniques were able to find subtle bugs in real digital and embedded software systems, as reported in the available literature~\cite{Clarke04,MerzFS12,CordeiroF11,Ivancic05,Cordeiro12}.
@@
 Due to that limitation, BMC tools are typically susceptible to exhaustion of time or memory limits, when checking complex circuit-implementations or programs with loops, whose bounds are too large or cannot be statically determined.
-Indeed, such an issue has pushed researchers to overcame the depth limitation and propose extensions capable of proving global correctness.
+Indeed, such an issue has pushed researchers to overcome the depth limitation and propose extensions capable of proving global correctness.
 
@@
 In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+%
 For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
@@
 %
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 
+\bibitem{Appel97}
+Andrew~W. Appel.
+\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
+\newblock Cambridge University Press, New York, NY, USA, 1997.
+
+\bibitem{Thomas01}
+Thomas Ball and Sriram~K. Rajamani.
+\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
+\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
+
+\bibitem{Strichman08}
+Daniel Kroening and Ofer Strichman.
+\newblock {\em Decision Procedures: An Algorithmic Point of View}.
+\newblock Springer Publishing Company, Incorporated, 2008.
+
+\bibitem{Tseitin83}
+G.~S. Tseitin.
+\newblock On the complexity of derivation in propositional calculus.
+\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
+  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
+
+ \bibitem{PatarinG97}
+Jacques Patarin and Louis Goubin.
+\newblock Trapdoor one-way permutations and multivariate polynominals.
+\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997.
+
 \end{thebibliography}

### 2017-06-21 paper.tex
fixed references
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
+%----------------------------------------------------------------------------
 \section{New Applications: Beyond Code Correctness}
 \label{Newapp}
-
-As already mentioned in section \ref{sec:model}, problems that are not directly related to code correctness may be tackled with formal methods, as long as the are conveniently modeled.
-Ara\'ujo {\it et.
-al} \cite{Araujo16} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
+%----------------------------------------------------------------------------
+
+As already mentioned in section \ref{sec:model}, problems that are not directly related to code correctness may be tackled with formal methods, as long as they are conveniently modeled.
+Ara\'ujo {\it et al.} \cite{Araujo16} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
-Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory, but that can be improved, through the use of simplification techniques and and problem constraints.
+Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory, but that can be improved, through the use of simplification techniques and problem constraints.
 
@@
 This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignments procedures, until a counterexample is not output.
-It is worth noticing that the timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et.
-al} \cite{Araujo16}, which can be seen as an advantage for such an implementation.
+It is worth noticing that the timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{Araujo16}, which can be seen as an advantage for such an implementation.
 
 Regarding behavior correctness, many other applications can be tackled, such as the ones related to automatic classification.
-Amoedo {\it et.
-al} proposed a modulation classification methodology, which the goal of spectrum sensing for opportunistic transmissions.
+Amoedo {\it et al.} proposed a modulation classification methodology, which the goal of spectrum sensing for opportunistic transmissions.
 Given that
+
+\textcolor{red}{Eddie: Acho que esta faltando texto aqui, certo?}
 %--------------------------------------
@@
 
+%9
 \bibitem{ldd}
@@
 
-%9
+%10
 \bibitem{mlembe}
@@
 
-%10
+%11
 \bibitem{Kroening15}
@@
 
-%11
+%12
 \bibitem{cudalucas}
@@
 
-%12
+%13
 \bibitem{stochmodel}
@@
 
-%13
+%14
 \bibitem{Shoukry}
@@
 
-%14
+%15
 \bibitem{cpsattack}
@@
 
-%15
+%16
 \bibitem{Araujo16}
@@
 
-%16
+%17
 \bibitem{Biere99}
@@
 
-%17
+%18
 \bibitem{handbook09}
@@
 
-%18
+%19
 \bibitem{BarrettSST09}
@@
 
-%19
+%20
 \bibitem{Armando06}
@@
 
-%20
+%21
+\bibitem{Bradley07}
+Bradley A., Manna Z.:
+\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
+\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
+
+%22
+\bibitem{Thomas01}
+Thomas Ball and Sriram~K. Rajamani.
+\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
+\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
+
+%23
+\bibitem{Strichman08}
+Daniel Kroening and Ofer Strichman.
+\newblock {\em Decision Procedures: An Algorithmic Point of View}.
+\newblock Springer Publishing Company, Incorporated, 2008.
+
+%24
+\bibitem{Appel97}
+Andrew~W. Appel.
+\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
+\newblock Cambridge University Press, New York, NY, USA, 1997.
+
+%25
+\bibitem{Tseitin83}
+G.~S. Tseitin.
+\newblock On the complexity of derivation in propositional calculus.
+\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
+  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
+
+ %26
+ \bibitem{PatarinG97}
+Jacques Patarin and Louis Goubin.
+\newblock Trapdoor one-way permutations and multivariate polynominals.
+\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997.
+
+%27
 \bibitem{Clarke04}
@@
 
-%21
+%28
 \bibitem{MerzFS12}
@@
 
-%22
+%29
 \bibitem{CordeiroF11}
@@
 
-%23
+%30
 \bibitem{Ivancic05}
@@
 
-%24
+%31
 \bibitem{Cordeiro12}
@@
 
-%25
+%32
 \bibitem{EenS03}
@@
 
-%26
+%33
 \bibitem{Sheera00}
@@
 
-%27
-\bibitem{Bradley07}
-Bradley A., Manna Z.:
-\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
-\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
-
-%28
+%34
 \bibitem{4610935}
@@
 
-%29
+%35
 \bibitem{Goldberg91whatevery}
@@
 
-%30
+%36
 \bibitem{pips:2013}
@@
 
-%31
+%37
 \bibitem{Henry:2012}
@@
 
-%32
+%38
 \bibitem{Beyer15}
@@
 
-%33
+%39
 \bibitem{Gadelha15}
@@
 
-%34
+%40
 \bibitem{Brain15}
@@
 
-%35
+%41
 \bibitem{Rocha15}
@@
 
-%36
+%42
 \bibitem{Kinductor}
@@
 
-%37
+%43
 \bibitem{Rocha17}
@@
 
-%38
+%44
 \bibitem{Donaldson10}
@@
 
-%39
+%45
 \bibitem{GrosseLD09}
@@
 
-%40
+%46
 \bibitem{David15}
@@
 
-%41
+%47
 \bibitem{Alur13}
@@
 
-%42
+%48
 \bibitem{sketch}
@@
 
-%43
+%49
 \bibitem{Sharma14}
@@
 
-%44
+%50
 \bibitem{dsv_spin2015}
@@
 
-%45
+%51
 \bibitem{Monteiro16}
@@
 
-%46
+%52
 \bibitem{Bessa17}
@@
 
-%47
+%53
 \bibitem{selfcar}
@@
 
-%48
+%54
 \bibitem{BMCml}
@@
 
-%49
+%55
 \bibitem{Behrend15}
@@
 
-%50
+%56
 \bibitem{Lettnin09}
@@
 
-%51
+%57
 \bibitem{Adouane16}
@@
 
-%52
+%58
 \bibitem{ECBS13}
@@
 
-%53
+%59
 \bibitem{Pereira15}
@@
 
-%54
+%60
 \bibitem{Sousa15}
@@
 
-%55
+%61
 \bibitem{Beyer14}
@@
 
-%56
+%62
 \bibitem{BeyerSVCOMP15}
@@
 
-%57
+%63
 \bibitem{KahlonWG09}
@@
 
-%58
+%64
 \bibitem{morse15}
@@
 
-%59
+%65
 \bibitem{Inverso14}
@@
 
-%60
+%66
 \bibitem{civl15}
@@
 
-%61
+%67
 \bibitem{Grumberg05}
@@
 
-%62
+%68
 \bibitem{McMillan11}
@@
 
-%63
+%69
 \bibitem{esbmc_controller}
@@
 
-%64
+%70
 \bibitem{esbmc_filter}
@@
 
-%65
+%71
 \bibitem{JMorse15}
@@
 
-%66
+%72
 \bibitem{Bessa16}
@@
 
-%67
+%73
 \bibitem{Chaves17}
@@
 
-%68
+%74
 \bibitem{sbrtdaniel}
@@
 
-%69
+%75
 \bibitem{Alves15}
@@
 
-%70
+%76
 \bibitem{Garcia16}
@@
 
-%71
+%77
 \bibitem{Jackson16}
@@
 
-%72
+%78
 \bibitem{MorseCNF13}
@@
 
-%73
+%79
 \bibitem{MorseRCN014}
@@
 
-%74
+%80
 \bibitem{Haran15}
@@
 
-%75
+%81
 \bibitem{Beyer16}
@@
 
-%76
+%82
 \bibitem{BeyerW15}
@@
 
-%77
+%83
 \bibitem{RochaIFM12}
@@
 
-%78
+%84
 \bibitem{antennasbrt}
@@
 
-\bibitem{Appel97}
-Andrew~W. Appel.
-\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
-\newblock Cambridge University Press, New York, NY, USA, 1997.
-
-\bibitem{Thomas01}
-Thomas Ball and Sriram~K. Rajamani.
-\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
-\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
-
-\bibitem{Strichman08}
-Daniel Kroening and Ofer Strichman.
-\newblock {\em Decision Procedures: An Algorithmic Point of View}.
-\newblock Springer Publishing Company, Incorporated, 2008.
-
-\bibitem{Tseitin83}
-G.~S. Tseitin.
-\newblock On the complexity of derivation in propositional calculus.
-\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
-  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
-
- \bibitem{PatarinG97}
-Jacques Patarin and Louis Goubin.
-\newblock Trapdoor one-way permutations and multivariate polynominals.
-\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997.
-
 \end{thebibliography}

### 2017-06-21 paper.tex
changes to section 7

@@
 Regarding behavior correctness, many other applications can be tackled, such as the ones related to automatic classification.
-Amoedo {\it et al.} proposed a modulation classification methodology, which the goal of spectrum sensing for opportunistic transmissions.
-Given that
-
-\textcolor{red}{Eddie: Acho que esta faltando texto aqui, certo?}
+Amoedo {\it et al.} \cite{amoedo} proposed a modulation classification methodology, with the goal of spectrum sensing for opportunistic transmissions.
+Besides a correct software implementation, {\it i.e.}, source code, which is already covered by formal methods already available in the literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
+This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals under a variety of condition, unifying source code and behavior verification.
+In addition, that kind of implementation usually employs digital filters with specific parameters, which could also be tackled.
+
+Finally, ML techniques are already being used in the software verification area \cite{mlswverif,nguyen}.
+For instance, Hamel \cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implement the respective test set.
+ML is already being employ in space state reduction, ML technique can be used also for invariant generation
+
 %--------------------------------------
@@
 
+\bibitem{amoedo}
+Amoedo D. A., da Silva J\'unior W. S., de Lima Filho E. A.:
+\newblock {\em{Parameter Selection for SVM in Automatic Modulation Classification of Analog and Digital Signals}}. 
+\newblock In: International Telecommunications Symposium (ITS), pp. 1-5, 2014.
+
+\bibitem{mlswverif}
+L Hamel:
+\newblock {\em {On the Use of Machine Learning in Formal Software Verification.}} 
+\newblock {Technical Report TR03-294, Dept. of Computer Science and Statistics, University of Rhode Island}, 2003.
+
+\bibitem{nguyen}
+L Hamel:
+\newblock {\em {The Application of Machine Learning Methods in Software Verification and Validation.}} 
+\newblock {University of Texas at Austin, master of science dissertation}, 2010.
+
 \end{thebibliography}

### 2017-06-21 paper.tex
minor changes to new text

@@
 
-As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program in single static assignment (SSA) form~\cite{Appel97} is shown in Figure~\ref{figure:verification-condition}(b).
-Note that the SSA form is an intermediate representation used by compilers to facilitate optimizations and transformations of the program code.
+As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b).
+One may notice that the SSA form is an intermediate representation, used by compilers, to facilitate optimizations and transformations of source code.
 The common property in SSA form is that every variable has only one definition in the program text.
-This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript) such that every assignment has a unique left hand side as shown in Figure~\ref{figure:verification-condition}(b).
+This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left hand side, as shown in Figure~\ref{figure:verification-condition}(b).
 %
@@
 \end{center}
-\caption{(a) A simple C program with a \textit{for} loop. (b) The corresponding unwound C program of (a) converted into SSA form.} 
+\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
 \label{figure:verification-condition}
@@
 
-Apart from that, this program has an exponential number of paths since each element of array $x$ can be either greater than one or less than or equal to one.
-Despite the large number of paths through the program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$ such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails.
-Note that BMC still encodes the states of the program with a size that grows linearly with N.
-More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic as follows~\cite{Bradley07}:
+Apart from that, this program has an exponential number of paths, since each element of $x$ can be either greater than one or less than or equal to one.
+Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails.
+One may also notice that BMC still encodes program stages with a size that grows linearly with $N$.
+More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
 %
@@
 The ternary operator $f \: \: ?
-\: \: t_1 : t_2$ shown in Figure~\ref{figure:verification-condition}(b) is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and depending on its value selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$).
-In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and we check whether the entire formula is satisfiable using an off-the-self SMT solver.
-Formula (\ref{verification-condition-example}) can simply be represented as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables.
-However, checking the validity of a first-order logic formula in a given background theory is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
+\: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$).
+In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
+Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
 

### 2017-06-22 paper.tex
improved RP5 description
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
-\textcolor{red}{nao deveriamos falar que no pior caso poderiamos traduzir para c?
-aqui nao deveriamos mencionar o trabalho do januario?} Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
+This approach could adopt the clang compiler~\cite{Lopes:2014:GSL:2692607}, which is already widely used in industry~\cite{clang-usage}, in order to produce an Abstract Syntax Tree (AST) for a given programming language ({\it e.g.}, C/C++/ObjectiveC/ObjectiveC++).
+This AST can then be converted either to an intermediate representation (IR) of the underlying verifier or to LLVM bitcode that is already produced by clang.
+Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
 
@@
 %77
+\bibitem{Lopes:2014:GSL:2692607}
+Bruno~Cardoso Lopes and Rafael Auler.
+\newblock {\em Getting Started with LLVM Core Libraries}.
+\newblock Packt Publishing, 2014.
+
+%78
+ \bibitem{clang-usage}
+Cade Metz.
+\newblock Why {A}pple?s swift language will instantly remake computer
+  programming, 204.
+\newblock http://www.wired.com/2014/07/apple-swift/.
+
+%79
 \bibitem{Jackson16}
@@
 
-%78
+%80
 \bibitem{MorseCNF13}
@@
 
-%79
+%81
 \bibitem{MorseRCN014}
@@
 
-%80
+%82
 \bibitem{Haran15}
@@
 
-%81
+%83
 \bibitem{Beyer16}
@@
 
-%82
+%84
 \bibitem{BeyerW15}
@@
 
-%83
+%85
 \bibitem{RochaIFM12}
@@
 
-%84
+%86
 \bibitem{antennasbrt}
@@
 
+%87
 \bibitem{amoedo}
@@
 
+%88
 \bibitem{mlswverif}
@@
 
+%89
 \bibitem{nguyen}

### 2017-06-22 paper.tex
wrote RP7
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-\textcolor{red}{We still have to mention RP7...} \textcolor{green}{I leave that to you...}
+\textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java).
+Given the availability of different programming languages for writing code for embedded systems, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications.
+The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated.
+In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need to maintain a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification.
+Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics.
+The clang compiler offers an industrial quality analyzer and an application programming interface to access and traverse its internal AST, which can then be used by software verifiers to generate their IR or alternatively analyze the LLVM bitcode produced by clang.
 

### 2017-06-22 paper.tex
minor corrections
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Finally, ML techniques are already being used in the software verification area \cite{mlswverif,nguyen}.
-For instance, Hamel \cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implement the respective test set.
-ML is already being employ in space state reduction, ML technique can be used also for invariant generation
+For instance, Hamel \cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implements the respective test set.
+ML is already being employed in space state reduction, ML technique can be used also for invariant generation.
 

### 2017-06-22 paper.tex
added paper about parameter tuning
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

--- 2017-06-22 paper.tex
@@
 Finally, ML techniques are already being used in the software verification area \cite{mlswverif,nguyen}.
-For instance, Hamel \cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implement the respective test set.
-ML is already being employ in space state reduction, ML technique can be used also for invariant generation
+For instance, Hamel~\cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implements the respective test set.
+ML is already being employed in space state reduction and to boost software verification by automatic tuning decision procedures~\cite{babic07}.
+ML technique can be used also for invariant generation.
 
@@
 
+ %90
+\bibitem{babic07}
+Hutter F., Babic D., Hoos H. H., Hu A. J.              
+\newblock {\em{Boosting Verification by Automatic Tuning of Decision Procedures.}}
+\newblock {In 7th International Conference Formal Methods in Computer-Aided Design {(FMCAD)}, pp. 27--34, 2007.}
+
 \end{thebibliography}

### 2017-06-24 paper.tex
changes to new applications section

@@
 
-Finally, ML techniques are already being used in the software verification area \cite{mlswverif,nguyen}.
-For instance, Hamel~\cite{mlswverif} proposes to induce a theory, through ML, that fits a test set for a program, in such a way that it implements the respective test set.
-ML is already being employed in space state reduction and to boost software verification by automatic tuning decision procedures~\cite{babic07}.
-ML technique can be used also for invariant generation.
+ML techniques are already being used in the software verification area \cite{mlswverif,nguyen,bridge,babic07}.
+For instance, Hamel~\cite{mlswverif} proposed to induce a theory, through ML, that fits a test set for a program, in such a way that it implements the respective test set.
+Phuc \cite{nguyen}, in turn, applied decision tree C4.5 and multilayer perceptron to annotated programs (metrics related to structure and semantics) and tackled the generation of test sets satisfying the desired constraints.
+Finally, Bridge studied the automation of heuristic selection based on features, regarding first-order logic theorem proving, and Hutter {\it et.
+al} \cite{babic07} showed that ML can be used for improving SAT solvers regarding huge verification tasks, when applied to parameter tuning~\cite{babic07}.
+Given what was presented, it seems worthwhile to pursue the integration of ML to BMC-based checkers, specifically ESBMC, mainly regarding solver tuning.
+In addition, ML can also be used for invariant computation and strengthening, which is a major bottleneck in \textit{k}-induction verification based on invariant algorithms.
 
@@
 Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
-Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
-Finally, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.
+There is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
+In addition, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.
+Finally, there are some bottlenecks in software verification that are possible to be handled with machine learning, given some initial studies carried out in the current literature, which may represent another step towards a unified and complete approach that combines code verification, behavior correctness and environment adaptation.
 
@@
 Special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification, probably through behavior models.
+Additionally, ML techniques will be incorporated to BMC frameworks, in order to provide flexibility and specific-problem tuning.
 
@@
 \bibitem{mlswverif}
-L Hamel:
+Hamel L.:
 \newblock {\em {On the Use of Machine Learning in Formal Software Verification.}} 
@@
 \bibitem{nguyen}
-L Hamel:
+Phuc N. V.,:
 \newblock {\em {The Application of Machine Learning Methods in Software Verification and Validation.}} 
@@
 %90
+\bibitem{bridge}
+Bridge J. P.:
+\newblock {\em {Machine learning and automated theorem proving.}} 
+\newblock {Technical Report 792, Computer Laboratory, University of Cambridge}, 2010.
+
+ %91
 \bibitem{babic07}

### 2017-07-23 paper.tex
general mods

@@
 The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
-Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
+Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
+Besides, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of embedded software: cyber-physical systems.
 As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
 Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with the target system, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
@@
 Besides, when physical interaction with the real world is needed, which happens in CPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
-Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation.
+Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation.
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
@@
 For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected, for instance.
-In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles the interaction with surrounding entities.
-
-Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may be change based on data acquired during previous actions, due to model updating.
-As a consequence, system correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data.
+In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles interaction with surrounding entities.
+
+Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may change based on data acquired during previous actions, due to model updating.
+As a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data.
 
 A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
-Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency)~\cite{Kroening15} or even new structures provided with the goal of ensuring more computational capacity~\cite{cudalucas} must be considered during verification.
+Additionally, increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency)~\cite{Kroening15} or even new structures provided with the goal of ensuring more computational capacity~\cite{cudalucas} must be considered during verification.
 Such observations expand the addressed issues and even complement what was mentioned in the previous paragraphs.
@@
 We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
-In Section~\ref{Verification-Challenges}, the main challenges regarding the verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section \ref{Newapp} then addresses new applications, which take into account the new trends mentioned here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems.
+In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
+Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems and Section \ref{Newapp} addresses new applications, which take into account the new trends mentioned here.
 Finally, we conclude and describe future work in Section~\ref{conclusions}.
@@
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent CPSs since the interaction among different entities (software and hardware components) presents several events that could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent CPSs, since interaction among different entities (software and hardware components) presents several events that could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 Rungger and Tabuada (\textcolor{red}{CITAR TAC 2016}) proposed a general and wide model for CPS.
-Following this proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
+Following their proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
 
@@
 al} \cite{Shoukry} tackled another important problem related to CPSs: security.
-Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated based on corrupted measurements.
+Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements.
 The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
-Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 
@@
 
-Note that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of the expressions.
+Note that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of expressions.
 A verification condition is a logical formula (constructed from the bounded program and desired correctness properties) whose validity implies that the program's behaviour agrees with its specification~\cite{Bradley07}.
 Correctness properties in programs can be specified by the user via \textit{assert} statements or automatically generated from a specification language~\cite{Thomas01}.
-If all of a bounded program's VCs are valid, then the program is in compliance with its specification up to the given bound.
+If all of a bounded program's VCs are valid, then a program is in compliance with its specification up to the given bound.
 
 As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b).
-One may notice that the SSA form is an intermediate representation, used by compilers, to facilitate optimizations and transformations of source code.
+One may notice that the SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code.
 The common property in SSA form is that every variable has only one definition in the program text.
@@
 Apart from that, this program has an exponential number of paths, since each element of $x$ can be either greater than one or less than or equal to one.
-Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails.
+Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $a<=N$ fails.
 One may also notice that BMC still encodes program stages with a size that grows linearly with $N$.
@@
 \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$).
-In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
-Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
+In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
+Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
 
@@
 If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we have initially $x=0.9$, then $x<0$ after the fourth iteration.
+This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
 As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
@@
 Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
-That happens due to the distinctive characteristics mentioned earlier, which influence embedded system development and also employed verification processes.
+That happens due to the distinctive characteristics mentioned earlier, which influence embedded system development and also verification processes.
 Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
@@
 The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part.
-The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is incrementally updated.
+The phases {\sc Synthesize} and {\sc Verify} interact via a finite set of test vectors {\sc inputs} that is incrementally updated.
 Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
-If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
-If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
+If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the phase {\sc verify}, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
+If this is the case, then the algorithm terminates; otherwise, additional information is provided to the phase {\sc synthesize}, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
 %
@@
 Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
-In order to avoid exploring all possible values, machine learning techniques can be used in the {\sc Synthesize} phase, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
-
-Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications, due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
+In order to avoid exploring all possible values, machine learning techniques can be used in the phase {\sc Synthesize}, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
+
+Nowadays, program synthesis engines that implement the CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications, due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.
 
@@
 That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable.
-As a consequence, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
+As a consequence, verification procedures have to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
 Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation.
 
-In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms) and other implementation restrictions, in order to explore the design space.
+In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms), and other implementation restrictions, in order to explore the design space.
 Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which may even be incorporated into a given system development procedure, by providing both system verification and model refinement.
@@
 DSSynth is able to successfully synthesize stable digital controllers for a set of intricate plant models, taken from the control literature, within minutes.
-Nonetheless, system correction is of paramount importance and even with such a synthesizing tool, a final verification procedure should be executed, with the goal of ensuring that the generated solution still complies with the provided specs and other issues, which may have not been initial tackled and have the potential to compromise system reliability, are not present.
+Nonetheless, system correction is of paramount importance and, even with such a synthesizing tool, a final verification procedure should be executed, with the goal of ensuring that the generated solution still complies with the provided specs and other issues, which may have not been initial tackled and have the potential to compromise system reliability, are not present.
 
@@
 
-In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
+In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
 Additionally, embedded system verification raises additional challenges, such as:
@@
 Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.
-The sixth aspect is related to the evolution of development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
+The sixth aspect is related to evolving development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
 The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correct address a property.
@@
 Finally, the latter also raises another question: an embedded system must be checked only when it is developed/created or also during its life cycle, in order to ensure that improved models and actions are sound and reliable?
-That is an issue that will probably concern researchers, when ML becomes mature and widely spread.
-
-Those nine challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where the controlled object ({\it e.g.}, physical plant) typically exhibits continuous behavior (which may eventually change), whereas the controller (usually implemented by a real-time computer system) operates in discrete time and over a quantized domain.
+That is an issue that will probably concern researchers, when ML becomes mature and widely spread in software verification.
+
+Those nine challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where controlled objects ({\it e.g.}, physical plants) typically exhibit continuous behavior (which may eventually change), whereas controllers (usually implemented by real-time computer systems) operate in discrete time and over a quantized domain.
 In particular, synthesizers for those systems need to consider the effects of quantizers (A/D and D/A converters), when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
@@
 
-\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend the background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
 
@@
 
-\textbf{(RP3)} proof of correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware.
-
-\textbf{(RP4)} incorporation of knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+\textbf{(RP3)} proof of correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware platforms.
+
+\textbf{(RP4)} incorporation knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
 
@@
 
-Section~\ref{achievements} outlines contributions toward these research problems.
+Section~\ref{achievements} outlines contributions toward such research problems.
 
@@
 This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
-Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
-One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific classes of programs.
-
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}.
+Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific program classes, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}.
+One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes.
+
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
@@
 In particular, a specific UAV application was tackled, with the goal of verifying its attitude controllers~\cite{Bessa16,Chaves17}.
-More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide a deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}.
+More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}.
 In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
@@
 Some work towards that was already presented by Sousa, Cordeiro, and Filho~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics.
-In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked.
+In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked.
 This approach could adopt the clang compiler~\cite{Lopes:2014:GSL:2692607}, which is already widely used in industry~\cite{clang-usage}, in order to produce an Abstract Syntax Tree (AST) for a given programming language ({\it e.g.}, C/C++/ObjectiveC/ObjectiveC++).
@@
 Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17}.
-Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representations ({\it e.g.}, state-space).
+Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, state-space).
 In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
@@
 The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated.
-In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need to maintain a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification.
+In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification.
 Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics.
@@
 Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15,Beyer16}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.
-The increasing number of verification tasks being added every year to the competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques.
-As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and increase further the reliability of the verified (embedded) system.
+The increasing number of verification tasks being added every year to this competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques.
+As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and further increase the reliability of verified (embedded) systems.
 
@@
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
-Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectory, but that can be improved, through the use of simplification techniques and problem constraints.
+Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectories, but that can be improved, through the use of simplification techniques and problem constraints.
 
@@
 In summary, a reference signal is evaluated, until an (approximately) error-free reception is acquired, which can be measured through a simple transport stream (TS) error-indication interface and further refined with other parameters, such as carrier-to-noise ratio and signal level.
-This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignments procedures, until a counterexample is not output.
+This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignment procedures, until a counterexample is not output.
 It is worth noticing that the timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{Araujo16}, which can be seen as an advantage for such an implementation.
@@
 Amoedo {\it et al.} \cite{amoedo} proposed a modulation classification methodology, with the goal of spectrum sensing for opportunistic transmissions.
-Besides a correct software implementation, {\it i.e.}, source code, which is already covered by formal methods already available in the literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
-This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals under a variety of condition, unifying source code and behavior verification.
+Besides a correct software implementation, {\it i.e.}, source code, which is already covered by formal methods already available in literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
+This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals under a variety of conditions, unifying source code and behavior verification.
 In addition, that kind of implementation usually employs digital filters with specific parameters, which could also be tackled.
@@
 There is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
-In addition, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.
-Finally, there are some bottlenecks in software verification that are possible to be handled with machine learning, given some initial studies carried out in the current literature, which may represent another step towards a unified and complete approach that combines code verification, behavior correctness and environment adaptation.
+In addition, it seems interesting to provide behavioral models when new application interfaces or programming language features are used, in order to extend capabilities of current verification tools, without changing the core BMC module.
+Finally, there are some bottlenecks in software verification that are possible to be handled with machine learning, given some initial studies carried out in the current literature, which may represent another step towards a unified and complete approach that combines code verification, behavior correctness, environment adaptation, and system evolution.
 

### 2017-08-02 paper.tex
TCPS format

@@
-\documentclass{acm_sen_article}
-
-\usepackage{color}
+\documentclass[format=acmsmall, review=false, screen=true]{acmart}
+
+\usepackage{booktabs} % For formal tables
+% \usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{listings}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-%\renewcommand{\baselinestretch}{0.93}
+\usepackage{hyperref}	
+\usepackage{afterpage}
+\usepackage{graphicx}
+\usepackage[caption=false]{subfig}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{epstopdf}
+\usepackage{xcolor,colortbl}
+% \usepackage{cite}
+\newtheorem{mydef}{Definition}
+\usepackage{multirow}
+\usepackage{xspace}
+\usepackage[ruled]{algorithm2e} % For algorithms
+\renewcommand{\algorithmcfname}{ALGORITHM}
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
+
+% Metadata Information
+\acmJournal{TCPS}
+\acmVolume{9}
+\acmNumber{9}
+\acmArticle{99}
+\acmYear{2017}
+\acmMonth{12}
+\copyrightyear{2017}
+%\acmArticleSeq{9}
+
+% Copyright
+%\setcopyright{acmcopyright}
+\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\acmDOI{0000001.0000001}
+
+% Paper history
+\received{February 2007}
+\received[revised]{March 2009}
+\received[accepted]{June 2009}
 
@@
 
+\usepackage{color}
+\usepackage{listings}
+
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
+%\renewcommand{\baselinestretch}{0.93}
+
 \begin{document}
@@
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{Lucas C. Cordeiro \\
-%Department of Computer Science\\
-University of Oxford, UK\\
-lucas.cordeiro@cs.ox.ac.uk
-\and
-Eddie B. de Lima Filho \\
-Samsung Electronics, Brazil \\
-eddie.l@samsung.com
-\and
-Iury V. de Bessa \\
-Federal University of Amazonas, Brazil \\
-iury.bessa@ufam.edu.br}
+\author{Lucas C. Cordeiro}
+\orcid{0000-0002-6235-4272}
+\affiliation{%
+  \institution{Department of Computer Science, University of Oxford}
+%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
+  \city{Oxford}
+%  \state{VA}
+%  \postcode{23185}
+  \country{UK}}
+
+\author{Eddie B. de Lima Filho}
+\orcid{0000-0002-2758-4638}
+\affiliation{%
+  \institution{TP Vision}
+%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
+  \city{Manaus}
+%  \state{VA}
+%  \postcode{23185}
+  \country{Brazil}}
+
+\author{Iury V. de Bessa}
+\orcid{0000-0002-2758-4638}
+\affiliation{%
+  \institution{Department of Electricity, Federal University of Amazonas}
+%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
+  \city{Manaus}
+%  \state{VA}
+%  \postcode{23185}
+  \country{Brazil}}
+
+%% author names and affiliations
+%% use a multiple column layout for up to three different
+%% affiliations
+%\author{Eddie B. de Lima Filho \\
+%%Department of Computer Science\\
+%University of Oxford, UK\\
+%lucas.cordeiro@cs.ox.ac.uk
+%\and
+%Eddie B. de Lima Filho \\
+%Samsung Electronics, Brazil \\
+%eddie.l@samsung.com
+%\and
+%Iury V. de Bessa \\
+%Federal University of Amazonas, Brazil \\
+%iury.bessa@ufam.edu.br}
 %\IEEEauthorblockN{Homer Simpson}
@@
 %\IEEEpeerreviewmaketitle
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10010405.10010432.10010439.10010440</concept_id>
+<concept_desc>Applied computing~Computer-aided design</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010520.10010553</concept_id>
+<concept_desc>Computer systems organization~Embedded and cyber-physical systems</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010178.10010213.10010214</concept_id>
+<concept_desc>Computing methodologies~Computational control theory</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10002944.10011122.10002945</concept_id>
+<concept_desc>General and reference~Surveys and overviews</concept_desc>
+<concept_significance>100</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Applied computing~Computer-aided design}
+\ccsdesc[500]{Computer systems organization~Embedded and cyber-physical systems}
+\ccsdesc[300]{Computing methodologies~Computational control theory}
+\ccsdesc[100]{General and reference~Surveys and overviews}
 
@@
 
-The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems}.
+The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs).
 Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).
@@
 %--------------------------------------
-\section{A model for Cyber-physical systems}
+\section{Automated Verification and Synthesis of Cyber-physical systems}
 \label{sec:model}
+%--------------------------------------
+
+Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS.
+A CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
+
+The overall process of formal and automated synthesis for CPS and HS involves three main steps: modeling, that is related to obtain a sound mathematical representation that is able to capture all the relevant aspects of continuous and discrete dynamics and their interaction, as well as formulate properties and invariants that describe the desirable behavior; verification, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller synthesis itself, that produces a controller or finds invariant control laws that are able to make the CPS, modeled according to the first step (modeling), meets the requirements and behaves as desired, as confirmed by the verification step.
+
+%--------------------------------------
+\subsection{A model for Cyber-physical systems}
+\label{ssec:model}
 %--------------------------------------
@@
 
-CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata (\textcolor{red}{CITAR}) and hybrid automata (\textcolor{red}{CITAR}).
-Rungger and Tabuada (\textcolor{red}{CITAR TAC 2016}) proposed a general and wide model for CPS.
-Following their proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,r)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
-
-\textcolor{red}{seria legal falar que nao determinismo pode ser tratado com model checking e ligar com o nosso trabalho...
+CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.
+
+Timed automata employ finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories with constant slope.
+Hybrid automata generalize timed automata, also employing finite automata with real variables whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
+Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
+
+Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for CPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for CPS.
+
+%Following their proposal, a CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,t)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, and $r$ is a transition function, such that $r:X\times U \rightarrow X$.
+
+%\textcolor{red}{seria legal falar que nao determinismo pode ser tratado com model checking e ligar com o nosso trabalho...
 inclusive , o tabuada tem um artigo falando sobre seguranca e SMT, que eu comentei abaixo}
+
+%--------------------------------------
+\subsection{Automated and Formal Verification of Cyber-physical systems}
+\label{ssec:verification}
+%--------------------------------------
+
+Given a class of CPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
+Decidability is a key issue in CPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model CPS) are infinite-state machines~\cite{Henzinger95}.
+Generally, the properties that must be hold for CPS are related to safety, liveness, and reachability.
+In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.} bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
 
@@
 %--------------------------------------
+\subsection{Control synthesis of Cyber-physical systems}
+\label{ssec:synthesis}
+%--------------------------------------
+
+%--------------------------------------
 \section{Bounded Model Checking (BMC)}
@@
 Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers.
-BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.
+BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}.
 
@@
 In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
-Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
+Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).
 
@@
 
-\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}.
+\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}.
 This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
@@
 
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
@@
 
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{refs}
+
 % conference papers do not normally have an appendix
@@
 %1
-\bibitem{Kopetz11}
-H.~Kopetz:
-\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
-\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
-
-%2
-\bibitem{xu15}
-Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
-\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
-\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
-
-%3
-\bibitem{leeCPS}
-Lee E.:
-\newblock {\em {Cyber-physical Systems: Design Challenges}}.
-\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
-
-%4
-\bibitem{leeCPS2}
-Lee E.:
-\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
-\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
-
-%5
-\bibitem{groza2015formal}
-Groza A., Letia I., Goron A., Zaporojan S.: 
-\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
-\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
-
-%6
-\bibitem{Cordeiro09}
-Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
-\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
-\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
-
-%7
-\bibitem{munir}
-Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
-\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
-\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
-
-%8
-\bibitem{Abate17}
-Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
-\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
-\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
-
-%9
-\bibitem{ldd}
-Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
-\newblock {\em {Model checking concurrent linux device drivers}}.
-\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
-
-%10
-\bibitem{mlembe}
-S.~Virtanen:
-\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
-\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.
-
-%11
-\bibitem{Kroening15}
-Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
-\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
-\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
-
-%12
-\bibitem{cudalucas}
-Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
-\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
-\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.
-
-%13
-\bibitem{stochmodel}
-Bargmann H.:
-\newblock {\em {The role of stochastic modelling in engineering science}}.
-\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.
-
-%14
-\bibitem{Shoukry}
-Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
-\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
-\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.
-
-%15
-\bibitem{cpsattack}
-Liu Y., Ning P., Reiter M. K.:
-\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
-\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.
-
-%16
-\bibitem{Araujo16}
-Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
-\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
-\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
-
-%17
-\bibitem{Biere99}
-Biere A., Cimatti A., Clarke E., Zhu Y.:
-\newblock {\em {Symbolic Model Checking without BDDs}}. 
-\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
-
-%18
-\bibitem{handbook09}
-Biere A., Heule M., van Maaren H., Walsh T., eds.:
-\newblock {\em {Handbook of Satisfiability}}.
-\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
-
-%19
-\bibitem{BarrettSST09}
-Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
-\newblock {\em {Satisfiability Modulo Theories}}. 
-\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
-
-%20
-\bibitem{Armando06}
-Armando A., Mantovani J., Platania L.:
-\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
-\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
-
-%21
-\bibitem{Bradley07}
-Bradley A., Manna Z.:
-\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
-\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
-
-%22
-\bibitem{Thomas01}
-Thomas Ball and Sriram~K. Rajamani.
-\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
-\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
-
-%23
-\bibitem{Strichman08}
-Daniel Kroening and Ofer Strichman.
-\newblock {\em Decision Procedures: An Algorithmic Point of View}.
-\newblock Springer Publishing Company, Incorporated, 2008.
-
-%24
-\bibitem{Appel97}
-Andrew~W. Appel.
-\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
-\newblock Cambridge University Press, New York, NY, USA, 1997.
-
-%25
-\bibitem{Tseitin83}
-G.~S. Tseitin.
-\newblock On the complexity of derivation in propositional calculus.
-\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
-  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
-
- %26
- \bibitem{PatarinG97}
-Jacques Patarin and Louis Goubin.
-\newblock Trapdoor one-way permutations and multivariate polynominals.
-\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997.
-
-%27
-\bibitem{Clarke04}
-Clarke E., Kroening D., Lerda F.:
-\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.
-
-%28
-\bibitem{MerzFS12}
-Merz F., Falke S., Sinz C.:
-\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
-\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
-
-%29
-\bibitem{CordeiroF11}
-Cordeiro L., Fischer B.:
-\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
-\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
-
-%30
-\bibitem{Ivancic05}
-Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
-\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
-\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
-
-%31
-\bibitem{Cordeiro12}
-Cordeiro L., Fischer B., Marques{-}Silva J.:
-\newblock {\em {{SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}}.
-\newblock {IEEE} Trans. Software Eng. 38(4), pp. 957--974, 2012.
-
-%32
-\bibitem{EenS03}
-E{\'{e}}n, N., S{\"{o}}rensson, N.:
-\newblock {\em {Temporal Induction by Incremental {SAT} Solving}}.
-\newblock Electronic Notes in Theoretical Computer Science 89(4), pp. 543 -- 560, 2003.
-
-%33
-\bibitem{Sheera00}
-Sheeran M., Singh S., St{\aa}lmarck G.:
-\newblock {\em {Checking Safety Properties using Induction and a {SAT}-solver}}.
-\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
-
-%34
-\bibitem{4610935}
-{IEEE} standard for floating-point arithmetic.
-\newblock {\em IEEE Std 754-2008}, pages 1--70, Aug 2008.
-
-%35
-\bibitem{Goldberg91whatevery}
-David Goldberg.
-\newblock What every computer scientist should know about floating point
-  arithmetic.
-\newblock {\em ACM Computing Surveys}, 23(1):5--48, 1991.
-
-%36
-\bibitem{pips:2013}
-ParisTech M.:
-\newblock {\em {PIPS: Automatic Parallelizer and Code Transformation Framework}}. 
-\newblock Accessed 21 February 2016
-
-%37
-\bibitem{Henry:2012}
-Henry J., Monniaux D., Moy M.: 
-\newblock {\em {PAGAI: A Path Sensitive Static Analyser}}. 
-\newblock In: Electronic Notes in Theoretical Computer Science, pp. 15--25, 2012.
-
-%38
-\bibitem{Beyer15}
-Beyer D., Dangl M., Wendler P.:
-\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
-\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
-
-%39
-\bibitem{Gadelha15}
-Gadelha M., Ismail H., Cordeiro L.:
-\newblock {\em {Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction}}.
-\newblock In: International Journal on Software Tools for Technology Transfer (to appear), 2015.
-\newblock http://dx.doi.org/10.1007/s10009-015-0407-9
-
-%40
-\bibitem{Brain15}
-Brain M., Joshi S., Kroening D., Schrammel P.:
-\newblock {\em {Safety Verification and Refutation by k-Invariants and \textit{k}-Induction}}.
-\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
-
-%41
-\bibitem{Rocha15}
-Rocha H., Ismail H., Cordeiro L., Barreto R.:
-\newblock {\em {Model Checking Embedded C Software using \textit{k}-Induction and Invariants}}.
-\newblock In: V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
-
-%42
-\bibitem{Kinductor}
-Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-\newblock {\em {Software Verification using \textit{k}-Induction}}.
-\newblock In: International Symposium on Static Analysis. LNCS 6887, pp. 351--368, 2011.
-
-%43
-\bibitem{Rocha17}
-Rocha, W., Rocha, H., Ismail H., Cordeiro, L., Fischer, B.:
-\newblock {\em {DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}}. 
-\newblock In: International Conference on Tools and Algorithms for the Construction and Analysis of Systems, 2017 (to appear).
-
-%44
-\bibitem{Donaldson10}
-Donaldson A., Kroening D., R{\"{u}}mmer P.:
-\newblock {\em {{SCRATCH}: A Tool for Automatic Analysis of {DMA} Races}}.
-\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
-
-%45
-\bibitem{GrosseLD09}
-Gro{\ss}e D., Le H., Drechsler R.:
-\newblock {\em {Induction-based Formal Verification of {SystemC TLM} Designs}}.
-\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
-
-%46
-\bibitem{David15}
-C.~David, D.~Kroening, and M.~Lewis.
-\newblock {\em {Using program synthesis for program analysis}}.
-\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
-
-%47
-\bibitem{Alur13}
-Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
-\newblock {\em {Syntax-guided synthesis}}. 
-\newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
-
-%48
-\bibitem{sketch}
-A.~Solar-Lezama.
-\newblock {\em {Program sketching}}.
-\newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
-
-%49
-\bibitem{Sharma14}
-R.~Sharma and A.~Aiken.
-\newblock {\em {From invariant checking to invariant inference using randomized search}}.
-\newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
-
-%50
-\bibitem{dsv_spin2015}
-Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-\newblock {\em {DSVerifier: A Bounded Model Checking Tool for Digital Systems}}. 
-\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
-
-%51
-\bibitem{Monteiro16}
-F.~Monteiro:
-\newblock {\em {Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}}.
-\newblock In: 24th {ACM} {SIGSOFT} International Symposium on Foundations of Software Engineering, pp. 1151--1153, 2016.
-
-%52
-\bibitem{Bessa17}
-Bessa, I., Ismail, H., Palhares, R., Cordeiro, L., Chaves Filho, J.:
-\newblock {\em {Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}}. 
-\newblock In: IEEE Transactions on Computers, 2017 (to appear).
-
-%53
-\bibitem{selfcar}
-Fairley P.:
-\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
-\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
-
-%54
-\bibitem{BMCml}
-L.~Bortolussi, D. Milios, G, Sanguinetti.:
-\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
-\newblock In:. Information and Computation, 2016.
-
-%55
-\bibitem{Behrend15}
-Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
-\newblock {\em {Scalable and Optimized Hybrid Verification of Embedded Software}}. 
-\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
-
-%56
-\bibitem{Lettnin09}
-Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
-\newblock {\em {Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}}. 
-\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
-
-%57
-\bibitem{Adouane16}
-Vilca J., Adouane L., Mezouar Y.:
-\newblock {\em {Optimal Multi-Criteria Waypoint Selection for Autonomous Vehicle Navigation in Structured Environment}}. 
-\newblock In: Journal of Intelligent and Robotic Systems 82(2): pp. 301--324, 2016.
-
-%58
-\bibitem{ECBS13}
-Ramalho M., Lopes M., Sousa F., Marques H., Cordeiro L., Fischer B.:
-\newblock {\em {SMT-Based Bounded Model Checking of C++ Programs} }.
-\newblock In: International Conference on Engineering of Computer-Based Systems, pp. 147--156, 2013.
-
-%59
-\bibitem{Pereira15}
-Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
-\newblock {\em {Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking}}.
-\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016.
-
-%60
-\bibitem{Sousa15}
-Sousa F., Cordeiro L., Lima Filho E.: 
-\newblock {\em { Bounded Model Checking of C++ Programs Based on the Qt Framework.}} 
-\newblock In: Global Conference on Consumer Electronics, pp. 179--180, 2015.
-
-%61
-\bibitem{Beyer14}
-D.~Beyer 
-\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
-
-%62
-\bibitem{BeyerSVCOMP15}
-D.~Beyer:
-\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
-\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
-
-%63
-\bibitem{KahlonWG09}
-Kahlon V., Wang C., Gupta A.:
-\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
-\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
-
-%64
-\bibitem{morse15}
-J.~Morse.
-\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
-\newblock {University of Southampton, PhD Thesis}, 2015.
-
-%65
-\bibitem{Inverso14}
-Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
-\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
-\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
-
-%66
-\bibitem{civl15}
-Zheng M, Rogers M, Luo Z, Dwyer M, Siegel S.
-\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
-\newblock In: International Conference on Automated Software Engineering, pp. 830--835, 2015.
-
-%67
-\bibitem{Grumberg05}
-Grumberg O., Lerda F., Strichman O., Theobald M.:
-\newblock {\em { Proof-guided underapproximation-widening for multi-process systems.}} 
-\newblock In: Symposium on Principles of Programming Languages, pp. 122--131, 2005.
-
-%68
-\bibitem{McMillan11}
-K.~McMillan:
-\newblock {\em {Widening and Interpolation.}} 
-\newblock In: International Symposium on Static Analysis, LCNS 6887, pp. 1, 2011.
-
-%69
-\bibitem{esbmc_controller}
-Bessa I., Abreu R., Cordeiro L., Filho J.:
-\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers}.}
-\newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.
-
-%70
-\bibitem{esbmc_filter}
-Abreu R., Cordeiro L., Filho E.:
-\newblock {\em {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}.}
-\newblock In: XXXI Brazilian Symposium on Telecommunications, 2013.
-\newblock http://dx.doi.org/10.14209/sbrt.2013.57
-
-%71
-\bibitem{JMorse15}
-Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {Model Checking LTL Properties over ANSI-C Programs with Bounded Traces.}} 
-\newblock In: Software and System Modeling 14(1): pp. 65--81, 2015.
-
-%72
-\bibitem{Bessa16}
-Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
-\newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
-\newblock In: Design Automation for Embedded Systems 20(2): pp. 95--126, 2016.
-
-%73
-\bibitem{Chaves17}
-Chaves L., Bessa I., Cordeiro L.:
-\newblock {\em { DSValidator: An Automated Counterexample Reproducibility Tool for Digital Systems}}. 
-\newblock In: CoRR abs/1610.07066, 2016.
-
-%74
-\bibitem{sbrtdaniel}
-de Mello D. P. M., de Freitas M. L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
-\newblock {\em{Verification of Magnitude and Phase Responses in Fixed-Point Digital Filters}}. 
-\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017.
-
-%75
-\bibitem{Alves15}
-Alves E. , Cordeiro L., Lima Filho E.:
-\newblock {\em{ Fault Localization in Multi-Threaded C Programs using Bounded Model Checking}}. 
-\newblock In: Brazilian Symposium on Computing Systems Engineering, pp. 96-101, 2015.
-
-%76
-\bibitem{Garcia16}
-Garcia M. , Sousa F., Cordeiro, L., Lima Filho E.: 
-\newblock {\em{ ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
-\newblock In: International SPIN symposium on Model Checking of Software (to appear), 2016.
-
-%77
-\bibitem{Lopes:2014:GSL:2692607}
-Bruno~Cardoso Lopes and Rafael Auler.
-\newblock {\em Getting Started with LLVM Core Libraries}.
-\newblock Packt Publishing, 2014.
-
-%78
- \bibitem{clang-usage}
-Cade Metz.
-\newblock Why {A}pple?s swift language will instantly remake computer
-  programming, 204.
-\newblock http://www.wired.com/2014/07/apple-swift/.
-
-%79
-\bibitem{Jackson16}
-Jackson D., Vaziri M.:
-\newblock {\em{ Correct or usable? the limits of traditional verification (impact paper award)}}. 
-\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 11, 2016.
-
-%80
-\bibitem{MorseCNF13}
-Morse J., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
-\newblock In: Tools and Algorithms for the Construction and Analysis of
-  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.
-
-%81
-\bibitem{MorseRCN014}
-Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
-\newblock {\em {ESBMC} 1.22.}
-\newblock In: Tools and Algorithms for the Construction and Analysis of
-  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.
-
-%82
-\bibitem{Haran15}
-Haran A., Carter M., Emmi M., Lal A., Qadeer S., Rakamaric Z.:
-\newblock {\em{ SMACK+Corral: A Modular Verifier - (Competition Contribution)}}. 
-\newblock In: 21st International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 451--454, 2015.
-
-%83
-\bibitem{Beyer16}
-Beyer, D.:
-\newblock {\em{ Reliable and Reproducible Competition Results with BenchExec and Witnesses (Report on SV-COMP 2016)}}. 
-\newblock In: 22nd International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9636, pp. 887--904, 2016.
-
-%84
-\bibitem{BeyerW15}
-Beyer D., Dangl M., Dietsch D., Heizmann M., Stahlbauer A.:
-\newblock {\em{ Witness validation and stepwise testification across software verifiers}}. 
-\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 721--733, 2015.
-
-%85
-\bibitem{RochaIFM12}
-Rocha H., Barreto R., Cordeiro L., Dias Neto A.:
-\newblock {\em{ Understanding Programming Bugs in ANSI-C Software Using Bounded Model Checking Counter-Examples}}. 
-\newblock In: International Conference on Integrated Formal Methods, LNCS 7321, pp. 128-142, 2012.
-
-%86
-\bibitem{antennasbrt}
-de Jesus A. S., rodrigues R. N., Ferreira A. N. G., Valente V. H., de Melo W. C., de Lima Filho E. B, da Silva J\'unior W. S.:
-\newblock {\em{Automatic Antenna Alignment System for Satellite Receivers Operating in C and Ku Bands}}. 
-\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017. (in Portuguese)
-
-%87
-\bibitem{amoedo}
-Amoedo D. A., da Silva J\'unior W. S., de Lima Filho E. A.:
-\newblock {\em{Parameter Selection for SVM in Automatic Modulation Classification of Analog and Digital Signals}}. 
-\newblock In: International Telecommunications Symposium (ITS), pp. 1-5, 2014.
-
-%88
-\bibitem{mlswverif}
-Hamel L.:
-\newblock {\em {On the Use of Machine Learning in Formal Software Verification.}} 
-\newblock {Technical Report TR03-294, Dept. of Computer Science and Statistics, University of Rhode Island}, 2003.
-
-%89
-\bibitem{nguyen}
-Phuc N. V.,:
-\newblock {\em {The Application of Machine Learning Methods in Software Verification and Validation.}} 
-\newblock {University of Texas at Austin, master of science dissertation}, 2010.
-
-%90
-\bibitem{bridge}
-Bridge J. P.:
-\newblock {\em {Machine learning and automated theorem proving.}} 
-\newblock {Technical Report 792, Computer Laboratory, University of Cambridge}, 2010.
-
- %91
-\bibitem{babic07}
-Hutter F., Babic D., Hoos H. H., Hu A. J.              
-\newblock {\em{Boosting Verification by Automatic Tuning of Decision Procedures.}}
-\newblock {In 7th International Conference Formal Methods in Computer-Aided Design {(FMCAD)}, pp. 27--34, 2007.}
+%\bibitem{Kopetz11}
+%H.~Kopetz:
+%\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
+%\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
+%
+%%2
+%\bibitem{xu15}
+%Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
+%\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
+%\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
+%
+%%3
+%\bibitem{leeCPS}
+%Lee E.:
+%\newblock {\em {Cyber-physical Systems: Design Challenges}}.
+%\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
+%
+%%4
+%\bibitem{leeCPS2}
+%Lee E.:
+%\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
+%\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
+%
+%%5
+%\bibitem{groza2015formal}
+%Groza A., Letia I., Goron A., Zaporojan S.: 
+%\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
+%\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
+%
+%%6
+%\bibitem{Cordeiro09}
+%Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
+%\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
+%\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
+%
+%%7
+%\bibitem{munir}
+%Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
+%\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
+%\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
+%
+%%8
+%\bibitem{Abate17}
+%Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
+%\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
+%\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
+%
+%%9
+%\bibitem{ldd}
+%Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
+%\newblock {\em {Model checking concurrent linux device drivers}}.
+%\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
+%
+%%10
+%\bibitem{mlembe}
+%S.~Virtanen:
+%\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
+%\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.
+%
+%%11
+%\bibitem{Kroening15}
+%Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
+%\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
+%\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
+%
+%%12
+%\bibitem{cudalucas}
+%Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
+%\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
+%\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.
+%
+%%13
+%\bibitem{stochmodel}
+%Bargmann H.:
+%\newblock {\em {The role of stochastic modelling in engineering science}}.
+%\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.
+%
+%%14
+%\bibitem{Shoukry}
+%Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
+%\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
+%\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.
+%
+%%15
+%\bibitem{cpsattack}
+%Liu Y., Ning P., Reiter M. K.:
+%\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
+%\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.
+%
+%%16
+%\bibitem{Araujo16}
+%Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
+%\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
+%\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
+%
+%%17
+%\bibitem{Biere99}
+%Biere A., Cimatti A., Clarke E., Zhu Y.:
+%\newblock {\em {Symbolic Model Checking without BDDs}}. 
+%\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
+%
+%%18
+%\bibitem{handbook09}
+%Biere A., Heule M., van Maaren H., Walsh T., eds.:
+%\newblock {\em {Handbook of Satisfiability}}.
+%\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
+%
+%%19
+%\bibitem{BarrettSST09}
+%Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
+%\newblock {\em {Satisfiability Modulo Theories}}. 
+%\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
+%
+%%20
+%\bibitem{Armando2009}
+%Armando A., Mantovani J., Platania L.:
+%\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
+%\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
+%
+%%21
+%\bibitem{Bradley07}
+%Bradley A., Manna Z.:
+%\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
+%\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
+%
+%%22
+%\bibitem{Thomas01}
+%Thomas Ball and Sriram~K. Rajamani.
+%\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
+%\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
+%
+%%23
+%\bibitem{Strichman08}
+%Daniel Kroening and Ofer Strichman.
+%\newblock {\em Decision Procedures: An Algorithmic Point of View}.
+%\newblock Springer Publishing Company, Incorporated, 2008.
+%
+%%24
+%\bibitem{Appel97}
+%Andrew~W. Appel.
+%\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
+%\newblock Cambridge University Press, New York, NY, USA, 1997.
+%
+%%25
+%\bibitem{Tseitin83}
+%G.~S. Tseitin.
+%\newblock On the complexity of derivation in propositional calculus.
+%\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
+%  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
+% 
+% %26
+% \bibitem{PatarinG97}
+%Jacques Patarin and Louis Goubin.
+%\newblock Trapdoor one-way permutations and multivariate polynominals.
+%\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997. 
+%
+%%27
+%\bibitem{Clarke04}
+%Clarke E., Kroening D., Lerda F.:
+%\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
+%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.
+%
+%%28
+%\bibitem{MerzFS12}
+%Merz F., Falke S., Sinz C.:
+%\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
+%\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
+%
+%%29
+%\bibitem{CordeiroF11}
+%Cordeiro L., Fischer B.:
+%\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
+%\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
+%
+%%30
+%\bibitem{Ivancic05}
+%Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
+%\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
+%\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
+%
+%%31
+%\bibitem{Cordeiro12}
+%Cordeiro L., Fischer B., Marques{-}Silva J.:
+%\newblock {\em {{SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}}.
+%\newblock {IEEE} Trans. Software Eng. 38(4), pp. 957--974, 2012.
+%
+%%32
+%\bibitem{EenS03}
+%E{\'{e}}n, N., S{\"{o}}rensson, N.:
+%\newblock {\em {Temporal Induction by Incremental {SAT} Solving}}.
+%\newblock Electronic Notes in Theoretical Computer Science 89(4), pp. 543 -- 560, 2003.
+%
+%%33
+%\bibitem{Sheera00}
+%Sheeran M., Singh S., St{\aa}lmarck G.:
+%\newblock {\em {Checking Safety Properties using Induction and a {SAT}-solver}}.
+%\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
+%
+%%34
+%\bibitem{4610935}
+%{IEEE} standard for floating-point arithmetic.
+%\newblock {\em IEEE Std 754-2008}, pages 1--70, Aug 2008.
+%
+%%35
+%\bibitem{Goldberg91whatevery}
+%David Goldberg.
+%\newblock What every computer scientist should know about floating point
+%  arithmetic.
+%\newblock {\em ACM Computing Surveys}, 23(1):5--48, 1991.
+%
+%%36
+%\bibitem{pips:2013}
+%ParisTech M.:
+%\newblock {\em {PIPS: Automatic Parallelizer and Code Transformation Framework}}. 
+%\newblock Accessed 21 February 2016
+%
+%%37
+%\bibitem{Henry:2012}
+%Henry J., Monniaux D., Moy M.: 
+%\newblock {\em {PAGAI: A Path Sensitive Static Analyser}}. 
+%\newblock In: Electronic Notes in Theoretical Computer Science, pp. 15--25, 2012.
+%
+%%38
+%\bibitem{Beyer15}
+%Beyer D., Dangl M., Wendler P.:
+%\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
+%\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
+%
+%%39
+%\bibitem{Gadelha15}
+%Gadelha M., Ismail H., Cordeiro L.:
+%\newblock {\em {Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction}}.
+%\newblock In: International Journal on Software Tools for Technology Transfer (to appear), 2015.
+%\newblock http://dx.doi.org/10.1007/s10009-015-0407-9
+%
+%%40
+%\bibitem{Brain15}
+%Brain M., Joshi S., Kroening D., Schrammel P.:
+%\newblock {\em {Safety Verification and Refutation by k-Invariants and \textit{k}-Induction}}.
+%\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
+%
+%%41
+%\bibitem{Rocha15}
+%Rocha H., Ismail H., Cordeiro L., Barreto R.:
+%\newblock {\em {Model Checking Embedded C Software using \textit{k}-Induction and Invariants}}.
+%\newblock In: V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
+%
+%%42
+%\bibitem{Kinductor}
+%Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
+%\newblock {\em {Software Verification using \textit{k}-Induction}}.
+%\newblock In: International Symposium on Static Analysis. LNCS 6887, pp. 351--368, 2011.
+%
+%%43
+%\bibitem{Rocha17}
+%Rocha, W., Rocha, H., Ismail H., Cordeiro, L., Fischer, B.:
+%\newblock {\em {DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}}. 
+%\newblock In: International Conference on Tools and Algorithms for the Construction and Analysis of Systems, 2017 (to appear).
+%
+%%44
+%\bibitem{Donaldson10}
+%Donaldson A., Kroening D., R{\"{u}}mmer P.:
+%\newblock {\em {{SCRATCH}: A Tool for Automatic Analysis of {DMA} Races}}.
+%\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
+%
+%%45
+%\bibitem{GrosseLD09}
+%Gro{\ss}e D., Le H., Drechsler R.:
+%\newblock {\em {Induction-based Formal Verification of {SystemC TLM} Designs}}.
+%\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
+%
+%%46
+%\bibitem{David15}
+%C.~David, D.~Kroening, and M.~Lewis.
+%\newblock {\em {Using program synthesis for program analysis}}.
+%\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
+%
+%%47
+%\bibitem{Alur13}
+%Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
+%\newblock {\em {Syntax-guided synthesis}}. 
+%\newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
+%
+%%48
+%\bibitem{sketch}
+%A.~Solar-Lezama.
+%\newblock {\em {Program sketching}}.
+%\newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
+%
+%%49
+%\bibitem{Sharma14}
+%R.~Sharma and A.~Aiken.
+%\newblock {\em {From invariant checking to invariant inference using randomized search}}.
+%\newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
+%
+%%50
+%\bibitem{dsv_spin2015}
+%Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
+%\newblock {\em {DSVerifier: A Bounded Model Checking Tool for Digital Systems}}. 
+%\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
+%
+%%51
+%\bibitem{Monteiro16}
+%F.~Monteiro:
+%\newblock {\em {Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}}.
+%\newblock In: 24th {ACM} {SIGSOFT} International Symposium on Foundations of Software Engineering, pp. 1151--1153, 2016.
+%
+%%52
+%\bibitem{Bessa17}
+%Bessa, I., Ismail, H., Palhares, R., Cordeiro, L., Chaves Filho, J.:
+%\newblock {\em {Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}}. 
+%\newblock In: IEEE Transactions on Computers, 2017 (to appear).
+%
+%%53
+%\bibitem{selfcar}
+%Fairley P.:
+%\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
+%\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
+%
+%%54
+%\bibitem{BMCml}
+%L.~Bortolussi, D. Milios, G, Sanguinetti.:
+%\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
+%\newblock In:. Information and Computation, 2016.
+%
+%%55
+%\bibitem{Behrend15}
+%Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
+%\newblock {\em {Scalable and Optimized Hybrid Verification of Embedded Software}}. 
+%\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
+%
+%%56
+%\bibitem{Lettnin09}
+%Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
+%\newblock {\em {Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}}. 
+%\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
+%
+%%57
+%\bibitem{Adouane16}
+%Vilca J., Adouane L., Mezouar Y.:
+%\newblock {\em {Optimal Multi-Criteria Waypoint Selection for Autonomous Vehicle Navigation in Structured Environment}}. 
+%\newblock In: Journal of Intelligent and Robotic Systems 82(2): pp. 301--324, 2016.
+%
+%%58
+%\bibitem{ECBS13}
+%Ramalho M., Lopes M., Sousa F., Marques H., Cordeiro L., Fischer B.:
+%\newblock {\em {SMT-Based Bounded Model Checking of C++ Programs} }.
+%\newblock In: International Conference on Engineering of Computer-Based Systems, pp. 147--156, 2013.
+%
+%%59
+%\bibitem{Pereira15}
+%Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
+%\newblock {\em {Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking}}.
+%\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016. 
+%
+%%60
+%\bibitem{Sousa15}
+%Sousa F., Cordeiro L., Lima Filho E.: 
+%\newblock {\em { Bounded Model Checking of C++ Programs Based on the Qt Framework.}} 
+%\newblock In: Global Conference on Consumer Electronics, pp. 179--180, 2015.
+%
+%%61
+%\bibitem{Beyer14}
+%D.~Beyer 
+%\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
+%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
+% 
+%%62
+%\bibitem{BeyerSVCOMP15}
+%D.~Beyer:
+%\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
+%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
+%
+%%63
+%\bibitem{KahlonWG09}
+%Kahlon V., Wang C., Gupta A.:
+%\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
+%\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
+%
+%%64
+%\bibitem{morse15}
+%J.~Morse.
+%\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
+%\newblock {University of Southampton, PhD Thesis}, 2015.
+%
+%%65
+%\bibitem{Inverso14}
+%Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
+%\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
+%\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
+%
+%%66
+%\bibitem{civl15}
+%Zheng M, Rogers M, Luo Z, Dwyer M, Siegel S.
+%\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
+%\newblock In: International Conference on Automated Software Engineering, pp. 830--835, 2015.
+%
+%%67
+%\bibitem{Grumberg05}
+%Grumberg O., Lerda F., Strichman O., Theobald M.:
+%\newblock {\em { Proof-guided underapproximation-widening for multi-process systems.}} 
+%\newblock In: Symposium on Principles of Programming Languages, pp. 122--131, 2005.
+%
+%%68
+%\bibitem{McMillan11}
+%K.~McMillan:
+%\newblock {\em {Widening and Interpolation.}} 
+%\newblock In: International Symposium on Static Analysis, LCNS 6887, pp. 1, 2011.
+%
+%%69
+%\bibitem{esbmc_controller}
+%Bessa I., Abreu R., Cordeiro L., Filho J.:
+%\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers}.}
+%\newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.
+%
+%%70
+%\bibitem{esbmc_filter}
+%Abreu R., Cordeiro L., Filho E.:
+%\newblock {\em {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}.}
+%\newblock In: XXXI Brazilian Symposium on Telecommunications, 2013.
+%\newblock http://dx.doi.org/10.14209/sbrt.2013.57
+%
+%%71
+%\bibitem{JMorse15}
+%Morse J., Cordeiro L., Nicole D., Fischer B.:
+%\newblock {\em {Model Checking LTL Properties over ANSI-C Programs with Bounded Traces.}} 
+%\newblock In: Software and System Modeling 14(1): pp. 65--81, 2015.
+%
+%%72
+%\bibitem{Bessa16}
+%Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
+%\newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
+%\newblock In: Design Automation for Embedded Systems 20(2): pp. 95--126, 2016.
+%
+%%73
+%\bibitem{Chaves17}
+%Chaves L., Bessa I., Cordeiro L.:
+%\newblock {\em { DSValidator: An Automated Counterexample Reproducibility Tool for Digital Systems}}. 
+%\newblock In: CoRR abs/1610.07066, 2016.
+%
+%%74
+%\bibitem{sbrtdaniel}
+%de Mello D. P. M., de Freitas M. L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
+%\newblock {\em{Verification of Magnitude and Phase Responses in Fixed-Point Digital Filters}}. 
+%\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017.
+%
+%%75
+%\bibitem{Alves15}
+%Alves E. , Cordeiro L., Lima Filho E.:
+%\newblock {\em{ Fault Localization in Multi-Threaded C Programs using Bounded Model Checking}}. 
+%\newblock In: Brazilian Symposium on Computing Systems Engineering, pp. 96-101, 2015.
+%
+%%76
+%\bibitem{Garcia16}
+%Garcia M. , Sousa F., Cordeiro, L., Lima Filho E.: 
+%\newblock {\em{ ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
+%\newblock In: International SPIN symposium on Model Checking of Software (to appear), 2016.
+%
+%%77
+%\bibitem{Lopes:2014:GSL:2692607}
+%Bruno~Cardoso Lopes and Rafael Auler.
+%\newblock {\em Getting Started with LLVM Core Libraries}.
+%\newblock Packt Publishing, 2014.
+%
+%%78
+% \bibitem{clang-usage}
+%Cade Metz.
+%\newblock Why {A}pple?s swift language will instantly remake computer
+%  programming, 204.
+%\newblock http://www.wired.com/2014/07/apple-swift/.
+%
+%%79
+%\bibitem{Jackson16}
+%Jackson D., Vaziri M.:
+%\newblock {\em{ Correct or usable? the limits of traditional verification (impact paper award)}}. 
+%\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 11, 2016.
+%
+%%80
+%\bibitem{MorseCNF13}
+%Morse J., Cordeiro L., Nicole D., Fischer B.:
+%\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
+%\newblock In: Tools and Algorithms for the Construction and Analysis of
+%  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.
+%
+%%81
+%\bibitem{MorseRCN014}
+%Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
+%\newblock {\em {ESBMC} 1.22.}
+%\newblock In: Tools and Algorithms for the Construction and Analysis of
+%  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.
+%
+%%82
+%\bibitem{Haran15}
+%Haran A., Carter M., Emmi M., Lal A., Qadeer S., Rakamaric Z.:
+%\newblock {\em{ SMACK+Corral: A Modular Verifier - (Competition Contribution)}}. 
+%\newblock In: 21st International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 451--454, 2015.
+%
+%%83
+%\bibitem{Beyer16}
+%Beyer, D.:
+%\newblock {\em{ Reliable and Reproducible Competition Results with BenchExec and Witnesses (Report on SV-COMP 2016)}}. 
+%\newblock In: 22nd International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9636, pp. 887--904, 2016.
+%
+%%84
+%\bibitem{BeyerW15}
+%Beyer D., Dangl M., Dietsch D., Heizmann M., Stahlbauer A.:
+%\newblock {\em{ Witness validation and stepwise testification across software verifiers}}. 
+%\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 721--733, 2015.
+%
+%%85
+%\bibitem{RochaIFM12}
+%Rocha H., Barreto R., Cordeiro L., Dias Neto A.:
+%\newblock {\em{ Understanding Programming Bugs in ANSI-C Software Using Bounded Model Checking Counter-Examples}}. 
+%\newblock In: International Conference on Integrated Formal Methods, LNCS 7321, pp. 128-142, 2012.
+%
+%%86
+%\bibitem{antennasbrt}
+%de Jesus A. S., rodrigues R. N., Ferreira A. N. G., Valente V. H., de Melo W. C., de Lima Filho E. B, da Silva J\'unior W. S.:
+%\newblock {\em{Automatic Antenna Alignment System for Satellite Receivers Operating in C and Ku Bands}}. 
+%\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017. (in Portuguese)
+%
+%%87
+%\bibitem{amoedo}
+%Amoedo D. A., da Silva J\'unior W. S., de Lima Filho E. A.:
+%\newblock {\em{Parameter Selection for SVM in Automatic Modulation Classification of Analog and Digital Signals}}. 
+%\newblock In: International Telecommunications Symposium (ITS), pp. 1-5, 2014.
+%
+%%88
+%\bibitem{mlswverif}
+%Hamel L.:
+%\newblock {\em {On the Use of Machine Learning in Formal Software Verification.}} 
+%\newblock {Technical Report TR03-294, Dept. of Computer Science and Statistics, University of Rhode Island}, 2003.
+%
+%
+%%89
+%\bibitem{nguyen}
+%Phuc N. V.,:
+%\newblock {\em {The Application of Machine Learning Methods in Software Verification and Validation.}} 
+%\newblock {University of Texas at Austin, master of science dissertation}, 2010.
+%
+%%90
+%\bibitem{bridge}
+%Bridge J. P.:
+%\newblock {\em {Machine learning and automated theorem proving.}} 
+%\newblock {Technical Report 792, Computer Laboratory, University of Cambridge}, 2010.
+%
+% 
+% %91
+%\bibitem{babic07}
+%Hutter F., Babic D., Hoos H. H., Hu A. J.              
+%\newblock {\em{Boosting Verification by Automatic Tuning of Decision Procedures.}}
+%\newblock {In 7th International Conference Formal Methods in Computer-Aided Design {(FMCAD)}, pp. 27--34, 2007.}
 

### 2017-08-02 paper.tex
updated title
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
added comment
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 % Do not put math or special symbols in the title.
-\title{Automated Verification and Synthesis of Embedded Systems: Advances and Challenges}
+\title{Automated Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges}
 
@@
 
-Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS.
+Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS \textcolor{red}{add citations}.
 A CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
 
-The overall process of formal and automated synthesis for CPS and HS involves three main steps: modeling, that is related to obtain a sound mathematical representation that is able to capture all the relevant aspects of continuous and discrete dynamics and their interaction, as well as formulate properties and invariants that describe the desirable behavior; verification, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller synthesis itself, that produces a controller or finds invariant control laws that are able to make the CPS, modeled according to the first step (modeling), meets the requirements and behaves as desired, as confirmed by the verification step.
+The overall formal and automated synthesis process for CPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the CPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.
 

### 2017-08-02 paper.tex
reviewed section 2.1
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 %--------------------------------------
-\subsection{A model for Cyber-physical systems}
+\subsection{A Model for Cyber-physical Systems}
 \label{ssec:model}
@@
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent CPSs, since interaction among different entities (software and hardware components) presents several events that could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent CPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible.
-On the other hand, that approach is not suitable for complex CPSs that present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
+On the other hand, that approach is not suitable for complex CPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
 
@@
 Hybrid automata generalize timed automata, also employing finite automata with real variables whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
-Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
+Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
 

### 2017-08-02 paper.tex
minor fixes
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Generally, the properties that must be hold for CPS are related to safety, liveness, and reachability.
-In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.} bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
-
-In addition, Shoukry {\it et.
-al} \cite{Shoukry} tackled another important problem related to CPSs: security.
+In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
+
+In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to CPSs: security.
 Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements.

### 2017-08-02 paper.tex
added Craig interpolants
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}.
-
+%
 In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. 
-Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents program correctness and/or a system's behavior, and an iteration bound \textit{k}, which limits loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, which is expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
-
+%
 \begin{figure}[h]
@@
 \end{figure}
+
+Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents program correctness and/or a system's behavior, and an iteration bound \textit{k}, which limits loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, which is expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
+The propositional problem associated with SAT-based BMC is formulated by constructing the following formula~\cite{Biere99}:
+%
+\begin{equation}
+\label{bounded-model-checking-biere}
+\psi_{k} = I\left(s_{0}\right) \wedge \bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right) \wedge \neg \phi_{k}
+\end{equation}
+
+\noindent Here, $\phi_{k}$ represents a safety property $\phi$ in step $k$, $I$ is the set of initial states of $M$, $R\left(s_{i},s_{i+1}\right)$ is the transition relation of $M$ at time steps $i$ and $i+1$.
+Hence, the formula $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$.
+$\neg \phi_{k}$ represents the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$.
+Finally, the resulting (bit-vector) formula is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability.
+Formula (\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}.
+Liveness properties (e.g., starvation, deadlock) that contain the LTL operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop.
+In this case, formula~\ref{bounded-model-checking-biere} can be rewritten as:
+
+%
+\begin{equation}
+\psi_{k} = I\left(s_{0}\right) \wedge \bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right) \wedge \left(\bigvee^{k}_{i=0} \neg \phi_{i}\right)
+\end{equation}
+
+\noindent where $\phi_{i}$ is the propositional variable $\phi$ at time step $i$.
+Thus, this formula can be satisfied if and only if for some $i$ ($i \leq k$) there exists a reachable state at time step $i$ in which $\phi$ is violated.
 
@@
 
+%------------------------------------------------------
+\subsection{Craig Interpolation}
+\label{sec:CraigInterpolationInModelChecking}
+%------------------------------------------------------
+
+Another feasible alternative to prove properties in BMC is to compute the Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, i.e., proofs that there is no counter-example of depth less than or equal to $k$.
+This proof does not ensure whether a given property holds in the model, but it contains information about the reachable states of the model.
+
+%\begin{prf}\label{interpolants}
+\begin{definition}\label{interpolants}
+\textit{Given a pair of formulae $\left(A,B\right)$, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
+
+\begin{itemize}
+	\item $A \Rightarrow F$
+	\item $F \wedge B$ \textit{is unsatisfiable}
+	\item $F$ \textit{refers only to the common variables of $A$ and $B$}
+\end{itemize}
+\end{definition}
+%\end{prf}
+
+As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
+Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$ since $x_2$ is also common to both $A$ and $B$.
+
+The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.
+In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define the Formula (\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
+Now suppose that $Q=I$ and we partition Formula (\ref{bounded-model-checking-biere}) so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$ as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.4]{interpolants}
+\caption{Computing image by interpolation~\cite{McMillan03}.}
+\label{figure:computing-image-by-interpolation}
+\end{figure}
+
+Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, i.e., we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
+From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
+According to Definition~\ref{interpolants}, $A$ must imply $F$ and since we defined $A$ to be the set of initial states and the first instance of $R$ (i.e., from Figure~\ref{figure:computing-image-by-interpolation}, $A=s_{0} \wedge s_{1}$), it follows that $F$ is \textit{true} in every state reachable from the initial state in one step.
+In other words, we can say that $F$ is an over-approximation of the forward image of $I$~\cite{McMillan03,McMillan05}.
+Also according to Definition~\ref{interpolants}, the formula $F \wedge B$ must be unsatisfiable (from Figure~\ref{figure:computing-image-by-interpolation}, $B=s_{2} \wedge s_{3} \wedge \ldots \wedge s_{k}$), which means that there is no state satisfying $F$ that can reach a final state $s_k$.
+After computing the interpolant $F$, we then check whether $F$ implies $Q$.
+If $F$ implies $Q$, then no reachable state can satisfy the property $\phi$ and we can thus conclude that the property holds.
+However, since $F$ is an approximation, we can falsely conclude that the final state is reachable.
+In this case, we update $Q = F \vee Q$ and $A = F \wedge R_{0}$, increase the value of $k+1$ and check whether $A \wedge B$ is unsatisfiable.
+If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state).
+Otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
+We stop this procedure when we have found a valid counter-example or have proved that the final state is not reachable (i.e., the property holds).
+The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+
 %--------------------------------------

### 2017-08-03 paper.tex
figures with grey scale
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
minor fix

@@
 % Do not put math or special symbols in the title.
-\title{Automated Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges}
+\title{Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges}
 
@@
 % make the title area
-\maketitle
 
@@
 
+\keywords{Embedded Systems, Cyber-Physical Systems, Formal Methods, Model Checking, Machine Learning}
+
+\maketitle
+
 %--------------------------------------
@@
 	\centering
-	\includegraphics[scale=0.35]{figure1.eps}
+	\includegraphics[scale=0.4]{figure1.jpg}
 	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
@@
 	\centering
-	\includegraphics[scale=0.35]{figure2.eps}
+	\includegraphics[scale=0.5]{figure2.jpg}
 	\caption{Bounded Model Checking.}
@@
 	\centering
-	\includegraphics[scale=0.12]{figure3.eps}
-	\caption{Counter-Example Guided Inductive Synthesis.}
+	\includegraphics[scale=0.5]{figure3.jpg}
+	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
 	\label{Counter-Example-Guided-Inductive-Synthesis}
@@
 	\centering
-	\includegraphics[scale=0.35]{figure4.eps}
+	\includegraphics[scale=0.5]{figure4.jpg}
 	\caption{Verification methodologies for embedded systems.}

### 2017-08-03 paper.tex
replaced CPS by ECPS
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time).
-They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances.
-Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.
+Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time)~\cite{Kopetz11}.
+They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances~\cite{xu15}.
+Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers.
 
 The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs).
-Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).
-
-The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, internet of things - IoT) in the latter.
-This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
-For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
-One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a CPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.
-
-Thus, CPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems.
+Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs).
+The presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
+For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
+One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.
+
+Thus, ECPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems.
 As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
@@
 
-Besides, when physical interaction with the real world is needed, which happens in CPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
+Besides, when physical interaction with the real world is needed, which happens in ECPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
 Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation.
@@
 
-The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by CPSs, IoT, and ML.
-We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
+The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML.
+We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
@@
 
-Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS \textcolor{red}{add citations}.
-A CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
-
-The overall formal and automated synthesis process for CPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the CPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.
+Automated and formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}.
+A ECPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
+
+The overall formal and automated synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.
 
@@
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent CPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible.
-On the other hand, that approach is not suitable for complex CPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
-
-CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.
+On the other hand, that approach is not suitable for complex ECPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
+
+ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.
 
@@
 
-Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for CPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for CPS.
+Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.
 
@@
 
-Given a class of CPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
-Decidability is a key issue in CPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model CPS) are infinite-state machines~\cite{Henzinger95}.
-Generally, the properties that must be hold for CPS are related to safety, liveness, and reachability.
+Given a class of ECPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
+Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPS) are infinite-state machines~\cite{Henzinger95}.
+Generally, the properties that must be hold for ECPS are related to safety, liveness, and reachability.
 In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
 
-In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to CPSs: security.
-Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements.
+In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security.
+Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements.
 The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.

### 2017-08-03 paper.tex
updated abstract
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
-Besides, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of embedded software: cyber-physical systems.
-As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
-The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems.
-Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
+Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPS).
+As a consequence, reliability of ECPS becomes a key issue during system development, which must be carefully addressed and assured.
+Normally, state-of-the-art verification methodologies for ECPS generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with the physical plant, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
+The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPS.
+Reliability issues, in the development of ECPS, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
 \end{abstract}

### 2017-08-03 paper.tex
reviewed introduction
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs).
-The presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
+In particular, the presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.
 For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
@@
 
-Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}.
-Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
+Figure~\ref{intelligent-product} shows some examples of ECPS, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}.
+Indeed, many current ECPS, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences.
-On the one hand, portable embedded systems are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics.
+On the one hand, portable ECPS are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics.
 On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health.
@@
 
-Besides, when physical interaction with the real world is needed, which happens in ECPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
+Besides, when physical interaction with the real world is needed, which happens in ECPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
 Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation.
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
-Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {\it i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
 
@@
 Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may change based on data acquired during previous actions, due to model updating.
-As a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data.
-
-A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
+As a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data, as recently reported in the literature~\cite{DBLP:conf/cav/HuangKWW17,DBLP:conf/cav/KatzBDJK17}, where formal verification techniques are applied to verify deep neural networks.
+
+A number of distinctive characteristics might influence the ECPS verification and synthesis process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy.
 Additionally, increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency)~\cite{Kroening15} or even new structures provided with the goal of ensuring more computational capacity~\cite{cudalucas} must be considered during verification.
-Such observations expand the addressed issues and even complement what was mentioned in the previous paragraphs.
+Such observations expand the addressed issues and even complement what was mentioned previously.
 
 The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML.
-We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
+We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems and Section \ref{Newapp} addresses new applications, which take into account the new trends mentioned here.
+Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPS and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
 Finally, we conclude and describe future work in Section~\ref{conclusions}.

### 2017-08-03 paper.tex
reviewed section 2
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Automated and formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}.
-A ECPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
-
-The overall formal and automated synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.
+Automated formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}.
+
+A ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
+
+The overall automated formal synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS to meet the requirements and behave as desired, which is confirmed by a verification oracle.
 
@@
 
-Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.
+Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.
 
@@
 Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPS) are infinite-state machines~\cite{Henzinger95}.
-Generally, the properties that must be hold for ECPS are related to safety, liveness, and reachability.
+Generally, the properties that must hold for ECPS are related to safety, liveness, and reachability.
 In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

### 2017-08-03 paper.tex
reviewed section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents program correctness and/or a system's behavior, and an iteration bound \textit{k}, which limits loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, which is expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.
-The propositional problem associated with SAT-based BMC is formulated by constructing the following formula~\cite{Biere99}:
+The propositional problem associated with SAT-based BMC is formulated by constructing the following equation~\cite{Biere99}:
 %
@@
 \noindent Here, $\phi_{k}$ represents a safety property $\phi$ in step $k$, $I$ is the set of initial states of $M$, $R\left(s_{i},s_{i+1}\right)$ is the transition relation of $M$ at time steps $i$ and $i+1$.
-Hence, the formula $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$.
+Hence, the equation $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$.
 $\neg \phi_{k}$ represents the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$.
-Finally, the resulting (bit-vector) formula is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability.
-Formula (\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}.
-Liveness properties (e.g., starvation, deadlock) that contain the LTL operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop.
-In this case, formula~\ref{bounded-model-checking-biere} can be rewritten as:
+Finally, the resulting (bit-vector) equation is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability.
+Eq.~(\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}.
+Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop.
+In this case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:
 
@@
 \noindent where $\phi_{i}$ is the propositional variable $\phi$ at time step $i$.
-Thus, this formula can be satisfied if and only if for some $i$ ($i \leq k$) there exists a reachable state at time step $i$ in which $\phi$ is violated.
+Thus, this equation can be satisfied if and only if for some $i$ ($i \leq k$) there exists a reachable state at time step $i$ in which $\phi$ is violated.
 

### 2017-08-03 paper.tex
reviewed section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 The common property in SSA form is that every variable has only one definition in the program text.
-This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left hand side, as shown in Figure~\ref{figure:verification-condition}(b).
+This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
 %
@@
 The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
-%
+In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+
 For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
@@
 This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
-As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.
+As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that\textit{k}-induction verifiers can automatically prove program correctness.
 In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
@@
 
-Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification.
-That happens due to the distinctive characteristics mentioned earlier, which influence embedded system development and also verification processes.
+Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification.
+That happens due to the distinctive characteristics mentioned earlier, which influence ECPS development and also verification processes.
 Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.

### 2017-08-03 paper.tex
reviewed section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Another feasible alternative to prove properties in BMC is to compute the Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
-This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, i.e., proofs that there is no counter-example of depth less than or equal to $k$.
+This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs that there is no counterexample of depth less than or equal to $k$.
 This proof does not ensure whether a given property holds in the model, but it contains information about the reachable states of the model.
@@
 The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.
-In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define the Formula (\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
-Now suppose that $Q=I$ and we partition Formula (\ref{bounded-model-checking-biere}) so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$ as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
+In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
+Now suppose that $Q=I$ and we partition Eq.~(\ref{bounded-model-checking-biere}) so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$ as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
 
@@
 
-Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, i.e., we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
+Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
 From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
@@
 Otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
-We stop this procedure when we have found a valid counter-example or have proved that the final state is not reachable (i.e., the property holds).
+We stop this procedure when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
 The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.

### 2017-08-03 paper.tex
reviewed section 8
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-This paper presented the main challenges related to the verification of design correctness, in embedded systems, and also raised some important side considerations about synthesis.
+This paper presented the main challenges related to the verification of design correctness, in ECPS, and also raised some important side considerations about synthesis.
 In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification and synthesis.
-Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
-extends BMC-based approaches from falsification to verification, can be applied to formally verify correctness and timeliness of embedded systems.
-
-Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
+Additionally, there is little evidence that model checking ECPS using \textit{k}-induction (and invariants), which extends BMC-based approaches from falsification to verification, can be applied to formally verify correctness and timeliness of ECPS.
+
+Given that software complexity has significantly increased in ECPS, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
 There is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
@@
 
-As future work, the main goal of this research is to extend BMC as a design, verification, and synthesis tool for achieving correct-by-construction embedded system implementations.
-Special attention will be given to cyber-physical systems and modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification, probably through behavior models.
+As future work, the main goal of this research is to extend BMC as a design, verification, and synthesis tool for achieving correct-by-construction ECPS implementations.
+Special attention will be given to modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification and synthesis, probably through behavior models.
 Additionally, ML techniques will be incorporated to BMC frameworks, in order to provide flexibility and specific-problem tuning.

### 2017-08-03 paper.tex
reviewed section 7
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
reviewed section 5
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-This position paper tackles seven major problems in computer-aided verification and synthesis for embedded systems, which are (partially) open in current published research.
-
-\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic embedded programs.
+This position paper tackles seven major problems in computer-aided verification and synthesis for ECPS, which are (partially) open in current published research.
+
+\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic ECPS.
 
@@
 
-\textbf{(RP3)} proof of correctness and timeliness of embedded systems, by taking into account stringent constraints imposed by hardware platforms.
+\textbf{(RP3)} proof of correctness and timeliness of ECPS, by taking into account stringent constraints imposed by hardware platforms.
 
@@
 
-\textbf{(RP7)} provision of a unified framework, which is able to tackle high and low level properties, as well as system behavior, with the potential to lead to another degree of completeness, in such a way that every implementation aspect of an embedded system is addressed and evaluated.
+\textbf{(RP7)} provision of a unified framework, which is able to tackle high and low level properties, as well as system behavior, with the potential to lead to another degree of completeness, in such a way that every implementation aspect of ECPS is addressed and evaluated.
 
@@
 al} \cite{babic07} showed that ML can be used for improving SAT solvers regarding huge verification tasks, when applied to parameter tuning~\cite{babic07}.
-Given what was presented, it seems worthwhile to pursue the integration of ML to BMC-based checkers, specifically ESBMC, mainly regarding solver tuning.
+Given what was presented, it seems worthwhile to pursue the integration of ML to BMC-based checkers, mainly regarding solver tuning.
 In addition, ML can also be used for invariant computation and strengthening, which is a major bottleneck in \textit{k}-induction verification based on invariant algorithms.

### 2017-08-03 paper.tex
added SC-square project reference
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes.
+In this particular research direction, the European Project SC$^2$ (Satisfiability Checking and Symbolic Computation: uniting two communities to solve real problems) aims to further extend background theories supported by SMT solvers, in order to solve problems related to security and safety of computer systems (or large mathematical problems) by the development of radically improved software tools~\cite{DBLP:journals/cca/AbrahamA0BBBCDE16}.
 

### 2017-08-03 paper.tex
added a reference about the Linux kernel
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
 One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal to remove redundant behaviour or analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.

### 2017-08-03 paper.tex
minor corrections
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}.
-In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication.
-Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.
+In general, however, there is still a lack of studies to verify system-level properties related to ECPS; emphasis should be given to micro-grids~\cite{xu15}, which require high-dependability requirements for computation, control, and communication.
+Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for ECPS engineers~\cite{Alves15}.
 
@@
 \textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java).
-Given the availability of different programming languages for writing code for embedded systems, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications.
+Given the availability of different programming languages for writing code for ECPS, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications.
 The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated.
@@
 The increasing number of verification tasks being added every year to this competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques.
-As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and further increase the reliability of verified (embedded) systems.
+As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and further increase the reliability of verified ECPS.
 

### 2017-08-07 paper.tex
addition to section 1

@@
 \begin{abstract}
-The dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
+Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
@@
 Normally, state-of-the-art verification methodologies for ECPS generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
-Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with the physical plant, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
+Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
 The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPS.
@@
 
-Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time)~\cite{Kopetz11}.
-They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances~\cite{xu15}.
+Generally, embedded computer systems perform dedicated functions with high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time)~\cite{Kopetz11}.
+They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home duties~\cite{xu15}.
 Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers.
@@
 Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models.
-Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {\it i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
-
-For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected, for instance.
+Specifically, such models can be used to evaluate and even synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {\it i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.
+
+For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services ({\it e.g.,} exported symbols and scheduling); however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if \textit{read} and \textit{write} operations are performed according to what is specified in the device's datasheet and setup periods are respected, for instance.
 In summary, the main goal would be to ensure that the developed driver is compliant with the underlying hardware, which goes further than code correctness and also tackles interaction with surrounding entities.
@@
 Additionally, increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, interfacing, and concurrency)~\cite{Kroening15} or even new structures provided with the goal of ensuring more computational capacity~\cite{cudalucas} must be considered during verification.
-Such observations expand the addressed issues and even complement what was mentioned previously.
+Such observations expand the addressed issues and even complement what was previously mentioned.
 
@@
 Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPS and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
-Finally, we conclude and describe future work in Section~\ref{conclusions}.
+Finally, we conclude and describe future work, in Section~\ref{conclusions}.
 

### 2017-08-07 paper.tex
additions to section 2

@@
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
-Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPS).
-As a consequence, reliability of ECPS becomes a key issue during system development, which must be carefully addressed and assured.
-Normally, state-of-the-art verification methodologies for ECPS generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
+Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPSs).
+As a consequence, reliability of ECPSs becomes a key issue during system development, which must be carefully addressed and assured.
+Normally, state-of-the-art verification methodologies for ECPSs generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation.
 Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
-The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPS.
-Reliability issues, in the development of ECPS, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
+The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs.
+Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
 \end{abstract}
@@
 For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}.
-One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.
+One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute an ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.
 
@@
 
-Figure~\ref{intelligent-product} shows some examples of ECPS, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}.
-Indeed, many current ECPS, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
+Figure~\ref{intelligent-product} shows some examples of ECPSs, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}.
+Indeed, many current ECPSs, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
 For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences.
-On the one hand, portable ECPS are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics.
+On the one hand, portable ECPSs are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics.
 On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health.
@@
 
-Besides, when physical interaction with the real world is needed, which happens in ECPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
+Besides, when physical interaction with the real world is needed, which happens in ECPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving.
 Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation.
@@
 The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML.
-We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
+We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPS and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
+Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPSs and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
 Finally, we conclude and describe future work, in Section~\ref{conclusions}.
@@
 
-Automated formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}.
-
-A ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
-
-The overall automated formal synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS to meet the requirements and behave as desired, which is confirmed by a verification oracle.
+Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis for ECPSs and HSs \textcolor{red}{add citations}.
+
+An ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
+
+The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS's behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
 
@@
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they should be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 On the other hand, that approach is not suitable for complex ECPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.
-
-ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.
-
-Timed automata employ finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories with constant slope.
-Hybrid automata generalize timed automata, also employing finite automata with real variables whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
-Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
-
-Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.
+Indeed, ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.
+
+Timed automata targets real-time systems and employs finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories (behaviors) with constant slope.
+Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
+Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
+
+Finally, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.
 
@@
 
-Given a class of ECPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
-Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPS) are infinite-state machines~\cite{Henzinger95}.
-Generally, the properties that must hold for ECPS are related to safety, liveness, and reachability.
-In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
+Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
+Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}.
+Generally, the properties that must hold for ECPSs are related to safety, liveness, and reachability.
+In particular, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
 

### 2017-08-07 paper.tex
first additions to section 3

@@
 
-Note that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of expressions.
+One may notice that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of expressions.
 A verification condition is a logical formula (constructed from the bounded program and desired correctness properties) whose validity implies that the program's behaviour agrees with its specification~\cite{Bradley07}.
@@
 As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b).
-One may notice that the SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code.
+The SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code.
 The common property in SSA form is that every variable has only one definition in the program text.
@@
 One may also notice that BMC still encodes program stages with a size that grows linearly with $N$.
-More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
+More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
 %
@@
 In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
-Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).
+Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).
 

### 2017-08-08 paper.tex
additional changes to section 3

@@
 
-One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}.
-The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs.
-In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.
+One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}; however, the main challenge regarding such an approach relies on computing and strengthening inductive invariants from programs.
+In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine invariance of a non-inductive assertion~\cite{Bradley07}.
 
@@
 If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
-As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that\textit{k}-induction verifiers can automatically prove program correctness.
+Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
+As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness.
 In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
@@
 
-Another feasible alternative to prove properties in BMC is to compute the Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
-This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs that there is no counterexample of depth less than or equal to $k$.
-This proof does not ensure whether a given property holds in the model, but it contains information about the reachable states of the model.
+Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs regarding the nonexistence of counterexamples of depth less than or equal to $k$.
+Those proofs do not ensure whether a given property holds in the model, but they contain information about reachable states of a model.
 
@@
 \begin{definition}\label{interpolants}
-\textit{Given a pair of formulae $\left(A,B\right)$, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
+\textit{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
 
 \begin{itemize}
-	\item $A \Rightarrow F$
-	\item $F \wedge B$ \textit{is unsatisfiable}
-	\item $F$ \textit{refers only to the common variables of $A$ and $B$}
+	\item $A \Rightarrow F$;
+	\item $F \wedge B$ \textit{is unsatisfiable};
+	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
 \end{itemize}
@@
 As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
-Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$ since $x_2$ is also common to both $A$ and $B$.
+Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$, since $x_2$ is a common part of both $A$ and $B$.
 
@@
 In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
-Now suppose that $Q=I$ and we partition Eq.~(\ref{bounded-model-checking-biere}) so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$ as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
+Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
 
@@
 Also according to Definition~\ref{interpolants}, the formula $F \wedge B$ must be unsatisfiable (from Figure~\ref{figure:computing-image-by-interpolation}, $B=s_{2} \wedge s_{3} \wedge \ldots \wedge s_{k}$), which means that there is no state satisfying $F$ that can reach a final state $s_k$.
-After computing the interpolant $F$, we then check whether $F$ implies $Q$.
-If $F$ implies $Q$, then no reachable state can satisfy the property $\phi$ and we can thus conclude that the property holds.
-However, since $F$ is an approximation, we can falsely conclude that the final state is reachable.
+After computing $F$, we then check whether $F$ implies $Q$.
+If $F$ implies $Q$, then no reachable state can satisfy the property $\phi$ and we can thus conclude that the property holds; however, since $F$ is an approximation, we can falsely conclude that the final state is reachable.
 In this case, we update $Q = F \vee Q$ and $A = F \wedge R_{0}$, increase the value of $k+1$ and check whether $A \wedge B$ is unsatisfiable.
-If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state).
-Otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
-We stop this procedure when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
+If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state); otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
+This procedure is stopped when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
 The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
@@
 In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms), and other implementation restrictions, in order to explore the design space.
-Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which may even be incorporated into a given system development procedure, by providing both system verification and model refinement.
+Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which can even be incorporated into a given system development procedure, by providing both system verification and model refinement.
 
@@
 For instance, regarding self-driving cars, an important property could be the detection of pedestrians, animals, bicycles, or any obstacle present on urban and rural roads.
-Indeed, for instance, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
+Indeed, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
 In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches \cite{BMCml}.

### 2017-08-08 paper.tex
additional paragraph to section 3.4

@@
 Indeed, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}.
-In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches \cite{BMCml}.
+In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised, incorporated to BMC approaches \cite{BMCml}, and directly applied to ECPS verification.
+
+In summary, current approaches already tackle ECPS characteristics ({\it e.g.}, interaction with real world entities and digital processing), which are included in verification and synthesis methodologies.
+Finally, the final complement may come from ML techniques, which have the potential to fill gaps ({\it e.g.}, lack of complete models including nondeterminism), bridge tools and models ({\it e.g.}, analog and digital domains in ECPSs and verification schemes), and provide model and methodology evolution.
 

### 2017-08-08 paper.tex
addition to section 4

@@
 
-Generally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use some assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
+Generally, state-of-the-art verification methodologies for embedded systems (including ECPSs) generate test vectors (with constraints) and use some assertion-based verification and high-level processor models, during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}.
 %
@@
 The fourth aspect deals with code that relies on existing infrastructures and must be compliant with those, such as Linux kernel modules.
-The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices.
-In particular, software developed for that type of embedded systems has been extensively tested and verified, and also optimized for efficiency over years of development.
+The fifth aspect, in turn, is inherent to a large number of embedded systems from  telecommunications, control systems, and medical devices, including ECPSs.
+In particular, software developed for those types of embedded system has been extensively tested and verified, and also optimized for efficiency over years of development.
 Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.

### 2017-08-08 paper.tex
final additions

@@
 One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes.
-In this particular research direction, the European Project SC$^2$ (Satisfiability Checking and Symbolic Computation: uniting two communities to solve real problems) aims to further extend background theories supported by SMT solvers, in order to solve problems related to security and safety of computer systems (or large mathematical problems) by the development of radically improved software tools~\cite{DBLP:journals/cca/AbrahamA0BBBCDE16}.
-
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
-Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}; however, the classes of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
-One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal to remove redundant behaviour or analyze interpolants~\cite{McMillan11} to prove non-interference of context switches.
-
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge remains still open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+In this particular research direction, the European Project SC$^2$ (Satisfiability Checking and Symbolic Computation: uniting two communities to solve real problems) aims to further extend background theories supported by SMT solvers, in order to solve problems related to security and safety of computer systems (or large mathematical problems), through the development of radically improved software tools~\cite{DBLP:journals/cca/AbrahamA0BBBCDE16}.
+
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
+Recent advances for verifying multi-threaded C programs have been proposed to speed up verification times, which significantly prune state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution, based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}, were made available; however, the set of concurrent-program classes ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
+One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal of removing redundant behaviour or analyzing interpolants~\cite{McMillan11} and then prove non-interference of context switches.
+
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge remains still open, {\it i.e.}, to compute and strengthen loop invariants for proving program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
@@
 Some work towards that was already presented by Sousa, Cordeiro, and Filho~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics.
-In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked.
+In summary, the BMC core (in that case, ESBMC) is not changed (it is still C/C++ code), but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked.
 This approach could adopt the clang compiler~\cite{Lopes:2014:GSL:2692607}, which is already widely used in industry~\cite{clang-usage}, in order to produce an Abstract Syntax Tree (AST) for a given programming language ({\it e.g.}, C/C++/ObjectiveC/ObjectiveC++).
 This AST can then be converted either to an intermediate representation (IR) of the underlying verifier or to LLVM bitcode that is already produced by clang.
+Obviously, a completely new language may present new structures and resources; however, if different inputs are parsed and converted into a form that preserves every aspect of their structure, that can then be analyzed and verified in an unified way.
 Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
@@
 The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated.
-In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification.
+In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main objective: formal program verification.
 Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics.
@@
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
-Nonetheless, the elapsed execution time is still a bottleneck for some applications, such as changes in UAV trajectories, but that can be improved, through the use of simplification techniques and problem constraints.
+Nonetheless, elapsed execution times are still a bottleneck for some applications, such as changes in UAV trajectories, but that can be improved, through the use of simplification techniques and problem constraints.
 
@@
 This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignment procedures, until a counterexample is not output.
-It is worth noticing that the timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{Araujo16}, which can be seen as an advantage for such an implementation.
+It is worth noticing that associated timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{Araujo16}, which can be seen as an advantage for such an implementation.
 
@@
 Besides a correct software implementation, {\it i.e.}, source code, which is already covered by formal methods already available in literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
-This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals under a variety of conditions, unifying source code and behavior verification.
+This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals, under a variety of conditions, and then unifying source code and behavior verification.
 In addition, that kind of implementation usually employs digital filters with specific parameters, which could also be tackled.
+Indeed, communication aspects of ECPSs could be tackled, as a module in system verification procedures.
 

### 2017-08-10 paper.tex
Section 2

@@
 Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
-
-Finally, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.
+Recently, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.
+In addition to HA and TA based models, there are a number of other ECPS modeling initiatives based on (exact or approximate) bisimulation~\cite{Girard11}, simulation~\cite{Rungger13} and interval analysis~\cite{Li16}.
+
+Another approach to ECPS modeling concerns quantized control systems and finite-word length (FWL) effects on the performance of these systems.
+Differently from bisimulations, quantized models are built from a countable subset of the (quantized) input space~\cite{Tabuada07}.
+In particular, Keel and Bhattacharyya showed that even the robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, the FWL representation of the digital controllers).
+Therefore, various works tried link sampled-data theory to hybrid systems models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.
+
+Finally, in addition to the modeling of the ECPS dynamics itself, it is often interesting to establish invariants on the state space of the ECPS, such information can be used in order to accelerate formal verification and synthesis processes for these systems.
+These invariants might be synthesized in order to point the boundaries of state-space~\cite{BenSassi12}, determine stability~\cite{Paul14}, or reachability of ECPS~\cite{Li16_2,Li16}.
 
@@
 Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}.
-Generally, the properties that must hold for ECPSs are related to safety, liveness, and reachability.
-In particular, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
+In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}.
+An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}.
+An example of barrier certificate is the stability, that ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside of this region.
+The stability may be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.
+
+Generally, the properties that must hold for ECPSs are related to safety/reachability, and liveness.
+Therefore, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking (BMC)~\cite{Veanes09}.
+
+Along with the CPS modeling and verification techniques purely based on HA, appeared logic-based approaches~\cite{Phan15} that may still use HA, but it is not based on a algebraic description of states and transitions.
+Some examples of logic-based verification are the theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among them the BMC stands out~\cite{handbook09}.
+
+%Much of the work related to the formal verification of ECPS seeks to ensure the safety, robustness, and reachability properties of an HA. 
+Recently, some of these works addressed the ECPS formal verification considering the discretization, switching and quantization effects.
+Dugirala and Viswanathan~\cite{Duggirala2015} verified embedded control softwares considering real time issues, e.g., delayed responses and possible loss of real time deadlines.
+Anta {\it et al.}~\cite{Anta10} presented a tool called Costan that finds errors in the implementation of a mathematical model and verifies whether the error is tolerated, considering the quantization effects and the fixed-point implementation focusing on the quantization error effects over stability of the system.
+Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} presets the DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends in order to provide support for digital system design and verification considering FWL effects.
+In particular, DSVerifier verifies if a designed digital controller and/or closed-loop ECPS presents the desired performance when it is implemented in a FWL format~\cite{Bessa16,Bessa17}.
+
+Finally, it is worth to mention that there are some other alternatives for modeling and analysis of ECPS, that are not based on verification, but in simulators~\cite{Phan2014,Nakajima12,Junjie12}.
+These approaches fails to ensure that some properties are hold but it presents a good scalability, what is ideal for large scale and complex systems that are hardly modeled through the automata- or logig-based approaches.
+Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for CPS.
+In order to achieve a trade off between the scalability of CPS simulators and the correctness of verification methods, the statistical model checking have been employed to formally verify ECPS~\cite{Clarke11}.
 
@@
 The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
-Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+The cyber-security is also a concern of Fiore {\it et al.} that aims to ensure the secure state estimation of UAV systems under sparse attacks~\cite{Fiore17}.
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 
@@
 %--------------------------------------
+
+The control synthesis for ECPS consists of finding an controller or an invariant control law $\mathcal{C}$, such that a closed-loop system composed of $(\mathcal{C},\Omega)$ satisfies every property $\mathcal{L}\in \Phi$.
+
+An efficient way to synthesize reliable ECPS is the Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}.
+The key idea is the definition (by the user) of the safe states set, {\it i.e.} the specification of $\Phi$, as well as a generic template for a control system that must satisfy these specification; the the CEGIS engine compute values for the open values in the template that hold $\Phi$~\cite{RienerKFB16}.
+Abate~{\it et al.}~\cite{Abate17,abatecav2017} presents a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}.
+
+% The mentioned authors developed a tool called DSSynth, which is able to synthesize digital controllers that are algorithmically and numerically sound, w.r.t.
+stability.
+DSSynth marks the first use of counterexample-guided inductive synthesis~\cite{sketch} to synthesize digital controllers, considering physical plants with uncertain models and FWL effects; however, low-level implementation errors ({\it e.g.}, limit cycles) are not further investigated.
 
@@
 
-\begin{thebibliography}{1}
+%\begin{thebibliography}{1}
 
@@
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2017-08-11 paper.tex
corrections

--- 2017-08-08 paper.tex
@@
 Timed automata targets real-time systems and employs finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories (behaviors) with constant slope.
-Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
+Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, but whose trajectories follow more general dynamical laws~\cite{Henzinger95}.
 Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.
-
-Finally, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.
+Recently, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.
+In addition to HA and TA based models, there are a number of other ECPS modeling initiatives based on (exact or approximate) bisimulation~\cite{Girard11}, simulation~\cite{Rungger13} and interval analysis~\cite{Li16}.
+
+Another approach to ECPS modeling tackles quantized control systems and finite-word length (FWL) effects on the performance of those systems.
+Differently from bisimulation, quantized models are built from a countable subset of the (quantized) input space~\cite{Tabuada07}.
+In particular, Keel and Bhattacharyya showed that even robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, FWL representations of digital controllers).
+Therefore, various studies tried link sampled-data theory to hybrid-system models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.
+
+Finally, in addition to ECPS-dynamics modeling, it is often interesting to establish invariants on a ECPS's state space, which can be used for accelerating formal verification and synthesis processes regarding those systems.
+Such invariants might be synthesized, in order to point out state-space boundaries~\cite{BenSassi12} and determine ECPS stability~\cite{Paul14} or reachability~\cite{Li16_2,Li16}.
 
@@
 Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}.
-Generally, the properties that must hold for ECPSs are related to safety, liveness, and reachability.
-In particular, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
+In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}.
+An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}.
+An example of barrier certificate is the stability, that ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside of this region.
+The stability may be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.
+
+Generally, properties that must hold for ECPSs are related to safety/reachability and liveness.
+Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.
+
+Along with CPS modeling and verification techniques purely based on HA, one can also highlight logic-based approaches~\cite{Phan15} that may still use HA, but are not based on a algebraic description of states and transitions.
+Some examples of logic-based verification are theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among which BMC stands out~\cite{handbook09}.
+
+%Much of the work related to the formal verification of ECPS seeks to ensure the safety, robustness, and reachability properties of an HA. 
+Recently, some studies have addressed ECPS formal verification considering discretization, switching, and quantization effects.
+Dugirala and Viswanathan~\cite{Duggirala2015} verified embedded control software considering real time issues, {\it e.g.}, delayed responses and possible loss of real time deadlines.
+Anta {\it et al.}~\cite{Anta10} presented a tool called Costan, which finds errors in mathematical-model implementations and verifies whether error is tolerated, considering quantization effects and fixed-point implementation, while focusing on quantization error effects over system stability.
+Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} preseted DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends, in order to provide support for digital system design and verification considering FWL effects.
+In particular, DSVerifier checks if a designed digital controller and/or closed-loop ECPS presents the desired performance, when it is implemented with a FWL format~\cite{Bessa16,Bessa17}.
+
+Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}.
+Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems that are hardly modeled through automata- or logig-based approaches.
+Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs.
+In order to achieve a balanced trade-off between ECPS-simulation scalability and correctness of verification methods, statistical model checking have been employed to formally verify ECPSs~\cite{Clarke11}.
 
@@
 Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements.
-The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
-Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+The mentioned authors proposed a methodology for estimating physical-system state, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands.
+Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}.
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 
@@
 %--------------------------------------
+
+Control synthesis for ECPSs consists in finding a controller or an invariant control law $\mathcal{C}$, such that a closed-loop system composed of $(\mathcal{C},\Omega)$ satisfies every property $\mathcal{L}\in \Phi$.
+
+An efficient way to synthesize reliable ECPSs is the Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}.
+Its key idea is the definition (by a user) of a set of safe states, {\it i.e.}, the specification of $\Phi$, as well as a generic template for a control system that must satisfy such a specification.
+The, a CEGIS engine compute values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
+Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}.
+
+% The mentioned authors developed a tool called DSSynth, which is able to synthesize digital controllers that are algorithmically and numerically sound, w.r.t.
+stability.
+DSSynth marks the first use of counterexample-guided inductive synthesis~\cite{sketch} to synthesize digital controllers, considering physical plants with uncertain models and FWL effects; however, low-level implementation errors ({\it e.g.}, limit cycles) are not further investigated.
 

### 2017-08-13 paper.tex
ECPS section

@@
 	\centering
-	\includegraphics[scale=0.4]{figure1.jpg}
+	\includegraphics[width=\textwidth]{figure1.jpg}
 	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
@@
 Generally, properties that must hold for ECPSs are related to safety/reachability and liveness.
+The earliest works on HS and ECPS proposed approaches based on CTL and LTL~\cite{Alur93} to specify properties and verify these systems.
+In the last decade, the Signal Temporal Logic (STL)~\cite{Maler04,Donze12} was proposed and have been used for this purpose, providing more suitable operators for dynamical systems and signal processing.
 Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.
@@
 The, a CEGIS engine compute values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
-Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}.
+Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}; the controllers must ensure non-fragile stability~\cite{Abate17} and safety~\cite{abatecav2017}.
+Ravanbakhsh and Sankaranarayanan~\cite{Ravanbakhsh15,Ravanbakhsh16} also employed CEGIS based on Lyapunov function to ensure the robust reach-while-stay property ({\it i.e.}, a safety specification)~\cite{Ravanbakhsh16} and non-zeno behavior~\cite{Ravanbakhsh15}.
+The CEGIS was also employed to synthesize model predictive controllers (MPC) for a wide varieties of applications, {\it e.g.} heating ventilation and air conditioning (HVAC) systems, autonomous vehicles control, and aircraft electric power system~\cite{Ghosh05,Rahman13}.
+
+CEGIS is an important, but not the unique, example of symbolic control synthesis for ECPS.
+The main advantage of using symbolic synthesis techniques ({\it e.g.}, SMT- and SAT-based) is the possibility of encoding (via logical formula) reachability/safety specifications on the original concrete systems.
+The references~\cite{Rungger13,AydinGol14,Holub16,Tabuada08,Zamani17,Zamani14,Zamani13,Zamani12,Alkhatib17,Lesser15,Girard13} are examples of works on symbolic control synthesis for ECPS.
+
+A common procedure for controller synthesis for ECPS is based on finite-state approximations of infinite-states systems, {\it i.e.}, the HA.
+This approach allows the fully automated synthesis of ECPS control systems that ensure the achievement of complex specifications.
+The weakness of this approach is that it often produces complex controllers with high implementation cost.
+To avoid this problem, refined controllers have been proposed aiming to provide some invariant properties for close-loop systems, {\it e.g.}: safety properties~\cite{Dallal17,Dallal13,Girard13}, reachability properties~\cite{Alkhatib17,Habets06}, stability~\cite{Zamani12,Tabuada08}, and safety/reachability properties considering the quantization effects~\cite{Reissig17}.
 
@@
 	\centering
-	\includegraphics[scale=0.5]{figure2.jpg}
+	\includegraphics[width=\textwidth]{figure2.jpg}
 	\caption{Bounded Model Checking.}
@@
 \centering
-\includegraphics[scale=0.4]{interpolants}
+\includegraphics[width=\textwidth]{interpolants}
 \caption{Computing image by interpolation~\cite{McMillan03}.}
@@
 	\centering
-	\includegraphics[scale=0.5]{figure3.jpg}
+	\includegraphics[width=\textwidth]{figure3.jpg}
 	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
@@
 	\centering
-	\includegraphics[scale=0.5]{figure4.jpg}
+	\includegraphics[width=\textwidth]{figure4.jpg}
 	\caption{Verification methodologies for embedded systems.}
@@
 In addition, ML can also be used for invariant computation and strengthening, which is a major bottleneck in \textit{k}-induction verification based on invariant algorithms.
+Seshia {\it et al.}~\cite{Seshia17} showed that control syntheses based on ML is also a trend, since it has been applied to different applications and model classes~\cite{Jha10,Sadigh14}.
 

### 2017-08-14 paper.tex
minor corrections

--- 2017-08-11 paper.tex
@@
 
-Generally, properties that must hold for ECPSs are related to safety/reachability and liveness.
+Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest works on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems.
+In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing.
 Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.
@@
 
-An efficient way to synthesize reliable ECPSs is the Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}.
+An efficient way to synthesize reliable ECPSs is Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}.
 Its key idea is the definition (by a user) of a set of safe states, {\it i.e.}, the specification of $\Phi$, as well as a generic template for a control system that must satisfy such a specification.
-The, a CEGIS engine compute values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
-Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}.
+Then, a CEGIS engine compute values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
+Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}, and the resulting systems must ensure non-fragile stability~\cite{Abate17} and safety~\cite{abatecav2017}.
+Ravanbakhsh and Sankaranarayanan~\cite{Ravanbakhsh15,Ravanbakhsh16} also employed CEGIS based on the Lyapunov's function to ensure the robust reach-while-stay property ({\it i.e.}, a safety specification)~\cite{Ravanbakhsh16} and non-zeno behavior~\cite{Ravanbakhsh15}.
+CEGIS was also employed to synthesize model predictive controllers for a wide varieties of applications, {\it e.g.} heating ventilation and air conditioning systems, autonomous vehicles control, and aircraft electric power system~\cite{Ghosh05,Rahman13}.
+
+CEGIS is an important, but not the unique, example of symbolic control synthesis for ECPSs.
+The main advantage of using symbolic synthesis techniques ({\it e.g.}, SMT- and SAT-based) is the possibility of encoding (via logic formulae) reachability/safety specifications on original concrete systems.
+Indeed, there are many studies~\cite{Rungger13,AydinGol14,Holub16,Tabuada08,Zamani17,Zamani14,Zamani13,Zamani12,Alkhatib17,Lesser15,Girard13} on symbolic control synthesis for ECPS, available in literature.
+
+A common procedure for controller synthesis for ECPS is based on finite-state approximations of infinite-states systems, {\it i.e.}, the HA.
+Such an approach allows the fully automated synthesis of ECPS control systems that ensure achievement of complex specifications; however, the main weakness of this approach is that it often produces complex controllers with high implementation cost.
+In order to avoid that problem, refined controllers have been proposed, which aim to provide some invariant properties for close-loop systems, {\it e.g.}: safety~\cite{Dallal17,Dallal13,Girard13}, reachability~\cite{Alkhatib17,Habets06}, stability~\cite{Zamani12,Tabuada08}, and safety/reachability properties, considering quantization effects~\cite{Reissig17}.
 
@@
 Phuc \cite{nguyen}, in turn, applied decision tree C4.5 and multilayer perceptron to annotated programs (metrics related to structure and semantics) and tackled the generation of test sets satisfying the desired constraints.
-Finally, Bridge studied the automation of heuristic selection based on features, regarding first-order logic theorem proving, and Hutter {\it et.
+Bridge studied the automation of heuristic selection based on features, regarding first-order logic theorem proving, and Hutter {\it et.
 al} \cite{babic07} showed that ML can be used for improving SAT solvers regarding huge verification tasks, when applied to parameter tuning~\cite{babic07}.
-Given what was presented, it seems worthwhile to pursue the integration of ML to BMC-based checkers, mainly regarding solver tuning.
-In addition, ML can also be used for invariant computation and strengthening, which is a major bottleneck in \textit{k}-induction verification based on invariant algorithms.
+Seshia {\it et al.}~\cite{Seshia17}, in turn, showed that control syntheses based on ML is also a trend, since it has been applied to different applications and model classes~\cite{Jha10,Sadigh14}.
+Finally, ML can also be used for invariant computation and strengthening, which is a major bottleneck in \textit{k}-induction verification based on invariant algorithms.
+As a consequence of what was presented, it seems worthwhile to pursue integration of ML into BMC-based checkers, given that it has the potential to tackle many open problems or at least enhance existing solutions.
 

### 2017-08-15 paper.tex
minor fix

@@
 
-Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis for ECPSs and HSs \textcolor{red}{add citations}.
+Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.
 

### 2017-08-16 paper.tex
removed section about interpolants
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML.
-We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
+We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
 In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
@@
 %------------------------------------------------------
-\subsection{Craig Interpolation}
-\label{sec:CraigInterpolationInModelChecking}
+%\subsection{Craig Interpolation}
+%\label{sec:CraigInterpolationInModelChecking}
 %------------------------------------------------------
 
-Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+%Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
 This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs regarding the nonexistence of counterexamples of depth less than or equal to $k$.
@@
 %\begin{prf}\label{interpolants}
-\begin{definition}\label{interpolants}
-\textit{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
-
-\begin{itemize}
-	\item $A \Rightarrow F$;
-	\item $F \wedge B$ \textit{is unsatisfiable};
-	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
-\end{itemize}
-\end{definition}
+%\begin{definition}\label{interpolants}
+%\textit{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
+
+%\begin{itemize}
+%	\item $A \Rightarrow F$;
+%	\item $F \wedge B$ \textit{is unsatisfiable};
+%	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
+%\end{itemize}
+%\end{definition}
 %\end{prf}
 
-As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
+%As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
 Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$, since $x_2$ is a common part of both $A$ and $B$.
 
-The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.
-In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
+%The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers. 
+%
+%In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
 Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\textwidth]{interpolants}
-\caption{Computing image by interpolation~\cite{McMillan03}.}
-\label{figure:computing-image-by-interpolation}
-\end{figure}
-
-Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
+%
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=\textwidth]{interpolants}
+%\caption{Computing image by interpolation~\cite{McMillan03}.}
+%\label{figure:computing-image-by-interpolation}
+%\end{figure}
+
+%Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
 From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
@@
 This procedure is stopped when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
-The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+%The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
 

### 2017-08-16 paper.tex
added citation
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs.
-The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs.
+The present survey discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs.
 Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
@@
 The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs).
-Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs).
+Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs)~\cite{leeCPS}.
 In particular, the presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain.

### 2017-08-16 paper.tex
fixed typos
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS's behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
+The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
 
@@
 In particular, Keel and Bhattacharyya showed that even robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, FWL representations of digital controllers).
-Therefore, various studies tried link sampled-data theory to hybrid-system models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.
+Therefore, various studies tried to connect sampled-data theory to hybrid-system models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.
 
@@
 An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}.
-An example of barrier certificate is the stability, that ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside of this region.
-The stability may be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.
-
-Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest works on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems.
+An example of barrier certificate is the stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region.
+The stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.
+
+Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest studies on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems.
 In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing.
@@
 Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}.
-Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems that are hardly modeled through automata- or logig-based approaches.
+Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems, which are hardly modeled through automata- or logig-based approaches.
 Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs.
@@
 Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}.
-Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
-
-%--------------------------------------
-\subsection{Control synthesis of Cyber-physical systems}
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+
+%--------------------------------------
+\subsection{Control Synthesis of Cyber-physical Systems}
 \label{ssec:synthesis}

### 2017-08-17 paper.tex
fixed typos and added references
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-An efficient way to synthesize reliable ECPSs is Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}.
+An possible approach to synthesize reliable ECPSs is Counterexample Guided Inductive Synthesis (CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}.
 Its key idea is the definition (by a user) of a set of safe states, {\it i.e.}, the specification of $\Phi$, as well as a generic template for a control system that must satisfy such a specification.
-Then, a CEGIS engine compute values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
+Then, a CEGIS engine computes values for the unknown symbols of the chosen template, which hold $\Phi$~\cite{RienerKFB16}.
 Abate~{\it et al.}~\cite{Abate17,abatecav2017} presented a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}, and the resulting systems must ensure non-fragile stability~\cite{Abate17} and safety~\cite{abatecav2017}.
 Ravanbakhsh and Sankaranarayanan~\cite{Ravanbakhsh15,Ravanbakhsh16} also employed CEGIS based on the Lyapunov's function to ensure the robust reach-while-stay property ({\it i.e.}, a safety specification)~\cite{Ravanbakhsh16} and non-zeno behavior~\cite{Ravanbakhsh15}.
-CEGIS was also employed to synthesize model predictive controllers for a wide varieties of applications, {\it e.g.} heating ventilation and air conditioning systems, autonomous vehicles control, and aircraft electric power system~\cite{Ghosh05,Rahman13}.
+CEGIS was also employed to synthesize model predictive controllers for a wide varieties of applications, {\it e.g.}, heating ventilation and air conditioning systems, autonomous vehicles control, and aircraft electric power system~\cite{Ghosh05,Rahman13}.
 
@@
 The main advantage of using symbolic synthesis techniques ({\it e.g.}, SMT- and SAT-based) is the possibility of encoding (via logic formulae) reachability/safety specifications on original concrete systems.
-Indeed, there are many studies~\cite{Rungger13,AydinGol14,Holub16,Tabuada08,Zamani17,Zamani14,Zamani13,Zamani12,Alkhatib17,Lesser15,Girard13} on symbolic control synthesis for ECPS, available in literature.
+Indeed, there are many studies~\cite{Rungger13,AydinGol14,Holub16,Tabuada08,Zamani17,Zamani14,Zamani13,Zamani12,Alkhatib17,Lesser15,Girard13} on symbolic control synthesis for ECPS, available in the literature.
 
 A common procedure for controller synthesis for ECPS is based on finite-state approximations of infinite-states systems, {\it i.e.}, the HA.
-Such an approach allows the fully automated synthesis of ECPS control systems that ensure achievement of complex specifications; however, the main weakness of this approach is that it often produces complex controllers with high implementation cost.
-In order to avoid that problem, refined controllers have been proposed, which aim to provide some invariant properties for close-loop systems, {\it e.g.}: safety~\cite{Dallal17,Dallal13,Girard13}, reachability~\cite{Alkhatib17,Habets06}, stability~\cite{Zamani12,Tabuada08}, and safety/reachability properties, considering quantization effects~\cite{Reissig17}.
+Such an approach allows the fully automated synthesis of ECPS control systems, which ensure complex specifications realization; however, the main weakness of this approach is that it often produces complex controllers with high implementation cost.
+In order to avoid that problem, refined controllers have been proposed, which aim to provide some invariant properties for close-loop systems, {\it e.g.}, safety~\cite{Dallal17,Dallal13,Girard13}, reachability~\cite{Alkhatib17,Habets06}, stability~\cite{Zamani12,Tabuada08}, and safety/reachability properties, considering quantization effects~\cite{Reissig17}.
 

### 2017-08-17 paper.tex
reviewed section 3.1
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}.
+Bounded Model Checking (BMC) based on SAT was originally proposed to verify hardware designs~\cite{Biere99,handbook09}.
 Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers.
@@
 Eq.~(\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}.
-Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop.
+Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop~\cite{JMorse15}.
 In this case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:
@@
 In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver.
-Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).
+Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}.
 
@@
 As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness.
-In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
+In this specific example, an inductive invariant would be $x>1$, given that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
 %
@@
 
-Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification.
+Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification and synhesis.
 That happens due to the distinctive characteristics mentioned earlier, which influence ECPS development and also verification processes.
-Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
+Additionally, there is still a lack of studies for software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
 

### 2017-08-17 paper.tex
reviewed section 3.2
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$.
-In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$, as starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}.
+In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$, as starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{David15,Abate17}.
 As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely.
@@
 Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$).
-If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the phase {\sc verify}, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
+If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the phase {\sc verify}, in order to check whether it is a proper solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$).
 If this is the case, then the algorithm terminates; otherwise, additional information is provided to the phase {\sc synthesize}, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again.
@@
 One may notice that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.
-Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
-In order to avoid exploring all possible values, machine learning techniques can be used in the phase {\sc Synthesize}, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
+Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate over a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$.
+In order to avoid exploring all possible values, machine learning techniques can be used in the phase {\sc Synthesize}, with the goal of learning from experience (input-output samples), {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
 

### 2017-08-17 paper.tex
reviewed section 3.3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects.
+Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering FWL effects.
 Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.

### 2017-08-17 paper.tex
reviewed section 3.3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants~\cite{Abate17}.
+In this respect, DSVerifier has been extended to automatically synthesize digital stabilizing controllers for continuous plants represented as transfer functions or state-space equations~\cite{Abate17,abatecav2017}.
 In particular, a CEGIS-based approach, implemented in a tool called digital system synthesizer (DSSynth), uses inductive synthesis in conjunction with the DSVerifier's algorithm for verifying robust closed-loop stability that addresses plant variations as interval sets, as well as FWL uncertainties in digital controllers.

### 2017-08-17 paper.tex
reviewed section 4
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 	\includegraphics[width=\textwidth]{figure1.jpg}
-	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
+	\caption{An ECPS is part of a well-specified larger one (intelligent product).}
 	\label{intelligent-product}
@@
 Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.
-That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable.
+That ultimately leads to the use of analog-to-digital (A/D) converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable.
 As a consequence, verification procedures have to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations.
@@
 %--------------------------------------
-\section{Verification and Synthesis \\ Challenges for Embedded Systems}
+\section{Verification and Synthesis \\ Challenges for Embedded \& Cyber-physical Systems}
 \label{Verification-Challenges} 
@@
 Indeed, the first two aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers, and the third one is essential in systems that implement device models, such as digital filters and controllers, which present a behavior that is highly dependent on signal inputs and outputs and whose deployment may be heavily affected by hardware restrictions.
+
 The fourth aspect deals with code that relies on existing infrastructures and must be compliant with those, such as Linux kernel modules.
@@
 Therefore, when a new product is derived from a given platform, a lot of legacy code is usually reused for improving development time and code quality.
+
 The sixth aspect is related to evolving development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
@@
 The eighth one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
+
 The ninth aspect is a consequence of changes in environment or different applications and scenarios, which may present new problems to embedded systems that did not exist or were not tackled during their development.

### 2017-08-17 paper.tex
reviewed section 5
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 intelligent product in Figure~\ref{intelligent-product}).
-Additionally,  finite-precision arithmetic and their related rounding errors need to be considered, when correct-by-construction code is generated for embedded systems.
+Additionally,  finite-precision arithmetic and their related rounding errors need to be considered, when correct-by-construction code is generated for ECPS.
 
@@
 
-This position paper tackles seven major problems in computer-aided verification and synthesis for ECPS, which are (partially) open in current published research.
+This survey tackles seven major problems in computer-aided verification and synthesis for ECPS, which are (partially) open in current published research.
 
@@
 
-\textbf{(RP2)} exploitation of SMT techniques to leverage bounded model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
+\textbf{(RP2)} exploitation of SMT techniques to leverage (bounded) model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
 

### 2017-08-17 paper.tex
reviewed section 6
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Thus, ECPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems.
+Thus, ECPSs demand short development cycles and high-level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems.
 As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.
@@
 
-\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}.
+\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}.
 This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}.
@@
 
-\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
+\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al.} was further developed to verify correct lock acquisition ordering and absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}.
 Recent advances for verifying multi-threaded C programs have been proposed to speed up verification times, which significantly prune state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution, based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}, were made available; however, the set of concurrent-program classes ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited.
@@
 Indeed, it would be interesting if a new programming language model could be loaded, which along with a BMC core could check different programs.
-Some work towards that was already presented by Sousa, Cordeiro, and Filho~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics.
+Some work towards that was already presented by Sousa {\it et al.}~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics.
 In summary, the BMC core (in that case, ESBMC) is not changed (it is still C/C++ code), but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked.
@@
 Obviously, a completely new language may present new structures and resources; however, if different inputs are parsed and converted into a form that preserves every aspect of their structure, that can then be analyzed and verified in an unified way.
-Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.
+Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in ECPS verification.
 
@@
 The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe), due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}.
-Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17}.
-Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, state-space).
+Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17,abatecav2017}.
+Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, Multiple-Input Multiple-Output).
 In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
@@
 Given the availability of different programming languages for writing code for ECPS, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications.
-The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated.
-In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main objective: formal program verification.
+The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an ECPS can be addressed and evaluated.
+In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main objective: formal program verification in ECPS.
 Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics.

### 2017-08-19 cover/Cover-Letter.tex
25 pages version and cover letter

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}

%\usepackage{mathtools}
%\mathtoolsset{showonlyrefs}

\usepackage{color}
\usepackage{natbib}
% \usepackage[style=authoryear-comp,dashed=false]{biblatex}
% \bibliography{piecewise_lyapunov}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\newcommand{\Real}{\mathbb{R}}

\pagestyle{empty}

\begin{document}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{oxford.jpg}}
\end{minipage}}} \hspace*{0.1cm}
\begin{minipage}{11.0cm}{\sf \large \textbf{UNIVERSITY OF OXFORD}}\\
{\sc Department of Computer Science}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Dr. Cordeiro\\
University of Oxford\\
Department of Computer Science\\
Wolfson Building, Parks Rd, Oxford OX1 3QD\\
{\em e-mail: lucas.cordeiro@cs.ox.ac.uk}
\end{minipage}
\end{flushright}

\vi
\hspace*{\fill}{\small Oxford, $19^{th}$ August, 2017.}
\vi

\begin{flushleft}
    Tei-Wei Kuo\\
    Editor-in-Chief\\
    ACM Transactions on Cyber-Physical Systems (TCPS)
    \end{flushleft}
\vi

\indent Dear Editor-in-Chief,
\vi

Please find enclosed our manuscript entitled as ``{\em Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges},'' by Lucas Cordeiro, Eddie Batista de Lima Filho, and Iury Bessa, which we would like to submit for publication in the journal of {\bf ACM Transactions on Cyber-Physical Systems (TCPS)}.

\vi

The present work is a survey paper that discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs. Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a correct-by-construction design.

\vi

The  authors  claim  that  none  of  the  material  in  the mentioned  paper  has  been  published  or  is  under  consideration for publication elsewhere.

\vi

Thank you for your time.

\vi\vi

\indent
~~~~~~~~~~~~~~~~~~~~~~~~~~~Sincerely,\\

\begin{quote}
\begin{quote}
\begin{flushright}

%\vfill

Lucas C. Cordeiro~~~~~~~
\end{flushright}
\end{quote}
\end{quote}

\end{document}

### 2018-02-02 paper.tex
25 pages version and cover letter
Added subsection about interpolants
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
+\newtheorem{prf}{\textbf{Definition}}
+
 % Metadata Information
-\acmJournal{TCPS}
-\acmVolume{9}
-\acmNumber{9}
-\acmArticle{99}
-\acmYear{2017}
-\acmMonth{12}
-\copyrightyear{2017}
+%\acmJournal{TCPS}
+%\acmVolume{9}
+%\acmNumber{9}
+%\acmArticle{99}
+%\acmYear{2017}
+%\acmMonth{12}
+%\copyrightyear{2017}
 %\acmArticleSeq{9}
@@
 % Paper history
-\received{February 2007}
-\received[revised]{March 2009}
-\received[accepted]{June 2009}
+\received{August 2017}
+%\received[revised]{March 2009}
+%\received[accepted]{June 2009}
 
@@
 %\renewcommand{\baselinestretch}{0.93}
+\renewcommand{\baselinestretch}{0.97}
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{figure1.jpg}
+	\includegraphics[width=0.6\textwidth]{figure1.jpg}
 	\caption{An ECPS is part of a well-specified larger one (intelligent product).}
@@
 %--------------------------------------
-\section{Bounded Model Checking (BMC)}
+\section{SAT-based and SMT-based Verification Techniques}
 \label{Preliminaries}
+%--------------------------------------
+
+%--------------------------------------
+\subsection{Bounded Model Checking (BMC)}
+\label{BMC}
 %--------------------------------------
@@
 	\centering
-	\includegraphics[width=\textwidth]{figure2.jpg}
+	\includegraphics[width=0.6\textwidth]{figure2.jpg}
 	\caption{Bounded Model Checking.}
@@
 \centering
-\begin{minipage}{0.45\textwidth}
-\begin{lstlisting}
+\begin{minipage}{0.4\textwidth}
+\begin{lstlisting}[basicstyle=\footnotesize]
 #include <assert.h>
@@
 \centering
-\begin{minipage}{0.45\textwidth}
-\begin{lstlisting}
+\begin{minipage}{0.4\textwidth}
+\begin{lstlisting}[basicstyle=\footnotesize]
 a1 = N
@@
 %--------------------------------------
-\subsection{Induction-based Verification of C Programs}
+\subsection{Completeness}
+\label{completeness}
+%--------------------------------------
+
+\textcolor{blue}{BMC can be used to find property violations up to the bound $k$ but not to prove properties, unless an upper bound is known on the depth of the state space, which is not generally the case.
+For software verification and synhtesis, we can adopt two different strategies in order to \emph{prove} properties: \textit{(i)} compute the \textit{completeness threshold}, which can be smaller than or equal to the maximum number of loop-iterations occurring in the program or \textit{(ii)} determine the high-level worst-case execution time (WCET), which also gives a bound on the maximum number of loop-iterations~\cite{handbook09,ckso04,Ganai08}.
+However, in practice, complex software systems involve large data-paths and complex expressions.
+Therefore, the verification conditions that arise from BMC of programs become harder to solve and require substantial amounts of memory to build.}
+
+%--------------------------------------
+\subsubsection{Induction-based Verification of C Programs}
 %--------------------------------------
@@
 \begin{minipage}{0.45\textwidth}
-\begin{lstlisting}
+\begin{lstlisting}[basicstyle=\footnotesize]
 #include <assert.h>
@@
 %------------------------------------------------------
-%\subsection{Craig Interpolation}
-%\label{sec:CraigInterpolationInModelChecking}
+\subsubsection{Craig Interpolation}
+\label{sec:CraigInterpolationInModelChecking}
 %------------------------------------------------------
 
-%Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+\textcolor{blue}{Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
 This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs regarding the nonexistence of counterexamples of depth less than or equal to $k$.
-Those proofs do not ensure whether a given property holds in the model, but they contain information about reachable states of a model.
-
-%\begin{prf}\label{interpolants}
-%\begin{definition}\label{interpolants}
-%\textit{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }
-
-%\begin{itemize}
-%	\item $A \Rightarrow F$;
-%	\item $F \wedge B$ \textit{is unsatisfiable};
-%	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
-%\end{itemize}
-%\end{definition}
-%\end{prf}
-
-%As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
-Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$, since $x_2$ is a common part of both $A$ and $B$.
-
-%The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers. 
-%
-%In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
-Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).
-%
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=\textwidth]{interpolants}
-%\caption{Computing image by interpolation~\cite{McMillan03}.}
-%\label{figure:computing-image-by-interpolation}
-%\end{figure}
-
-%Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
+Those proofs do not ensure whether a given property holds in the model, but they contain information about reachable states of a model.}
+
+\begin{prf}\label{interpolants}
+\begin{definition}\label{interpolants}
+\textit{\textcolor{blue}{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties}~\cite{McMillan03,McMillan05}: }
+
+\begin{itemize}
+	\item $A \Rightarrow F$;
+	\item $F \wedge B$ \textit{is unsatisfiable};
+	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
+\end{itemize}
+\end{definition}
+\end{prf}
+
+\textcolor{blue}{As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$.
+Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$, since $x_2$ is a common part of both $A$ and $B$.}
+
+\textcolor{blue}{The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.}
+%
+\textcolor{blue}{In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
+Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).}
+%
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{interpolants}
+\caption{Computing image by interpolation~\cite{McMillan03}.}
+\label{figure:computing-image-by-interpolation}
+\end{figure}
+
+\textcolor{blue}{Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
 From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
@@
 This procedure is stopped when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
-%The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}}.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{figure3.jpg}
+	\includegraphics[width=0.4\textwidth]{figure3.jpg}
 	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figure4.jpg}
+	\includegraphics[width=0.5\textwidth]{figure4.jpg}
 	\caption{Verification methodologies for embedded systems.}

### 2018-02-05 paper.tex
introduction to section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 % Do not put math or special symbols in the title.
-\title{Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges}
+\title{Automated Program Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}
 
@@
 %--------------------------------------
-\subsection{Automated and Formal Verification of Cyber-physical systems}
+\subsection{Automated and Formal Verification of Cyber-Physical Systems}
 \label{ssec:verification}
@@
 %--------------------------------------
-\section{SAT-based and SMT-based Verification Techniques}
+\section{SAT- and SMT-based Verification and Synthesis Techniques}
 \label{Preliminaries}
 %--------------------------------------
+
+%The amount of software embedded in CPS is increasing, partly due to increased design complexity, but also as a result of shift toward software-centric implementations, which designers prefer because software allows them to flexibly and efficiently implement new features.
+Synthesis is one of the key automation techniques for improving productivity and developing efficient implementations for a design specification.
+
+\textcolor{blue}{Due to the advent of sophisticated SMT solvers built over efficient SAT solvers, an increasingly number of software verifiers and synthesizers are being developed in the literature, in order to improve  productivity and produce efficient implementations for a ECPS design specification.
+Here, we describe recent advances in symbolic model checking of software, including BMC, \textit{k}-induction, and Craig interpolation techniques.
+This section also overviews software synthesis techniques that also use logic to describe states and transformations between system states.}
 

### 2018-02-23 paper.tex
improved introduction of section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-\textcolor{blue}{Due to the advent of sophisticated SMT solvers built over efficient SAT solvers, an increasingly number of software verifiers and synthesizers are being developed in the literature, in order to improve  productivity and produce efficient implementations for a ECPS design specification.
-Here, we describe recent advances in symbolic model checking of software, including BMC, \textit{k}-induction, and Craig interpolation techniques.
-This section also overviews software synthesis techniques that also use logic to describe states and transformations between system states.}
+\textcolor{blue}{Due to the advent of sophisticated SMT solvers built over efficient SAT solvers, an increasingly number of software verifiers as well as synthesizers have emerged in the literature, in order to verify design correctness and produce reliable implementations for ECPS.
+Here, we describe some achievements in symbolic model checking of software, including BMC, \textit{k}-induction, ``property-based reachability'' (or IC3), and Craig interpolation, which are currently used to the safety verification of ECPS.
+This section also overviews software synthesis techniques that are applicable to write correct-by-construction implementations of ECPS.}
 

### 2018-02-23 paper.tex
added section about IC3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 Indeed, such an issue has pushed researchers to overcome the depth limitation and propose extensions capable of proving global correctness.
-
-%--------------------------------------
-\subsection{Completeness}
-\label{completeness}
-%--------------------------------------
-
-\textcolor{blue}{BMC can be used to find property violations up to the bound $k$ but not to prove properties, unless an upper bound is known on the depth of the state space, which is not generally the case.
-For software verification and synhtesis, we can adopt two different strategies in order to \emph{prove} properties: \textit{(i)} compute the \textit{completeness threshold}, which can be smaller than or equal to the maximum number of loop-iterations occurring in the program or \textit{(ii)} determine the high-level worst-case execution time (WCET), which also gives a bound on the maximum number of loop-iterations~\cite{handbook09,ckso04,Ganai08}.
+\textcolor{blue}{In particular, we can adopt two possible strategies to prove properties using BMC: \textit{(i)} compute the \textit{completeness threshold}, which can be smaller than or equal to the maximum number of loop-iterations occurring in the program or \textit{(ii)} determine the high-level worst-case execution time (WCET), which also gives a bound on the maximum number of loop-iterations~\cite{handbook09,ckso04,Ganai08}.
 However, in practice, complex software systems involve large data-paths and complex expressions.
@@
 %--------------------------------------
-\subsubsection{Induction-based Verification of C Programs}
+\subsection{Induction-based Verification of C Programs}
 %--------------------------------------
@@
 
+%--------------------------------------
+\subsection{Property-based Reachability (or IC3)}
+%--------------------------------------
+
+\textcolor{blue}{The application of the \textit{k}-induction method is gaining popularity in the software verification community.
+Bradley {\it et al.} introduced  ``property-based reachability'' (or IC$3$) procedure for the safety verification of systems~\cite{Bradley12,Bradley13}; the authors have shown that IC$3$ can scale on certain benchmarks, where \textit{k}-induction fails to succeed.}
+
 %------------------------------------------------------
-\subsubsection{Craig Interpolation}
+\subsection{Craig Interpolation}
 \label{sec:CraigInterpolationInModelChecking}

### 2018-02-23 paper.tex
added section about IC3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \usepackage{amsmath}
+\usepackage{centernot}
 \usepackage{epstopdf}
@@
   language={C},
-  morekeywords={assert,stdio,math,printf,assume,then,end-if,end,input,output}
+  morekeywords={stdio,math,printf,assume,then,end-if,end,input,output}
 }
@@
 For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
-If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
+The star-notation indicates nondeterminism.
+If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
 Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
@@
 \centering
-\begin{minipage}{0.45\textwidth}
+\begin{minipage}{0.35\textwidth}
 \begin{lstlisting}[basicstyle=\footnotesize]
@@
 int main(void) {
- float x=2;
+ float x=*;
  while (1) {
@@
 
-\textcolor{blue}{The application of the \textit{k}-induction method is gaining popularity in the software verification community.
-Bradley {\it et al.} introduced  ``property-based reachability'' (or IC$3$) procedure for the safety verification of systems~\cite{Bradley12,Bradley13}; the authors have shown that IC$3$ can scale on certain benchmarks, where \textit{k}-induction fails to succeed.}
+\textcolor{blue}{Bradley {\it et al.} introduced  ``property-based reachability'' (or IC$3$) procedure for the safety verification of systems~\cite{Bradley12,Bradley13}; the authors have shown that IC$3$ can scale on certain benchmarks, where \textit{k}-induction fails to succeed.
+In particular, the success of IC3 over \textit{k}-induction procedures is due to the ability of the former to guide the search for inductive instances with counterexamples to the inductiveness (CTIs) of the given property~\cite{Bradley18}.
+Consider again the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, that $\rho : x > 0$ is an invariant.}
+
+\begin{equation}
+x > 0 \wedge x' = 2*x - 1 \centernot\implies x > 0. \nonumber
+\end{equation}
+
+\textcolor{blue}{In this specific example, a CTI returned by an SMT solver is $x=0$.
+If this state is not eliminated, then the invariant $\rho$ cannot be established.
+The generated inductive assertion should establish that the CTI $x=0$  is unreachable.
+If no such inductive assertion exists, then other CTIs can be examined instead ({\it e.g.}, $0.1$, $0.2$, $\dots$, $0.9$).
+As a result, the lemmas should be strong enough that consecutively revisiting a finite set of CTIs will eventually end up in an assertion, which is inductive relative to them, thus eliminating the CTI.
+In this example, the instance $x > 1$ ($a=1$, $b=0$) is inductive and eliminates all possible CTIs.}
+
+\textcolor{blue}{Jovanovi\'{c} {\it et al.}~\cite{Jovanovic:2016:PK:3077629.3077648} present a reformulation of IC3, separating the reachability checking from the inductive reasoning.
+They further replace the regular induction algorithm by the \textit{k}-induction algorithm and show that it provides more concise invariants.
+The authors implemented the algorithm in the SALLY model checker using Yices2 to do a forward search and MathSAT5 to do a backward search.
+They showed that the new algorithm is able to solve a number of real-world benchmarks, at least as fast as other approaches.}
 

### 2018-03-02 paper.tex
reviewed the k-induction section
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}.
-The star-notation indicates nondeterminism.
-If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration.
-As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness.
+For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}; \textcolor{blue}{the star-notation indicates nondeterminism.
+Suppose that one wants to prove that $P : x > 0$ is invariant.
+To attempt to prove the invariant property P, one can apply induction:}
+
+\begin{itemize} 
+\item \textcolor{blue}{It holds initially because}
+\end{itemize}
+
+\begin{equation}
+\underbrace{x = 2}_{\text{initial condition}} \implies \underbrace{x > 0}_{\text{P}}. \nonumber
+\end{equation}
+
+\begin{itemize} 
+\item \textcolor{blue}{Whenever $P$ holds for $k$ loop unwindings, it also holds for $k + 1$ steps.}
+\end{itemize}
+
+\begin{equation}
+\underbrace{x > 0}_{\text{P}} \wedge \underbrace{x = 2 \wedge x' = 2*x - 1}_{\text{transition relation}} \implies \underbrace{x' > 0}_{\text{P'}}. \nonumber
+\end{equation}
+
+\textcolor{blue}{If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
+Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; in particular, if we initially assign $x=0.9$, then $x<0$ after the fourth iteration.} As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove correctness.
 In this specific example, an inductive invariant would be $x>1$, given that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
@@
 int main(void) {
- float x=*;
- while (1) {
+ float x=2;
+ while (*) {
    x = ((2*x) - 1);
@@
 In particular, the success of IC3 over \textit{k}-induction procedures is due to the ability of the former to guide the search for inductive instances with counterexamples to the inductiveness (CTIs) of the given property~\cite{Bradley18}.
-Consider again the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, that $\rho : x > 0$ is an invariant.}
+Consider again the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, that $P : x > 0$ is an invariant.}
 
 \begin{equation}
-x > 0 \wedge x' = 2*x - 1 \centernot\implies x > 0. \nonumber
+\underbrace{x > 0}_{\text{P}} \wedge \underbrace{x = * \wedge x' = 2*x - 1}_{\text{transition relation}} \centernot\implies \underbrace{x > 0}_{\text{P'}}. \nonumber
 \end{equation}
@@
 \textcolor{blue}{In this specific example, a CTI returned by an SMT solver is $x=0$.
-If this state is not eliminated, then the invariant $\rho$ cannot be established.
+If this state is not eliminated, then the invariant $P$ cannot be established.
 The generated inductive assertion should establish that the CTI $x=0$  is unreachable.

### 2018-03-02 paper.tex
added example for the CEGIS section
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 \usepackage{xcolor,colortbl}
+\usepackage{tikz}
+\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
+\usetikzlibrary{patterns}
 % \usepackage{cite}
@@
 \newtheorem{prf}{\textbf{Definition}}
+
+\renewcommand{\baselinestretch}{0.986}
+\setlength{\belowcaptionskip}{-2.8 pt}
 
@@
 
+%---------------------------------------------------
+\subsubsection{CEGIS for Embedded \& CPS}
+\label{sec:CEGISforECPS}
+%---------------------------------------------------
+
+\textcolor{blue}{A typical synthesizer for Embedded \& CPS iterates between two main phases, 
+an inductive synthesis phase, {\sc synthesise}, and a validation phase, {\sc verify}~\cite{AbateBCCCDKKP17}.  
+The two phases interact via a finite set of test vectors, which is updated
+incrementally.  Given a specification of the ECPS ({\it e.g.}, stability and safety), the inductive
+synthesis procedure tries to find a candidate solution satisfying the
+specification for the given set of test inputs.}
+%
+\textcolor{blue}{If the synthesis phase succeeds in finding a witness, this witness is a
+candidate solution to the full synthesis formula.  This candidate
+solution is passed to the validation phase, which checks whether it is a proper solution
+({\it i.e.}, satisfies the specification for all possible inputs).  If this is the
+case, then the algorithm terminates; otherwise, additional information is
+provided to the inductive synthesis phase in the form of a new
+counterexample, C-ex, which is added to the set of test inputs, and the loop
+iterates again.}
+
+\textcolor{blue}{One possible architecture for a synthesiser is 
+presented in Fig.~\ref{fig:CEGIS-precision-increment}, considering the 
+stability and safety specification. It starts by devising a digital controller that 
+stabilizes the physical model while remaining safe for a pre-selected time 
+horizon ($k$) and a single initial state; then, it employs a multi-staged verification 
+process as follows: (i)~The first verification stage ({\sc safety}) checks that the candidate
+solution, which we synthesized to be safe for at least one initial
+state, is safe for \emph{all} possible initial states, {\it i.e.}, does not reach
+an unsafe state within $k$ steps.
+(ii)~The second verification stage ({\sc precision})
+ restores soundness with respect to the plant precision
+by using interval arithmetic~\cite{moore1966interval} to validate the 
+operations performed by the previous stage. 
+(iii)~The third verification stage ({\sc complete}) checks that the current
+$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
+completeness threshold $\overline{k}$ for the current candidate controller 
+$K$~\cite{abatecav2017}, {\it i.e.}, the number of
+iterations required to sufficiently unwind the closed-loop state-space
+model such that the boundaries are not violated for any larger number of
+iterations, and check that $k{\geq}\overline{k}$}.
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+{\scriptsize
+\centering
+\resizebox{.5\textwidth}{!}
+{
+  \begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
+  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center
+},
+          row sep=.6cm, column sep=.9cm] {
+   \coordinate (aux1);
+   \& \coordinate (aux2);
+   \&;\\
+   \coordinate (aux3);
+   \& \coordinate (aux4);
+   \&;\\
+   \coordinate (aux5);
+   \& \coordinate (aux6);
+   \&;\\
+   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc Synthesize}};
+   \&
+   complexnode/.pic={ 
+     \node[rectangle,draw,dotted,
+	minimum width=6cm,
+	minimum height=1cm,
+        pattern=north west lines, pattern color=gray!20,
+	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
+     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 1.Safety}};
+     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 2.Precision}};
+     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 3.Complete}};
+     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
+   } 
+   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
+   \& \\
+   \node[minimum height=0cm] (gp) {\sf Program Search};
+   \&
+   complexnode/.pic={ 
+     \coordinate (aux);
+   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
+   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
+   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
+   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
+   }   
+    \\
+  };
+
+   \path
+    ([yshift=2em]synth.east) edge node[xshift=-0.5em,align=center] {K} ([yshift=2em]verif1.west)
+    ([yshift=-2em]verif1.west) edge node {C-ex} ([yshift=-2em]synth.east)
+    ([xshift=-5em]fp.north) edge node[align=center]  {} ([xshift=-5em]verif2.south)
+    ([xshift=-5em]sv.north) edge node[align=center]  {} ([xshift=-5em]verif3.south)
+    %([xshift=-5em]cv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif4.south)
+    ([xshift=5em]verif1.south) edge node[align=center] {} ([xshift=5em]bmc.north)
+    ([xshift=5em]verif2.south) edge node[align=center] {} ([xshift=5em]fp.north)
+    ([xshift=5em]verif3.south) edge node[align=center] {} ([xshift=5em]sv.north)
+    %([xshift=5em]verif4.south) edge node[align=center] {$K$} ([xshift=5em]cv.north)
+    ([xshift=-5em]bmc.north) edge node[align=center]  {} ([xshift=-5em]verif1.south)
+    (verif) edge node {PASS} (done)
+    ([xshift=5em]synth.south) edge node[align=center] {} ([xshift=5em]gp.north)
+    ([xshift=-5em]gp.north) edge node[align=center] {} ([xshift=-5em]synth.south)
+    (aux3) edge (synth.north);
+   \path[-]
+   (verif2.north) edge node[align=center] {} ([xshift=0cm]aux6)
+   ([xshift=0cm]aux6) edge node[align=center] {Increase Precision} (aux5)
+   (verif3.north) edge node[align=center] {} ([xshift=6.7cm]aux4)
+   ([xshift=6.7cm]aux4) edge node[align=center] {Increase Unfolding Bound} (aux3);
+   %(verif4.north) edge node[align=center] {} ([xshift=10.5cm]aux2)
+   %([xshift=10.5cm]aux2) edge node[align=center] {Increase Sampling Rate} (aux1);
+
+ \end{tikzpicture}
+}}
+\caption{CEGIS with multi-staged verification}
+\label{fig:CEGIS-precision-increment}
+\end{figure}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %--------------------------------------

### 2018-03-08 paper.tex
fixed typo
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
updated title
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
mods to section 2

@@
 % Do not put math or special symbols in the title.
-\title{Automated Program Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}
+\title{Automated Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}
 
@@
 \affiliation{%
-  \institution{TP Vision}
+  \institution{TPV Technology}
 %  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
@@
 
-Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.
+Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as provide controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.
 
@@
 
+\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho...
+parece que estamos misturando tudo...
+sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo muito nebuloso?}
 The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
@@
 The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
-Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they should be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled, and they should be better represented by nondeterministic approaches~\cite{leeCPS2}.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
@@
 An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}.
-An example of barrier certificate is the stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region.
-The stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.
+An example of barrier certificate is stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region.
+Stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification} or linear system stability theory for discrete simulations~\cite{Bessa17}.
 
 Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest studies on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems.
-In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing.
+In the last decade, signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing.
 Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.
 
-Along with CPS modeling and verification techniques purely based on HA, one can also highlight logic-based approaches~\cite{Phan15} that may still use HA, but are not based on a algebraic description of states and transitions.
+Along with CPS modeling and verification techniques purely based on HA, one can also highlight logic-based approaches~\cite{Phan15} that may still use HA, but which are not based on a algebraic description of states and transitions.
 Some examples of logic-based verification are theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among which BMC stands out~\cite{handbook09}.
@@
 Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} preseted DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends, in order to provide support for digital system design and verification considering FWL effects.
-In particular, DSVerifier checks if a designed digital controller and/or closed-loop ECPS presents the desired performance, when it is implemented with a FWL format~\cite{Bessa16,Bessa17}.
+In particular, DSVerifier checks if a designed digital controller and/or closed-loop ECPS presents the desired performance, when it is implemented with a given FWL format~\cite{Bessa16,Bessa17}.
 
 Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}.
-Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems, which are hardly modeled through automata- or logig-based approaches.
+Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems that are hardly modeled through automata- or logig-based approaches.
 Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs.
@@
 }}
-\caption{CEGIS with multi-staged verification}
+\caption{\textcolor{blue}{CEGIS with multi-staged verification for ECPS.}}
 \label{fig:CEGIS-precision-increment}

### 2018-03-09 paper.tex
changes to section 3

@@
 parece que estamos misturando tudo...
-sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo muito nebuloso?}
+sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo um pouco nebuloso?}
 The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
@@
 
-\textcolor{blue}{Due to the advent of sophisticated SMT solvers built over efficient SAT solvers, an increasingly number of software verifiers as well as synthesizers have emerged in the literature, in order to verify design correctness and produce reliable implementations for ECPS.
+\textcolor{blue}{Due to the advent of sophisticated SMT solvers built over efficient SAT solvers, an increasingly number of software verifiers as well as synthesizers have emerged in the literature, in order to verify design correctness and produce reliable implementations for ECPS, which was later reinforced by the introduction of knowledge regarding their behavior.
 Here, we describe some achievements in symbolic model checking of software, including BMC, \textit{k}-induction, ``property-based reachability'' (or IC3), and Craig interpolation, which are currently used to the safety verification of ECPS.
@@
 \noindent Here, $\phi_{k}$ represents a safety property $\phi$ in step $k$, $I$ is the set of initial states of $M$, $R\left(s_{i},s_{i+1}\right)$ is the transition relation of $M$ at time steps $i$ and $i+1$.
-Hence, the equation $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$.
-$\neg \phi_{k}$ represents the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$.
+Hence, the equation $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$ and $\neg \phi_{k}$ means the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$.
 Finally, the resulting (bit-vector) equation is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability.
@@
 Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop~\cite{JMorse15}.
-In this case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:
+In that case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:
 
@@
 The SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code.
-The common property in SSA form is that every variable has only one definition in the program text.
+The common property in SSA form is that every variable state has only one definition in the program text.
 This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
@@
                 \wedge \neg \left(a_{N+1} \leq N\right) \\
-              \end{array} \right ]  \\
+              \end{array} \right ].  \\
               \\           
@@
 However, in practice, complex software systems involve large data-paths and complex expressions.
-Therefore, the verification conditions that arise from BMC of programs become harder to solve and require substantial amounts of memory to build.}
+Therefore, the verification conditions that arise from BMC of programs become harder to solve and require substantial amounts of memory to build.
+Finally, another strategy is to use a complementary technique, with the potential to fill the gap between the adopted depth and all subsequent ones, such as induction, which is explained below.} \textcolor{red}{colocar uma ref}.
 
@@
 
-For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}; \textcolor{blue}{the star-notation indicates nondeterminism.
-Suppose that one wants to prove that $P : x > 0$ is invariant.
-To attempt to prove the invariant property P, one can apply induction:}
+For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, \textcolor{blue}{where the star-notation indicates nondeterminism.
+Suppose that one wants to prove that $P : x > 0$ is invariant and, in order to do that, induction can be applied:}
 
@@
 \begin{itemize} 
-\item \textcolor{blue}{Whenever $P$ holds for $k$ loop unwindings, it also holds for $k + 1$ steps.}
+\item \textcolor{blue}{Whenever $P$ holds for $k$ loop unwindings, it also holds for $k + 1$ steps, given that}
 \end{itemize}
@@
 
-\textcolor{blue}{If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations.
-Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; in particular, if we initially assign $x=0.9$, then $x<0$ after the fourth iteration.} As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove correctness.
-In this specific example, an inductive invariant would be $x>1$, given that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
+\textcolor{blue}{If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the invariant $x>0$ holds initially and after each iteration and $x$ tends to infinity after $128$ iterations.
+Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration.
+In particular, if we initially assign $x=0.9$, then $x<0$ after the fourth iteration.} As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove correctness.
+In that specific example, an inductive invariant would be $x>1$, given that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
 %
@@
 
-\textcolor{blue}{Bradley {\it et al.} introduced  ``property-based reachability'' (or IC$3$) procedure for the safety verification of systems~\cite{Bradley12,Bradley13}; the authors have shown that IC$3$ can scale on certain benchmarks, where \textit{k}-induction fails to succeed.
-In particular, the success of IC3 over \textit{k}-induction procedures is due to the ability of the former to guide the search for inductive instances with counterexamples to the inductiveness (CTIs) of the given property~\cite{Bradley18}.
-Consider again the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, that $P : x > 0$ is an invariant.}
+\textcolor{blue}{Bradley {\it et al.} introduced  the ``property-based reachability'' (or IC$3$) procedure for the safety verification of systems~\cite{Bradley12,Bradley13} and have shown that IC$3$ can scale on certain benchmarks, where \textit{k}-induction fails to succeed.
+In particular, the success of IC3 over \textit{k}-induction procedures is due to the ability of the former to guide the search for inductive instances with counterexamples for the inductiveness (CTIs) of a given property~\cite{Bradley18}.
+Consider again the C-code fragment shown in Figure~\ref{figure:inductive-invariant}, where $P : x > 0$ is an invariant.}
 
@@
 
-\textcolor{blue}{Jovanovi\'{c} {\it et al.}~\cite{Jovanovic:2016:PK:3077629.3077648} present a reformulation of IC3, separating the reachability checking from the inductive reasoning.
-They further replace the regular induction algorithm by the \textit{k}-induction algorithm and show that it provides more concise invariants.
-The authors implemented the algorithm in the SALLY model checker using Yices2 to do a forward search and MathSAT5 to do a backward search.
+\textcolor{blue}{Jovanovi\'{c} {\it et al.}~\cite{Jovanovic:2016:PK:3077629.3077648} presented a reformulation of IC3, by separating reachability checking from inductive reasoning.
+They further replace the regular induction algorithm by the \textit{k}-induction one and show that it provides more concise invariants.
+Those authors implemented the mentioned algorithm in the SALLY model checker using Yices2, in order to perform a forward search, and MathSAT5, which executes a backward search.
 They showed that the new algorithm is able to solve a number of real-world benchmarks, at least as fast as other approaches.}
@@
 %
-\textcolor{blue}{In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
+\textcolor{blue}{In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries}, where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$.
 Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).}
@@
 
-\textcolor{blue}{Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$.
-From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
+\textcolor{blue}{Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$, from which we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$.
 According to Definition~\ref{interpolants}, $A$ must imply $F$ and since we defined $A$ to be the set of initial states and the first instance of $R$ (i.e., from Figure~\ref{figure:computing-image-by-interpolation}, $A=s_{0} \wedge s_{1}$), it follows that $F$ is \textit{true} in every state reachable from the initial state in one step.
@@
 In this case, we update $Q = F \vee Q$ and $A = F \wedge R_{0}$, increase the value of $k+1$ and check whether $A \wedge B$ is unsatisfiable.
-If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state); otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
-This procedure is stopped when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds).
+If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final one); otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$.
+This procedure is stopped when we have found a valid counterexample or proved that the final state is not reachable (i.e., the property holds).
 The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}}.
@@
 \textcolor{blue}{A typical synthesizer for Embedded \& CPS iterates between two main phases, 
-an inductive synthesis phase, {\sc synthesise}, and a validation phase, {\sc verify}~\cite{AbateBCCCDKKP17}.  
+an inductive synthesis phase, that is, {\sc synthesise}, and a validation one, that is, {\sc verify}~\cite{AbateBCCCDKKP17}.  
 The two phases interact via a finite set of test vectors, which is updated
-incrementally.  Given a specification of the ECPS ({\it e.g.}, stability and safety), the inductive
-synthesis procedure tries to find a candidate solution satisfying the
-specification for the given set of test inputs.}
-%
-\textcolor{blue}{If the synthesis phase succeeds in finding a witness, this witness is a
-candidate solution to the full synthesis formula.  This candidate
-solution is passed to the validation phase, which checks whether it is a proper solution
-({\it i.e.}, satisfies the specification for all possible inputs).  If this is the
+incrementally.  Given a specification of a ECPS ({\it e.g.}, stability and safety), the inductive
+synthesis procedure tries to find a candidate solution satisfying that
+specification, for the given set of test inputs.}
+%
+\textcolor{blue}{If the synthesis phase succeeds in finding a witness, then the latter is a
+candidate solution for the full synthesis formula. This candidate is passed to the validation phase, which checks whether it is a proper solution
+({\it i.e.}, it satisfies the specification for all possible inputs).  If this is the
 case, then the algorithm terminates; otherwise, additional information is
-provided to the inductive synthesis phase in the form of a new
+provided to the inductive synthesis phase, in the form of a new
 counterexample, C-ex, which is added to the set of test inputs, and the loop
@@
 stability and safety specification. It starts by devising a digital controller that 
-stabilizes the physical model while remaining safe for a pre-selected time 
-horizon ($k$) and a single initial state; then, it employs a multi-staged verification 
-process as follows: (i)~The first verification stage ({\sc safety}) checks that the candidate
+stabilizes the physical model, while remaining safe for a pre-selected time 
+horizon ($k$) and a single initial state. Then, it employs a multi-staged verification 
+process as follows: (i)~The first verification stage ({\sc safety}) checks if the candidate
 solution, which we synthesized to be safe for at least one initial
-state, is safe for \emph{all} possible initial states, {\it i.e.}, does not reach
+state, is safe for \emph{all} possible initial states, {\it i.e.}, it does not reach
 an unsafe state within $k$ steps.
 (ii)~The second verification stage ({\sc precision})
- restores soundness with respect to the plant precision
+ restores soundness with respect to plant precision,
 by using interval arithmetic~\cite{moore1966interval} to validate the 
 operations performed by the previous stage. 
-(iii)~The third verification stage ({\sc complete}) checks that the current
+(iii)~The third verification stage ({\sc complete}) checks if the current
 $k$ is large enough to ensure safety for any $k'{>}k$. Here, we compute the
@@
 iterations required to sufficiently unwind the closed-loop state-space
-model such that the boundaries are not violated for any larger number of
-iterations, and check that $k{\geq}\overline{k}$}.
+model, such that the boundaries are not violated for any larger number of
+iterations, and check if $k{\geq}\overline{k}$}.
 

### 2018-03-10 paper.tex
initial mods to challenges and research problems

@@
 We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
-In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research.
-Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPSs and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
+In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{achievements} tackles those challenges that are (partially) open in current published research and later describes current achievements and future trends in verification and synthesis of ECPSs, and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here.
 Finally, we conclude and describe future work, in Section~\ref{conclusions}.
@@
 
-\begin{prf}\label{interpolants}
+\begin{prf}
 \begin{definition}\label{interpolants}
@@
 	\item Handling of non-linear and non-convex optimization problems;
-	\item Incorporation of new and adapted checks, due to system or environment evolution.
+	\item Incorporation of new and adapted checks, due to system or environment change;
+	\item \textcolor{blue}{Provide a form of system evolution, in such a way that a version more adapted to the current scenario is obtained.}
 \end{enumerate}
@@
 The sixth aspect is related to evolving development processes and technologies, which may delay the application of suitable verification and synthesis approaches, if verifiers and synthesizers do not support different programming languages and interfaces.
-The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correct address a property.
+The seventh aspect highlights a subjective task in verification: where one must insert an evaluation point or assertion, in order to correctly address a property.
 The eighth one is related to the widespread use of embedded systems in autonomous-vehicle navigation systems~\cite{Adouane16}, which demands optimization solving during their execution for a wide range of functions, including non-linear and non-convex problems using fixed- and floating-point arithmetic.
@@
 
-Those nine challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where controlled objects ({\it e.g.}, physical plants) typically exhibit continuous behavior (which may eventually change), whereas controllers (usually implemented by real-time computer systems) operate in discrete time and over a quantized domain.
+\textcolor{blue}{The tenth aspect is related to the current development and test strategies adopted for embedded systems, which aim at simply providing robust structures, instead of antifragile ones.
+In summary, it is not feasible to account for every scenario or error, but it would be more interesting to provide a module or structure capable of thriving on problems and getting more robust.
+Such an aspect shares some similarity with the ninth one and may also be achieved through ML techniques.}
+
+Those ten challenges place additional difficulties for developing reliable synthesizers for embedded systems, especially for cyber-physical systems and micro-grids, where controlled objects ({\it e.g.}, physical plants) typically exhibit continuous behavior (which may eventually change), whereas controllers (usually implemented by real-time computer systems) operate in discrete time and over a quantized domain.
 In particular, synthesizers for those systems need to consider the effects of quantizers (A/D and D/A converters), when a digital equivalent of the controlled object is considered, {\it i.e.}, a model of their physical environment (cf.
@@
 %--------------------------------------
-\section{Research Problem (RP)}
-\label{Research-Problem}
-%--------------------------------------
-
-This survey tackles seven major problems in computer-aided verification and synthesis for ECPS, which are (partially) open in current published research.
-
-\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic ECPS.
-
-\textbf{(RP2)} exploitation of SMT techniques to leverage (bounded) model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
-
-\textbf{(RP3)} proof of correctness and timeliness of ECPS, by taking into account stringent constraints imposed by hardware platforms.
-
-\textbf{(RP4)} incorporation knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
-
-\textbf{(RP5)} provision of tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
-
-\textbf{(RP6)} development of automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with physical environments, by considering uncertain models and FWL effects.
-
-\textbf{(RP7)} provision of a unified framework, which is able to tackle high and low level properties, as well as system behavior, with the potential to lead to another degree of completeness, in such a way that every implementation aspect of ECPS is addressed and evaluated.
-
-Section~\ref{achievements} outlines contributions toward such research problems.
-
-%--------------------------------------
 \section{Current Achievements and \\ Future Trends}
@@
 %--------------------------------------
+
+\textcolor{blue}{In the preceding sections, we identified and discussed many aspects regarding ECPSs, which included design, synthesis, behavior, verification, and bottlenecks.
+Nonetheless, from the point of view of computer-aided verification and synthesis for ECPS, it is possible to highlight seven major problems, which are tackled here and can be regarded as (partially) open in current published research.}
+
+\textbf{(RP1)} provision of suitable encoding into SMT \cite{BarrettSST09}, which may extend background theories typically supported by SMT solvers, with the goal of reasoning accurately and effectively about realistic ECPS.
+
+\textbf{(RP2)} exploitation of SMT techniques to leverage (bounded) model checking of multi-threaded software, in order to mitigate the state-explosion problem due to thread interleaving.
+
+\textbf{(RP3)} proof of correctness and timeliness of ECPS, by taking into account stringent constraints imposed by hardware platforms.
+
+\textbf{(RP4)} incorporation knowledge about system purpose and associated features, which aims to detect system-level and behavior failures.
+
+\textbf{(RP5)} provision of tools and approaches capable of addressing different programming languages and application interfaces, with the goal of reducing the time needed to adapt current verification techniques to new developments and technologies.
+
+\textbf{(RP6)} development of automated synthesis approaches that are algorithmically and numerically sound, in order to handle (control) software that is tightly coupled with physical environments, by considering uncertain models and FWL effects.
+
+\textbf{(RP7)} provision of a unified framework, which is able to tackle high and low level properties, as well as system behavior, with the potential to lead to another degree of completeness, in such a way that every implementation aspect of ECPS is addressed and evaluated.
+
+\textcolor{blue}{Due to the fact that the mentioned research problems are the focus of the present technology area, many works tackled them and other researchers still address them.
+As a consequence, we can cite the followinf contributions toward them.}
 
@@
 
+\textcolor{blue}{In addition, a aspect closely related to ML techiniques, as mentioned in Section \cite{Verification-Challenges}, is system fragility.
+Currently, most design strategies try to produce robust systems, which are able to deal with a limited set of common errors that happen when they are operating, {\it i.e., they are robust}.
+Nonetheless, that is effective only if problems foreseen during design phases happen; otherwise, operation may be compromised.}
+
+\textcolor{blue}{Indeed, depending on the employed point of view, fragility and robustness may be regarded as the same property.
+A system that is more robust is less fragile, but it is still prone to problems, which depends on operating scenarios, use cases, and uncertainties.
+In summary, if the chosen model does not tackle every possible problem or error that might occur in a given environment, system-failure events are still possible, which is true even for the Ariane 5's failure \cite{antifragile}.}
+
+\textcolor{blue}{As a result, it may be more interesting to develop methodologies that are not simply robust, regarding a predetermined condition, but instead thrive on problems and grow stronger (as the human body, which gets stronger when exposed to germs), which leads to antifragile systems.
+In engineering, we are only beginning to develop solutions that can be classified as antifragile, such as multiple-input/multiple-output (MIMO) systems, which use multiple signal copies together with multipath to achieve performance enhancement~\cite{spectrum2013}.}
+
+\textcolor{blue}{Systems that could greatly benefit from being antifragile are the cyber-physical ones, which interact with other external systems and use their internal representations to interpret physical properties.
+As faults occur, which may be treated with some sort of system elasticity, and external environments are normally dynamic, given that it naturally changes with time, variations in operating conditions happen and should be tolerated, in order to keep behavior quality \cite{antifragile}.
+Besides, when discussing the fundamentals of antifragility, it is difficult not to tackle machine learning \cite{antifragile}, which has the potential to provide some wisdom regarding correct behavior and internal representations, in order to keep a system working in a volatile environment.}
+
+\textcolor{blue}{For instance, we could think about a system including an adaptation module composed by two elements: a machine learning engine, capable of capturing current behaviors and identify different trens, and a SMT-based verification method, capable of recursively performing retuning procedures, which could test new parameters suggested by the machine learning module and then judge theis suitability.
+As a consequence, we would be able to design systems with the potential to being antifragile, given that they would learn from new challenges and then modify its behavior model, in order to account for new errors and threats.}
+
 %--------------------------------------
@@
 
-Given that software complexity has significantly increased in ECPS, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
+Given that software complexity has significantly increased in ECPS, there are still some (recent) advances to stress and exhaustively cover system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers.
 There is a trend towards incorporating knowledge about the system to be verified, which may take software verification and synthesis one step further, where not only code correctness will be addressed, but also full system reliability.
@@
 Special attention will be given to modern micro-grids, considering small-scale versions of a distributed system, so that reliability and other system-level properties ({\it e.g.}, carbon emission reduction in smart cities) are amenable to automated verification and synthesis, probably through behavior models.
-Additionally, ML techniques will be incorporated to BMC frameworks, in order to provide flexibility and specific-problem tuning.
+Additionally, ML techniques will be incorporated to BMC frameworks, in order to provide flexibility and specific-problem tuning, or even allow system adaptation and behavior tuning.
 

### 2018-03-10 paper.tex
corrections to the new text

@@
 
-In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allow specific checks regarding system behavior and not only code correctness.
-Additionally, embedded system verification raises additional challenges, such as:
+In particular, the main challenges regarding verification of embedded systems lie on improving coverage, where more system functions are verified, reducing verification time, {\it i.e.}, pruning state-space exploration during verification, providing completeness, {\it i.e.},  if all possible states can be reached and evaluated, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness.
+Additionally, embedded-system verification raises additional challenges, such as:
 %
@@
 	\item Incorporation of new and adapted checks, due to system or environment change;
-	\item \textcolor{blue}{Provide a form of system evolution, in such a way that a version more adapted to the current scenario is obtained.}
+	\item \textcolor{blue}{Provision of a form of system evolution, in such a way that a version more adapted to a given scenario is obtained.}
 \end{enumerate}
@@
 
-\textcolor{blue}{The tenth aspect is related to the current development and test strategies adopted for embedded systems, which aim at simply providing robust structures, instead of antifragile ones.
-In summary, it is not feasible to account for every scenario or error, but it would be more interesting to provide a module or structure capable of thriving on problems and getting more robust.
-Such an aspect shares some similarity with the ninth one and may also be achieved through ML techniques.}
+\textcolor{blue}{The tenth aspect is related to the current development and test strategies adopted for embedded systems, which aim at simply providing robust structures, instead of antifragile ones \cite{antifragile}.
+In summary, it is not feasible to account for every error or faulty scenario and then it would be more beneficial to provide a module, architecture, or methodology capable of thriving on problems and getting stronger.
+Such an aspect shares some similarity with the ninth one and may also be achieved through ML techniques.
+Indeed, this kind of scheme consists in a completely different approach, which would profoundly change the way we design embedded systems.}
 
@@
 
-\textcolor{blue}{Due to the fact that the mentioned research problems are the focus of the present technology area, many works tackled them and other researchers still address them.
-As a consequence, we can cite the followinf contributions toward them.}
+\textcolor{blue}{Due to the fact that the mentioned research problems are the some of the main issues found in the technology area addressed in this article, many works tackled them and other researchers still perform experiments and suggest new methodologies and schemes for mitigating or solving their consequences.
+As a result, we can cite the following contributions toward them, which already provide some answers or at least present promissing approaches.}
 
@@
 Amoedo {\it et al.} \cite{amoedo} proposed a modulation classification methodology, with the goal of spectrum sensing for opportunistic transmissions.
-Besides a correct software implementation, {\it i.e.}, source code, which is already covered by formal methods already available in literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
+Besides a correct software implementation, {\it i.e.}, correct source code, which is already covered by formal methods already available in literature \cite{esbmc_filter,Cordeiro12,Clarke04}, one may argue if the proposed methodology is able to recognize modulations under some reception condition, such as severe analog interference.
 This way, it would be interesting to develop a verification module for transmission and reception software, which could be capable of creating inputs related to sampled radio frequency signals, under a variety of conditions, and then unifying source code and behavior verification.
-In addition, that kind of implementation usually employs digital filters with specific parameters, which could also be tackled.
-Indeed, communication aspects of ECPSs could be tackled, as a module in system verification procedures.
+In addition, that kind of implementation usually employs digital filters with specific parameters, which could also be addressed.
+Indeed, communication aspects of ECPSs could be tackled, as properties in system verification procedures.
 
@@
 For instance, Hamel~\cite{mlswverif} proposed to induce a theory, through ML, that fits a test set for a program, in such a way that it implements the respective test set.
-Phuc \cite{nguyen}, in turn, applied decision tree C4.5 and multilayer perceptron to annotated programs (metrics related to structure and semantics) and tackled the generation of test sets satisfying the desired constraints.
-Bridge studied the automation of heuristic selection based on features, regarding first-order logic theorem proving, and Hutter {\it et.
+Phuc \cite{nguyen}, in turn, applied decision tree C4.5 and multilayer perceptron to annotated programs (metrics related to structure and semantics) and tackled the generation of test sets satisfying desired constraints.
+Bridge, Holden, and Paulson \cite{} studied the automation of heuristic selection based on features, regarding first-order logic theorem proving, and Hutter {\it et.
 al} \cite{babic07} showed that ML can be used for improving SAT solvers regarding huge verification tasks, when applied to parameter tuning~\cite{babic07}.
@@
 
-\textcolor{blue}{In addition, a aspect closely related to ML techiniques, as mentioned in Section \cite{Verification-Challenges}, is system fragility.
-Currently, most design strategies try to produce robust systems, which are able to deal with a limited set of common errors that happen when they are operating, {\it i.e., they are robust}.
-Nonetheless, that is effective only if problems foreseen during design phases happen; otherwise, operation may be compromised.}
+\textcolor{blue}{In addition, another aspect closely related to ML techniques, as mentioned in Section \ref{Verification-Challenges}, is system fragility.
+Currently, most design strategies try to produce robust systems, which are able to deal with a number of errors that happen when they are operating.
+Nonetheless, that is effective only if problems foreseen during design phases happen; otherwise, operation may be severely compromised and even cause loss of lives.}
 
@@
 
-\textcolor{blue}{As a result, it may be more interesting to develop methodologies that are not simply robust, regarding a predetermined condition, but instead thrive on problems and grow stronger (as the human body, which gets stronger when exposed to germs), which leads to antifragile systems.
-In engineering, we are only beginning to develop solutions that can be classified as antifragile, such as multiple-input/multiple-output (MIMO) systems, which use multiple signal copies together with multipath to achieve performance enhancement~\cite{spectrum2013}.}
-
-\textcolor{blue}{Systems that could greatly benefit from being antifragile are the cyber-physical ones, which interact with other external systems and use their internal representations to interpret physical properties.
+\textcolor{blue}{As a result, it may be more interesting to develop methodologies that are not simply robust, regarding predetermined conditions, but instead thrive on problems and grow stronger (as the human body, which gets stronger when exposed to germs), which leads to antifragile systems.
+In engineering, we are only beginning to develop solutions that can be classified as antifragile, such as multiple-input/multiple-output (MIMO) systems, which use multiple signal copies together with multipath to achieve performance enhancement~\cite{spectrum2013}.
+In that case, signal copies would initially result in a destructive effect; however, with the use of a methodology that takes advantage on them, the net result is performance improvement.}
+
+\textcolor{blue}{Systems that could greatly benefit from being antifragile are the cyber-physical ones, which interact with other external elements and use their internal representations to interpret physical properties.
 As faults occur, which may be treated with some sort of system elasticity, and external environments are normally dynamic, given that it naturally changes with time, variations in operating conditions happen and should be tolerated, in order to keep behavior quality \cite{antifragile}.
@@
 
-\textcolor{blue}{For instance, we could think about a system including an adaptation module composed by two elements: a machine learning engine, capable of capturing current behaviors and identify different trens, and a SMT-based verification method, capable of recursively performing retuning procedures, which could test new parameters suggested by the machine learning module and then judge theis suitability.
-As a consequence, we would be able to design systems with the potential to being antifragile, given that they would learn from new challenges and then modify its behavior model, in order to account for new errors and threats.}
+\textcolor{blue}{For instance, we could think of a system including an antifragility module composed by two elements: a machine learning engine, capable of capturing current behaviors and identifying different trends, and a SMT-based verification method, capable of recursively performing retuning procedures, which could test new parameters estimated from the data provided by the machine learning module and then test new behaviors based on them, through counterexamples.
+In addition, there could also be a measurement module, which would also judge some metrics regarding the new behavior ({e.g.}, signal-to-noise-error).
+As a consequence, we would be able to design systems with the potential of being antifragile, given that they would learn from new challenges and then modify its behavior model, in order to account for new errors and threats and even benefit from them.}
 

### 2018-03-12 paper.tex
updated title
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 % Do not put math or special symbols in the title.
-\title{Automated Verification and Synthesis of Embedded \& Cyber-Physical Systems: Advances and Challenges}
+\title{Automated Verification and Synthesis of Cyber-Physical Systems: Advances and Challenges}
 
@@
 %--------------------------------------
-\section{Automated Verification and Synthesis of Cyber-physical systems}
+\section{Automated Verification and Synthesis of Cyber-Physical Systems}
 \label{sec:model}
@@
 Therefore, the verification conditions that arise from BMC of programs become harder to solve and require substantial amounts of memory to build.
-Finally, another strategy is to use a complementary technique, with the potential to fill the gap between the adopted depth and all subsequent ones, such as induction, which is explained below.} \textcolor{red}{colocar uma ref}.
+Finally, another strategy is to use a complementary technique, with the potential to fill the gap between the adopted depth and all subsequent ones, such as induction, which is explained below~\cite{EenS03,Sheera00}.}
 

### 2018-03-12 paper.tex
fixed references
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction and in a completely automatic way ({\it i.e.}, users do not need to provide loop invariants), were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15,Kinductor,Rocha17}.
+Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction and in a completely automatic way ({\it i.e.}, users do not need to provide loop invariants), were recently proposed~\cite{Gadelha15,Beyer15,Brain15,Kinductor,Rocha17}.
 Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10}, considering the Cell BE processor, and (2) do respect time constraints, which are specified during system design phases~\cite{EenS03}.
@@
 
-It is worth noticing that \textit{k}-induction with invariants has the potential to be directly integrated into existing BMC approaches, given that the induction algorithm itself can be seen as an extension after \textit{k} unwindings and it is possible to generate program invariants with other software modules, which are then translated and instrumented into an input program \cite{Rocha15}.
+It is worth noticing that \textit{k}-induction with invariants has the potential to be directly integrated into existing BMC approaches, given that the induction algorithm itself can be seen as an extension after \textit{k} unwindings and it is possible to generate program invariants with other software modules, which are then translated and instrumented into an input program \cite{Rocha17}.
 
@@
 
-\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10,Rocha17}; however, the main challenge remains still open, {\it i.e.}, to compute and strengthen loop invariants for proving program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
+\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha17,Donaldson10}; however, the main challenge remains still open, {\it i.e.}, to compute and strengthen loop invariants for proving program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches.
 In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t.
@@
 
-%\begin{thebibliography}{1}
-
-%1
-%\bibitem{Kopetz11}
-%H.~Kopetz:
-%\newblock {\em {Real-Time Systems - Design Principles for Distributed Embedded Applications}}.
-%\newblock Real-Time Systems Series, Springer, ISBN 978-1-4419-8236-0, pp. 1--376, 2011.
-%
-%%2
-%\bibitem{xu15}
-%Xua X., Jiaa H., Wanga D., Yub D., Chiangc H.:
-%\newblock {\em {Hierarchical energy management system for multi-source multi-product microgrids}}.
-%\newblock In: Renewable Energy, v. 78, pp. 621--630, 2015.
-%
-%%3
-%\bibitem{leeCPS}
-%Lee E.:
-%\newblock {\em {Cyber-physical Systems: Design Challenges}}.
-%\newblock In: International Symposium on Object Oriented Real-Time Distributed Computing, pp. 363--369, 2008.
-%
-%%4
-%\bibitem{leeCPS2}
-%Lee E.:
-%\newblock {\em {The Past, Present and Future of Cyber-Physical Systems: A Focus on Models}}. 
-%\newblock In: Sensors 15(3): pp. 4837--4869, 2015.
-%
-%%5
-%\bibitem{groza2015formal}
-%Groza A., Letia I., Goron A., Zaporojan S.: 
-%\newblock {\em {A formal approach for identifying assurance deficits in unmanned aerial vehicle software}}.
-%\newblock In: Progress in Systems Engineering, Springer, pp. 233--239, 2015.
-%
-%%6
-%\bibitem{Cordeiro09}
-%Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
-%\newblock {\em {Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints}}.
-%\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.
-%
-%%7
-%\bibitem{munir}
-%Munir S., Stankovic J. A., Liang C.-J. M., Lin S.:
-%\newblock {\em {Cyber Physical System Challenges for Human-in-the-Loop Control}}. 
-%\newblock In: 8th International Workshop on Feedback Computing, pp. 1--4, 2013.
-%
-%%8
-%\bibitem{Abate17}
-%Abate A., Bessa I., Cattaruzza D., Cordeiro L., David C., Kesseli P., Kroening D.:
-%\newblock {\em {Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants}}.
-%\newblock In: 20th ACM International Conference on Hybrid Systems: Computation and Control, 2017 (to appear).
-%
-%%9
-%\bibitem{ldd}
-%Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
-%\newblock {\em {Model checking concurrent linux device drivers}}.
-%\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
-%
-%%10
-%\bibitem{mlembe}
-%S.~Virtanen:
-%\newblock {\em {Advancing Embedded Systems and Real-Time Communications with Emerging Technologies}}.
-%\newblock IGI Global, ISBN 978-1466660342, pp. 1--502, 2014.
-%
-%%11
-%\bibitem{Kroening15}
-%Kroening D., Liang L., Melham T., Schrammel P., Tautschnig M.:
-%\newblock {\em {Effective Verification of Low-Level Software with Nested Interrupts}}. 
-%\newblock In: Design, Automation and Test in Europe, pp. 229--234, 2015.
-%
-%%12
-%\bibitem{cudalucas}
-%Monteiro F. R., da Silva Alves E. H., Silva I. S., Ismail H., Cordeiro L. C., de Lima Filho E. B.:
-%\newblock {\em {ESBMC-GPU A Context-Bounded Model Checking Tool to Verify CUDA Programs}}. 
-%\newblock In: Science of Computer Programming xx(y): pp. xxxx--xxxx, 2017.
-%
-%%13
-%\bibitem{stochmodel}
-%Bargmann H.:
-%\newblock {\em {The role of stochastic modelling in engineering science}}.
-%\newblock In: Acta Mecanica 125(1), pp. 63--71, 1997.
-%
-%%14
-%\bibitem{Shoukry}
-%Shoukry Y., Nuzzo P., Puggelli A., Sangiovanni-Vincentelli A. L., Seshia S. A., Tabuada P.:
-%\newblock {\em {Secure State Estimation for Cyber Physical Systems Under Sensor Attacks: A Satisfiability Modulo Theory Approach}}.
-%\newblock {IEEE} Trans. Automatic Control PP(99), pp. 1--16, 2017.
-%
-%%15
-%\bibitem{cpsattack}
-%Liu Y., Ning P., Reiter M. K.:
-%\newblock {\em {“False data injection attacks against state estimation in electric power grid}}.
-%\newblock In: ACM Conference on Computer and Communications Security, pp. 21--32, 2009.
-%
-%%16
-%\bibitem{Araujo16}
-%Araujo, R., Bessa, I., Cordeiro, L., Chaves Filho, J.:
-%\newblock {\em{ SMT-based Verification Applied to Non-convex Optimization Problems}}. 
-%\newblock In: VI Brazilian Symposium on Computing Systems Engineering, 2016 (to appear).
-%
-%%17
-%\bibitem{Biere99}
-%Biere A., Cimatti A., Clarke E., Zhu Y.:
-%\newblock {\em {Symbolic Model Checking without BDDs}}. 
-%\newblock In: Tools and Algorithms for Construction and Analysis of Systems, LNCS 1579, pp. 193--207, 1999.
-%
-%%18
-%\bibitem{handbook09}
-%Biere A., Heule M., van Maaren H., Walsh T., eds.:
-%\newblock {\em {Handbook of Satisfiability}}.
-%\newblock Volume 185 of Frontiers in Artificial Intelligence and Applications., {IOS} Press, 2009.
-%
-%%19
-%\bibitem{BarrettSST09}
-%Barrett C., Sebastiani R., Seshia S.A., Tinelli C.:
-%\newblock {\em {Satisfiability Modulo Theories}}. 
-%\newblock In: Volume 185 of Frontiers in Artificial Intelligence and Applications. IOS Press, pp. 825--885, 2009.
-%
-%%20
-%\bibitem{Armando2009}
-%Armando A., Mantovani J., Platania L.:
-%\newblock {\em {Bounded Model Checking of Software Using SMT Solvers Instead of SAT Solvers}}. 
-%\newblock In: {SPIN} Workshop on Model Checking Software, LNCS 3925, pp. 146-162, 2006.
-%
-%%21
-%\bibitem{Bradley07}
-%Bradley A., Manna Z.:
-%\newblock {\em {The calculus of computation - decision procedures with applications to verification}}. 
-%\newblock In: Springer, pp. I-XV, pp. 1--366, 2007.
-%
-%%22
-%\bibitem{Thomas01}
-%Thomas Ball and Sriram~K. Rajamani.
-%\newblock {\em SLIC: A Specification Language for Interface Checking (of C)}.
-%\newblock Technical Report MSR-TR-2001-21, Microsoft Research, 2001.
-%
-%%23
-%\bibitem{Strichman08}
-%Daniel Kroening and Ofer Strichman.
-%\newblock {\em Decision Procedures: An Algorithmic Point of View}.
-%\newblock Springer Publishing Company, Incorporated, 2008.
-%
-%%24
-%\bibitem{Appel97}
-%Andrew~W. Appel.
-%\newblock {\em Modern Compiler Implementation in C: Basic Techniques}.
-%\newblock Cambridge University Press, New York, NY, USA, 1997.
-%
-%%25
-%\bibitem{Tseitin83}
-%G.~S. Tseitin.
-%\newblock On the complexity of derivation in propositional calculus.
-%\newblock In {\em In J. Siekmann and G. Wrightson, editors, Automation of
-%  Reasoning 2: Classical Papers on Computational Logic 1967-1970}.
-% 
-% %26
-% \bibitem{PatarinG97}
-%Jacques Patarin and Louis Goubin.
-%\newblock Trapdoor one-way permutations and multivariate polynominals.
-%\newblock In {\em ICICS}, {\em LNCS} 1334, pages 356--368. Springer, 1997. 
-%
-%%27
-%\bibitem{Clarke04}
-%Clarke E., Kroening D., Lerda F.:
-%\newblock {\em {A Tool for Checking {ANSI-C} Programs}}.
-%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems. LNCS 2988, Springer Berlin Heidelberg, pp. 168--176, 2004.
-%
-%%28
-%\bibitem{MerzFS12}
-%Merz F., Falke S., Sinz C.:
-%\newblock {\em {{LLBMC}: Bounded Model Checking of {C} and {C++} Programs using a Compiler {IR}}}.
-%\newblock In: International Conference on Verified Software: Theories, Tools, Experiments. LNCS 7152, pp. 146--161, 2012.
-%
-%%29
-%\bibitem{CordeiroF11}
-%Cordeiro L., Fischer B.:
-%\newblock {\em {Verifying Multi-threaded Software using {SMT}-based Context-Bounded Model Checking}}.
-%\newblock In: International Conference on Software Engineering. pp. 331--340, 2011.
-%
-%%30
-%\bibitem{Ivancic05}
-%Ivanicic F., Shlyakhter I., Gupta A., Ganai, M.K.:
-%\newblock {\em {Model Checking {C} Programs using {F-Soft}}}.
-%\newblock In: International Conference on Computer Design: VLSI in Computers and Processors, pp. 297--308, 2005.
-%
-%%31
-%\bibitem{Cordeiro12}
-%Cordeiro L., Fischer B., Marques{-}Silva J.:
-%\newblock {\em {{SMT}-based Bounded Model Checking for Embedded {ANSI-C} Software}}.
-%\newblock {IEEE} Trans. Software Eng. 38(4), pp. 957--974, 2012.
-%
-%%32
-%\bibitem{EenS03}
-%E{\'{e}}n, N., S{\"{o}}rensson, N.:
-%\newblock {\em {Temporal Induction by Incremental {SAT} Solving}}.
-%\newblock Electronic Notes in Theoretical Computer Science 89(4), pp. 543 -- 560, 2003.
-%
-%%33
-%\bibitem{Sheera00}
-%Sheeran M., Singh S., St{\aa}lmarck G.:
-%\newblock {\em {Checking Safety Properties using Induction and a {SAT}-solver}}.
-%\newblock In: International Conference on Formal Methods in Computer-Aided Design. Springer-Verlag, pp. 108--125, 2000.
-%
-%%34
-%\bibitem{4610935}
-%{IEEE} standard for floating-point arithmetic.
-%\newblock {\em IEEE Std 754-2008}, pages 1--70, Aug 2008.
-%
-%%35
-%\bibitem{Goldberg91whatevery}
-%David Goldberg.
-%\newblock What every computer scientist should know about floating point
-%  arithmetic.
-%\newblock {\em ACM Computing Surveys}, 23(1):5--48, 1991.
-%
-%%36
-%\bibitem{pips:2013}
-%ParisTech M.:
-%\newblock {\em {PIPS: Automatic Parallelizer and Code Transformation Framework}}. 
-%\newblock Accessed 21 February 2016
-%
-%%37
-%\bibitem{Henry:2012}
-%Henry J., Monniaux D., Moy M.: 
-%\newblock {\em {PAGAI: A Path Sensitive Static Analyser}}. 
-%\newblock In: Electronic Notes in Theoretical Computer Science, pp. 15--25, 2012.
-%
-%%38
-%\bibitem{Beyer15}
-%Beyer D., Dangl M., Wendler P.:
-%\newblock {\em {Boosting \textit{k}-Induction with Continuously-Refined Invariants}}.
-%\newblock In: International Conference on Computer-Aided Verification, LNCS 9206, pp. 622--640, 2015.
-%
-%%39
-%\bibitem{Gadelha15}
-%Gadelha M., Ismail H., Cordeiro L.:
-%\newblock {\em {Handling Loops in Bounded Model Checking of C Programs via \textit{k}-Induction}}.
-%\newblock In: International Journal on Software Tools for Technology Transfer (to appear), 2015.
-%\newblock http://dx.doi.org/10.1007/s10009-015-0407-9
-%
-%%40
-%\bibitem{Brain15}
-%Brain M., Joshi S., Kroening D., Schrammel P.:
-%\newblock {\em {Safety Verification and Refutation by k-Invariants and \textit{k}-Induction}}.
-%\newblock In: International Symposium on Static Analysis, LNCS 9291, pp. 145--161, 2015.
-%
-%%41
-%\bibitem{Rocha15}
-%Rocha H., Ismail H., Cordeiro L., Barreto R.:
-%\newblock {\em {Model Checking Embedded C Software using \textit{k}-Induction and Invariants}}.
-%\newblock In: V Brazilian Symposium on Computing Systems Engineering, pp. 90--95, 2015.
-%
-%%42
-%\bibitem{Kinductor}
-%Donaldson A., Haller L., Kroening D., R{\"{u}}mmer, P.:
-%\newblock {\em {Software Verification using \textit{k}-Induction}}.
-%\newblock In: International Symposium on Static Analysis. LNCS 6887, pp. 351--368, 2011.
-%
-%%43
-%\bibitem{Rocha17}
-%Rocha, W., Rocha, H., Ismail H., Cordeiro, L., Fischer, B.:
-%\newblock {\em {DepthK: A \textit{k}-Induction Verifier Based on Invariant Inference for C Programs}}. 
-%\newblock In: International Conference on Tools and Algorithms for the Construction and Analysis of Systems, 2017 (to appear).
-%
-%%44
-%\bibitem{Donaldson10}
-%Donaldson A., Kroening D., R{\"{u}}mmer P.:
-%\newblock {\em {{SCRATCH}: A Tool for Automatic Analysis of {DMA} Races}}.
-%\newblock In: ACM Symposium on Principles and Practice of Parallel Programming. ACM, pp. 311--312, 2011.
-%
-%%45
-%\bibitem{GrosseLD09}
-%Gro{\ss}e D., Le H., Drechsler R.:
-%\newblock {\em {Induction-based Formal Verification of {SystemC TLM} Designs}}.
-%\newblock In: International Workshop on Microprocessor Test and Verification, pp. 101--106, 2009.
-%
-%%46
-%\bibitem{David15}
-%C.~David, D.~Kroening, and M.~Lewis.
-%\newblock {\em {Using program synthesis for program analysis}}.
-%\newblock In: Logic for Programming, Artificial Intelligence, and Reasoning, LNCS 9450, pp. 483--498, 2015.
-%
-%%47
-%\bibitem{Alur13}
-%Alur R., Bod\'{\i}k R., Juniwal G., Martin M., Raghothaman M., Seshia S., Singh R., Solar-Lezama A., Torlak E., Udupa A.:
-%\newblock {\em {Syntax-guided synthesis}}. 
-%\newblock In: Formal Methods in Computer-Aided Design, pp. 1--8, 2013.
-%
-%%48
-%\bibitem{sketch}
-%A.~Solar-Lezama.
-%\newblock {\em {Program sketching}}.
-%\newblock In: International Journal on Software Tools for Technology Transfer, 15(5-6):475--495, 2013.
-%
-%%49
-%\bibitem{Sharma14}
-%R.~Sharma and A.~Aiken.
-%\newblock {\em {From invariant checking to invariant inference using randomized search}}.
-%\newblock In: Computer Aided Verification, LNCS 8559, pp. 88--105, 2014.
-%
-%%50
-%\bibitem{dsv_spin2015}
-%Ismail H., Bessa I., Cordeiro L., Lima Filho E., Chaves Filho J.:
-%\newblock {\em {DSVerifier: A Bounded Model Checking Tool for Digital Systems}}. 
-%\newblock In: International SPIN Symposium on Model Checking of Software, LNCS 9232, pp. 126--131, 2015.
-%
-%%51
-%\bibitem{Monteiro16}
-%F.~Monteiro:
-%\newblock {\em {Bounded model checking of state-space digital systems: the impact of finite word-length effects on the implementation of fixed-point digital controllers based on state-space modeling}}.
-%\newblock In: 24th {ACM} {SIGSOFT} International Symposium on Foundations of Software Engineering, pp. 1151--1153, 2016.
-%
-%%52
-%\bibitem{Bessa17}
-%Bessa, I., Ismail, H., Palhares, R., Cordeiro, L., Chaves Filho, J.:
-%\newblock {\em {Formal Non-Fragile Stability Verification of Digital Control Systems with Uncertainty}}. 
-%\newblock In: IEEE Transactions on Computers, 2017 (to appear).
-%
-%%53
-%\bibitem{selfcar}
-%Fairley P.:
-%\newblock {\em {Self-Driving Cars Have a Bicycle Problem}}.
-%\newblock http://spectrum.ieee.org/transportation/self-driving/selfdriving-cars-have-a-bicycle-problem.
-%
-%%54
-%\bibitem{BMCml}
-%L.~Bortolussi, D. Milios, G, Sanguinetti.:
-%\newblock {\em {Smoothed model checking for uncertain Continuous-Time Markov Chains}}.
-%\newblock In:. Information and Computation, 2016.
-%
-%%55
-%\bibitem{Behrend15}
-%Behrend J., Lettnin D., Gruenhage A., Ruf J., Kropf T., Rosenstiel W.:
-%\newblock {\em {Scalable and Optimized Hybrid Verification of Embedded Software}}. 
-%\newblock In: J. Electronic Testing 31(2): pp. 151--166, 2015.
-%
-%%56
-%\bibitem{Lettnin09}
-%Lettnin D., Nalla P. K., Behrend J., Ruf J., Gerlach J., Kropf T., Rosenstiel W., Sch{\"o}nknecht V., Reitemeyer S.:
-%\newblock {\em {Semiformal Verification of Temporal Properties in Automotive Hardware Dependent Software}}. 
-%\newblock In: Design, Automation and Test in Europe, pp. 1214--1217, 2009.
-%
-%%57
-%\bibitem{Adouane16}
-%Vilca J., Adouane L., Mezouar Y.:
-%\newblock {\em {Optimal Multi-Criteria Waypoint Selection for Autonomous Vehicle Navigation in Structured Environment}}. 
-%\newblock In: Journal of Intelligent and Robotic Systems 82(2): pp. 301--324, 2016.
-%
-%%58
-%\bibitem{ECBS13}
-%Ramalho M., Lopes M., Sousa F., Marques H., Cordeiro L., Fischer B.:
-%\newblock {\em {SMT-Based Bounded Model Checking of C++ Programs} }.
-%\newblock In: International Conference on Engineering of Computer-Based Systems, pp. 147--156, 2013.
-%
-%%59
-%\bibitem{Pereira15}
-%Pereira P.  Albuquerque H., Marques H., Silva I., Carvalho C., Santos V., Ferreira R., Cordeiro L.: 
-%\newblock {\em {Verifying CUDA Programs using SMT-Based Context-Bounded Model Checking}}.
-%\newblock In: ACM Symposium on Applied Computing, Software Verification and Testing Track, pp. 1648-1653, 2016. 
-%
-%%60
-%\bibitem{Sousa15}
-%Sousa F., Cordeiro L., Lima Filho E.: 
-%\newblock {\em { Bounded Model Checking of C++ Programs Based on the Qt Framework.}} 
-%\newblock In: Global Conference on Consumer Electronics, pp. 179--180, 2015.
-%
-%%61
-%\bibitem{Beyer14}
-%D.~Beyer 
-%\newblock {\em {Status report on software verification - (competition summary {SV-COMP})}}. 
-%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 8413, pp. 373--388, 2014.
-% 
-%%62
-%\bibitem{BeyerSVCOMP15}
-%D.~Beyer:
-%\newblock {\em {Software Verification and Verifiable Witnesses - (Report on SV-COMP 2015).}} 
-%\newblock In: Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 401-416, 2015.
-%
-%%63
-%\bibitem{KahlonWG09}
-%Kahlon V., Wang C., Gupta A.:
-%\newblock {\em{Monotonic Partial Order Reduction: An Optimal Symbolic Partial Order Reduction Technique}}.
-%\newblock In: International Conference on Computer-Aided Verification, LNCS 5643, pp. 398--413, 2009.
-%
-%%64
-%\bibitem{morse15}
-%J.~Morse.
-%\newblock {\em {Expressive and Efficient Bounded Model Checking of Concurrent Software}}.
-%\newblock {University of Southampton, PhD Thesis}, 2015.
-%
-%%65
-%\bibitem{Inverso14}
-%Inverso O., Tomasco E., Fischer B., La Torre S., Parlato G.:
-%\newblock {\em {Bounded Model Checking of Multi-threaded C Programs via Lazy Sequentialization.}}
-%\newblock In: International Conference on Computer-Aided Verification, LNCS 8559, pp. 585-602, 2014.
-%
-%%66
-%\bibitem{civl15}
-%Zheng M, Rogers M, Luo Z, Dwyer M, Siegel S.
-%\newblock {\em{{CIVL}: Formal Verification of Parallel Programs}}.
-%\newblock In: International Conference on Automated Software Engineering, pp. 830--835, 2015.
-%
-%%67
-%\bibitem{Grumberg05}
-%Grumberg O., Lerda F., Strichman O., Theobald M.:
-%\newblock {\em { Proof-guided underapproximation-widening for multi-process systems.}} 
-%\newblock In: Symposium on Principles of Programming Languages, pp. 122--131, 2005.
-%
-%%68
-%\bibitem{McMillan11}
-%K.~McMillan:
-%\newblock {\em {Widening and Interpolation.}} 
-%\newblock In: International Symposium on Static Analysis, LCNS 6887, pp. 1, 2011.
-%
-%%69
-%\bibitem{esbmc_controller}
-%Bessa I., Abreu R., Cordeiro L., Filho J.:
-%\newblock {\em {SMT-Based Bounded Model Checking of Fixed-Point Digital Controllers}.}
-%\newblock In: Annual Conference of the Industrial Electronics Society, pp. 295--301, 2014.
-%
-%%70
-%\bibitem{esbmc_filter}
-%Abreu R., Cordeiro L., Filho E.:
-%\newblock {\em {Verifying Fixed-Point Digital Filters using SMT-Based Bounded Model Checking}.}
-%\newblock In: XXXI Brazilian Symposium on Telecommunications, 2013.
-%\newblock http://dx.doi.org/10.14209/sbrt.2013.57
-%
-%%71
-%\bibitem{JMorse15}
-%Morse J., Cordeiro L., Nicole D., Fischer B.:
-%\newblock {\em {Model Checking LTL Properties over ANSI-C Programs with Bounded Traces.}} 
-%\newblock In: Software and System Modeling 14(1): pp. 65--81, 2015.
-%
-%%72
-%\bibitem{Bessa16}
-%Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
-%\newblock {\em { Verification of Fixed-Point Digital Controllers Using Direct and Delta Forms Realizations.}} 
-%\newblock In: Design Automation for Embedded Systems 20(2): pp. 95--126, 2016.
-%
-%%73
-%\bibitem{Chaves17}
-%Chaves L., Bessa I., Cordeiro L.:
-%\newblock {\em { DSValidator: An Automated Counterexample Reproducibility Tool for Digital Systems}}. 
-%\newblock In: CoRR abs/1610.07066, 2016.
-%
-%%74
-%\bibitem{sbrtdaniel}
-%de Mello D. P. M., de Freitas M. L., Cordeiro L., da Silva J\'unior W. S., de Bessa I. V., de Lima Filho E. B, Clavier L.:
-%\newblock {\em{Verification of Magnitude and Phase Responses in Fixed-Point Digital Filters}}. 
-%\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017.
-%
-%%75
-%\bibitem{Alves15}
-%Alves E. , Cordeiro L., Lima Filho E.:
-%\newblock {\em{ Fault Localization in Multi-Threaded C Programs using Bounded Model Checking}}. 
-%\newblock In: Brazilian Symposium on Computing Systems Engineering, pp. 96-101, 2015.
-%
-%%76
-%\bibitem{Garcia16}
-%Garcia M. , Sousa F., Cordeiro, L., Lima Filho E.: 
-%\newblock {\em{ ESBMC$^{QtOM}$: A Bounded Model Checking Tool to Verify Qt Applications}}. 
-%\newblock In: International SPIN symposium on Model Checking of Software (to appear), 2016.
-%
-%%77
-%\bibitem{Lopes:2014:GSL:2692607}
-%Bruno~Cardoso Lopes and Rafael Auler.
-%\newblock {\em Getting Started with LLVM Core Libraries}.
-%\newblock Packt Publishing, 2014.
-%
-%%78
-% \bibitem{clang-usage}
-%Cade Metz.
-%\newblock Why {A}pple?s swift language will instantly remake computer
-%  programming, 204.
-%\newblock http://www.wired.com/2014/07/apple-swift/.
-%
-%%79
-%\bibitem{Jackson16}
-%Jackson D., Vaziri M.:
-%\newblock {\em{ Correct or usable? the limits of traditional verification (impact paper award)}}. 
-%\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 11, 2016.
-%
-%%80
-%\bibitem{MorseCNF13}
-%Morse J., Cordeiro L., Nicole D., Fischer B.:
-%\newblock {\em {Handling unbounded loops with {ESBMC} 1.20}}.
-%\newblock In: Tools and Algorithms for the Construction and Analysis of
-%  Systems. LNCS 7795, Springer Berlin Heidelberg, pp. 619--622, 2013.
-%
-%%81
-%\bibitem{MorseRCN014}
-%Morse J., Ramalho M., Cordeiro L., Nicole D., Fischer B.:
-%\newblock {\em {ESBMC} 1.22.}
-%\newblock In: Tools and Algorithms for the Construction and Analysis of
-%  Systems. LNCS 8413, Springer Berlin Heidelberg, pp. 405--407, 2014.
-%
-%%82
-%\bibitem{Haran15}
-%Haran A., Carter M., Emmi M., Lal A., Qadeer S., Rakamaric Z.:
-%\newblock {\em{ SMACK+Corral: A Modular Verifier - (Competition Contribution)}}. 
-%\newblock In: 21st International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9035, pp. 451--454, 2015.
-%
-%%83
-%\bibitem{Beyer16}
-%Beyer, D.:
-%\newblock {\em{ Reliable and Reproducible Competition Results with BenchExec and Witnesses (Report on SV-COMP 2016)}}. 
-%\newblock In: 22nd International Conference Tools and Algorithms for the Construction and Analysis of Systems, LNCS 9636, pp. 887--904, 2016.
-%
-%%84
-%\bibitem{BeyerW15}
-%Beyer D., Dangl M., Dietsch D., Heizmann M., Stahlbauer A.:
-%\newblock {\em{ Witness validation and stepwise testification across software verifiers}}. 
-%\newblock In: ESEC/SIGSOFT Foundations of Software Engineering, pp. 721--733, 2015.
-%
-%%85
-%\bibitem{RochaIFM12}
-%Rocha H., Barreto R., Cordeiro L., Dias Neto A.:
-%\newblock {\em{ Understanding Programming Bugs in ANSI-C Software Using Bounded Model Checking Counter-Examples}}. 
-%\newblock In: International Conference on Integrated Formal Methods, LNCS 7321, pp. 128-142, 2012.
-%
-%%86
-%\bibitem{antennasbrt}
-%de Jesus A. S., rodrigues R. N., Ferreira A. N. G., Valente V. H., de Melo W. C., de Lima Filho E. B, da Silva J\'unior W. S.:
-%\newblock {\em{Automatic Antenna Alignment System for Satellite Receivers Operating in C and Ku Bands}}. 
-%\newblock To appear in: Brazilian Symposium on Telecommunications and Signal Processing, pp. 1-5, 2017. (in Portuguese)
-%
-%%87
-%\bibitem{amoedo}
-%Amoedo D. A., da Silva J\'unior W. S., de Lima Filho E. A.:
-%\newblock {\em{Parameter Selection for SVM in Automatic Modulation Classification of Analog and Digital Signals}}. 
-%\newblock In: International Telecommunications Symposium (ITS), pp. 1-5, 2014.
-%
-%%88
-%\bibitem{mlswverif}
-%Hamel L.:
-%\newblock {\em {On the Use of Machine Learning in Formal Software Verification.}} 
-%\newblock {Technical Report TR03-294, Dept. of Computer Science and Statistics, University of Rhode Island}, 2003.
-%
-%
-%%89
-%\bibitem{nguyen}
-%Phuc N. V.,:
-%\newblock {\em {The Application of Machine Learning Methods in Software Verification and Validation.}} 
-%\newblock {University of Texas at Austin, master of science dissertation}, 2010.
-%
-%%90
-%\bibitem{bridge}
-%Bridge J. P.:
-%\newblock {\em {Machine learning and automated theorem proving.}} 
-%\newblock {Technical Report 792, Computer Laboratory, University of Cambridge}, 2010.
-%
-% 
-% %91
-%\bibitem{babic07}
-%Hutter F., Babic D., Hoos H. H., Hu A. J.              
-%\newblock {\em{Boosting Verification by Automatic Tuning of Decision Procedures.}}
-%\newblock {In 7th International Conference Formal Methods in Computer-Aided Design {(FMCAD)}, pp. 27--34, 2007.}
-
-%\end{thebibliography}
-
 % that's all folks

### 2018-03-12 paper.tex
fixed references
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

@@
 
-\textcolor{blue}{Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.
+\textcolor{blue}{Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05}.
 This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs regarding the nonexistence of counterexamples of depth less than or equal to $k$.
@@
 This procedure is stopped when we have found a valid counterexample or proved that the final state is not reachable (i.e., the property holds).
-The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}}.
+The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05}}.
 

### 2018-03-25 IET CPS/SurveyECPS.tex
IET CPSTA

--- 2018-03-12 paper.tex
@@
-\documentclass[format=acmsmall, review=false, screen=true]{acmart}
+\documentclass{cta-author}
+
+\newtheorem{theorem}{Theorem}{}
+\newtheorem{corollary}{Corollary}{}
+\newtheorem{remark}{Remark}{}
+\newtheorem{definition}{Definition}{}
 
@@
 
-\newtheorem{prf}{\textbf{Definition}}
+%\newtheorem{prf}{\textbf{Definition}}
 
@@
 
-% Metadata Information
-%\acmJournal{TCPS}
-%\acmVolume{9}
-%\acmNumber{9}
-%\acmArticle{99}
-%\acmYear{2017}
-%\acmMonth{12}
-%\copyrightyear{2017}
-%\acmArticleSeq{9}
-
-% Copyright
-%\setcopyright{acmcopyright}
-\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{0000001.0000001}
-
-% Paper history
-\received{August 2017}
-%\received[revised]{March 2009}
-%\received[accepted]{June 2009}
-
-\lstset{language=C,basicstyle=\small}
-\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
-\lstset{tabsize=2}
-\lstset{firstnumber=1}
-\lstset{frame=single}
-\lstset{
-  language={C},
-  morekeywords={stdio,math,printf,assume,then,end-if,end,input,output}
-}
-
-\usepackage{color}
-\usepackage{listings}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-%\renewcommand{\baselinestretch}{0.93}
-\renewcommand{\baselinestretch}{0.97}
-
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+
+%\supertitle{Brief Paper}
+
 \title{Automated Verification and Synthesis of Cyber-Physical Systems: Advances and Challenges}
 
-\author{Lucas C. Cordeiro}
-\orcid{0000-0002-6235-4272}
-\affiliation{%
-  \institution{Department of Computer Science, University of Oxford}
-%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
-  \city{Oxford}
-%  \state{VA}
-%  \postcode{23185}
-  \country{UK}}
-
-\author{Eddie B. de Lima Filho}
-\orcid{0000-0002-2758-4638}
-\affiliation{%
-  \institution{TPV Technology}
-%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
-  \city{Manaus}
-%  \state{VA}
-%  \postcode{23185}
-  \country{Brazil}}
-
-\author{Iury V. de Bessa}
-\orcid{0000-0002-2758-4638}
-\affiliation{%
-  \institution{Department of Electricity, Federal University of Amazonas}
-%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
-  \city{Manaus}
-%  \state{VA}
-%  \postcode{23185}
-  \country{Brazil}}
-
-%% author names and affiliations
-%% use a multiple column layout for up to three different
-%% affiliations
-%\author{Eddie B. de Lima Filho \\
-%%Department of Computer Science\\
-%University of Oxford, UK\\
-%lucas.cordeiro@cs.ox.ac.uk
-%\and
-%Eddie B. de Lima Filho \\
-%Samsung Electronics, Brazil \\
-%eddie.l@samsung.com
-%\and
-%Iury V. de Bessa \\
-%Federal University of Amazonas, Brazil \\
-%iury.bessa@ufam.edu.br}
-%\IEEEauthorblockN{Homer Simpson}
-%\IEEEauthorblockA{Twentieth Century Fox\\
-%Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\and
-%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-%\IEEEauthorblockA{Starfleet Academy\\
-%San Francisco, California 96678--2391\\
-%Telephone: (800) 555--1212\\
-%Fax: (888) 555--1212}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% make the title area
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
+\author{\au{Lucas C. Cordeiro$^{1\corr}$} \au{Eddie B. de Lima Filho$^2$} \au{Iury V. de Bessa$^3$}}
+
+\address{\add{1}{Department of Computer Science, University of Oxford, Oxford, United Kingdom}
+\add{2}{TPV Technology, Manaus, Brazil}
+\add{3}{Department of Electricity, Federal University of Amazonas, Manaus, Brazil}
+\email{lucasccordeiro@gmail.com}}
+
 \begin{abstract}
-Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
+\looseness=-1 Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics.
 Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors.
@@
 
-%\IEEEpeerreviewmaketitle
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10010405.10010432.10010439.10010440</concept_id>
-<concept_desc>Applied computing~Computer-aided design</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10010520.10010553</concept_id>
-<concept_desc>Computer systems organization~Embedded and cyber-physical systems</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10010147.10010178.10010213.10010214</concept_id>
-<concept_desc>Computing methodologies~Computational control theory</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002944.10011122.10002945</concept_id>
-<concept_desc>General and reference~Surveys and overviews</concept_desc>
-<concept_significance>100</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Applied computing~Computer-aided design}
-\ccsdesc[500]{Computer systems organization~Embedded and cyber-physical systems}
-\ccsdesc[300]{Computing methodologies~Computational control theory}
-\ccsdesc[100]{General and reference~Surveys and overviews}
-
-\keywords{Embedded Systems, Cyber-Physical Systems, Formal Methods, Model Checking, Machine Learning}
-
 \maketitle
@@
 
+{\color{blue}
 Figure~\ref{intelligent-product} shows some examples of ECPSs, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}.
 Indeed, many current ECPSs, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks.
-For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
+For instance, UAVs are a trend on military missions and civil applications, since they can easily achieve places that cannot be accessed by humans without on-board pilots with different degrees of autonomy; however, an incorrect plan execution may cost civilian lives, which is unacceptable.
 In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences.
@@
 On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health.
+}
 %
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figure1.jpg}
+	\includegraphics[width=\columnwidth]{figure1.jpg}
 	\caption{An ECPS is part of a well-specified larger one (intelligent product).}
@@
 
+{\color{blue}
 Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may change based on data acquired during previous actions, due to model updating.
-As a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data, as recently reported in the literature~\cite{DBLP:conf/cav/HuangKWW17,DBLP:conf/cav/KatzBDJK17}, where formal verification techniques are applied to verify deep neural networks.
+Indeed, ML algorithms have a dual role on software verification development.
+With the dissemination of the use of these algorithms, several ECPS have embedded ML software, and as a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data, as recently reported in the literature~\cite{DBLP:conf/cav/HuangKWW17,DBLP:conf/cav/KatzBDJK17}, where formal verification techniques are applied to verify deep neural networks.
+Otherwise, model checking mechanisms have also incorporated ML algorithms to increase the efficiency of verification~\cite{BUCCAFURRI199957,Bortolussi2015323,Braazdil201498} and synthesis~\cite{Jha2017693} procedures.
+}
 
@@
 
-The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML.
+{\color{blue}
+The present article is an survey study whose main goals are described as follows:
+\begin{itemize}
+	\item briefly present part of the established research about ECPS modelling, verification and synthesis;
+	\item provide an overview about the technical issues and advances in symbolic formal verification and synthesis methodologies that are currently employed for ECPS;
+	\item present the open challenges on the ECPS verification and synthesis and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. 
+\end{itemize}
+
+The remaining of this survey is organized as follows.
 We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}.
@@
 Finally, we conclude and describe future work, in Section~\ref{conclusions}.
+}
 
@@
 
+{\color{blue}
+
 Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as provide controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.
-
-An ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
-
-\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho...
+The reliability of an ECPS is related to its ability to maintain the continuity of correct execution of specific function.
+In particular, the reliability of an ECPSs depends on the correctness of their embedded software and the compatibility with hardware.
+The correctness is asserted with respect to a given specification, that may be a safety specification or a reachability specification~\cite{Jhala:2009:SMC:1592434.1592438}.
+A safety specification is related to  a set of undesirable states that must be avoided by the system, and a reachability is related to a desired state, that must be achieved.
+In both cases, it is needed a model to compute the reachability of ECPS an than decide about the correctness and reliability.
+
+A general model for a ECPS is defined as follows.
+
+\begin{definition}
+	An ECPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
+\end{definition}
+
+%\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho...
 parece que estamos misturando tudo...
 sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo um pouco nebuloso?}
-The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
+
+The synthesis of reliable ECPS is related to the synthesis of controllers that are able to handle the complex dynamic of these systems.
+Indeed, the so-called ECPS can be considered HS, since they present a mixture of continuous and discrete dynamics.
+Thus, the control synthesis for ECPS is based in the hybrid control system theory, that is still in development, since the control theory for continuous and discrete systems were developed without connexions~\cite{lunze_lamnabhi-lagarrigue_2009}.
+The formal synthesis of ECPS includes but is not limited to the control synthesis problem, and must consider the formal verification of correctness and reliability.
+
+Summarizing, the overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.
+
+}
+
+%The classification of approaches in terms of their deterministic vs stochastic nature should be definitely revised: a model is non-deterministic when it accepts different behaviors without providing a measure of probability (e.g.
+this is the case of Timed Automata, which are instead mentioned as a deterministic model); a model is stochastic when the space of behaviors is provided with a measure of probability.
+By the way, a large literature now develops on mixed models that combine non-deterministic and stochastic behavior (e.g.
+Markov Decision Processes), but this class of models is not even mentioned.
 
@@
 %--------------------------------------
+{
+\color{blue}
 
@@
 Specially, deterministic and stochastic models are the basis of many recent engineering advances.
-The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}.
+The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness which is described by probability distributions, i.e., some of these variable are random variables whose behavior are described by probability density functions~\cite{stochmodel}.
 Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled, and they should be better represented by nondeterministic approaches~\cite{leeCPS2}.
+Non-deterministic models is able to model different behaviors without providing a measure of probability, i.e., their variables are assumed to be different values but they are not random variables.
 Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
+}
 
@@
 
-Finally, in addition to ECPS-dynamics modeling, it is often interesting to establish invariants on a ECPS's state space, which can be used for accelerating formal verification and synthesis processes regarding those systems.
+{\color{blue}
+In addition to ECPS-dynamics modeling, it is often interesting to establish invariants on a ECPS's state space, which can be used for accelerating formal verification and synthesis processes regarding those systems.
 Such invariants might be synthesized, in order to point out state-space boundaries~\cite{BenSassi12} and determine ECPS stability~\cite{Paul14} or reachability~\cite{Li16_2,Li16}.
+Finally, it is worth to mention that there are still some recent modelling approaches that are mixtures of stochastic and non-deterministic models for ECPS, e.g., stochastic nondeterministic automata~\cite{Fisher2014}, and piecewise-deterministic Markov chains~\cite{Julius09}.
+}
 
@@
 
+{\color{blue}
+
+An ECPS verification problem is defined as follows:
+
+\begin{definition}
 	Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
+\end{definition}
+
  Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}.
@@
 Stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification} or linear system stability theory for discrete simulations~\cite{Bessa17}.
+}
 
@@
 Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}.
-Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
+Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{ARAUJO2017,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
 
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figure2.jpg}
+	\includegraphics[width=\columnwidth]{figure2.jpg}
 	\caption{Bounded Model Checking.}
@@
 	\centering
-\begin{minipage}{0.4\textwidth}
+	\begin{minipage}{\columnwidth}
 		\begin{lstlisting}[basicstyle=\footnotesize]
@@
 
-\begin{prf}
+%\begin{prf}
 	\begin{definition}\label{interpolants}
@@
 	\end{definition}
-\end{prf}
+%\end{prf}
 
@@
 	\centering
-\includegraphics[width=\textwidth]{interpolants}
+	\includegraphics[width=\columnwidth]{interpolants}
 	\caption{Computing image by interpolation~\cite{McMillan03}.}
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{figure3.jpg}
+	\includegraphics[width=0.7\columnwidth]{figure3.jpg}
 	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
@@
 		\centering
-\resizebox{.5\textwidth}{!}
+		\resizebox{\columnwidth}{!}
 		{
@@
 Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware.
-Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering FWL effects.
+Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering FWL effects.
 Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{figure4.jpg}
+	\includegraphics[width=\columnwidth]{figure4.jpg}
 	\caption{Verification methodologies for embedded systems.}
@@
 Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, Multiple-Input Multiple-Output).
-In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}.
+In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{ARAUJO2017}.
 As a consequence, verifiers and synthesizers need to handle a wide range of functions, including non-linear and non-convex optimization problems based on both fixed- and floating-point arithmetic.
@@
 		As already mentioned in section \ref{sec:model}, problems that are not directly related to code correctness may be tackled with formal methods, as long as they are conveniently modeled.
-Ara\'ujo {\it et al.} \cite{Araujo16} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
+Ara\'ujo {\it et al.} \cite{ARAUJO2017} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.
@@
 This way, by using an SMT-based verification method, one may map the TS error indication and other performance figures as properties and then recursively perform realignment procedures, until a counterexample is not output.
-It is worth noticing that associated timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{Araujo16}, which can be seen as an advantage for such an implementation.
+It is worth noticing that associated timing requirements are more relaxed than what was mentioned in the work developed by Ara\'ujo {\it et al.}~\cite{ARAUJO2017}, which can be seen as an advantage for such an implementation.
 
@@
 A system that is more robust is less fragile, but it is still prone to problems, which depends on operating scenarios, use cases, and uncertainties.
+Keel and Bhatactacharyya~\cite{bhattacharyya} showed that robust and optimal control system can still be fragile with respect to the implementation issues.
 			In summary, if the chosen model does not tackle every possible problem or error that might occur in a given environment, system-failure events are still possible, which is true even for the Ariane 5's failure \cite{antifragile}.}
@@
 		\textcolor{blue}{As a result, it may be more interesting to develop methodologies that are not simply robust, regarding predetermined conditions, but instead thrive on problems and grow stronger (as the human body, which gets stronger when exposed to germs), which leads to antifragile systems.
-In engineering, we are only beginning to develop solutions that can be classified as antifragile, such as multiple-input/multiple-output (MIMO) systems, which use multiple signal copies together with multipath to achieve performance enhancement~\cite{spectrum2013}.
-In that case, signal copies would initially result in a destructive effect; however, with the use of a methodology that takes advantage on them, the net result is performance improvement.}
+In engineering, we are only beginning to develop solutions that can be classified as antifragile, that can be achieved via analytical or physical redundancy to achieve performance enhancement~\cite{spectrum2013}.
+For example, multiple-input/multiple-output (MIMO) systems, which use multiple signal copies together with multipath signal copies would initially result in a destructive effect; however, with the use of a methodology that takes advantage on them, the net result is performance improvement.}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
+%\section{Acknowledgments}
+%
+%Shan Cheng is indebted to the China Scholarship Council (CSC)
+%for its financial support as a visiting scholar at the University of
+%Technology, Sydney. This work was partially supported by the
+%National Science Foundation of China (Grant no. 10972129), the
+%Innovation Program of Shanghai Municipal Education Commission (Grant
+%no. 10ZZ61), the Shanghai Leading Academic Discipline Project (Project no.
+%S30106), and the Graduate Innovation Foundation of Shanghai
+%University (Project no. SHUCX 101080).
+
+%\begin{thebibliography}{99}
+%
+%\bibitem{1}
+%Vicsek T., Czir\'{o}k A., Ben-Jacob E., Cohen I.: `Novel type of phase transition in a system of self-driven particles', \textit{Phys. Rev. Lett.}, 1995, \textbf{75}, pp.~1226--1229
+%
+%\bibitem{2}
+%Jadbabaie A., Lin J., Morse A.S.: `Coordination of groups of
+%mobile autonomous agents using nearest neighbor rules', \textit{IEEE
+%Trans. Autom. Control}, 2003, \textbf{48}, pp.~988--1001
+%
+%\bibitem{3}
+%Olfati-Saber R., Murray R.M.: `Consensus problems in networks
+%of agents with switching topology and time-delays', \textit{IEEE Trans.
+%Autom. Control}, 2004, \textbf{49}, pp.~1520--1533
+%
+%\bibitem{4}
+%Ren W., Atkins E.: `Distributed multi-vehicle coordinated
+%control via local information exchange', \textit{Int. J. Robust
+%Nonlinear Control}, 2007, \textbf{17}, pp.~1002--1033
+%
+%\bibitem{5}
+%Ren W.: `On consensus algorithms for double-integrator dynamics',
+%\textit{IEEE Trans. Autom. Control}, 2008, \textbf{53}, pp.~1503--1509
+%
+%\bibitem{6}
+%Xie G., Wang L.: `Consensus control for a class of networks of
+%dynamic agents', \textit{Int. J. Robust Nonlinear Control}, 2007,
+%\textbf{17}, \hbox{pp.~941--959}
+%
+%\bibitem{7}
+%Gao Y., Wang L.: `Consensus of multiple dynamic agents with sampled information', \textit{IET Control
+%Theory Appl.}, 2010, \textbf{4}, pp.~945--956
+%
+%\bibitem{8}
+%Tanner H.G., Jadbabaie A., Pappas G.J.: `Flocking in fixed
+%and switching networks', \textit{IEEE Trans. Autom. Control}, 2007,
+%\textbf{52}, \hbox{pp.~863--868}
+%
+%\bibitem{9}
+%Lee D.J., Spong M.W.: `Stable flocking of multiple inertial
+%agents on balanced graphs', \textit{IEEE Trans. Autom. Control}, 2007,
+%\textbf{52}, \hbox{pp.~1469--1475}
+%
+%\bibitem{10}
+%Hu J., Lin Y.S.: `Consensus control for multi-agent systems with double-integrator dynamics and time delays', \textit{IET Control
+%Theory Appl.}, 2010, \textbf{4}, pp.~109--118
+%
+%\bibitem{11}
+%Lin P., Jia Y.: `Further results on decentralised coordination
+%in networks of agents with second-order dynamics', \textit{IET Control
+%Theory Appl.}, 2009, \textbf{3}, pp.~957--970
+%
+%\bibitem{12}
+%Lin P., Jia Y.: `Robust $H_{\infty}$ consensus analysis of a class of second-order multi-agent systems with uncertainty', \textit{IET Control
+%Theory Appl.}, 2010, \textbf{4}, pp.~487--498
+%
+%\bibitem{13}
+%He W., Cao J.: `Consensus control for high-order multi-agent
+%systems', \textit{IET Control Theory Appl.}, 2011, \textbf{1}, pp.~231--238
+%
+%\bibitem{14}
+%Hui Q., Haddad W.M.: `Distributed nonlinear control algorithms
+%for network consensus', \textit{Automatica}, 2008, \textbf{44}, pp.~2375--2381
+%
+%\bibitem{15}
+%Liu X., Chen T., Lu W.: `Consensus problem in directed
+%networks of multi-agents via nonlinear protocols', \textit{Phys. Lett.
+%A}, 2009, \textbf{373}, pp.~3122--3127
+%
+%\bibitem{16}
+%Zheng Y., Zhu Y., Wang L.: `Consensus of heterogeneous multi-agent systems', \textit{IET Control
+%Theory Appl.}, 2011, \textbf{5}, pp.~1881--1888
+%
+%\bibitem{17}
+%Zhang W., Zeng D., Qu S.: `Dynamic feedback consensus control of a class of high-order multi-agent systems',
+%\textit{IET Control Theory Appl.}, 2010, \textbf{4}, pp.~2219--2222
+%
+%\bibitem{18}
+%Su H.S., Wang X.F., Lin Z.L.: `Flocking of multi-agents with
+%a virtual leader', \textit{IEEE Trans. Autom. Control}, 2009,
+%\textbf{54}, pp.~293--307
+%
+%\bibitem{19}
+%Song Q., Cao J., Yu W.: `Second-order leader-following
+%consensus of nonlinear multi-agent system via pinning control', \textit{Syst. Control Lett.}, 2010, \textbf{59}, pp.~553--562
+%
+%\bibitem{20}
+%Hong Y.G., Chen G.R., Bushnell L.D.: `Distributed observers
+%design for leader-following control of multi-agent networks', \textit{Automatica}, 2008, \textbf{44}, pp.~846--850
+%
+%\bibitem{21}
+%Hu J., Hong Y.: `Leader-following coordination of multi-agent
+%systems with coupling time delays', \textit{Physica A}, 2007,
+%\textbf{374}, pp.~853--863
+%
+%\bibitem{22}
+%Ren W.: `Consensus strategies for cooperative control of vehicle
+%formations', \textit{IET Control Theory Appl.}, 2007, \textbf{1}, pp.~505--512
+%
+%\bibitem{23}
+%Khalil H.K.: `Nonlinear systems' (Prentice-Hall, New
+%Jersey,  2002, 3rd edn.)
+%
+%\end{thebibliography}
+
+\bibliographystyle{iet}
+
 \bibliography{refs}
 
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-
-%The authors would like to thank...
-%The authors thank M. Dangl for reviewing a draft version of this paper.~This research was supported by CNPq $475647$/$2013$-$0$ grant.
-
-% that's all folks
 \end{document}

### 2018-03-25 paper.tex
IET CPSTA

@@
 As already mentioned in section \ref{sec:model}, problems that are not directly related to code correctness may be tackled with formal methods, as long as they are conveniently modeled.
-Ara\'ujo {\it et al.} \cite{Araujo16} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
+Ara\'ujo {\it et al.} \cite{ARAUJO2017} employed SMT for handling non-convex optimization problems, with the goal of ensuring optimal solutions.
 Indeed, it is possible to recursively re-constrains a model checking procedure, by using the current counterexample as a seed for the next run, until a global minimum is achieved.

### 2018-03-25 IET CPS/example.c

@@ -1,0 +1,13 @@
+#include <assert.h>
+int main(void) {
+  unsigned int i=0, n;
+  float x=3;
+  __VERIFIER_assume(n>1);
+  while (i<=n) {
+    x = ((2*x) - 1);
+    i++;
+  }
+  assert(x>0);
+  return 0;
+}
+

### 2018-03-25 IET CPS/example2.c

@@ -1,0 +1,20 @@
+#include <stdio.h>
+#include <assert.h>
+#include <math.h>
+int main(void) {
+ unsigned int i=1;
+ float x=0.9;
+ while (1) {
+   x = ((2*x) - 1);
+   if(isinf(x)) {
+     printf("\%d\n", i);
+     break;
+   }
+   printf("iii: \%d\n", i);
+   assert(x>0);
+   ++i;
+ }
+ assert(x>0);
+ return 0;
+}
+
